### 2018-06-29 Introduction.tex
Created the project!

\section{Introduction}
Let $\sg$ be a smooth orientable surface of genus $g$ with one boundary component $\partial\sg$, and let $\sgm$ be $\sg$ with
a choice of $m$ distinct points in the interior, called \emph{punctures}.

Let $\gg$ be the mapping class group of $\sg$, i.e. the group of isotopy classes of diffeomorphisms of $\sg$:
diffeomorphisms are required to fix $\partial\sg$ pointwise; similarly let $\ggm$ be the mapping class group of $\sgm$, i.e.
the group of isotopy classes of diffeomorphisms of $\sgm$ that fix $\partial\sgm$ pointwise and \emph{permute} the $m$ punctures.

Forgetting punctures yields a surjective map $\ggm\to\gg$ with kernel $\bms$,
the $m-$th \emph{braid group} of the surface $\sg$: we get the Birman exact sequence
\begin{equation}
\label{eqn:Birman}
1\to\bms\to\ggm\to\gg\to 1.
\end{equation}
% This short exact sequence corresponds to 
% \begin{equation}
%  \cms\to\mgm\to\mg.
% \end{equation}
% Here $\cms$ is the \emph{unordered} configuration space of $m$ points in the interior of $\Sigma$, i.e.
% \begin{equation}
%  \cms=\fms/\S_m=\set{(p_1,\dots,p_m)\in\pa{\mathring\sg}^m|p_i\neq p_j \;\forall i\neq j}/\S_m,
% \end{equation}
% the quotient of the \emph{ordered} configuration space of $m$ points in the interior of $\Sigma$ by the natural
% action of the symmetric group $\S_m$ that permutes the (ordered list of) points in the configuration; $\mg$ and $\mgm$
% are the moduli spaces of Riemann structures on $\sg$ and $\sgm$ respectively.

The associated Leray-Serre spectral sequence in rational homology has a second page $E^2_{p,q}=H_p(\gg;H_q(\bms;\Q))$,
and converges to $H_{p+q}(\ggm;\Q)$.

Our main theorem is that the spectral sequence has already collapsed in its second page.
\begin{thm}
\label{thm:main}
For each $l\geq 0$ there is an isomorphism of $\Q$-vector spaces
\begin{equation}
\label{eqn:main}
 H_l(\ggm;\Q)\cong \bigoplus_{p+q=l} H_p(\gg;H_q(\bms;\Q)).
\end{equation}
\end{thm}
Thus the computation of $H_*(\ggm;\Q)$ reduces to the computation of the homology of $\gg$ with
twisted coefficients in the representation $H_*(\bms;\Q)$: the action of $\gg$ is quite explicit
and in particular is symplectic. The group $\gg$ is significantly smaller than $\ggm$, so the
computation of the right-hand side of equation \ref{eqn:main} should be more tractable. From now on $\Q$-coefficients will always be understood.

### 2018-06-29 MAIN.tex
Created the project!

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage{graphicx}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}

\input{macros.tex}

\begin{document}

\title{Splitting of the homology of the puntured mapping class group.}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

%\keywords{Topological complexity, aspherical spaces, Lusternik-Schnirelmann category, cohomological dimension, topological robotics, braid groups}
%\subjclass[2010]{55M99, 55P20 (Primary); 55M30, 20J06, 68T40 (Secondary).}

\begin{abstract}
Let $\ggm$ be the mapping class group of an orientable surface of genus $g$ with one parametrised
boundary component and $m$ permutable punctures; let $\gg=\gg^0$; let $\bms$ be the braod group 
of the surface $\sg$ on $m$ strands; let $\F$ be $\Z_2$ or $\Q$. We prove that $H_*(\ggm;\F)\cong H_*(\gg;H_*(\bms;\F))$.
We compute (?) the right hand side explicitly for $g=1$.
\end{abstract}

%\thanks{}

\maketitle

\input{Introduction.tex}

\input{Preliminaries.tex}

\input{HBraidSurf.tex}

\input{ThmModTwo.tex}

\input{Sink.tex}

\input{ThmQ.tex}

\input{GenusOne.tex}

\input{Conclusions.tex}

\bibliography{bibliography.bib}{}
\bibliographystyle{plain}

\end{document}

### 2018-10-19 HBraidSurf.tex
Changing the cmstwo construction to one that involves the sink

\section{Homology of configuration spaces of surfaces}
In this section we compute the homology $H_*(\bms)=H_*(C_m(\S))$; we
are interested in these groups as they appear in the Leray-Serre spectral sequence associated to
the Birman exact sequence \ref{eq:Birman} or, which is the same, to the bundle
\ref{eq:Birmanbundle}. In particular we are interested in $H_*(\bms)=H_*(C_m(\S))$
\emph{as a $\Z_2-$representation of the group $\gg$.} From now on $\Z_2-$coefficients in homology
and cohomology will always be understood.

In \cite{LM} L\"offler and Milgram proved that this must be a symplectic
representation of the mapping class group. By symplectic we mean the following:
\begin{defn}
 \label{def:symplrep}
 Let $\H=H_1(\S)\simeq\Z_2^{2g}$.
 The natural action of $\gg$ on $\H$ induces a surjective map
 $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{symplectic}
 if it is a pull-back of a representation of $Sp_{2g}(\Z_2)$ along this map.
\end{defn}

In \cite{BCM} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\S))$ \emph{as a graded $\Z_2-$vector space}.
Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced, as
their descripition of $H_*(C_m(\S))$ depends on a handle decomposition of $\S$.

We prove here the following theorem that, to the best of the author's knowledge, doesn't appear
in the literature.
\begin{thm}
 \label{thm:Hbms*as*ggrep}
 There is an isomorphism of bigraded $\Z_2-$representations of $\gg$
 \[
  H_*\pa{\coprod_{m\geq 0} C_m(\S)}\simeq \Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]\otimes\Sym_{\bullet}(\H).
 \]
 Here we mean the following:
 \begin{itemize}
  \item the bigrading is given on left by homological degree $*$ and by the connected component
  in the disjoint union on which the homology class is supported, i.e. by the number $m$ of points
  involved in constructing the homology class; we call $*$ the \emph{degree} and $m$ the \emph{weight};
  \item for $i\geq 0$ $Q^i\epsilon$ is the image in $H_{2^i-1}(C_{2^i}(\S))$ of a generator
  of the group $H_{2^i-1}(C_{2^i}(\Sigma_{0,1}))\simeq \Z_2$ 
  under the natural map induced by an embedding $\Sigma_{0,1}\hookrightarrow\S$,
  and $\Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]$ is the polynomial ring on
  infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
  \item $\H$ is identified with $H_1(C_1(\S))$ in a natural way, and $\Sym_{\bullet}(\H)$ is the
  symmetric algebra on $\H$;
  \item degrees and weights are extended on right by the usual multiplicativity-additivity rule;
  \item the action of $\gg$ on right is the diagonal action on the product: it is a trivial action
  on the factor $\Z_2[Q^i\epsilon\,|\, i\geq 0]$ and it is the symplectic action induced by the action on $\H$ on the
  factor $\Sym_{\bullet}(\H)$.
  \end{itemize}

  Notice that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
  the degree.
\end{thm}
For the rest of the section we will write $\S=\S$.

\subsection{Cohomology of $C_m(\gg)$.} Our first aim is to compute the \emph{cohomology} $H^*(C_m(\gg))$ as a graded
$\Z_2-$vector space (not as a $\gg-$representation yet); we will mimic the approach of Fuchs (\cite{Fuchs:CohomBraidModtwo}).
We will turn back to homology of $\cms$ in the next subsection.

As already said this computation recovers a known result, but it has the advantage of
being quite elementary and of providing a part of
the geometric insight that we will need later.

We first introduce a space $\T(\S)$ which is homeomorphic to the interior of $\S$. The construction
corresponds to a handle decomposition of $\S$ with one $0-$handle and $2g$ $1-$handles.
\begin{defn}
\label{def:Tsg}
If $g=0$, hence $\S=\sg$ is the disc, we set $\T(\S)=(0,1)^2$, the interior of the unit square. Assume
now $g\geq 1$.

Dissect the interval $[0,1]$ into $2g$ equal subintervals through the points $P_i=\frac{i}{2g}$ for $0\leq i\leq 2g$
(for $i=0,2g$ we get the two endpoints of $[0,1]$).

Let $Q$ be the subspace of $[0,1]^2$ obtained by the union of $(0,1)^2$ and the vertical open intervals
$I_i^l=\set{0}\times (P_i,P_{i+1})$ and $I_i^r=\set{1}\times (P_i,P_{i+1})$ for $1\leq i\leq 2g$;
equivalently, we remove from $[0,1]^2$ the two horizontal sides
$[0,1]\times\set{0,1}$ and the $4g-2$ points $\set{0,1}\times \set{P_1,\dots,P_{2g-1}}$.

Notice that all intervals $I_i^l$'s and $I_j^r$'s are
canonically diffeomorphic
to $(0,1)$ by projecting on the second coordinate, rescaling linearly by $2g$
(without changing orientation!) and translating; therefore we will specify a bijection
between the two sets of left and right intervals, and then $\T(\S)$ will be obtained from $Q$
by identifying in the canonical way the two intervals in each couple.

For $1\leq i\leq g$, we identify $I^l_{2i-1}$ with $I^r_{2i}$, obtaining an open interval $\U_i\subset\T(\S)$,
and we identify $I^r_{2i-1}$ with $I^l_{2i}$, obtaining an open interval $\V_i\subset\T(\S)$.
Each interval $\U_i,\V_i$ has a natural parametrisation by $(0,1)$.
\end{defn}
The space $\T(\S)$ is homeomorphic to $\mrS$; from now
on we will identify the two open surfaces and in particular we will identify $\cms$ with the
space of configurations of $m$ points in $\T(\S)$.

We consider the one-point-compactification $\cms^c$ of $\cms$; since the space $\cms$ is the interior of a compact
$2m-$manifold with boundary, by Poincaré-Lefschetz
duality we have
\[
 H^*(\cms)\simeq \tilde H_{2m-*}(\cms^c).
\]
Our next aim is to define a cell structure on the space $\cms^c$.
\begin{defn}
\label{def:ehopen}
A \emph{tuple} $\tup$ is a choice of the following set of data:
 \begin{itemize}
  \item a natural number $0\leq l\leq m$;
  \item integers $x_1,\dots,x_l\geq 1$;
  \item integers $u_1,\dots,u_g$ and $v_1,\dots,v_g$, all $\geq 0$.
 \end{itemize}
with the condition that
\[
 m=\sum_{i=1}^lx_i+\sum_{i=1}^g(u_i+v_i).
\]
% We will generically use the letter $m$ to denote any of the numbers $l$, $x_i$,$y_i$,$z$,$u_i$,$v_i$ or $w_i$
% in the above sum, so $m\in\Z$; instead .
% Similarly we will use the letter $\M$ to denote any of the intervals $\U_i,\V_i,\W_i$;
% they will correspond to the letters $U_i,V_i,W_i$, so we also define a subset
% $\bsymb=\set{(U_i,V_i)_{i\leq g},(W_i)_{i\leq n-1}}\subset\symb$.

The \emph{degree} of $\tup$ is defined as $m+l$.

For a tuple $\tup$ let $e^{\tup}$ be the subspace
of $\cms$ of configurations of $m$ points in $\mrS$ such that
\begin{itemize}
 \item for all $1\leq i\leq g$, exactly $u_i$ points lie on $\U_i$
 and exactly $v_i$ points lie on $\V_i$;
 \item there are exactly $l$ vertical lines in $(0,1)^2\subset\mrS$ of the
 form $\set{s_i}\times(0,1)$ for some $0<s_1<\dots<s_l<1$, containing at least one
 point of the configuration. From left to right, these lines contain exactly $x_1,\dots,x_l$ points
 respectively.
\end{itemize}
\end{defn}
The space $e^{\tup}$ is homeomorphic to \emph{the interior} of the following product of simplices:

% We want to show that the collection of the $e_h$ for varying $h$, together with
% the 0-cell $*$, give a cell decomposition of $C_k(\S)^c$. Denote by $\Delta^h$ the following product of simplices
\[
 \Delta^{\tup}\colon =
 \Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}},
%  =\prod_{m\in\symb_l}\Delta^m.
\]
where the simplex $\Delta^r$ is the subspace of $[0,1]^r$ of sequences $0\leq \tau_1\leq\dots\leq\tau_r\leq 1$
(the numbers $\tau_1,\dots,\tau_r$ are the \emph{local coordinates} of the simplex). The homeomorphism is given
as follows:
\begin{itemize}
 \item the local coordinates of the $\Delta^l-$factor correspond to the positions $s_1,\dots,s_l$ of the vertical
 lines in $(0,1)^2$ containing points of the configuration;
 \item the local coordinates of the $\Delta^{x_i}-$factor correspond to the positions of the $x_i$ points
 lying on the vertical line $\set{s_i}\times(0,1)$;
 \item the local coordinates of the $\Delta^{u_i}-$factor correspond to the positions of the $u_i$ points
 lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}-$factor,
 with $v_i$ and $\V_i$ instead of $\U_i$ and $u_i$.
\end{itemize}
Notice that $\Delta^{\tup}$ has dimension equal to the degree of $\tup$.
This embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^c$ extends to a continuous map
$\phi^{\tup}\colon\Delta^{\tup}\to\cms^c$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
subspaces $e^{\tup'}$ for tuples $\tup'$ with \emph{lower} degree than $\tup$, together with the point
at infinity.

The construction of the map $\phi^{\tup}$ is as follows:
\begin{enumerate}
\item we see the one-point-compactification
$\mrS^c$ of $\mrS=\T(\S)$ as the quotient of $[0,1]^2$ that identifies left and right intervals as in
in definition \ref{def:Tsg} of $\T(\S)$, and moreover collapses $[0,1]^2\setminus Q$ to a point, which
is the point at infinity;
\item we consider the $m-$fold symmetric product $\SP^m(\mrS^c)$: it contains 
as an \emph{open} subspace $\cms$, so we can see $\cms^c$ as the quotient of $\SP^m(\mrS^c)$ by
the subspace $\SP^m(\mrS^c)\setminus\cms$;
\item the homeomorphism $\mathring{\Delta}^{\tup}\to e^{\tup}\subset\cms$ extends now to a map
$\Delta^{\tup}\to \SP^m(\mrS^c)$, that we can then further project to $\cms^c$: the composition
is the map $\phi^{\tup}$.
\end{enumerate}

Therefore the collection of the $e^{\tup}$, together with the $0-$cell given by the point at infinity,
give a cell decomposition of $\cms^c$, with characteristic maps of cells $\phi^{\tup}$.

We can easily use these characteristic maps to compute the \emph{reduced} cellular chain complex of $\cms^c$
with coefficients in $\Z_2$, which
reads as follows.
\begin{lem}
\label{lem:doperatoropenmodtwo}
Let $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq 2g})$ and
$\tup'=(l-1,(x'_i)_{i\leq l-1},(u'_i,v'_i)_{i\leq 2g})$
be tuples in consecutive degrees $m+l$ and $m+l-1$, and let $[\partial \tup\colon\tup']\in\Z_2$ denote the coefficient
of $\tup'$ in $\partial \tup$ in the chain complex $\tilde\Ch_*(\cms^c)$.
Then $[\partial \tup:\tup']=0$ unless one (and exactly one) of the following situations occur:
\begin{itemize}
 \item $l\geq 2$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, setting $x'_i=x_i+x_{i+1}$
 for one value $1\leq i\leq l-1$, and shifting the values
 $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case
 \[
  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}.
 \]
 This binomial coefficient may be anyway equal to $0\in\Z_2$.
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
 in integers $\delta u_i,\delta v_i\geq 0$
 \[
  x_1=\sum_{i=1}^g \pa{\delta u_i+\delta v_i},
 \]
 setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and shifting the values
 $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\pa{\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}}.
 \]
 This binomial coefficient may be anyway equal to $0\in\Z_2$.
 \item $l\geq 1$ and $\tup h$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
 in integers $\delta u_i,\delta v_i\geq 0$
 \[
  x_l=\sum_{i=1}^g \pa{\delta u_i+\delta v_i},
 \]
 setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
 Also in this case
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\pa{\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}}.
 \]
 This binomial coefficient may be anyway equal to $0\in\Z_2$.
\end{itemize}
\end{lem}
\begin{proof}
 If we restrict the map $\phi^{\tup}\colon\Delta^{\tup}\to\cms$ to any face of the multisimplex
 $\Delta^{\tup}$ coming from a factor different from $\Delta^l$,
 we obtain either a constant map to the point at infinity, hence these faces don't contribute
 to the boundary of $e^{\tup}$ in the reduced cellular chain complex. Instead for $0\leq i\leq l$ the restriction
 \[
  \phi^{\tup}\colon\partial_i\Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}}\to\cms^c
 \]
 hits the cell $e^{\tup'}$ exactly as many times as specified in the statement of the lemma for the cases $1\leq i\leq l-1$,
 $i=0$ and $i=l$ respectively. We are working in $\Z_2$ so we don't have to be careful about orientations of simplices
 and therefore about signs.
\end{proof}

We can filter the chain complex $\tilde\Ch_*(\cms^c)$ by giving weight $\sum_{i=1}^lx_i$ to
the cell $e^{\tup}$, with $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$. By lemma \ref{lem:doperatoropenmodtwo}
the weight is weakly decreasing along boundaries. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
of weight $\leq p$, and let $F_p/F_{p-1}$ be the $p-$th filtration stratum.

Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1}))$,
one copy for each way of splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$, where the isomorphism
shifts degrees by $p$. Indeed in $F_p/F_{p-1}$ we can compute differentials only according to the first
of the three possibilities listed in lemma \ref{lem:doperatoropenmodtwo}, in particular the numbers $u_i,v_i$
don't change along boundaries in $F_p/F_{p-1}$; it is then immediate to identify the \emph{internal} differentials
with the ones one would get in the case $g=0$.

We notice that $\tCh_*(C_p(\Sigma_{0,1}))$ is exactly the chain complex described by Fuchs in
\cite{Fuchs:CohomBraidModtwo}; in particular its homology is generated by symmetric chains,
where, for a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_i$,
the associated symmetric chain is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
$(l,(x_i)_{i\leq l})$ such that
\begin{itemize}
 \item $l=\sum_{i=1}^{\infty}\alpha_i$;
 \item every $x_i$ is a power of 2;
 \item for all $j\geq 0$ there are exactly $\alpha_j$ indices $i$ such that $x_i=2^j$.
\end{itemize}

If we run the Leray spectral sequence associated to the filtered chain complex $\tCh_*(\cms)$,
the $E^1-$page contains many copies of the homology of $\tCh_*(C_p(\Sigma_{0,1}))$,
each generated by a symmetric chain. It is straightforward to check that these symmetric
(arising as cycles in $F_p/F_{p-1}$), are also cycles when lifted, as a sum of the same
cells, to the complex $\tCh_*(\cms^c)$: one has only to check that the \emph{outer}
differentials cancel out, and this is easy to see as every \emph{left} differential
of one cell in the sum of a symmetric chain corresponds to a \emph{right} differential
of possibly another cell in the same symmetric chain.

In particular the spectral sequence collapses on its first page and the $\Z_2$-homology of
$\cms^c$ has a $\Z_2$-basis given by all choices of the following set of data:
\begin{itemize}
 \item a number $0\leq p\leq m$;
 \item a splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$;
 \item a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ with $\alpha_j\geq 0$.
\end{itemize}
The class $(p,(u_i,v_i),(\alpha_j))\in H_*(\cms^c)$ has homological degree
$m+\sum_{j=1}^{\infty}\alpha_i$.

% To interpret this class
% as a monomial in the tensor product of theorem \ref{thm:Hbms*as*ggrep} we need the following definition.
\begin{defn}
 \label{defn:dualHbasis}
We can see the $\U_i$'s and $\V_i$'s as properly embedded $1-$manifolds in $\mrS=\T(\S)$;
by Poincaré-Lefschetz duality they represent classes in $H^1(\S)$, and in particular they form a basis
of this cohomology group. We fix simple closed curves $\u_i,\v_i$ on $\S$, for $1\leq i\leq g$, whose fundamental classes
form the dual basis of $H_1(\S)$. We assume that, apart from the following necessary exceptions,
all the curves $\U_i,\V_i,\u_i,\v_i$ for $1\leq i\leq g$ are disjoint:
\begin{itemize}
 \item $\u_i$ and $\v_i$ intersect once, transversely;
 \item $\u_i$ and $\U_i$ intersect once, transversely;
 \item $\v_i$ and $\V_i$ intersect once, transversely;
 \end{itemize}
SEE PICTURE.
\end{defn}
We establish a bijection between monomials in the tensor product of theorem  \ref{thm:Hbms*as*ggrep}
and the basis of $H^*(\cms)\simeq H_*(\cms^c)$ that we have just found.

Then the class 
\[
(p,(u_i,v_i),(\alpha_j))\in H_{m+\sum\alpha_j}(\cms^c)\simeq H^{m-\sum\alpha_j}(\cms)
\]
is associated with the monomial $\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g([\u_i]^{u_i}[\v_i]^{v_i})$.

This shows an isomorphism of bigraded vector spaces
\begin{equation}\label{eq:isovectorspaces}
  \bigoplus_{m\geq 0} H^*(\cms)\simeq \Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]\otimes\Sym_{\bullet}(\H),
\end{equation}

and using the canonical identification of $\Z_2-$cohomology as the dual of homology, we can
conclude that there exists an isomorphism as in theorem \ref{thm:Hbms*as*ggrep}
at least \emph{as bigraded $\Z_2-$vector spaces}.

\subsection{Action of $\gg$} We now turn back to \emph{homology} of $\cms$;
our first task is to construct a map
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2-$vector spaces have the same dimension:
for the dimension of $H_m(\cms)$ we have to count the monomials of degree $m$ and weight $m$
in the right-hand side of equation \ref{eq:isovectorspaces}; a monomial whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$, i.e. it should lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
is $m$ it should lie in $\Sym_m(\H)$.

The map $\psi_m$ that we construct will turn out to be a $\gg-$equivariant isomorphism.
We will first construct a map $\tpsi_m\colon \H^{\otimes m}\to H_m(\cms)$ and then show
that factors through $\Sym_m(\H)$.

% \begin{defn}
% \label{defn:cCm}
% Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of closed curves $c_1,\dots,c_m\subset\mrS$
% such that:
% \begin{itemize}
%  \item each $c_i\colon\Sone\to\mrS$ is smooth and a local embedding;
%  \item the curves $\set{c_i}_{i\leq m}$
%  are pairwise transverse, no three of them intersect at a common point in $\mrS$.
% \end{itemize}
% Two such $m-$tuples are isotopic if there is an ambient isotopy of $\S$
% transforming one $m-$tuple into the other. For each isotopy class we
% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
% also transverse to all $\U_i$'s and $\V_i$'s.
% 
% We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
% There is a canonical surjective map $\pi\colon\ZcC{m}\to\Sym_m(\H)$ given by
% \[
%  \pi(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m].
% \]
% The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
% on $\Sym_m(\H)$, symplectically. The map $\pi$ is $\gg-$equivariant.
% \end{defn}
% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pi$ to give the desired isomorphism $\psi$.
% 
% The rough idea is to map the basis element
% $(c_1,\dots,c_m)$ first to $[c_1]\otimes\dots\otimes[c_m]\in H_m((\S)^{\times m})$, then map
% this further to $[c_1]\cdot\ldots\cdot[c_m]\in H_m(\SP^m(\S))$ and then lift this to an element in
% $H_m(\cms)$ along the inclusion $\cms\subset \SP^m(\S)$.
% 
We construct $\tpsi_m$ by induction on $m$. For $m=0$ both source and target are
isomorphic to $\Z_2$; for $m=1$
both source and target are canonically identified with $\H$: we take $\tpsi_0$ and $\tpsi_1$
to be these canonical isomorphisms.
% we simply map $c_1$ to its fundamental class in $\H=H_1(\S)=H_1(C_1(\S))$.
Let now $m\geq 1$ and suppose that we have defined a map $\tpsi_m\colon \H^{\otimes m}\to H_m(\cms)$.

\begin{defn}\label{defn:variationsCm}
Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$, using the model $\T(\S)$:
$\D$ is an open disc in $\mrS$ near the boundary and disjoint from all $\U_i,\V_i$.

We introduce a few variations of the notion of configuration spaces.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ of elements
 $(q;\set{p_1,\dots,p_m})$ with $q\neq p_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{red} (meaning that
 it is special), whereas the other are \emph{black}.
 \item The space $\cmstwo$ is the subspace of $\SP^m(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point with multiplicity $2$ and $m-1$ other points with multiplicity $1$:
 in the latter case we require the double point to lie in $\D$.
\end{itemize}
There is a natural map $\pi\colon\mrS\times \cms\to \cmstwo$ , which converts the red point
into a black point. This map restricts to a covering map $C_{1,m}(\S)\to C_{m+1}(\S)$; we have a commutative
 diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \mrS\times \cms \\
   @V\pi VV @V\pi VV\\
   C_{m+1}(\S) @>\subset >> \cmstwo
   \end{CD}
\end{equation}
We notice that all these spaces are $2(m+1)-$dimensional manifolds, arising as open subsets
of $\mrS\times\SP^m(\mrS)$ or $\SP^{m+1}(\mrS)$, which are manifolds of dimension $2(m+1)$.

Let $\Delta=\cmstwo\setminus C_{m+1}(\S)$ and similarly $\tilde\Delta=\mrS\times \cms\setminus C_{1,m}(\S)$.
We notice that $\pi$ restricts to a homeomorphism $\tilde\Delta\to\Delta$. Moreover both $\Delta\subset\cmstwo$
and $\tilde\Delta\subset\mrS\times \cms$ are embedded submanifolds of codimension $2$, and the map $\pi$ restricts to
a $2-$ramified covering between the normal bundles, which are bundles with fiber a $2-$disc.
\end{defn}

Diagram \ref{eq:cmstwodiagram} induces a commutative diagram in homology
\[
\minCDarrowwidth15pt
 \begin{CD}
  @. H_{m+1}\pa{\mrS\times \cms} @>>> H_{m+1}\pa{\mrS\times \cms, C_{1,m}(\S)}\\
  @. @V\pi_*VV @V\pi_*VV\\
  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwo} @>>> H_{m+1}\pa{\cmstwo,C_{m+1}(\S)}
 \end{CD}
\]

We can map $\H^{\otimes m+1}$ to $H_{m+1}\pa{\mrS\times \cms}$ through the map
$\Id_{\H}\otimes\tpsi_m$, and then further down to $H_{m+1}\pa{\cmstwo}$.

We claim that
we land in the image of the map
\[
H_{m+1}\pa{ C_{m+1}(\S)}\to H_{m+1}\pa{\cmstwo}.
\]
To check this we further
map towards right to $H_{m+1}\pa{\cmstwo,C_{m+1}(\S)}$ and prove that we get the zero map.
Indeed the other equivalent composition
involves the map
\[
\pi_*\colon H_{m+1}\pa{\mrS\times \cms, C_{1,m}(\S)}\to H_{m+1}\pa{\cmstwo,C_{m+1}(\S)};
\]
this map can be rewritten, after using the Thom isomorphism, as a map
\[
 H_{m-1}(\tilde\Delta)\to H_{m-1}(\Delta)
\]
and the latter map is multiplication by $2$, as the normal bundle of $\tilde\Delta$ is a
double covering of the normal bundle of $\Delta$. Therefore the claim is proved, and
as we are working with $\Z_2-$vector spaces there exists a lift
\[
 \tpsi_{m+1}\colon\H^{\otimes m+1}\to H_{m+1}(C_{m+1}(\S))
\]
of the map $\pi_*\circ(\Id_{\H}\otimes\tpsi_m)\colon \H^{\otimes m+1}\to H_{m+1}\pa{\cmstwo}$.

Thus we have constructed, for each $m\geq 0$, a map $\tpsi_m\colon \H^{\otimes m}$.

\begin{lem}
 \label{lem:cmstocmstwoinjective}
 The map
 \[
H_m\pa{\cms}\to H_m\pa{C_m^{(2)}}
\]
is injective
 \end{lem}
 \begin{proof}
  We can equivalently check that the map $H^m\pa{\cmstwo}\to H^m\pa{\cms}$ is surjective,
  or, again, that the map $H_m(\cmstwo^c)\to H_m(\cms^c)$ is surjective.

  The last statement is clear
 \end{proof}

### 2018-10-19 Introduction.tex
Changing the cmstwo construction to one that involves the sink

@@
 Forgetting punctures yields a surjective map $\ggm\to\gg$ with kernel $\bms$,
-the $m-$th \emph{braid group} of the surface $\sg$: we get the Birman exact sequence
+the $m-$th \emph{braid group} of the surface $\sg$: we get the Birman exact sequence (see \cite{Birman:mcgbr}):
 \begin{equation}
-\label{eqn:Birman}
+\label{eq:Birman}
 1\to\bms\to\ggm\to\gg\to 1.
@@
 
-The associated Leray-Serre spectral sequence in rational homology has a second page $E^2_{p,q}=H_p(\gg;H_q(\bms;\Q))$,
-and converges to $H_{p+q}(\ggm;\Q)$.
+The associated Leray-Serre spectral sequence in $\Z_2-$homology has a second page $E^2_{p,q}=H_p(\gg;H_q(\bms;\Z_2))$,
+and converges to $H_{p+q}(\ggm;\Z_2)$.
 
@@
 \label{thm:main}
-For each $l\geq 0$ there is an isomorphism of $\Q$-vector spaces
+For each $l\geq 0$ there is an isomorphism
 \begin{equation}
-\label{eqn:main}
- H_l(\ggm;\Q)\cong \bigoplus_{p+q=l} H_p(\gg;H_q(\bms;\Q)).
+\label{eq:main}
+ H_l(\ggm;\Z_2)\cong \bigoplus_{p+q=l} H_p(\gg;H_q(\bms;\Z_2)).
 \end{equation}
 \end{thm}
-Thus the computation of $H_*(\ggm;\Q)$ reduces to the computation of the homology of $\gg$ with
-twisted coefficients in the representation $H_*(\bms;\Q)$: the action of $\gg$ is quite explicit
-and in particular is symplectic. The group $\gg$ is significantly smaller than $\ggm$, so the
-computation of the right-hand side of equation \ref{eqn:main} should be more tractable. From now on $\Q$-coefficients will always be understood.
+Thus the computation of $H_*(\ggm;\Z_2)$ reduces to the computation of the homology of $\gg$ with
+twisted coefficients in the representation $H_*(\bms;\Z_2)$: as we will see the $\ggm-$representation
+$H_*(\bms;\Z_2)$ splits as a direct sum of symmetric powers of $H_1(\sg)$.
+In particular this representation is symplectic: this had already benn stated in \cite{LM}.
+
+The group $\gg$ is significantly smaller than $\ggm$, so the
+computation of the right-hand side of equation \ref{eq:main}
+should be more tractable.
+WE DO THIS COMPUTATION FOR $g=1$.
+
+From now on $\Z_2$-coefficients will always be understood if no coefficients are specified.

### 2018-10-19 MAIN.tex
Changing the cmstwo construction to one that involves the sink

@@
 
-\title{Splitting of the homology of the puntured mapping class group.}
+\title{Splitting of the homology of the punctured mapping class group}
 
@@
 \begin{abstract}
-Let $\ggm$ be the mapping class group of an orientable surface of genus $g$ with one parametrised
-boundary component and $m$ permutable punctures; let $\gg=\gg^0$; let $\bms$ be the braod group 
-of the surface $\sg$ on $m$ strands; let $\F$ be $\Z_2$ or $\Q$. We prove that $H_*(\ggm;\F)\cong H_*(\gg;H_*(\bms;\F))$.
+Let $\ggm$ be the mapping class group of an orientable surface $\sgm$ of genus $g$ with one parametrised
+boundary component and $m$ permutable punctures, and let $\bms$ be the braid group on $m$ strands
+of the surface $\sg$. We prove that $H_*(\ggm;\Z_2)\cong H_*(\gg;H_*(\bms;\Z_2))$.
 We compute (?) the right hand side explicitly for $g=1$.
@@
 
-\bibliography{bibliography.bib}{}
+\bibliography{Bibliography.bib}{}
 \bibliographystyle{plain}

### 2018-10-19 Preliminaries.tex
Changing the cmstwo construction to one that involves the sink

\section{Preliminaries}
In this section we recollect some classical definitions and results about braid groups and mapping class groups.
By $\sg$ we will always denote a compact, smooth, orientable surface of genus $g$ with one parametrised
boundary component $\partial \sg$; by $\sgm$ we denote a surface $\sg$ with a choice of $m$ distinct
points in the interior: these points are called \emph{punctures} and are not assumed to be ordered
or labelled.

\begin{defn}
 The $m$-th \emph{ordered configuration space} of a surface $\sg$ is the space
\[
 F_m(\sg)=\set{(p_1,\dots,p_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  p_i\neq p_j  \,\forall i\neq j}
\]
 Notice that we require the points of the configuration to lie in the interior of $\sg$.
 Thus $F_m(\sg)$ is naturally endowed with the structure of a smooth, orientable $2m-$dimensional
 manifold.

 The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the points of a configuration;
 the orbit space
 \[
 F_m(\sg)/\mathfrak{S}_m
 \]
 is called the \emph{$m-$th unordered configuration space}
 of $\sg$ and denoted by $C_m(\sg)$; this space is also a $2m-$dimensional orientable manifold.

%  
%  We denote by $C_k(\sg)^c$ the one-point compactification of the space $C_k(\sg)$; the point
%  at infinity is denoted by $*$ and serves as basepoint.
\end{defn}

A classical result by Fadell and Neuwirth (\cite{FadellNeuwirth}) ensures
that $C_m(\sg)$ is aspherical; the fundamental group $\pi_1(C_m(\sg))$ is
called the \emph{braid group on $m$ strands of $\sg$} and is denoted by $\bms$.

\begin{defn}
 \label{def:mcg}
 Let $\Diff(\sg)$ be the group of diffeomorphisms of $\sg$ that fix $\partial\sg$ pointwise,
 endowed with the Whitney $C^{\infty}-$topology. We denote by $\gg=\pi_0(\Diff(\sg))$ its group of connected
 components, called also the \emph{mapping class group} of $\sg$.

 Similarly let $\Diff(\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
 pointwise and permute the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm))$ be its
 group of connected components, called the mapping class group of $\sgm$.
 \end{defn}
 A classical result by Earle and Schatz (\cite{EarleSchatz}) ensures that the connected components
 of $\Diff(\sg)$ are contractible. In particular $B\Diff(\sg)\simeq B\gg$ is a classifying space for
 the group $\gg$ and we call $\cF_{g,1}\to B\gg$ the universal $\sg-$bundle over $B\gg$, corresponding to
 the bundle
 \[
  \sg\times_{\Diff(\sg)}E\Diff(\sg)\to B\Diff(\sg).
 \]
Applying the construction of the $m-$th unordered configuration space fiberwise we obtain a bundle
$C_m(\cF_{g,1})\to B\gg$ with fiber $C_m(\sg)$.

The space $C_m(\cF_{g,1})$ is a classifying space for the
group $\ggm$ and the Birman exact sequence \ref{eq:Birman} is obtained by taking fundamental
groups of the spaces
\begin{equation}\label{eq:Birmanbundle}
C_m(\sg)\to C_m(\cF_{g,1})\to B\gg.
\end{equation}

% \begin{defn}
%  \label{def:modulispaces}
% We denote by $\mg$ the moduli space of Riemann surfaces $\sg$ of genus $g$ 
%  with one parametrised
%  boundary component, and by $\mgm$ the moduli space of Riemann surfaces $\sgm$ with one
%  parametrised boundary component and $m$ unordered punctures.
%  There is a natural fiber bundle map
%  \[
%   \pi\colon \mgm\to\mg
%  \]
% which forgets the position of the punctures. The fiber of $\pi$ is the configuration space of $m$ unordered
% points in the interior of $\sg$, i.e.
% \[
%  C_m(\sg)=\set{(p_1,\dots,p_m) \in \mathring{\sgigma}_{g,1}^m  \,|\,  p_i\neq p_j  \,\forall i\neq j}
% \]
% \end{defn}
% 
% The Birman exact sequence \ref{eq:Birman} can be obtained from the bundle
% \begin{equation}\label{eq:Birmanbundle}
%  C_m(\sg)\to\mgm\to\mg
% \end{equation}
% by computing all fundamental groups: indeed the three spaces appearing in \ref{eq:Birmanbundle} are
% the classifying spaces of the groups in \ref{eq:Birman}.

### 2018-10-25 Actiongg.tex
Finished proving the  gg-equivariant isomorphism

\section{Action of $\gg$}
We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of theorem
\ref{thm:Hbms*as*ggrep} in bidigrees $(*,m)$ with $*=m$, and then extend it to other bidegrees.

\subsection{Bidigrees of the form $(m,m)$}
In this subsection we construct a $\gg-$equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2-$vector spaces have the same dimension:
for the dimension of $H_m(\cms)$ we have to count the monomials of degree $m$ and weight $m$
in the right-hand side of equation \ref{eq:isovectorspaces}; a monomial whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$, i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m\pa{\mrS}
% \]
% which carries the fundamental class of the $m-$fold torus to some class in
% $H_m(\SP^m\pa{\mrS})$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

The construction of $\psi_m$ is rather long and technical and involves a few definitions.
\begin{defn}
 \label{defn:D}
 Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$, using the model $\T(\S)$:
$\D$ is an open disc in $\mrS$ near $\partial\S$ and disjoint from all $\U_i$'s and $\V_i$'s.

In the following it is useful to redefine $\gg$ as the group of isotopy classes
of diffeomorphisms of $\S$ that fix both $\partial \S$ and $\D$ pointwise: this is 
not a problem, since the inclusion
\[
\Diff(\S;\D\cup\partial\S)\subset \Diff(\S;\partial\S)
\]
induces a bijection on the set of path-components.
\end{defn}

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that:
\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
 \item the curves $\set{c_i}_{i\leq m}$ intersect each other only inside $\D$;
 \item each curve $c_i$ is transverse to $\partial\D$.
\end{itemize}
Two such $m-$tuples are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.
% For each isotopy class we
% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
% also transverse to all $\U_i$'s and $\V_i$'s. We write
% $[c_1,\dots,c_m]$ for the isotopy class whose preferred representative is
% $(c_1,\dots,c_m)$.

We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m([c_1,\dots,c_m])=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg-$equivariant.
\end{defn}
% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg-$equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 The space $\cmsD$ is the subspace of $\SP^{m}\pa{\mrS}$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$. The space $\cmsD$ is an open subset
 of $\SP^m\pa{\mrS}$, which is a $2m-$dimensional manifold, hence $\cmsD$ is also a manifold.

 There is a natural inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class in $\cC^m$ represented
  by a $m-$tuple $(c_1,\dots,c_m)$, the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> \pa{\mrS}^{\times m} @>>> \SP^m\pa{\mrS}
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple which is isotopic relative
to $\D\cup \partial\S$.

We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in $\cmsD$ of the fundamental cycle of the
$m-$fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
on the union $c_1\cup\dots\cup c_m$, meaning that it hits configurations in $\cmsD$ of points of $\mrS$
lying in this union.

The group $\Diff(\S,\D\cup\partial\S)$ acts on $\cmsD$, and there is an induced action of $\gg$
on $H_m(\cmsD)$. The map $\j_m$ is $\gg-$equivariant.
\end{defn}

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $H_m(\cmsD^c)\to H_m(\cms^c)$
 is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
 $\cmsD^c\setminus\cms$ to $\infty$.

 Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $H_m(\cms^c)$
 is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
 and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$.

 The class $[0,(u_i,v_i)]$ is represented by a generalised symmetric chain consisting of only one cell,
 the cell $e^{\tup}$
 with $\tup=(0,(u_i,v_i))$.

 In particular there is a map of pairs
 $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^c,\infty)$, and $[0,(u_i,v_i)]$
 is the image along this map of the fundamental class of $H_m\pa{\Delta^{\tup},\partial\Delta^{\tup}}$.

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $H_m(\cmsD^c)\to H_m(\cms^c)$ follows.
\end{proof}

We have the following diagram of $\gg$-equivariant maps, where the full arrows are those
that we have already constructed, and we still have to prove the existence of the dashed arrows
\begin{equation}
\label{eq:fulldasheddiagram}
\begin{tikzcd}[column sep=8em,row sep=5em]
  \ZcC{m} \ar[r,"\pr_m",two heads]\ar[d,dashed, swap, "\tpsi_m"]\ar[dr,"\j_m",near start]
  & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*"hook] & H_m(\cmsD)
 \end{tikzcd}
\end{equation}

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg-$equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and the last statement doesn't depend on how $\gg$ acts on these groups.

We will prove by induction on $m$ the following technical lemma:

\begin{lem}
 \label{lem:tpsiwithproperties}
For each $m-$tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
and for each open neighborhood $\N\subset\mrS$ of
$\D\cup c_1\cup\dots\cup c_m$, there is a singular cycle $\fc=\tpsi_m(c_1,\dots,c_m)$
in $\cms$ with the following properties:
\begin{itemize}
 \item the cycle $\fc$ is \emph{supported} on $\N$, i.e.,
this singular cycle only hits configurations of $m$ distinct points of $\mrS$ that actually lie in $\N$;
\item $\pa{\iota_m}_*(\fc)$ represents the homology class $\j_m(c_1,\dots,c_m)\in H_m\pa{\cmsD}$;
\item the two cycles $\pa{\iota_m}_*(\fc)$ and $[c_1]\cdot\ldots\cdot[c_m]$
are connected by a homology in $\cmsD$ which is supported on $\N$ (the word
\emph{homology} denotes here a $(m+1)-$singular chain whose boundary is the difference between the two cycles).
\end{itemize}
\end{lem}

For $m=0$ both $\ZcC{0}$ and $H_0(C_0(\S))$ are
isomorphic to $\Z_2$ and there is nothing to show; for $m=1$
we have a canonical identification $H_1(C_1(\S))\simeq\H\simeq\Sym_1(\H)$, so we take $\tpsi_1=\pr_1$;
obviously for all $c_1$ representing a class in $\cC^1$, the homology class $\pr_1(c_1)\in\H$
is represented by a cycle supported on $c_1$,
and in this case the cycles $\iota_*\pa{\tpsi(c_1)}$ and $\pr_1(c_1)=\j_m(c_1)$ coincide.

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^m$.
% and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
% with the aforementioned property.

\begin{defn}
\label{defn:variationsCm}
We introduce a few variations of the notion of configuration space.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{red} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is the red point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the red point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}\pa{\mrS}$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\S$, with multiplicity $1$.
\end{itemize}
 \end{defn}

We the following natural inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m\pa{\mrS};
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}\pa{\mrS}.
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

In particular there is a sequence of maps
\[
 H_{m+1}\pa{C_{m+1}(\S)}\to H_{m+1}\pa{\cmstwoD}\to H_{m+1}\pa{C_{m+1}^{\D}(\S)}
\]
and we will first lift the homology class $j_{m+1}(c_1,\dots,c_{m+1})$ to $H_m\pa{\cmstwoD}$ and
then to $H_{m+1}(C_{m+1}(\S))$, each time controlling the support of our representing
cycles and of the homologies between them.

Fix a small neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.

For the first lift, let $\N'=\pa{\N\setminus c_1}\cup\D$, which is an open
neighborhood of $\D\cup c_2\cup\dots\cup c_{m+1}$ in $\mrS$. By inductive hypothesis
there is a cycle $\fc=\tpsi_m(c_2,\dots,c_{m+1})$ in $\cms$ which is supported on $\N'$,
and such that $(\iota_m)_*(\fc)$ is homologous to $[c_2]\cdot\ldots\cdot[c_{m+1}]$
along a homology in $\cmsD$ supported on $\N'$ as well.

We can multiply both
cycles and the homology between them by the cycle $[c_1]$: the result are the two homologous cycles
$[c_1]\cdot \fc$ and $[c_1]\cdot\ldots\cdot[c_{m+1}]$ in $C_{m+1}^{\D}(\S)$: both cycles and the
homology between them are supported on $\N$. Notice now that the
cycle $[c_1]\cdot\fc$ lives in $\cmstwoD$, so the first lift is done and we can now
deal with the second lift.

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the red point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$;
% the latter being even a $m+1-$fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are properly embedded submanifolds of codimension $2$, and the map $\p$ restricts to
a $2-$ramified covering between their respective normal bundles.
\end{defn}

Diagram \ref{eq:cmstwodiagram} induces a commutative diagram in homology
\begin{equation}
 \label{eq:fivediagram}
\minCDarrowwidth15pt
 \begin{CD}
  @. H_{m+1}\pa{\fmstwoD} @>>> H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)}\\
  @. @V\p_*VV @V\p_*VV\\
  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwoD} @>>> H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
 \end{CD}
\end{equation}

Recall that we want lift the homology class represented by the cycle $[c_1]\cdot\fc$
from the bottom central group to the bottom left group.

We first notice that there is a lift $[c_1]\otimes\fc$ to a cycle in $\pa{\fmstwoD}$,
which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be red. We then notice that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)};
\]
can be rewritten, after using excision to a tubular neighborhood of $\Donem$ and $\Dmone$ and
the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone);
\]
the latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

Therefore the image of the cycle $[c_1]\otimes \fc$ along the diagonal of the square
in diagram \ref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
is zero, hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$: more
precisely there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
to $[c_1]\cdot\fc$.

To prove lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
a good homology, namely two that are supported on $\N$: a priori both $\fc'$ and the homology
between $\pa{\iota_{m+1}}_*(\fc')$ and $[c_1]\cdot\fc$ are only supported on $\mrS$.

But this is easily done by considering, in the whole argument of the proof, the surface $\N$ instead of $\mrS$:
we can define configuration spaces as in definition \ref{defn:variationsCm} using the open surface
$\N$ instead of $\mrS$, and repeat the whole argument considering this as the \emph{ambient surface}:
indeed we only need $\N$ to contain $\D$ and all curves $c_1,\dots,c_{m+1}$.

Here it is crucial that the action of $\gg$ is not involved in the statement of lemma
\ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
by diffeomorphisms of $\S$.

Lemma \ref{lem:tpsiwithproperties} is proved.

We now have to prove the following lemma to conclude the proof of theorem \ref{thm:Hbms*as*ggrep}
in bidigrees $(m,m)$.
\begin{lem}
 \label{lem:tpsi->psi}
The map $\tpsi_m\colon\ZcC{m}\to H_m(\cms)$ is surjective and factors through the map $\pr_m$.
\end{lem}
\begin{proof}
The factorisation is equivalent to the inclusion $\ker\pr_m\subseteq\ker\tpsi_m$: since both
$\pr_m$ and $\tpsi_m$
are $\gg-$equivariant, also the induced map $\Sym_m(\H)\to H_m(\cms)$ will automatically be
$\gg-$equivariant.

Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $H_m(\cms^c)\simeq H^m(\cms)$, which is
the dual of $H_m(\cms)$, is given by the classes $[0,(u_i,v_i)]$, represented by generalised
symmetric chains consisting of only one cell $e^{\tup}$, with $\tup=(0,(u_i,v_i))$.
The corresponding linear functionals are the algebraic intersection products
coming from Poincaré-Lefschetz duality:
\[
 (\cdot\cap e^{\tup})\colon H_m(\cms)\to\Z_2.
\]
Therefore
\[
 \ker\tpsi_m=\bigcap_{\tup}\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m},
\]
and it suffices to check that $\ker\pr_m\subseteq\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m}$
for all $\tup$ of the form $(0,(u_i,v_i))$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.

Recall from the proof of lemma \ref{lem:cms->cmsDinj} that the cohomology class $(\cdot\cap e^{\tup})$ is a pullback
of a cohomology class of $\cmsD$, that by abuse of notation we still call $(\cdot\cap e^{\tup})$:
then we can compute the map $(\cdot\cap e^{\tup})\circ\tpsi_m$ as the map $(\cdot\cap e^{\tup})\circ\j_m$.

Looking at the geometric intersection between the cycle $[c_1]\cdot\dots[c_m]$ in
$\cmsD$, and the cycle $e^{\tup}$ in $\cmsD^c$, we see that the map
\[
(\cdot\cap e^{\tup})\circ\j_m\colon \ZcC{m}\to\Z_2
\]
coincides with the composition
\[
\pa{ \prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}}\circ\pr_m,
\]
where $\prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}\in\Sym_m(\Hom(\H;\Z_2))=\Hom\pa{\Sym_m(\H);\Z_2}$;
in particular it factors through $\pr_m$.

To show surjectivity of $\tpsi_m$, choose a tuple $\tup$ of the form $(0,(u_i,v_i))$
and an $m-$tuple of curves $(c_1,\dots,c_m)$ containing, for every $1\leq i\leq g$, $u_i$ parallel
copies of some curve representing $\u_i$ and $v_i$ parallel copies of some curve representing
$\v_i$ (see definition \ref{defn:dualHbasis}), such that all intersections between these curves
lie in $\D$.

Then $j_m(c_1,\dots,c_m)\cap e^{\tup}=1$ and for all
other tuples $\tup'=(0,(u'_i,v'_i))$ we have instead $j_m(c_1,\dots,c_m)\cap e^{\tup'}=0$.

This shows that $\psi_m(c_1,\dots,c_m)$ must be $[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
one of the generating monomials.
\end{proof}

Theorem \ref{thm:Hbms*as*ggrep} is then proved in all bidegrees $(*,m)$ with $*=m$.

\subsection{General bidegrees $(m-l,m)$.} In this subsection we fix a bidegree $(m-l,m)$
with $0\leq l\leq m$.
Our aim is to prove theorem \ref{thm:Hbms*as*ggrep} for this bidegree $(m-l,m)$.
\begin{defn}
%  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
%  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
%  $\D\subset\D'$.
 Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
 Notice that $\S'$ is also a surface of type $\sg$.

 For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
 using $\D$ and $\mrS'$ instead of $\mrS$, respectively.

 For every splitting $m=p+(m-p)$ there is an open embedding
 \[
 \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
 \]
 given by taking the union of configurations:
 \[
  \mu(\set{P_1,\dots,P_p};\set{P'_1,\dots,P_{m-p}})=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
 \]
%  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
%  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
\end{defn}
For all $0\leq p\leq m$, the group $\Diff(\S;\partial\S\cup\D)$ acts both on $C_p(\D)\times C_{m-p}(\S')$
and on $\cms$, and the map $\mu$ is equivariant with respect to this action;
hence, using the K\"{u}nneth formula, there is an induced $\gg-$equivariant map in homology
\begin{equation}
 \label{eq:mu*}
 \mu_*\colon H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}\to H_{m-l}\pa{\cms}.
\end{equation}

Notice that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
of the $\gg-$representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
section is isomorphic to the symplectic representation $\Sym_{m-p}(\H)$.

We will prove the following lemma, from which theorem \ref{thm:Hbms*as*ggrep} easily follows:
\begin{lem}
 \label{lem:oplussplitting}
For all $l\leq p\leq m$ the map $\mu_*$ in equation \ref{eq:mu*} is injective, and the collection
of all these maps yields a splitting
\[
 H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}
\]
\end{lem}
\begin{proof}
Notice that the statement of the lemma doesn't depend on the the action of $\gg$:
we have a map between two $\Z_2-$vector spaces, we already know that it is $\gg-$equivariant,
we only need to show that it is a bijection.
% 
% We can equivalently consider the maps in cohomology
%  \[
%    \mu^*\colon  H^{m-l}\pa{\cms}\to H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)},
%  \]
% and show that they are surjective for $0\leq p\leq m-l$ and together exhibit a product splitting
% \[
%   H^{m-l}(\cms)=\prod_{p=0}^{m-l} H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)}.
% \]
% Notice that the natural target for the cohomology map $\mu^*$ is the whole group
% \[
%  H^{m-l}\pa{C_{m-p}(\S')\times C_p(\D)}\simeq \prod_{q=0}^p H^q\pa{C_{m-p}(\S')}\otimes H^{m-l-q}\pa{C_p(\D)},
% \]
% where we have used the K\"unneth formula; we consider here, by abuse of notation,
% the composition of this map with the projection
% on the factor corresponding to $q=p$.
% 
% Using Poincaré-Lefschetz duality and again the K\"unneth formula we reduce to show that the maps
% \[
% \mu^c_* \colon \tH_{m-l}(\cms^c)\to \tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}
% \]
% are surjective and exhibit $\tH_{m-l}(\cms^c)$ as a product
% \[
% \tH_{m-l}(\cms^c)=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}.
% \]
% Here a few explanations are required:
% we are considering the map $\mu^c\colon \cms^c\to \pa{C_{m-p}(\S')\times C_p(\D)}^c$ which collapses the complement
% of the open subspace $C_{m-p}(\S')\times C_p(\D)\subset\cms$ to the point at infinity; it induces a map
% $\mu^c_*\colon \tH_{m-l}(\cms^c) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}$;
% applying Poincaré-Lefschetz duality
% and the K\"unneth formula to the second group we get an isomorphism
% \[
%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
% \]
% we are considering the composition of the map $\mu^c_*$ with the projection onto the factor $q=p$, and by
% abuse of notation we call this map simply $\mu^c_*$.
%

Fix $l\leq p\leq m$ and let $[a]=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
$H_{p-l}(C_p(\D))$, i.e., $\sum_j\alpha_j=l$ and $\sum_j2^j\alpha_j=p$; let also
$\beta=\prod_{i=1}^g \u_i^{u_i}\v_i^{v_i}$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
isomorphism proved in the previous subsection, i.e., $\sum_i(u_i+v_i)=m-p$. Here $a$ and
$b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
and $b$ supported on $\S'$.

Then $a\otimes b$ represents a generator of $H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))$,
and we are interested in the homology class represented by $\mu_*(a\otimes b)$.
We will study
the intersection of $\mu_*([a]\otimes [b])$ with cohomology classes of $\cms$ represented by
generalised symmetric chains in $\cms^c$.

% $[a]\otimes [b]$ has a dual class in $\tH_{p-l}(C_p(\D)^c)\otimes \tH_p(C_{m-p}(\S')^c)$
% represented by the product of generalised symmetric chains
% \[
%  (\alpha_j)_j\otimes (p,(u_i,v_i)_{i\leq g}).
% \]
% If $\tup=(l,(x_i)_{i\leq l})$ is a tuple such that $e^{\tup}$ appears in the symmetric
% chain $(\alpha_j)_j$, and if $\tup'=(0,(u_i,v_i)_{i\leq g})$ is the only tuple such that
% $e^{\tup'}$ appears in the generalised symmetric chain $(p,(u_i,v_i)_{i\leq g})$, then
% the product $e^{\tup}\times e^{\tup'}\subset C_p(\D)\times C_{m-p}(\S')\subset\cms$ is
% contained in the closure of the generalised symmetric chain $e^{\tup}$, where...

% If $p<p'$ then
% the \emph{geometric} intersection between the cycle $|p',(\alpha'_j),(u'_i,v'_i)|$
% in $\cms^c$ and the cycle $\mu_*(a\otimes b)$ in $\cms$ is empty: indeed the first cycle is supported
% on $\infty\in\cms^c$ and on configurations where at least $p'$ points lie in the union
% $\bigcup_i(\U_i\cup\V_i)\subset\S'$,
% whereas the second cycle is supported on configurations where at least $m-p$ points lie
% inside $\D$. The algebraic intersection between the homology classes $\mu_*([a]\otimes [b])$
% and $[p',(\alpha'_j),(u'_i,v'_i)]$ is therefore also zero.
% 
% Consider now the case $p=p'$.

To compute the algebraic intersection
between $\mu_*(a\otimes b)$ and $[p',(\alpha'_j),(u'_i,v'_i)]$ we
consider the map
\[
 \mu^c\colon \cms^c\to \pa{C_p(\D)\times C_{m-p}(\S')}^c
\]
which collapses to $\infty$ the complement in $\cms^c$ of $C_p(\D)\times C_{m-p}(\S')$.

By Poincaré-Lefschetz duality the map $\mu^c_*$ in reduced homology corresponds to the cohomology map
\[
 \mu^*\colon H^*(\cms)\to H^*(C_p(\D)\times C_{m-p}(\S'))=H^*(C_p(\D))\otimes H^*(C_{m-p}(\S')).
\]

We give $\pa{C_p(\D)\times C_{m-p}(\S')}^c$ the cell structure of the smash product
$C_p(\D)^c\wedge C_{m-p}(\S')^c$. Here $C_p(\D)^c$ is given the cell structure of $C_p((0,1)^2)$
along the natural identification $\D=(1/4,3/4)\times(1/2,1)\cong(0,1)^2$ obtained by rescaling
and translating; moreover we choose any diffeomorphism $\mrS'\cong\mrS$ that restricts to the identity
on all $\U_i$'s and $\V_i$'s, and give $C_{m-p}(\S')^c$ the cell structure of $C_{m-p}(\S)^c$.

Recall that $\cms^c$ can be filtered according to the norm of cells: a cell $e^{\tup}$ with
$\tup=(l,(x_i),(u_i,v_i))$ has norm $\sum x_i$. In the previous section we just considered
the associated filtration of the reduced chain complex $\tCh_*(\cms^c)$, but now we need
to consider the closed subcomplex $F_p\cms^c\subset\cms^c$ consisting of all cells of norm $\leq p$.

The crucial remark is that $\mu^c$ restricts to a cellular map 
\[
F_p\cms^c\to \pa{C_p(\D)\times C_{m-p}(\S')}^c.
\]
Consider an open cell
cell $e^{\tup}$ for some tuple $\tup=(l, (x_i),(u'_i,v'_i))$ of norm $p'\leq p$:
its intersection with the subspace
$C_p(\D)\times C_{m-p}(\S')$ is empty if $p'<p$, and it is exactly the open product of cells
$e^{\tup'}\times e^{\tup''}$,
where $\tup'=(l,(x_i))$ and $\tup''=(0,(u'_i,v'_i))$, if $p'=p$; therefore
$\mu^c(e^{\tup})$ is $\set{\infty}$ in the first case, and in the second case it
is contained in the union $\set{\infty}\cup e^{\tup'}\times e^{\tup''}$, which is also
contained in the $p'-$skeleton of $\pa{C_p(\D)\times C_{m-p}(\S')}^c$.

We replace the map $\mu^c$ by a cellular approximation that agrees with it on the $p-$skeleton,
and by abuse of notation we still call $\mu^c$ the new map.

Consider now any generalised symmetric chain $|p',(\alpha'_j),(u'_i,v'_i)|$ representing
a class in $H_{m+l}(\cms^c)=H^{m-l}(\cms)$, hence $\sum_j\alpha'_j=l$, and suppose $p'\leq p$.

If $p'<p$, the previous argument shows that $\mu^c_*\pa{|p',(\alpha'_j),(u'_i,v'_i)|}=\infty$,
and in particular the corresponding homology class is mapped to zero.

Suppose now $p'=p$: then the homology class $[p,(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$
is mapped along $\mu^c_*$ to the class
$[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]\in\tH\pa{C_p(\D)^c\wedge C_{m-p}(\S')^c}$
Indeed the previous
argument shows that each cell appearing in the cycle $|p,(\alpha'_j),(u'_i,v'_i)|$ covers with degree
1 a corresponding cell of the cycle $|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|$, and no other cell
of the same dimension: hence at the level of chains we have 
\[
\mu^c_*\pa{|p,(\alpha'_j),(u'_i,v'_i)|}=|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|.
\]

We can now compute the algebraic intersection of $\mu_*([a]\otimes [b])$ with
any cohomology class $[p',(\alpha'_j),(u'_i,v'_i)]$ as the intersection between
$[a]\otimes [b]$ and $\mu^c_*[p',(\alpha'_j),(u'_i,v'_i)]$.

For $p'<p$ the previous argument show that this intersection is zero.

For $p'=p$ we see that the intersection between the homology classes
$a\otimes b$ and $\mu^c_*([p,(\alpha'_j),(u'_i,v'_i)])=[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]$
is $1\in\Z_2$ exactly when $u_i=u'_i$ and $v_i=v'_i$ for all $i$, and $\alpha_j=\alpha'_j$
for all $j$; otherwise it is $0$.

We consider the collection of all strings of the form $(p,(u_i,v_i)_{i\leq g},(\alpha_j)_j)$
satisfying $\sum\alpha_j=l$ and $\sum \alpha_j2^j=p$; we order these strings lexicographically,
so that in particular $p$ is weakly increasing; we associate
to each string its corresponding class in $H_{m-l}(\cms)$ of the form $\mu_*(a\otimes b)$ and its
corresponding class in $\tH_{m+l}(\cms)$ represented by the corresponding generalised symmetric chain.

Then the matrix of algebraic intersections is an upper-triangular matrix
with $1$'s on the diagonal, and in particular
it is invertible. This shows that the set of classes $\mu_*(a\otimes b)$
is a basis for $H_{m-l}(\cms)$.

% 
% 
% 
% , and let $e^{\tup}$
% be one of the cells appearing in this generalised symmetric chain, with
% $\tup=(l,(x_i)_{i\leq l}, (u_i,v_i)_{i\leq g})$, and consider the restriction
% of $\mu^c\colon\cms^c\to \pa{C_{m-p}(\S')\times C_p(\D)}^c$
% to $e^{\tup}\subset\cms^c$
% 
% If $p\neq p'$, the whole cell $e^{\tup}$ is mapped
% to $\infty$: indeed if $p<p'$, then a configuration in $e^{\tup}$ contains
% at least $p'$ points inside $\S'$, so it can't lie in $C_p(\D)\times C_{m-p}(\S')$;
% if $p'<p$ then a configuration in $e^{\tup}$ contains at least $m-p$ points 
% 
% 
% then the algebraic
% intersection between $(p',(u'_i,v'_i),(\alpha'_j))$ and $\mu_*(a\otimes b)$ is $0\in\Z_2$,
% unless $p=p'$, $\alpha_j=\alpha'_j$ for all $j$, and $u_i=u'_i$ and $v_i=v'_i$ for all $1\leq i\leq g$.
% 
% Indeed the generalised symmetric chain $(p',(u'_i,v'_i),(\alpha'_j))$ hits only configurations
% in $\cms$ where for all $1\leq i\leq g$ at least $u'_i$ points lie on $\U_i$ and at least $v'_i$
% points lie on $\V_i$
\end{proof}

One could expect that the basis given by all elements $[a]\otimes [b]\in H_{m-l}(\cms)$
is dual to the basis of classes $[p',(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$, i.e., the
matrix considered in the end of the previous proof is not only upper-triangular but also
diagonal. This is however not true, as the following example shows...

### 2018-10-25 HBraidSurf.tex
Finished proving the  gg-equivariant isomorphism

@@
 \section{Homology of configuration spaces of surfaces}
-In this section we compute the homology $H_*(\bms)=H_*(C_m(\S))$; we
-are interested in these groups as they appear in the Leray-Serre spectral sequence associated to
+We want to understand the homology groups $H_*(\bms)=H_*(C_m(\sg))$: we
+are interested in these groups as they appear as the homology of the fiber
+in the Leray-Serre spectral sequence associated to
 the Birman exact sequence \ref{eq:Birman} or, which is the same, to the bundle
-\ref{eq:Birmanbundle}. In particular we are interested in $H_*(\bms)=H_*(C_m(\S))$
-\emph{as a $\Z_2-$representation of the group $\gg$.} From now on $\Z_2-$coefficients in homology
-and cohomology will always be understood.
+\ref{eq:Birmanbundle}. In particular we are interested in $H_*(C_m(\S))$
+\emph{as a $\Z_2-$representation of the group $\gg$.}
+% From now on $\Z_2-$coefficients in homology
+% and cohomology will always be understood.
 
 In \cite{LM} L\"offler and Milgram proved that this must be a symplectic
-representation of the mapping class group. By symplectic we mean the following:
+representation of the mapping class group. By \emph{symplectic} we mean the following:
 \begin{defn}
- \label{def:symplrep}
+ \label{defn:symplrep}
  Let $\H=H_1(\S)\simeq\Z_2^{2g}$.
@@
 
-In \cite{BCM} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\S))$ \emph{as a graded $\Z_2-$vector space}.
+In \cite{BCM} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\sg))$ \emph{as a graded $\Z_2-$vector space}.
 Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced, as
-their descripition of $H_*(C_m(\S))$ depends on a handle decomposition of $\S$.
+their descripition of $H_*(C_m(\sg))$ depends on a handle decomposition of $\sg$, which is not preserved,
+not even up to isotopy, by diffeomorphisms of $\sg$.
 
-We prove here the following theorem that, to the best of the author's knowledge, doesn't appear
-in the literature.
+In this section and in the next one we will prove the following theorem; to the best of the author's knowledge
+it doesn't appear in the literature.
 \begin{thm}
@@
  \[
-  H_*\pa{\coprod_{m\geq 0} C_m(\S)}\simeq \Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]\otimes\Sym_{\bullet}(\H).
+  \bigoplus_{m\geq 0} H_*\pa{C_m(\S)}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
  \]
@@
  \begin{itemize}
-  \item the bigrading is given on left by homological degree $*$ and by the connected component
-  in the disjoint union on which the homology class is supported, i.e. by the number $m$ of points
-  involved in constructing the homology class; we call $*$ the \emph{degree} and $m$ the \emph{weight};
-  \item for $i\geq 0$ $Q^i\epsilon$ is the image in $H_{2^i-1}(C_{2^i}(\S))$ of a generator
-  of the group $H_{2^i-1}(C_{2^i}(\Sigma_{0,1}))\simeq \Z_2$ 
+  \item the bidegree is given on left by homological degree $*$ and by the direct summand,
+  indexed by $m$,
+  on which the homology class is supported, i.e. by the number $m$ of points
+  involved in constructing the homology class; we call $*$ the \emph{degree} and $m$ the \emph{weight},
+  and write $(*,m)$ for the bidegree;
+  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
+  of the group $H_{2^j-1}(C_{2^j}(\Sigma_{0,1}))\simeq \Z_2$ 
   under the natural map induced by an embedding $\Sigma_{0,1}\hookrightarrow\S$,
-  and $\Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]$ is the polynomial ring on
+  and $\Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]$ is the polynomial ring on
   infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
-  \item $\H$ is identified with $H_1(C_1(\S))$ in a natural way, and $\Sym_{\bullet}(\H)$ is the
+  \item $\H=H_1(\sg)$ is identified with $H_1(C_1(\S))$ in a natural way, and $\Sym_{\bullet}(\H)$ is the
   symmetric algebra on $\H$;
@@
   \item the action of $\gg$ on right is the diagonal action on the product: it is a trivial action
-  on the factor $\Z_2[Q^i\epsilon\,|\, i\geq 0]$ and it is the symplectic action induced by the action on $\H$ on the
+  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$ and it is the symplectic action induced by the action on $\H$ on the
   factor $\Sym_{\bullet}(\H)$.
@@
   Notice that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
-  the degree.
+  the degree: indeed factors of the form $Q^j\epsilon$ have weight strictly higher than the degree,
+  whereas factors in $\H$ and its symmetric powers have equal weight and degree.
 \end{thm}
-For the rest of the section we will write $\S=\S$.
+In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2-$vector spaces}
+as in theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
+$\gg$. In these two sections we abbreviate $\S=\sg$.
 
-\subsection{Cohomology of $C_m(\gg)$.} Our first aim is to compute the \emph{cohomology} $H^*(C_m(\gg))$ as a graded
-$\Z_2-$vector space (not as a $\gg-$representation yet); we will mimic the approach of Fuchs (\cite{Fuchs:CohomBraidModtwo}).
-We will turn back to homology of $\cms$ in the next subsection.
+Since we work with $\Z_2-$coefficients, it is equivalent to compute homology or cohomology,
+and in this section we will prefer to compute, for each bidegree $(*,m)$, $H^*(\cms)$ as a graded
+$\Z_2-$vector space; the dual graded vector space $H_*(\cms)$ is of course abstractly isomorphic,
+i.e. it has the same dimension.
 
+We will mimic the method used by Fuchs (\cite{Fuchs:CohomBraidModtwo}) to compute the $\Z_2-$cohomology
+of the configuration space of the plane.
 As already said this computation recovers a known result, but it has the advantage of
 being quite elementary and of providing a part of
-the geometric insight that we will need later.
+the geometric insight that we will need in the next section.
 
-We first introduce a space $\T(\S)$ which is homeomorphic to the interior of $\S$. The construction
+In the whole section we assume $m\geq 0$ to be fixed.
+
+We first introduce a space $\T(\S)$ which is homeomorphic to $\mrS$, the interior of $\S$. The construction
 corresponds to a handle decomposition of $\S$ with one $0-$handle and $2g$ $1-$handles.
 \begin{defn}
-\label{def:Tsg}
+\label{defn:Tsg}
 If $g=0$, hence $\S=\sg$ is the disc, we set $\T(\S)=(0,1)^2$, the interior of the unit square. Assume
@@
 
-Dissect the interval $[0,1]$ into $2g$ equal subintervals through the points $P_i=\frac{i}{2g}$ for $0\leq i\leq 2g$
+Dissect the interval $[0,1]$ into $2g$ equal subintervals through the points $\cP_i=\frac{i}{2g}$ for $0\leq i\leq 2g$
 (for $i=0,2g$ we get the two endpoints of $[0,1]$).
 
-Let $Q$ be the subspace of $[0,1]^2$ obtained by the union of $(0,1)^2$ and the vertical open intervals
-$I_i^l=\set{0}\times (P_i,P_{i+1})$ and $I_i^r=\set{1}\times (P_i,P_{i+1})$ for $1\leq i\leq 2g$;
+Let $Q\subset[0,1]^2$ be the union of $(0,1)^2$ and the vertical open intervals
+$I_i^l=\set{0}\times (\cP_i,\cP_{i+1})$ and $I_i^r=\set{1}\times (\cP_i,\cP_{i+1})$ for $1\leq i\leq 2g$;
 equivalently, we remove from $[0,1]^2$ the two horizontal sides
-$[0,1]\times\set{0,1}$ and the $4g-2$ points $\set{0,1}\times \set{P_1,\dots,P_{2g-1}}$.
+$[0,1]\times\set{0,1}$ and the $4g-2$ points $\set{0,1}\times \set{\cP_1,\dots,\cP_{2g-1}}$.
 
-Notice that all intervals $I_i^l$'s and $I_j^r$'s are
+Notice that all intervals $I_i^l$ and $I_j^r$ are
 canonically diffeomorphic
-to $(0,1)$ by projecting on the second coordinate, rescaling linearly by $2g$
-(without changing orientation!) and translating; therefore we will specify a bijection
+to $(0,1)$ by projecting on the second coordinate, rescaling linearly by a factor $2g$
+and translating; therefore we will specify a bijection
 between the two sets of left and right intervals, and then $\T(\S)$ will be obtained from $Q$
@@
 \end{defn}
-The space $\T(\S)$ is homeomorphic to $\mrS$; from now
+The space $\T(\S)$ is homeomorphic to $\mrS$. From now
 on we will identify the two open surfaces and in particular we will identify $\cms$ with the
@@
 
-We consider the one-point-compactification $\cms^c$ of $\cms$; since the space $\cms$ is the interior of a compact
+In the following we will consider the one-point compactification of some spaces that
+happen to be open manifolds. For a space $\mathcal{M}$ we denote by $\mathcal{M}^c$ its one-point
+compactification; the basepoint is the point at infinity, that we always denote by $\infty$, even
+if it is a different point for different spaces $\mathcal{M}^c$.
+
+We consider in particular the one-point-compactification $\cms^c$ of $\cms$.
+Since the space $\cms$ is homeomorphic to the interior of a compact
 $2m-$manifold with boundary, by Poincaré-Lefschetz
@@
 \begin{defn}
-\label{def:ehopen}
+\label{defn:ehopen}
 A \emph{tuple} $\tup$ is a choice of the following set of data:
@@
  lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}-$factor,
- with $v_i$ and $\V_i$ instead of $\U_i$ and $u_i$.
+ with $v_i$ and $\V_i$ instead of $u_i$ and $\U_i$.
 \end{itemize}
-Notice that $\Delta^{\tup}$ has dimension equal to the degree of $\tup$.
-This embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^c$ extends to a continuous map
+Notice that the dimension of $\Delta^{\tup}$ is equal to the degree of $\tup$.
+The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^c$ extends to a continuous map
 $\phi^{\tup}\colon\Delta^{\tup}\to\cms^c$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
-subspaces $e^{\tup'}$ for tuples $\tup'$ with \emph{lower} degree than $\tup$, together with the point
-at infinity.
+all subspaces $e^{\tup'}$ for tuples $\tup'$ with \emph{lower} degree than $\tup$, together with the point $\infty$.
 
@@
 \begin{enumerate}
-\item we see the one-point-compactification
-$\mrS^c$ of $\mrS=\T(\S)$ as the quotient of $[0,1]^2$ that identifies left and right intervals as in
-in definition \ref{def:Tsg} of $\T(\S)$, and moreover collapses $[0,1]^2\setminus Q$ to a point, which
-is the point at infinity;
-\item we consider the $m-$fold symmetric product $\SP^m(\mrS^c)$: it contains 
-as an \emph{open} subspace $\cms$, so we can see $\cms^c$ as the quotient of $\SP^m(\mrS^c)$ by
-the subspace $\SP^m(\mrS^c)\setminus\cms$;
-\item the homeomorphism $\mathring{\Delta}^{\tup}\to e^{\tup}\subset\cms$ extends now to a map
-$\Delta^{\tup}\to \SP^m(\mrS^c)$, that we can then further project to $\cms^c$: the composition
+\item we identify the one-point compactification
+$(\mrS)^c$ of $\mrS=\T(\S)$ as the quotient of $[0,1]^2$ that identifies left and right intervals as in
+in definition \ref{defn:Tsg} of $\T(\S)$, and moreover collapses $[0,1]^2\setminus Q$ to the point $\infty$;
+\item we consider the $m-$fold symmetric product $\SP^m\pa{(\mrS)^c}$: it contains 
+as an \emph{open} subspace $\cms$, so we can identify $\cms^c$ as the quotient of $\SP^m\pa{(\mrS)^c}$ collapsing
+the subspace $\SP^m\pa{(\mrS)^c}\setminus\cms$ to $\infty$;
+\item the homeomorphism $\mathring{\Delta}^{\tup}\to e^{\tup}\subset\cms$ extends
+now to a map $\Delta^{\tup}\to\SP^m\pa{[0,1]^2}$, that we can then further project to
+$ \SP^m\pa{(\mrS)^c}$ and then to $\cms^c$: the composition
 is the map $\phi^{\tup}$.
@@
 
-Therefore the collection of the $e^{\tup}$, together with the $0-$cell given by the point at infinity,
+Therefore the collection of the $e^{\tup}$, together with the $0-$cell given by $\infty$,
 give a cell decomposition of $\cms^c$, with characteristic maps of cells $\phi^{\tup}$.
@@
 \label{lem:doperatoropenmodtwo}
-Let $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq 2g})$ and
-$\tup'=(l-1,(x'_i)_{i\leq l-1},(u'_i,v'_i)_{i\leq 2g})$
+Let $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$ and
+$\tup'=(l-1,(x'_i)_{i\leq l-1},(u'_i,v'_i)_{i\leq g})$
 be tuples in consecutive degrees $m+l$ and $m+l-1$, and let $[\partial \tup\colon\tup']\in\Z_2$ denote the coefficient
-of $\tup'$ in $\partial \tup$ in the chain complex $\tilde\Ch_*(\cms^c)$.
+of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^c)$.
 Then $[\partial \tup:\tup']=0$ unless one (and exactly one) of the following situations occur:
@@
  for one value $1\leq i\leq l-1$, and shifting the values
- $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case
+ $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. We call this way to obtain $\tup'$ from $\tup$ an \emph{inner differential}.
+ In this case
  \[
@@
  setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and shifting the values
- $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case
+ $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. We call this a \emph{left, outer differential}. In this case
  \[
-  [\partial \tup\colon\tup']=\prod_{i=1}^g\pa{\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}}.
+  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}.
  \]
  This binomial coefficient may be anyway equal to $0\in\Z_2$.
- \item $l\geq 1$ and $\tup h$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
+ \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
  in integers $\delta u_i,\delta v_i\geq 0$
@@
  setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
- Also in this case
+ We call this a \emph{right, outer differential}. Also in this case
  \[
-  [\partial \tup\colon\tup']=\prod_{i=1}^g\pa{\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}}.
+  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}.
  \]
@@
 \begin{proof}
- If we restrict the map $\phi^{\tup}\colon\Delta^{\tup}\to\cms$ to any face of the multisimplex
+ If we restrict the map $\phi^{\tup}\colon\Delta^{\tup}\to\cms^c$ to any face of the multisimplex
  $\Delta^{\tup}$ coming from a factor different from $\Delta^l$,
- we obtain either a constant map to the point at infinity, hence these faces don't contribute
+ we obtain a constant map to $\infty$, hence these faces don't contribute
  to the boundary of $e^{\tup}$ in the reduced cellular chain complex. Instead for $0\leq i\leq l$ the restriction
@@
  \]
- hits the cell $e^{\tup'}$ exactly as many times as specified in the statement of the lemma for the cases $1\leq i\leq l-1$,
- $i=0$ and $i=l$ respectively. We are working in $\Z_2$ so we don't have to be careful about orientations of simplices
+ hits homeomorphically the open cell $e^{\tup'}$
+ exactly as many times as specified in the statement
+ of the lemma for the cases $1\leq i\leq l-1$,
+ $i=0$ and $i=l$ respectively, and no other open cell.
+ We are working in $\Z_2$ so we don't have to be careful about orientations of simplices
  and therefore about signs.
@@
 
-We can filter the chain complex $\tilde\Ch_*(\cms^c)$ by giving weight $\sum_{i=1}^lx_i$ to
+We can filter the chain complex $\tilde\Ch_*(\cms^c)$ by giving norm $\sum_{i=1}^lx_i$ to
 the cell $e^{\tup}$, with $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$. By lemma \ref{lem:doperatoropenmodtwo}
-the weight is weakly decreasing along boundaries. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
-of weight $\leq p$, and let $F_p/F_{p-1}$ be the $p-$th filtration stratum.
+the norm is weakly decreasing along boundaries. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
+of norm $\leq p$, and let $F_p/F_{p-1}$ be the $p-$th filtration stratum.
 
 Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1}))$,
-one copy for each way of splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$, where the isomorphism
-shifts degrees by $p$. Indeed in $F_p/F_{p-1}$ we can compute differentials only according to the first
-of the three possibilities listed in lemma \ref{lem:doperatoropenmodtwo}, in particular the numbers $u_i,v_i$
-don't change along boundaries in $F_p/F_{p-1}$; it is then immediate to identify the \emph{internal} differentials
-with the ones one would get in the case $g=0$.
+one copy for each way of splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$; the isomorphism
+doesn't preserve the degrees but shifts them by $p$.
+
+Indeed in $F_p/F_{p-1}$ all outer
+differentials vanish (see lemma \ref{lem:doperatoropenmodtwo}
+for the definition), in particular the numbers $u_i,v_i$
+don't change along the boundary map of $F_p/F_{p-1}$, and this latter chain complex splits as a direct sum
+of chain complexes indexed by splittings $m-p=\sum_{i=1}^g (u_i+v_i)$ as above.
+It is then immediate to identify the inner differentials
+with the ones one would have in the case $g=0$.
 
 We notice that $\tCh_*(C_p(\Sigma_{0,1}))$ is exactly the chain complex described by Fuchs in
-\cite{Fuchs:CohomBraidModtwo}; in particular its homology is generated by symmetric chains,
-where, for a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_i$,
-the associated symmetric chain is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
-$(l,(x_i)_{i\leq l})$ such that
+\cite{Fuchs:CohomBraidModtwo}; in particular its homology is generated by symmetric chains:
+\begin{defn}
+\label{defn:symchain}
+Given a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_i$,
+the associated symmetric chain in $\tCh_*(C_p(\Sigma_{0,1}))$
+is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
+$\pa{l,(x_i)_{i\leq l}}$ such that
 \begin{itemize}
@@
 \end{itemize}
+Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
+\end{defn}
 
 If we run the Leray spectral sequence associated to the filtered chain complex $\tCh_*(\cms)$,
-the $E^1-$page contains many copies of the homology of $\tCh_*(C_p(\Sigma_{0,1}))$,
-each generated by a symmetric chain. It is straightforward to check that these symmetric
-(arising as cycles in $F_p/F_{p-1}$), are also cycles when lifted, as a sum of the same
-cells, to the complex $\tCh_*(\cms^c)$: one has only to check that the \emph{outer}
-differentials cancel out, and this is easy to see as every \emph{left} differential
-of one cell in the sum of a symmetric chain corresponds to a \emph{right} differential
-of possibly another cell in the same symmetric chain.
+the $E^1-$page contains on the $p-$th column the homology of $F_p/F_{p-1}$; as we have seen this consists of
+many copies of the homology of $\tCh_*(C_p(\Sigma_{0,1}))$, each generated by symmetric chains.
 
-In particular the spectral sequence collapses on its first page and the $\Z_2$-homology of
-$\cms^c$ has a $\Z_2$-basis given by all choices of the following set of data:
-\begin{itemize}
- \item a number $0\leq p\leq m$;
- \item a splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$;
- \item a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ with $\alpha_j\geq 0$.
-\end{itemize}
-The class $(p,(u_i,v_i),(\alpha_j))\in H_*(\cms^c)$ has homological degree
-$m+\sum_{j=1}^{\infty}\alpha_i$.
+\begin{defn}
+\label{defn:gensymchain}
+For a splitting
+\[
+ m=p+(m-p)=\pa{\sum_{j=0}^{\infty}\alpha_j2^j}+\pa{\sum_{i=1}^g(u_i+v_i)}
+\]
+we define a chain in $\tCh_{m+l}(\cms)$, denoted by $|p,(\alpha_j)_j,(u_i,v_i)_{i\leq g}|$: it
+is obtained by summing all $e^{\tup}$ where $\tup$
+ranges among all tuples of the form $\pa{l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g}}$
+satisfying the three properties listed in definition \ref{defn:symchain}.
 
+We call such a chain a
+\emph{generalised symmetric chain}, and its homology class in $\tH_{m+l}(\cms^c)$ is denoted
+by $[p,(\alpha_j),(u_i,v_i)]$
+\end{defn}
+
+It is straightforward to check that a generalised symmetric chain $|p,(\alpha_j),(u_i,v_i)|$
+is not only
+a cycle when projected to its filtration quotient $F_p/F_{p-1}$, as the $E^1-$page
+of the spectral sequence tells us, but also
+in the chain complex $\tCh_*(\cms^c)$ itself. Indeed also all outer
+differentials of a generalised symmetric chain cancel out, since
+every left differential
+of one cell $e^{\tup}$ in the generalised symmetric chain cancels out with a right differential
+of possibly another cell $e^{\tup'}$.
+
+In particular the spectral sequence collapses on its first page and $H_*(\cms^c)$
+has a $\Z_2$-basis given by generalised symmetric chains. We have the following lemma:
+\begin{lem}
+\label{lem:gensymchain}
+The homology $H_*(\cms^c)$ has a graded basis given by the classes $[p,(\alpha_j),(u_i,v_i)]$
+associated to generalised symmetric chains.
+% this basis is bijection with
+% the choices of the following set of data:
+% \begin{itemize}
+%  \item a number $0\leq p\leq m$;
+%  \item a splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$;
+%  \item a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ with $\alpha_j\geq 0$.
+% \end{itemize}
+% The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^c)$ has homological degree
+% $m+\sum_{j=1}^{\infty}\alpha_i$.
+\end{lem}
 % To interpret this class
 % as a monomial in the tensor product of theorem \ref{thm:Hbms*as*ggrep} we need the following definition.
+
+% We are almost ready to prove an isomorphism of bigraded $\Z_2-$vector spaces as in theorem \ref{thm:Hbms*as*ggrep}.
+
 \begin{defn}
  \label{defn:dualHbasis}
-We can see the $\U_i$'s and $\V_i$'s as properly embedded $1-$manifolds in $\mrS=\T(\S)$;
-by Poincaré-Lefschetz duality they represent classes in $H^1(\S)$, and in particular they form a basis
-of this cohomology group. We fix simple closed curves $\u_i,\v_i$ on $\S$, for $1\leq i\leq g$, whose fundamental classes
-form the dual basis of $H_1(\S)$. We assume that, apart from the following necessary exceptions,
-all the curves $\U_i,\V_i,\u_i,\v_i$ for $1\leq i\leq g$ are disjoint:
-\begin{itemize}
- \item $\u_i$ and $\v_i$ intersect once, transversely;
- \item $\u_i$ and $\U_i$ intersect once, transversely;
- \item $\v_i$ and $\V_i$ intersect once, transversely;
- \end{itemize}
-SEE PICTURE.
+We can see the $\U_i$'s and $\V_i$'s as properly embedaded $1-$manifolds in $\mrS$;
+by Poincaré-Lefschetz duality they represent classes in $H_1\pa{(\mrS)^c}\simeq H^1(\S)$,
+and in particular they form a basis
+of this cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
+
+% We fix simple closed curves $\u_i,\v_i$ on $\S$, for $1\leq i\leq g$, whose fundamental classes
+% $[\u_i],[\v_i]$ form the dual basis of $H_1(\S)$. We assume that, apart from the following necessary exceptions,
+% all curves $\U_i,\V_i,\u_i,\v_i$ for $1\leq i\leq g$ are disjoint:
+% \begin{itemize}
+%  \item $\u_i$ and $\v_i$ intersect once, transversely;
+%  \item $\u_i$ and $\U_i$ intersect once, transversely;
+%  \item $\v_i$ and $\V_i$ intersect once, transversely;
+%  \end{itemize}
+% SEE PICTURE.
 \end{defn}
 We establish a bijection between monomials in the tensor product of theorem  \ref{thm:Hbms*as*ggrep}
-and the basis of $H^*(\cms)\simeq H_*(\cms^c)$ that we have just found.
-
-Then the class 
+and the basis of $H^*(\cms)\simeq H_*(\cms^c)$ in lemma \ref{lem:gensymchain}:
+the class 
 \[
-(p,(u_i,v_i),(\alpha_j))\in H_{m+\sum\alpha_j}(\cms^c)\simeq H^{m-\sum\alpha_j}(\cms)
+[p,(u_i,v_i),(\alpha_j)]\in H_{m+\sum\alpha_j}(\cms^c)\simeq H^{m-\sum\alpha_j}(\cms)
 \]
@@
 
-This shows an isomorphism of bigraded vector spaces
+This shows an isomorphism of bigraded $\Z_2-$vector spaces
 \begin{equation}\label{eq:isovectorspaces}
@@
 at least \emph{as bigraded $\Z_2-$vector spaces}.
-
-\subsection{Action of $\gg$} We now turn back to \emph{homology} of $\cms$;
-our first task is to construct a map
-\[
-\psi_m\colon \Sym_m(\H))\to H_m(\cms).
-\]
-% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
-% both by Poincaré-Lefschetz duality.
-
-We already know that these $\Z_2-$vector spaces have the same dimension:
-for the dimension of $H_m(\cms)$ we have to count the monomials of degree $m$ and weight $m$
-in the right-hand side of equation \ref{eq:isovectorspaces}; a monomial whose
-weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
-$Q^i\epsilon$, i.e. it should lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
-is $m$ it should lie in $\Sym_m(\H)$.
-
-The map $\psi_m$ that we construct will turn out to be a $\gg-$equivariant isomorphism.
-We will first construct a map $\tpsi_m\colon \H^{\otimes m}\to H_m(\cms)$ and then show
-that factors through $\Sym_m(\H)$.
-
-% \begin{defn}
-% \label{defn:cCm}
-% Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of closed curves $c_1,\dots,c_m\subset\mrS$
-% such that:
-% \begin{itemize}
-%  \item each $c_i\colon\Sone\to\mrS$ is smooth and a local embedding;
-%  \item the curves $\set{c_i}_{i\leq m}$
-%  are pairwise transverse, no three of them intersect at a common point in $\mrS$.
-% \end{itemize}
-% Two such $m-$tuples are isotopic if there is an ambient isotopy of $\S$
-% transforming one $m-$tuple into the other. For each isotopy class we
-% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
-% also transverse to all $\U_i$'s and $\V_i$'s.
-% 
-% We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
-% There is a canonical surjective map $\pi\colon\ZcC{m}\to\Sym_m(\H)$ given by
-% \[
-%  \pi(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m].
-% \]
-% The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
-% on $\Sym_m(\H)$, symplectically. The map $\pi$ is $\gg-$equivariant.
-% \end{defn}
-% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
-% factors through $\pi$ to give the desired isomorphism $\psi$.
-% 
-% The rough idea is to map the basis element
-% $(c_1,\dots,c_m)$ first to $[c_1]\otimes\dots\otimes[c_m]\in H_m((\S)^{\times m})$, then map
-% this further to $[c_1]\cdot\ldots\cdot[c_m]\in H_m(\SP^m(\S))$ and then lift this to an element in
-% $H_m(\cms)$ along the inclusion $\cms\subset \SP^m(\S)$.
-% 
-We construct $\tpsi_m$ by induction on $m$. For $m=0$ both source and target are
-isomorphic to $\Z_2$; for $m=1$
-both source and target are canonically identified with $\H$: we take $\tpsi_0$ and $\tpsi_1$
-to be these canonical isomorphisms.
-% we simply map $c_1$ to its fundamental class in $\H=H_1(\S)=H_1(C_1(\S))$.
-Let now $m\geq 1$ and suppose that we have defined a map $\tpsi_m\colon \H^{\otimes m}\to H_m(\cms)$.
-
-\begin{defn}\label{defn:variationsCm}
-Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$, using the model $\T(\S)$:
-$\D$ is an open disc in $\mrS$ near the boundary and disjoint from all $\U_i,\V_i$.
-
-We introduce a few variations of the notion of configuration spaces.
-\begin{itemize} 
- \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ of elements
- $(q;\set{p_1,\dots,p_m})$ with $q\neq p_i$ for all $i$; in other words it is
- the space of configurations of $m+1$ points, one of which is \emph{red} (meaning that
- it is special), whereas the other are \emph{black}.
- \item The space $\cmstwo$ is the subspace of $\SP^m(\mrS)$ of configurations where either all $m+1$ points
- are distinct, or there is exactly one point with multiplicity $2$ and $m-1$ other points with multiplicity $1$:
- in the latter case we require the double point to lie in $\D$.
-\end{itemize}
-There is a natural map $\pi\colon\mrS\times \cms\to \cmstwo$ , which converts the red point
-into a black point. This map restricts to a covering map $C_{1,m}(\S)\to C_{m+1}(\S)$; we have a commutative
- diagram
- \begin{equation}\label{eq:cmstwodiagram}
-  \begin{CD}
-   C_{1,m}(\S) @>\subset >> \mrS\times \cms \\
-   @V\pi VV @V\pi VV\\
-   C_{m+1}(\S) @>\subset >> \cmstwo
-   \end{CD}
-\end{equation}
-We notice that all these spaces are $2(m+1)-$dimensional manifolds, arising as open subsets
-of $\mrS\times\SP^m(\mrS)$ or $\SP^{m+1}(\mrS)$, which are manifolds of dimension $2(m+1)$.
-
-Let $\Delta=\cmstwo\setminus C_{m+1}(\S)$ and similarly $\tilde\Delta=\mrS\times \cms\setminus C_{1,m}(\S)$.
-We notice that $\pi$ restricts to a homeomorphism $\tilde\Delta\to\Delta$. Moreover both $\Delta\subset\cmstwo$
-and $\tilde\Delta\subset\mrS\times \cms$ are embedded submanifolds of codimension $2$, and the map $\pi$ restricts to
-a $2-$ramified covering between the normal bundles, which are bundles with fiber a $2-$disc.
-\end{defn}
-
-Diagram \ref{eq:cmstwodiagram} induces a commutative diagram in homology
-\[
-\minCDarrowwidth15pt
- \begin{CD}
-  @. H_{m+1}\pa{\mrS\times \cms} @>>> H_{m+1}\pa{\mrS\times \cms, C_{1,m}(\S)}\\
-  @. @V\pi_*VV @V\pi_*VV\\
-  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwo} @>>> H_{m+1}\pa{\cmstwo,C_{m+1}(\S)}
- \end{CD}
-\]
-
-We can map $\H^{\otimes m+1}$ to $H_{m+1}\pa{\mrS\times \cms}$ through the map
-$\Id_{\H}\otimes\tpsi_m$, and then further down to $H_{m+1}\pa{\cmstwo}$.
-
-We claim that
-we land in the image of the map
-\[
-H_{m+1}\pa{ C_{m+1}(\S)}\to H_{m+1}\pa{\cmstwo}.
-\]
-To check this we further
-map towards right to $H_{m+1}\pa{\cmstwo,C_{m+1}(\S)}$ and prove that we get the zero map.
-Indeed the other equivalent composition
-involves the map
-\[
-\pi_*\colon H_{m+1}\pa{\mrS\times \cms, C_{1,m}(\S)}\to H_{m+1}\pa{\cmstwo,C_{m+1}(\S)};
-\]
-this map can be rewritten, after using the Thom isomorphism, as a map
-\[
- H_{m-1}(\tilde\Delta)\to H_{m-1}(\Delta)
-\]
-and the latter map is multiplication by $2$, as the normal bundle of $\tilde\Delta$ is a
-double covering of the normal bundle of $\Delta$. Therefore the claim is proved, and
-as we are working with $\Z_2-$vector spaces there exists a lift
-\[
- \tpsi_{m+1}\colon\H^{\otimes m+1}\to H_{m+1}(C_{m+1}(\S))
-\]
-of the map $\pi_*\circ(\Id_{\H}\otimes\tpsi_m)\colon \H^{\otimes m+1}\to H_{m+1}\pa{\cmstwo}$.
-
-Thus we have constructed, for each $m\geq 0$, a map $\tpsi_m\colon \H^{\otimes m}$.
-
-\begin{lem}
- \label{lem:cmstocmstwoinjective}
- The map
- \[
-H_m\pa{\cms}\to H_m\pa{C_m^{(2)}}
-\]
-is injective
- \end{lem}
- \begin{proof}
-  We can equivalently check that the map $H^m\pa{\cmstwo}\to H^m\pa{\cms}$ is surjective,
-  or, again, that the map $H_m(\cmstwo^c)\to H_m(\cms^c)$ is surjective.
-
-  The last statement is clear
- \end{proof}

### 2018-10-29 Actiongg.tex
Finished proving the main theorem!

@@
 The construction of $\psi_m$ is rather long and technical and involves a few definitions.
-\begin{defn}
- \label{defn:D}
- Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$, using the model $\T(\S)$:
-$\D$ is an open disc in $\mrS$ near $\partial\S$ and disjoint from all $\U_i$'s and $\V_i$'s.
-
-In the following it is useful to redefine $\gg$ as the group of isotopy classes
-of diffeomorphisms of $\S$ that fix both $\partial \S$ and $\D$ pointwise: this is 
-not a problem, since the inclusion
-\[
-\Diff(\S;\D\cup\partial\S)\subset \Diff(\S;\partial\S)
-\]
-induces a bijection on the set of path-components.
-\end{defn}
 
@@
   & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
-  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*"hook] & H_m(\cmsD)
+  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*"hook] & H_m\pa{\cmsD}
  \end{tikzcd}
@@
 
-Theorem \ref{thm:Hbms*as*ggrep} is then proved in all bidegrees $(*,m)$ with $*=m$.
+Theorem \ref{thm:Hbms*as*ggrep} is then proved in all bidegrees of the form $(m,m)$.
 
@@
 with $0\leq l\leq m$.
-Our aim is to prove theorem \ref{thm:Hbms*as*ggrep} for this bidegree $(m-l,m)$.
-\begin{defn}
-%  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
-%  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
-%  $\D\subset\D'$.
- Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
- Notice that $\S'$ is also a surface of type $\sg$.
+Our aim is to prove theorem \ref{thm:Hbms*as*ggrep} for this bidegree.
 
- For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
- using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
+% \begin{defn}
+% %  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
+% %  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
+% %  $\D\subset\D'$.
+% %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
+% %  Notice that $\S'$ is also a surface of type $\sg$.
+% %  
+%  Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
+%  $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
+%  
+%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
+%  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
+%  
+%  For every splitting $m=p+(m-p)$ there is an open embedding
+%  \[
+%  \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
+%  \]
+%  given by taking the union of configurations:
+% 
+% %  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
+% %  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
+% \end{defn}
+Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
+$\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
 
- For every splitting $m=p+(m-p)$ there is an open embedding
- \[
- \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
- \]
- given by taking the union of configurations:
- \[
-  \mu(\set{P_1,\dots,P_p};\set{P'_1,\dots,P_{m-p}})=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
- \]
-%  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
-%  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
-\end{defn}
 For all $0\leq p\leq m$, the group $\Diff(\S;\partial\S\cup\D)$ acts both on $C_p(\D)\times C_{m-p}(\S')$
-and on $\cms$, and the map $\mu$ is equivariant with respect to this action;
+and on $\cms$, and the map $\mu$ is equivariant with respect to this action
+(see definition \ref{defn:universalSbundle});
 hence, using the K\"{u}nneth formula, there is an induced $\gg-$equivariant map in homology
@@
 
-We will prove the following lemma, from which theorem \ref{thm:Hbms*as*ggrep} easily follows:
+We will prove the following lemma, from which theorem \ref{thm:Hbms*as*ggrep} follows:
 \begin{lem}
@@
 Notice that the statement of the lemma doesn't depend on the the action of $\gg$:
-we have a map between two $\Z_2-$vector spaces, we already know that it is $\gg-$equivariant,
-we only need to show that it is a bijection.
+we have a map between two vector spaces, we already know that it is $\gg-$equivariant,
+we only need to show that it is a bijection. Notice also that lemma \ref{lem:gensymchain}
+implies that the two vector spaces have the same dimension.
 % 
@@
 Fix $l\leq p\leq m$ and let $[a]=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
-$H_{p-l}(C_p(\D))$, i.e., $\sum_j\alpha_j=l$ and $\sum_j2^j\alpha_j=p$; let also
-$\beta=\prod_{i=1}^g \u_i^{u_i}\v_i^{v_i}$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
-isomorphism proved in the previous subsection, i.e., $\sum_i(u_i+v_i)=m-p$. Here $a$ and
+$H_{p-l}(C_p(\D))$, with $\sum_j\alpha_j=l$ and $\sum_j2^j\alpha_j=p$. Let also
+$[b]=\prod_{i=1}^g \u_i^{u_i}\v_i^{v_i}$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
+isomorphism proved in the previous subsection, with $\sum_i(u_i+v_i)=m-p$. Here $a$ and
 $b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
@@
 
-Then $a\otimes b$ represents a generator of $H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))$,
-and we are interested in the homology class represented by $\mu_*(a\otimes b)$.
+Then $[a]\otimes [b]$ is a generator of $H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))$,
+and we are interested in the homology class $\mu_*([a]\otimes [b])$: there is one such
+class for any choice of $[a]$ and $[b]$ as above, that is, for any choice of a string of numbers
+$(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
+such that $\sum_j\alpha_j=l$, $\sum_j\alpha_j2^j=p$ and $\sum_i(u_i+v_i)=m-p$; our aim
+is to show that the collection of all these classes of the form $\mu_*([a]\otimes [b])$
+is a basis for $H_{m-l}(\cms)$.
+
 We will study
@@
 To compute the algebraic intersection
-between $\mu_*(a\otimes b)$ and $[p',(\alpha'_j),(u'_i,v'_i)]$ we
+between $\mu_*([a]\otimes [b])$ and $[p',(\alpha'_j),(u'_i,v'_i)]$ we
 consider the map
@@
 Recall that $\cms^c$ can be filtered according to the norm of cells: a cell $e^{\tup}$ with
-$\tup=(l,(x_i),(u_i,v_i))$ has norm $\sum x_i$. In the previous section we just considered
-the associated filtration of the reduced chain complex $\tCh_*(\cms^c)$, but now we need
-to consider the closed subcomplex $F_p\cms^c\subset\cms^c$ consisting of all cells of norm $\leq p$.
+$\tup=(l,(x_i),(u_i,v_i))$ has norm $\sum x_i$, and the norm is weakly decreasing along boundaries.
+In the previous section we just considered
+the associated filtration of the reduced chain complex $\tCh_*(\cms^c)$, whereas now we
+consider the closed subcomplex $F_p\cms^c\subset\cms^c$ which is the union
+of all cells of norm $\leq p$.
 
@@
 \]
-Consider an open cell
+To see this, consider an open cell
 cell $e^{\tup}$ for some tuple $\tup=(l, (x_i),(u'_i,v'_i))$ of norm $p'\leq p$:
@@
 
-We replace the map $\mu^c$ by a cellular approximation that agrees with it on the $p-$skeleton,
-and by abuse of notation we still call $\mu^c$ the new map.
-
+% We replace the map $\mu^c$ by a cellular approximation that agrees with it on the $p-$skeleton,
+% and by abuse of notation we still call $\mu^c$ the new map.
+% 
 Consider now any generalised symmetric chain $|p',(\alpha'_j),(u'_i,v'_i)|$ representing
@@
 
-We consider the collection of all strings of the form $(p,(u_i,v_i)_{i\leq g},(\alpha_j)_j)$
-satisfying $\sum\alpha_j=l$ and $\sum \alpha_j2^j=p$; we order these strings lexicographically,
+We consider the collection of all strings of the form $(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
+satisfying $\sum\alpha_j=l$,$\sum \alpha_j2^j=p$ and $\sum_i(u_i+v_i)=m-p$;
+we order these strings lexicographically,
 so that in particular $p$ is weakly increasing; we associate
-to each string its corresponding class in $H_{m-l}(\cms)$ of the form $\mu_*(a\otimes b)$ and its
-corresponding class in $\tH_{m+l}(\cms)$ represented by the corresponding generalised symmetric chain.
+to each string its corresponding class in $H_{m-l}(\cms)$ of the form $\mu_*([a]\otimes [b])$ and its
+corresponding class $[p,(\alpha_j),(u_i,v_i)]\in\tH_{m+l}(\cms^c)$.
 
-Then the matrix of algebraic intersections is an upper-triangular matrix
+Then the matrix of algebraic intersections between these two sets of
+classes is an upper-triangular matrix
 with $1$'s on the diagonal, and in particular
-it is invertible. This shows that the set of classes $\mu_*(a\otimes b)$
+it is invertible. This shows that the set of classes $\mu_*([a]\otimes [b])$
 is a basis for $H_{m-l}(\cms)$.
@@
 matrix considered in the end of the previous proof is not only upper-triangular but also
-diagonal. This is however not true, as the following example shows...
+diagonal. This is however not true, as the following example shows.
+
+Let $g=1$, $m=2$, $p=1$, $p'=2$ and consider the classes $[a]=\epsilon\in H_1(C_1(\D))$,
+$[b]=\u_i\in \H=H_1(C_1(\S'))$; moreover let $[p',(\alpha'_j),u'_1,v'_1]$ be defined by
+$p'=2$, $\alpha'_1=1$, $u'_1=v'_1=0$. Represent $[a]$ by a point in $\D$, for example
+the point $(1/2,3/4)$; represent $[b]$ by a simple closed curve $b\in\S'$
+that intersects only once, transversely, the vertical segment passing through $a$, namely
+$\set{1/2}\times(0,1)$.
+
+Then the cycle $|p',(\alpha'_j),u'_1,v'_1|$ consists only of the cell $e^{\tup'}$ with
+$\tup'=(1,(x'_1=2),(u'_1=v'_1=0))$, and this cell intersects once, transversely,
+the cycle $\mu_*(a\otimes b)$, which is represented by the curve of configurations of two
+points in $\mrS$, one of which is fixed at $a$ whereas the other runs along $b$: indeed
+there is exactly one position for the second point to lie under $a$.
+
+Hence the algebraic intersection between these two classes is $1$, and since $p'>p$
+this is an entry strictly above the diagonal in the matrix considered in the proof
+of lemma \ref{lem:oplussplitting}.

### 2018-10-29 HBraidSurf.tex
Finished proving the main theorem!

@@
 \section{Homology of configuration spaces of surfaces}
+\label{sec:HBraidSurf}
 We want to understand the homology groups $H_*(\bms)=H_*(C_m(\sg))$: we
@@
 in the Leray-Serre spectral sequence associated to
-the Birman exact sequence \ref{eq:Birman} or, which is the same, to the bundle
-\ref{eq:Birmanbundle}. In particular we are interested in $H_*(C_m(\S))$
-\emph{as a $\Z_2-$representation of the group $\gg$.}
-% From now on $\Z_2-$coefficients in homology
-% and cohomology will always be understood.
+\ref{eq:Birman}, \ref{eq:Birmanbundle} or \ref{eq:BirmanbundleD}.
+In particular we are interested in $H_*(C_m(\S))$
+as a $\Z_2-$representation of the group $\gg$.
 
-In \cite{LM} L\"offler and Milgram proved that this must be a symplectic
-representation of the mapping class group. By \emph{symplectic} we mean the following:
+In \cite{LM} L\"offler and Milgram proved that this must be a $\Z_2-$symplectic
+representation of the mapping class group. By \emph{$\Z_2-$symplectic} we mean the following:
 \begin{defn}
@@
  The natural action of $\gg$ on $\H$ induces a surjective map
- $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{symplectic}
+ $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{$\Z_2-$symplectic}
  if it is a pull-back of a representation of $Sp_{2g}(\Z_2)$ along this map.
@@
 
-In \cite{BCM} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\sg))$ \emph{as a graded $\Z_2-$vector space}.
+In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\sg))$ \emph{as a graded $\Z_2-$vector space}.
 Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced, as
@@
  \[
-  \bigoplus_{m\geq 0} H_*\pa{C_m(\S)}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
+  \bigoplus_{m\geq 0} H_*\pa{C_m(\sg)}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
  \]
@@
   and write $(*,m)$ for the bidegree;
-  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
-  of the group $H_{2^j-1}(C_{2^j}(\Sigma_{0,1}))\simeq \Z_2$ 
-  under the natural map induced by an embedding $\Sigma_{0,1}\hookrightarrow\S$,
+  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\sg))$ of a generator
+  of the group $H_{2^j-1}(C_{2^j}(\D))\simeq \Z_2$ 
+  under the natural map induced by the embedding $\D\hookrightarrow\mrS$,
   and $\Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]$ is the polynomial ring on
@@
   symmetric algebra on $\H$;
-  \item degrees and weights are extended on right by the usual multiplicativity-additivity rule;
+  %\item degrees and weights are extended on right by the usual multiplicativity-additivity rule;
   \item the action of $\gg$ on right is the diagonal action on the product: it is a trivial action
-  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$ and it is the symplectic action induced by the action on $\H$ on the
-  factor $\Sym_{\bullet}(\H)$.
+  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$ and it is the $\Z_2-$symplectic action on the
+  factor $\Sym_{\bullet}(\H)$ induced by the $\Z_2-$symplectic action on $\H$.
   \end{itemize}
-
+\end{thm}
 Notice that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
@@
 whereas factors in $\H$ and its symmetric powers have equal weight and degree.
-\end{thm}
+
+Notice that in the case $g=0$ the group $\Gamma_{0,1}$ is trivial and the previous theorem
+reduces to equation \ref{eq:Cohen}.
+% isomorphism of bigraded vector spaces
+% \begin{equation}
+% \label{eq:genuszero}
+% \bigoplus_{m\geq 0} H_*\pa{C_m(\Sigma_{0,1})}\simeq \Z_2[Q^j\epsilon\,|\, j\geq 0].
+% \end{equation}
+% This isomorphism was essentially proved by Fuchs in \ref{Fuchs:CohomBraidModtwo}. Moreover the left hand
+% side is the homology of the space $\coprod_{m\geq 0} C_m(\Sigma_{0,1})$, which has a natural structure
+% of algebra over the operad $E^2$ of little $2-$cubes. The isomorphism in equation \ref{eq:genuszero}
+% is an isomorphism of rings (using the Pontryagin product on the left hand side), and $Q$ represents here
+% the Araki-Kudo-Dyer-Lashof operation $Q\colon H_*(X;\Z_2)\to H_{2*+1;\Z_2}(X)$,
+% which is defined for any $E^2-$algebra $X$; $\epsilon\in H_0(C_1(\Sigma_{0,1}))$ is the generator.
+% This explains our notation; we refer to Cohen \cite[Chap.3]{CLM} for more details.
+
 In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2-$vector spaces}
 as in theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
-$\gg$. In these two sections we abbreviate $\S=\sg$.
+$\gg$.
+%In these two sections we abbreviate $\S=\sg$.
 
-Since we work with $\Z_2-$coefficients, it is equivalent to compute homology or cohomology,
-and in this section we will prefer to compute, for each bidegree $(*,m)$, $H^*(\cms)$ as a graded
-$\Z_2-$vector space; the dual graded vector space $H_*(\cms)$ is of course abstractly isomorphic,
-i.e. it has the same dimension.
+Since we work with coefficients in the field $\Z_2$, it is equivalent to compute homology or cohomology,
+and in this section we will prefer to compute $H^*(\cms)$ for all bidegrees $(*,m)$.
+% the dual graded vector space $H_*(\cms)$ is abstractly isomorphic,
+% i.e. it has the same dimension.
 
 We will mimic the method used by Fuchs (\cite{Fuchs:CohomBraidModtwo}) to compute the $\Z_2-$cohomology
-of the configuration space of the plane.
+of $C_m(\D)$.
 As already said this computation recovers a known result, but it has the advantage of
@@
 corresponds to a handle decomposition of $\S$ with one $0-$handle and $2g$ $1-$handles.
+
 \begin{defn}
@@
 Each interval $\U_i,\V_i$ has a natural parametrisation by $(0,1)$.
-\end{defn}
+
 The space $\T(\S)$ is homeomorphic to $\mrS$. From now
@@
 space of configurations of $m$ points in $\T(\S)$.
+
+Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$:
+$\D$ is an open disc in $\mrS$ near $\partial\S$ and 
+it is disjoint from all $\U_i$'s and $\V_i$'s.
+\end{defn}
 
@@
 It is then immediate to identify the inner differentials
-with the ones one would have in the case $g=0$.
+with the ones one would have in the case $g=0$, i.e. for the surface $\Sigma_{0,1}$.
 
-We notice that $\tCh_*(C_p(\Sigma_{0,1}))$ is exactly the chain complex described by Fuchs in
-\cite{Fuchs:CohomBraidModtwo}; in particular its homology is generated by symmetric chains:
+We notice that $\tCh_*(C_p(\Sigma_{0,1})^c)$ is exactly the chain complex described by Fuchs in
+\cite{Fuchs:CohomBraidModtwo}, so we pause for a moment and
+recall some classical facts about the cohomology of configuration
+spaces of the disc.
 \begin{defn}
 \label{defn:symchain}
-Given a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_i$,
-the associated symmetric chain in $\tCh_*(C_p(\Sigma_{0,1}))$
+Given a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_j$,
+the associated symmetric chain in $\tCh_*(C_p(\Sigma_{0,1})^c)$
 is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
@@
 \begin{itemize}
- \item $l=\sum_{i=1}^{\infty}\alpha_i$;
+ \item $l=\sum_{j=1}^{\infty}\alpha_j$;
  \item every $x_i$ is a power of 2;
@@
 Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
+
+We denote by $|(\alpha_j)_{j\geq 0}|\in\tCh_{p+l}(C_p(\Sigma_{0,1})^c)$ the symmetric chain
+and by $[(\alpha_j)_{j\geq 0}]\in\tH_{p+l}(C_p(\Sigma_{0,1})^c)$ the associated homology class.
 \end{defn}
+In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\Sigma_{0,1})^c)$
+is given by the collection of all classes $[(\alpha_j)_{j\geq 0}]$ satisfying $\sum\alpha_j2^j$.
 
-If we run the Leray spectral sequence associated to the filtered chain complex $\tCh_*(\cms)$,
+Using Poincaré-Lefschetz duality we obtain a basis for the cohomology of $C_p(\Sigma_{0,1})$;
+the dual basis of $H_*(C_p(\Sigma_{0,1}))$
+happens to be the basis of monomials
+$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\Sigma_{0,1}))$ of weight $m$, using
+the isomorphism \ref{eq:Cohen} in its full meaning (i.e. as an isomorphism of rings,
+where $Q$ denotes the first Dyer-Lashof operation).
+
+% This notation is compatible, after interpreting equation \ref{eq:genuszero} as an isomorphism
+% of rings, with the results of \cite[Chap.3]{CLM}: the monomials constructed there, using
+% the Pontryagin product and the operation $Q$, are really the dual basis of the basis of
+% symmetric chains for cohomology.
+
+We will not need this finer result in this article, so the expression
+$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_{p-l}(C_p(\Sigma_{0,1}))$ will only denote
+the (unique) homology class whose algebraic intersection with $[(\alpha'_j)]\in\tH_{p+l}(C_p(\Sigma_{0,1})^c)$
+is $1\in\Z_2$ if and only if the equality $\alpha_j=\alpha'_j$ holds for all $j\geq 0$.
+% i.e. we \emph{define} monomials as the dual basis of the basis of symmetric chains, and keep considering
+% \ref{eq:Cohen} only as an isomorphism of bigraded vector spaces.
+
+We now go back to the filtered chain complex $\tCh_*(\cms^c)$:
+if we run the associated Leray spectral sequence,
 the $E^1-$page contains on the $p-$th column the homology of $F_p/F_{p-1}$; as we have seen this consists of
@@
 
-In particular the spectral sequence collapses on its first page and $H_*(\cms^c)$
-has a $\Z_2$-basis given by generalised symmetric chains. We have the following lemma:
+In particular the spectral sequence collapses on its first page and
+% $H_*(\cms^c)$
+% has a $\Z_2$-basis given by generalised symmetric chains.
+we have the following lemma:
 \begin{lem}
@@
 % The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^c)$ has homological degree
-% $m+\sum_{j=1}^{\infty}\alpha_i$.
+% $m+\sum_{j=1}^{\infty}\alpha_j$.
 \end{lem}
@@
 and in particular they form a basis
-of this cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
+of the latter cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
 

### 2018-10-29 Introduction.tex
Finished proving the  gg-equivariant isomorphism
Finished proving the main theorem!

@@
 
-The associated Leray-Serre spectral sequence in $\Z_2-$homology has a second page $E^2_{p,q}=H_p(\gg;H_q(\bms;\Z_2))$,
-and converges to $H_{p+q}(\ggm;\Z_2)$.
+The associated Leray-Serre spectral sequence $E(m)$ in $\Z_2-$homology has a second page $E(m)^2_{k,q}=H_k(\gg;H_q(\bms;\Z_2))$,
+and converges to $H_{k+q}(\ggm;\Z_2)$.
 
@@
 \label{thm:main}
-For each $l\geq 0$ there is an isomorphism
+For all $l\geq 0$ there is an isomorphism
 \begin{equation}
 \label{eq:main}
- H_l(\ggm;\Z_2)\cong \bigoplus_{p+q=l} H_p(\gg;H_q(\bms;\Z_2)).
+ H_l(\ggm;\Z_2)\cong \bigoplus_{k+q=l} H_k(\gg;H_q(\bms;\Z_2)).
 \end{equation}
@@
 WE DO THIS COMPUTATION FOR $g=1$.
-
-From now on $\Z_2$-coefficients will always be understood if no coefficients are specified.

### 2018-10-29 Preliminaries.tex
Finished proving the  gg-equivariant isomorphism
Finished proving the main theorem!

@@
 \section{Preliminaries}
+\label{sec:Preliminaries}
+In the whole article $\Z_2$-coefficients for homology and cohomology will be understood,
+unless explicitly stated otherwise.
+
 In this section we recollect some classical definitions and results about braid groups and mapping class groups.
@@
 \begin{defn}
+\label{defn:cms}
  The $m$-th \emph{ordered configuration space} of a surface $\sg$ is the space
 \[
- F_m(\sg)=\set{(p_1,\dots,p_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  p_i\neq p_j  \,\forall i\neq j}
+ F_m(\sg)=\set{(P_1,\dots,P_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  P_i\neq P_j  \,\forall i\neq j}
 \]
@@
  \label{def:mcg}
- Let $\Diff(\sg)$ be the group of diffeomorphisms of $\sg$ that fix $\partial\sg$ pointwise,
- endowed with the Whitney $C^{\infty}-$topology. We denote by $\gg=\pi_0(\Diff(\sg))$ its group of connected
+ Let $\Diff(\sg;\partial\sg)$ be the group of diffeomorphisms of $\sg$ that fix $\partial\sg$ pointwise,
+ endowed with the Whitney $C^{\infty}-$topology. We denote by $\gg=\pi_0(\Diff(\sg;\partial\sg))$ its group of connected
  components, called also the \emph{mapping class group} of $\sg$.
 
- Similarly let $\Diff(\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
- pointwise and permute the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm))$ be its
+ Similarly let $\Diff(\sgm;\partial\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
+ pointwise and permute the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$ be its
  group of connected components, called the mapping class group of $\sgm$.
@@
  A classical result by Earle and Schatz (\cite{EarleSchatz}) ensures that the connected components
- of $\Diff(\sg)$ are contractible. In particular $B\Diff(\sg)\simeq B\gg$ is a classifying space for
- the group $\gg$ and we call $\cF_{g,1}\to B\gg$ the universal $\sg-$bundle over $B\gg$, corresponding to
- the bundle
+ of $\Diff(\sg)$ are contractible. In particular $B\Diff(\sg;\partial\sg)\simeq B\gg$ is
+ a classifying space for
+ the group $\gg$, i.e. a space of type $K(\gg,1)$; we call $\cF_{g,1}\to B\Diff(\sg;\partial\sg)$ the universal
+ $\sg-$bundle
  \[
-  \sg\times_{\Diff(\sg)}E\Diff(\sg)\to B\Diff(\sg).
+  \sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
  \]
 Applying the construction of the $m-$th unordered configuration space fiberwise we obtain a bundle
-$C_m(\cF_{g,1})\to B\gg$ with fiber $C_m(\sg)$.
+$C_m(\cF_{g,1})\to B\Diff(\sg;\partial\sg)$ with fiber $C_m(\sg)$.
 
@@
 \begin{equation}\label{eq:Birmanbundle}
-C_m(\sg)\to C_m(\cF_{g,1})\to B\gg.
+C_m(\sg)\to C_m(\cF_{g,1})\to B\Diff(\sg;\partial\sg).
 \end{equation}
 
-% \begin{defn}
-%  \label{def:modulispaces}
-% We denote by $\mg$ the moduli space of Riemann surfaces $\sg$ of genus $g$ 
-%  with one parametrised
-%  boundary component, and by $\mgm$ the moduli space of Riemann surfaces $\sgm$ with one
-%  parametrised boundary component and $m$ unordered punctures.
-%  There is a natural fiber bundle map
-%  \[
-%   \pi\colon \mgm\to\mg
-%  \]
-% which forgets the position of the punctures. The fiber of $\pi$ is the configuration space of $m$ unordered
-% points in the interior of $\sg$, i.e.
-% \[
-%  C_m(\sg)=\set{(p_1,\dots,p_m) \in \mathring{\sgigma}_{g,1}^m  \,|\,  p_i\neq p_j  \,\forall i\neq j}
-% \]
-% \end{defn}
-% 
-% The Birman exact sequence \ref{eq:Birman} can be obtained from the bundle
-% \begin{equation}\label{eq:Birmanbundle}
-%  C_m(\sg)\to\mgm\to\mg
-% \end{equation}
-% by computing all fundamental groups: indeed the three spaces appearing in \ref{eq:Birmanbundle} are
-% the classifying spaces of the groups in \ref{eq:Birman}.
+In the whole article the genus $g$ of the surfaces that we consider is supposed to be fixed,
+unless explicitly stated otherwise, and we will abbreviate $\S=\sg$.
+We denote by $\D$ the open disc $\mathring{\Sigma}_{0,1}$.
+
+It will be useful, for many constructions, to choose an embedding
+$\D\hookrightarrow\mrS$ near $\partial\S$ and to replace
+$\Diff(\S;\partial\S)$ with its subgroup $\Diff(\S;\partial\S\cup\D)$ of diffeomorphisms
+of $\S$ that fix pointwise both $\partial\S$ and $\D$: here when we say that $\D$ is embedded
+\emph{near $\partial\S$} we mean that there is a compact subsurface $\S'\subset\S$
+such that $\S$ is the boundary connected sum $\S'\natural\bar\D$.
+From now on we suppose such an embedding to be fixed and we consider $\D$ as a subspace
+of $\mrS$. In section \ref{sec:HBraidSurf} we will specify an
+embedding in a particular model for $\mrS$ that we will define.
+
+The construction of definition \ref{defn:cms} can be specialised to the open
+surfaces $\D$ and $\S'$, yielding spaces $C_m(\D)$ and $C_m(\S')$ respectively.
+
+The inclusion $\Diff(\S;\partial\S\cup\D)\subset\Diff(\S;\partial\S)$ is a homotopy
+equivalence, hence the previous constructions are equivalent to the ones that we
+give in the next definition.
+\begin{defn}
+\label{defn:universalSbundle}
+We call $\cF_{\S,\D}\to B\Diff(\S;\partial\S\cup\D)$
+the universal $\S-$bundle
+ \[
+  \S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
+ \]
+The group $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms;
+the space $\cF_{\S,\D}$ contains subspaces
+\[
+ \cF_{\S'}= \S'\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)
+\]
+and $\D\times B\Diff(\S;\partial\S\cup\D)$, fibering over $B\Diff(\S;\partial\S\cup\D)$
+with fibers $\S'$ and $\D$ respectively.
+
+Applying the construction of the $m-$th unordered configuration space fiberwise, we obtain spaces
+$C_m(\cF_{\S,\D})$, $C_m(\cF_{\S'})$ and $C_m(\D)\times B\Diff(\S;\partial\S\cup\D)$,
+all fibering over $B\Diff(\S;\partial\S\cup\D)$ fibers, respectively, $\cms$,
+$C_m(\S')$ and $C_m(\D)$.
+
+For all $0\leq p\leq m$ there is a natural map
+\[
+ \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
+\]
+which consists on the union of configurations:
+ \[
+  \mu(\set{P_1,\dots,P_p};\set{P'_1,\dots,P'_{m-p}})=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
+ \]
+We can apply this construction to each couple of fibers of $C_m(\cF_{\S,\D})$ and $C_{m-p}(\cF_{\S'})$
+over the same point of $ B\Diff(\S;\partial\S\cup\D)$, obtaining a map $\mu^{\cF}$:
+\[
+ \mu^{\cF}\colon C_p(\D)\times C_{m-p}(\cF_{\S'})\to C_m(\cF_{\S,\D}).
+\]
+If we see the domain of $\mu^{\cF}$ as a fibered product
+\[
+ \pa{C_m(\D)\times B\Diff(\S;\partial\S\cup\D)}\times_{B\Diff(\S;\partial\S\cup\D)}C_{m-p}(\cF_{\S'}),
+\]
+then the map $\mu^{\cF}$ is a map of bundles over $B\Diff(\S;\partial\S\cup\D)$, and corresponding
+map on fibers is precisely $\mu$.
+\end{defn}
+The fiber bundle \ref{eq:Birmanbundle} corresponding to the Birman exact sequence \ref{eq:Birman}
+can now be replaced by the following one
+\begin{equation}\label{eq:BirmanbundleD}
+\cms\to C_m(\cF_{\S,\D})\to B\Diff(\S;\partial\S\cup\D).
+\end{equation}
+We conclude this section by recalling the structure of $H_*(C_m(\D))$.
+
+The cohomology of $C_m(\D)$ with coefficients in $\Z_2$ was first computed by Fuchs in
+\cite{Fuchs:CohomBraidModtwo}: we will recall this computation more in detail in section
+\ref{sec:HBraidSurf}, where we will make in detail a similar computation.
+
+In \cite[Chap.III]{CLM} Cohen considers $\coprod_{m\geq 0}C_m(\D)$ as an
+algebra over the operad of little $2-$cubes, and in particular describes its
+homology with coefficients in $\Z_2$ as follows:
+\begin{equation}
+ \label{eq:Cohen}
+H_*\pa{\coprod_{m\geq 0}C_m(\D)}\simeq \Z_2\left[Q^j\epsilon \, |\, j\geq 0\right].
+\end{equation}
+
+Here $\epsilon\in H_0(C_1(\D))$ is the fundamental class, and for all
+$k,m\geq 0$ we denote by $Q\colon H_k(C_m(\D))\to H_{2k+1}(C_{2m}(\D))$
+the first Dyer-Lashof operation. In particular $Q^j\epsilon$ is the generator of
+$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$.
+
+The isomorphism in equation \ref{eq:Cohen} is
+an isomorphism of bigraded rings: the left-hand side is a ring with the Pontryagin product,
+and the right-hand side is a polynomial ring in the infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
+the bidegree is given by
+$*$, called \emph{degree}, which is the homological degree,
+and $m$, called \emph{weight}, which is the index of the connected component
+on which the homology class is supported, or informally the number of points
+involved in the construction of a homology class.
+
+In this article we will only need the isomorphism in equation \ref{eq:Cohen} to hold as
+an isomorphism of bigraded $\Z_2-$vector spaces.
+In particular for all choices of natural numbers $(\alpha_j)_{j\geq 0}$ with all but finitely
+many $\alpha_j=0$, we will consider the monomial $\prod_j\pa{Q^j\epsilon}^{\alpha_j}$, which is
+a homology class in some bidegree $(*,m)$: all that
+we are going to need is that the set of these monomials is a bigraded basis for the
+left-hand side of equation \ref{eq:Cohen}.

### 2018-10-29 Rational.tex
Finished proving the main theorem!

\section{A rational counterexample}
In this section we prove that a statement as in theorem \ref{thm:Hbms*as*ggrep}
cannot hold if we consider homology with coefficients in $\Q$.

We recall that $H_*(\cms;\Q)$ has been computed \emph{as a bigraded $\Q-$vector space}
in \cite{BCM} and, more recently, in \cite{Knudson}; a description of it as a
$\gg-$representation seems to be still missing in the literature.

In particular we prove the following fact.
\begin{thm}
 \label{thm:counterexample}
 Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\S);\Q)$ is not a symplectic
 representation of $\gg$.
\end{thm}
\begin{proof}
 We will use the following strategy:
 \begin{itemize}
  \item we define a homology class $[a]\in H_2(C_2(\S);\Q)$ represented by a cycle $a$;
  \item we prove that $[a]\neq 0$ by computing the algebraic intersection of
  $[a]$ with a homology class
  in $\tH_2(\cms^c;\Q)$; here it is worth stressing that for all $m$ the manifold
  $\cms$ is orientable: therefore we can apply Poincaré-Lefschetz duality
  also with rational coefficients;
  \item we define another homology class $[b]\in H_2(\cms;\Q)$ and show that
  $[b]$ is mapped to $[b]+2[a]$ by some element in the Torelli group $\mathcal{I}_{g,1}$.
 \end{itemize}
Recall that the Torelli group $\mathcal{I}_{g,1}$ is the kernel of the surjective homomorphism
$\gg\to Sp_{2g}(\Z)$; if an element of the Torelli group acts non-trivially on
some class in $H_2(C_2(\S);\Q)$, then this is not a symplectic representation of $\gg$.

We consider again $\T(\S)$ as model for $\mrS$. Let $c$ be an simple closed curve
representing the homology class $\u_1$, and assume that $c$ intersects $\U_1$ once
transversely and is disjoint from all other $\U_i$'s and from all $\V_i$'s. Let $c'$
be a parallel copy of $c$.

We consider the torus $a=c\times c'$ of configurations in $C_2(\S)$ having one point lying on $c$ and
one lying on $c'$: by abuse of notation, denote also one of its fundamental cycles by $a$
(there are two possible choices, corresponding to the two possible orientations of $a$).

Let $\tup=(0,(u_i,v_i))$ with $u_1=2$ and all other $u_i$'s, as well as all $v_i$'s, equal to zero;
then the map
\[
\phi_{\tup}\colon (\Delta^{\tup},\partial\Delta^{\tup})\to (C_2(\S)^c,\infty)
\]
maps the fundamental class in $H_2(\Delta^{\tup},\partial\Delta^{\tup};\Q)$ to a homology
class in $H_2(C_2(\S);\Q)$; if we see $e^{\tup}$ as a properly embedded, orientable submanifold
of $C_2(\S)$, then we are just considering its fundamental class $[e^{\tup}]\in H_2(C_2(\S)^c)$.

The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore $[a]\cap[e^{\tup}]=\pm 1$,
where the signs depends on how we have chosen orientations on $a$, $e^{\tup}$ and $C_2(\S)$ itself;
in particular $[a]\neq 0$.

The action of $\gg$ on isotopy classes of simple closed curves is transitive, so we can repeat
the construction of the torus $a$ with any other copy of parallel, non-separating simple closed curves
$c$ and $c'$, and the resulting class $[a]$ will always be non-zero in $H_2(C_2(\S);\Q)$.

Let now $d$ and $d'$ be disjoint, non-separating simple closed curves such that
cutting $\S$ along $d$ and $d'$ we obtain a subsurface $\S'\simeq\Sigma_{1,2}$ of genus 1, with boundary
components $d$ and $d'$ (here we need that the genus of $\S$ is at least 2);
suppose moreover that $c$ is a non-separating simple closed curve in $\S'$,
and let $c',c''$ be two parallel copies of $c$ in $\Sigma_{1,2}$, one on each side of a small tubular
neighborhood of $c$.
Then $d,d',c',c''$ are the boundary of a subsurface $\S''\simeq\Sigma_{0,4}\subset\S'$.
Orient all curves $d,d',c,c',c''$ in such a way that the following equalities hold:
\begin{itemize}
 \item $[d]=[d']\in H_1(\S;\Q)$, as witnessed by the subsurface $\S'$;
 \item $[c]=[c']=[c'']\in H_1(\S;\Q)$;
 \item $[d]-[d']+[c']-[c'']=0\in H_1(\S\setminus c;\Q)$, as witnessed by the subsurface $\S''$.
\end{itemize}
See picture PICTURE.

The four tori $c\times c'$, $c\times c''$, $c\times d$ and $c\times d'$ are contained in $C_2(\S)$
and the equality
\[
 [c\times d]-[c\times d']+[c\times c']-[c\times c'']=0\in H_2(C_2(\S))
\]
holds, as witnessed by the homology $c\times \S''\subset C_2(\S)$.

Moreover the classes $[c\times c']$ and $-[c\times c'']$ are \emph{equal}:
indeed there is an isotopy of $\S$ mapping $c$ to $c'$ and $c''$ to $c$,
as oriented curves; the class $[c\times c'']$ is mapped to the class $[c'\times c]=-[c\times c']$.

Let again $a=[c\times c']$ and let $b=[c\times d]$: then the class $b'=[c\times d']$
is equal to $b+2a\in H_2(C_2(\S);\Q)$.

Consider now an element of the Torelli group that fixes $c$ and maps $d$ to $d'$ preserving
the orientation, for example the bounding pair represented in picture PICTURE:
then the class $b$ is mapped to $b'=b+2a\neq b$.
\end{proof}

### 2018-10-29 Ssequence.tex
Finished proving the main theorem!

\section{Proof of theorem \ref{thm:main}}
We will prove theorem \ref{thm:main} by induction on $m$.
The case $m=0$ is trivial. Suppose now $m\geq 1$.

For all $m\geq$ we let $E(m)$ be the Leray-Serre spectral sequence associated with
the bundle \ref{eq:BirmanbundleD}: its second page has the form
\[
 E(m)^2_{k,q}=H_k(B\Diff(\S;\partial\S\cup\D);H_q(\cms)).
\]
From theorem \ref{thm:Hbms*as*ggrep} we know that this spectral sequence is concentrated
on the rows $q=0,\dots, m$.

We want to prove the vanishing of all differentials appearing in the pages
$E(m)^r$ with $r\geq 2$; the $r-$th differential takes the form
\[
 \partial_r\colon E(m)^r_{k,q}\to E(m)^r_{k-r,q+r-1}.
\]
In particular any differential $\partial_r$ exiting from the row $q=m$ is trivial,
because it lands in a higher, hence trivial row.

Fix now $q=m-l<m$, in particular $l\geq 1$; by theorem \ref{thm:Hbms*as*ggrep}, and
in particular by lemma \ref{lem:oplussplitting}, we have a splitting of $H_k(\gg;H_{m-l}(\cms))$ as
\begin{equation}
\label{eq:Hksplitting}
 \bigoplus_{p=l}^m H_k\pa{\gg;\mu_*\pa{H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}};
\end{equation} 
we will now fix $l\leq p\leq m$ and show the vanishing of all differentials $\partial_r$ exiting from the
summand with label $p$ in equation \ref{eq:Hksplitting}.

Consider the map $\mu^{\cF}$ from definition \ref{defn:universalSbundle} as a map of
bundles over the space $B\Diff(\S;\partial\S\cup\D)$:
\[
 \mu^{\cF}\colon C_p(\D)\times C_{m-p}(\cF_{\S'})\to C_m(\cF_{\S,\D}).
\]
Notice that the first bundle $C_p(\D)\times C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$
is the product of the space $C_p(\D)$ with the bundle $C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$;
therefore the spectral sequence associated with $C_p(\D)\times C_{m-p}(\cF_{\S'})$ is isomorphic,
from the second page on, to the tensor product
of $H_*(C_p(\D))$ and the spectral sequence associated with the bundle $(\cF_{\S'})$; the
latter is isomorphic, in our notation, to the spectral sequence $E(m-p)$.
In particular $\mu^{\cF}$ induces a map of spectral
sequences
\[
\mu^{\cF}_*\colon H_*(C_p(\D))\otimes E(m-p)\to E(m);
\]
that in the second page, on the $(m-l)-$th row, restricts to
the inclusion of one of the direct summands in equation \ref{eq:Hksplitting}:
\[
H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}\subset H_k\pa{\gg;H_{m-l}(\cms)}.
\]

In particular if we prove the vanishing of all differentials $\partial_r$
% exiting from the summand
% $H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}$
in the first spectral sequence, then also all differentials
$\partial_r$ exiting from the summand $H_k\pa{\gg;H_{m-l}(\cms)}$ in the
second spectral sequence must vanish.

% We observe that there is an isomorphism
% \[
%  H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}\simeq H_{p-l}(C_p(\D))\otimes H_k\pa{\gg; H_{m-p}(C_{m-p}(\S'))};
% \]
The differentials in the spectral sequence $H_*(C_p(\D))\otimes E(m-p)$
are obtained by tensoring the identity of $H_*(C_p(\D))$ with the differentials
of the spectral sequence $E(m-p)$; as $p\geq l\geq 1$ we know by inductive hypothesis
that the latter vanish. Theorem \ref{thm:main} is proved.

### 2018-11-05 Actiongg.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 \subsection{Bidigrees of the form $(m,m)$}
-In this subsection we construct a $\gg-$equivariant isomorphism 
+In this subsection we construct a $\gg-$~equivariant isomorphism 
 \[
@@
 We already know that these $\Z_2-$vector spaces have the same dimension:
-for the dimension of $H_m(\cms)$ we have to count the monomials of degree $m$ and weight $m$
-in the right-hand side of equation \ref{eq:isovectorspaces}; a monomial whose
+indeed $\Sym_m(\H)$ is precisely the summand in bidegree $(m,m)$ in equation
+\ref{eq:isovectorspaces}, using that a monomial
+ whose
 weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
-$Q^i\epsilon$, i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
-is $m$ it must lie in $\Sym_m(\H)$.
+$Q^i\epsilon$.
+% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
+% is $m$ it must lie in $\Sym_m(\H)$.
 % 
@@
 % \[
-% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m\pa{\mrS}
+% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
 % \]
 % which carries the fundamental class of the $m-$fold torus to some class in
-% $H_m(\SP^m\pa{\mrS})$; we then lifts this class, in a nice way, to a class
+% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
 % in $H_m(\cms)$.
@@
 Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
-such that:
-\begin{itemize}
+such that
+%\begin{itemize}
 %  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
- \item the curves $\set{c_i}_{i\leq m}$ intersect each other only inside $\D$;
- \item each curve $c_i$ is transverse to $\partial\D$.
-\end{itemize}
-Two such $m-$tuples are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
+% \item
+ any two curves $c_i,c_j$ intersect each other only inside $\D$.
+
+% \item each curve $c_i$ is transverse to $\partial\D$.
+%\end{itemize}
+Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
+images.
+
+Two $m-$tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
 transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.
+
+We call \emph{multicurve} an element of $\cC^m$; by abuse of notation
+$(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
+See picture \ref{fig:defcCm}
 % For each isotopy class we
@@
 \[
- \pr_m([c_1,\dots,c_m])=[c_1]\cdot\ldots\cdot[c_m],
+ \pr_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m],
 \]
@@
 \end{defn}
+
+\begin{figure}\centering
+ \includegraphics[scale=0.7]{figures/defcCm.png}
+ \caption{A multicurve in $\cC^5$.}
+\label{fig:defcCm}
+\end{figure}
+
 % Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
@@
  \label{defn:cmsD}
- The space $\cmsD$ is the subspace of $\SP^{m}\pa{\mrS}$ of configurations where all points
- having multiplicity $\geq 2$ lie inside $D$. The space $\cmsD$ is an open subset
- of $\SP^m\pa{\mrS}$, which is a $2m-$dimensional manifold, hence $\cmsD$ is also a manifold.
+ The space $\cmsD$ is the subspace of $\SP^{m}(\mrS)$ of configurations where all points
+ having multiplicity $\geq 2$ lie inside $D$.
 
- There is a natural inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
-  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class in $\cC^m$ represented
-  by a $m-$tuple $(c_1,\dots,c_m)$, the composition
+ Recall that the $m-$fold symmetric product of an open surface is a $2m-$dimensional manifold;
+ the space $\cmsD$ is an open subset
+ of $\SP^m(\mrS)$, hence $\cmsD$ is also a manifold.
+ See figure \ref{fig:defcmsD}.
+
+ There is a natural, open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
+  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
+  the composition
   \[
    \begin{CD}
-    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> \pa{\mrS}^{\times m} @>>> \SP^m\pa{\mrS}
+    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> (\mrS)^{\times m} @>>> \SP^m(\mrS)
    \end{CD}
@@
 of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
-if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple which is isotopic relative
-to $\D\cup \partial\S$.
+if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple in the same multicurve.
 
@@
  It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
- or, using Poincaré-Lefschetz duality, that the map $H_m(\cmsD^c)\to H_m(\cms^c)$
+ or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$
  is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
@@
 
- Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $H_m(\cms^c)$
+ Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $\tH_m(\cms^c)$
  is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
- and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$.
+ and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$ for
+ such classes.
 
@@
  It is straightforward to check that the map $\phi^{\tup}$ factors through
- $(\cmsD,\infty)$, as a map of pairs; surjectivity of $H_m(\cmsD^c)\to H_m(\cms^c)$ follows.
+ $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$ follows.
 \end{proof}
@@
 Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
-lands in the image of $(\iota_m)_*$, and the last statement doesn't depend on how $\gg$ acts on these groups.
+lands in the image of $(\iota_m)_*$, and this last statement doesn't depend on how $\gg$ acts on these groups.
 
@@
 \label{defn:variationsCm}
-We introduce a few variations of the notion of configuration space.
+We introduce a few variations of the notion of configuration space; see figure \ref{fig:defcmsD}.
 \begin{itemize} 
@@
  $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
- the space of configurations of $m+1$ points, one of which is \emph{red} (meaning that
+ the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
  it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
@@
 %  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
-%  Again $q$ is the red point
+%  Again $q$ is thewhite point
  \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
@@
  $\bar P\not\in \D$ must be distinct from all $P_i$'s.
- Again $\bar P$ is called the red point.
+ Again $\bar P$ is called the white point.
  %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
- \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}\pa{\mrS}$ of configurations where either all $m+1$ points
+ \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
  are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
- somewhere in $\S$, with multiplicity $1$.
+ somewhere in $\mrS$, with multiplicity $1$.
 \end{itemize}
@@
 
-We the following natural inclusions:
+\begin{figure}\centering
+ \includegraphics{figures/defcmsD.png}
+ \caption{A configuration in each of the space introduced in definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
+ Whenever a multiplicity is not specified, it is equal to 1.}
+\label{fig:defcmsD}
+\end{figure}
+
+We have the following natural inclusions:
 \[
-C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m\pa{\mrS};
+C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
 \]
 \[
-C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}\pa{\mrS}.
+C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}(\mrS).
 \]
@@
 
-Fix a small neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.
+Fix a neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.
 
@@
 
-There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the red point
+There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
 into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
-a map $C_{1,m}(\S)\to C_{m+1}(\S)$;
+a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
 % the latter being even a $m+1-$fold covering map;
@@
 We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
-and $\Donem\subset\fmstwoD$ are properly embedded submanifolds of codimension $2$, and the map $\p$ restricts to
-a $2-$ramified covering between their respective normal bundles.
+and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
+a $2-$fold ramified covering between their respective normal bundles.
 \end{defn}
@@
 which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
-to be red. We then notice that the right vertical map
+to be white. We then notice that the right vertical map
 \[
-\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)};
+\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
 \]
-can be rewritten, after using excision to a tubular neighborhood of $\Donem$ and $\Dmone$ and
-the Thom isomorphism, as a map
+can be rewritten, after using excision to tubular neighborhoods of $\Donem$ and $\Dmone$ respectively,
+and the Thom isomorphism, as a map
 \[
- H_{m-1}(\Donem)\to H_{m-1}(\Dmone);
+ H_{m-1}(\Donem)\to H_{m-1}(\Dmone).
 \]
-the latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
+The latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
 indeed the normal bundle of $\Donem$ is a
@@
 
-But this is easily done by considering, in the whole argument of the proof, the surface $\N$ instead of $\mrS$:
-we can define configuration spaces as in definition \ref{defn:variationsCm} using the open surface
-$\N$ instead of $\mrS$, and repeat the whole argument considering this as the \emph{ambient surface}:
-indeed we only need $\N$ to contain $\D$ and all curves $c_1,\dots,c_{m+1}$.
+This can be done by replacing, in the whole argument of the proof, the surface $\mrS$ with the surface $\N$.
+We can define configuration spaces as in definition \ref{defn:variationsCm} also for the open surface
+$\N$, and we can repeat the argument considering $\N$ as the \emph{ambient surface}:
+indeed we only needed a surface containing $\D$ and all curves $c_1,\dots,c_{m+1}$.
 
-Here it is crucial that the action of $\gg$ is not involved in the statement of lemma
+It is crucial that the action of $\gg$ is not involved in the statement of lemma
 \ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
 by diffeomorphisms of $\S$.
-
 Lemma \ref{lem:tpsiwithproperties} is proved.
@@
 
-Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $H_m(\cms^c)\simeq H^m(\cms)$, which is
+Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $\tH_m(\cms^c)\simeq H^m(\cms)$, which is
 the dual of $H_m(\cms)$, is given by the classes $[0,(u_i,v_i)]$, represented by generalised
@@
 
-This shows that $\psi_m(c_1,\dots,c_m)$ must be $[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
+This shows that $\psi_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
 one of the generating monomials.
@@
 
-Theorem \ref{thm:Hbms*as*ggrep} is then proved in all bidegrees of the form $(m,m)$.
+Theorem \ref{thm:Hbms*as*ggrep} is now proved in all bidegrees of the form $(m,m)$.
 
@@
 % \end{defn}
-Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
-$\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
+Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
+$\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.
 
@@
 is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
-of the $\gg-$representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
+of the representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
 section is isomorphic to the symplectic representation $\Sym_{m-p}(\H)$.
@@
 of all these maps yields a splitting
-\[
- H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}
-\]
+\begin{equation}
+  \label{eq:musplitting}
+ H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}.
+\end{equation}
 \end{lem}
@@
 Notice that the statement of the lemma doesn't depend on the the action of $\gg$:
-we have a map between two vector spaces, we already know that it is $\gg-$equivariant,
-we only need to show that it is a bijection. Notice also that lemma \ref{lem:gensymchain}
+we have a map from the right-hand side to the left-hand side of equation \ref{eq:musplitting},
+we already know that it is $\gg-$equivariant,
+we only need to show that it is a linear isomorphism. Notice also that lemma \ref{lem:gensymchain}
 implies that the two vector spaces have the same dimension.
@@
 Fix $l\leq p\leq m$ and let $[a]=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
-$H_{p-l}(C_p(\D))$, with $\sum_j\alpha_j=l$ and $\sum_j2^j\alpha_j=p$. Let also
-$[b]=\prod_{i=1}^g \u_i^{u_i}\v_i^{v_i}$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
-isomorphism proved in the previous subsection, with $\sum_i(u_i+v_i)=m-p$. Here $a$ and
+$H_{p-l}(C_p(\D))$, hence $l=\sum_j\alpha_j$ and $p=\sum_j\alpha_j2^j$. Let also
+$[b]=\prod_{i=1}^g (\u_i^{u_i}\v_i^{v_i})$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
+isomorphism proved in the previous subsection, with $(m-p)=\sum_i(u_i+v_i)$. Here $a$ and
 $b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
-and $b$ supported on $\S'$.
+and $b$ supported on $\S'$. See figure \ref{fig:axb}
+
+\begin{figure}\centering
+ \includegraphics[scale=0.65]{figures/axb.png}
+ \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_3(\D))$;
+ the class $[b]=\v_1\cdot\u_2\cdot\v_3^2\in H_4(C_4(\S'))$; and the product class
+ $\mu_*([a]\otimes[b])\in H_6(C_9(\S))$.}
+\label{fig:axb}
+\end{figure}
 
@@
 $(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
-such that $\sum_j\alpha_j=l$, $\sum_j\alpha_j2^j=p$ and $\sum_i(u_i+v_i)=m-p$; our aim
-is to show that the collection of all these classes of the form $\mu_*([a]\otimes [b])$
-is a basis for $H_{m-l}(\cms)$.
+such that $l=\sum_j\alpha_j$, $p=\sum_j\alpha_j2^j$ and $(m-p)=\sum_i(u_i+v_i)$; our aim
+is to show that the collection of all the corresponding classes of the form $\mu_*([a]\otimes [b])$
+gives a basis for $H_{m-l}(\cms)$.
 
@@
 
-By Poincaré-Lefschetz duality the map $\mu^c_*$ in reduced homology corresponds to the cohomology map
+By Poincaré-Lefschetz duality, the map $\mu^c_*$ in reduced homology corresponds to the cohomology map
 \[
@@
 
-We give $\pa{C_p(\D)\times C_{m-p}(\S')}^c$ the cell structure of the smash product
-$C_p(\D)^c\wedge C_{m-p}(\S')^c$. Here $C_p(\D)^c$ is given the cell structure of $C_p((0,1)^2)$
-along the natural identification $\D=(1/4,3/4)\times(1/2,1)\cong(0,1)^2$ obtained by rescaling
-and translating; moreover we choose any diffeomorphism $\mrS'\cong\mrS$ that restricts to the identity
+We give $\pa{C_p(\D)\times C_{m-p}(\S')}^c$ the cell complex structure of the smash product
+$C_p(\D)^c\wedge C_{m-p}(\S')^c$. Here $C_p(\D)^c$ is given the cell structure of $C_p((0,1)^2)^c$
+coming from the natural identification $\D=(1/4,3/4)\times(1/2,1)\cong(0,1)^2$, which is obtained by rescaling
+and translating. Moreover we choose any diffeomorphism $\mrS'\cong\mrS$ that restricts to the identity
 on all $\U_i$'s and $\V_i$'s, and give $C_{m-p}(\S')^c$ the cell structure of $C_{m-p}(\S)^c$.
@@
 the associated filtration of the reduced chain complex $\tCh_*(\cms^c)$, whereas now we
-consider the closed subcomplex $F_p\cms^c\subset\cms^c$ which is the union
+consider the closed subcomplex $F_p\cms^c\subset\cms^c$, which is the union
 of all cells of norm $\leq p$.
 
-The crucial remark is that $\mu^c$ restricts to a cellular map 
+The crucial observation is that $\mu^c$ restricts to a cellular map 
 \[
@@
 \]
+
 To see this, consider an open cell
-cell $e^{\tup}$ for some tuple $\tup=(l, (x_i),(u'_i,v'_i))$ of norm $p'\leq p$:
-its intersection with the subspace
-$C_p(\D)\times C_{m-p}(\S')$ is empty if $p'<p$, and it is exactly the open product of cells
-$e^{\tup'}\times e^{\tup''}$,
-where $\tup'=(l,(x_i))$ and $\tup''=(0,(u'_i,v'_i))$, if $p'=p$; therefore
+cell $e^{\tup}\subset\cms$ for some tuple $\tup=(l', (x_i),(u'_i,v'_i))$ of norm $p'\leq p$.
+If $p'<p$, then
+$e^{\tup}\cap (C_p(\D)\times C_{m-p}(\S'))$ is empty. If $p'=p$, then
+\[
+e^{\tup}\cap (C_p(\D)\times C_{m-p}(\S'))=e^{\tup'}\times e^{\tup''},
+\]
+where $\tup'=(l',(x_i))$ and $\tup''=(0,(u'_i,v'_i))$.
+
+Therefore
 $\mu^c(e^{\tup})$ is $\set{\infty}$ in the first case, and in the second case it
@@
 Consider now any generalised symmetric chain $|p',(\alpha'_j),(u'_i,v'_i)|$ representing
-a class in $H_{m+l}(\cms^c)=H^{m-l}(\cms)$, hence $\sum_j\alpha'_j=l$, and suppose $p'\leq p$.
+a class in $\tH_{m+l}(\cms^c)=H^{m-l}(\cms)$, hence $l=\sum_j\alpha'_j$; and suppose $p'\leq p$.
 
@@
 is mapped along $\mu^c_*$ to the class
-$[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]\in\tH\pa{C_p(\D)^c\wedge C_{m-p}(\S')^c}$
-Indeed the previous
-argument shows that each cell appearing in the cycle $|p,(\alpha'_j),(u'_i,v'_i)|$ covers with degree
-1 a corresponding cell of the cycle $|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|$, and no other cell
-of the same dimension: hence at the level of chains we have 
+\[
+[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]\in\tH\pa{C_p(\D)^c\wedge C_{m-p}(\S')^c}.
+\]
+Indeed each cell appearing in the cycle $|p,(\alpha'_j),(u'_i,v'_i)|$ is mapped with degree
+1 onto a corresponding cell of the cycle $|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|$, and hits no other cell
+in dimension $m+l$: hence even at the level of chains we have 
 \[
@@
 
-For $p'=p$ we see that the intersection between the homology classes
-$a\otimes b$ and $\mu^c_*([p,(\alpha'_j),(u'_i,v'_i)])=[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]$
+For $p'=p$ we see that the intersection between
+$[a]\otimes [b]$ and $\mu^c_*([p,(\alpha'_j),(u'_i,v'_i)])=[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]$
 is $1\in\Z_2$ exactly when $u_i=u'_i$ and $v_i=v'_i$ for all $i$, and $\alpha_j=\alpha'_j$
@@
 We consider the collection of all strings of the form $(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
-satisfying $\sum\alpha_j=l$,$\sum \alpha_j2^j=p$ and $\sum_i(u_i+v_i)=m-p$;
+satisfying $l=\sum_j\alpha_j$,$p=\sum_j \alpha_j2^j$ and $(m-p)=\sum_i(u_i+v_i)$;
 we order these strings lexicographically,
@@
 One could expect that the basis given by all elements $[a]\otimes [b]\in H_{m-l}(\cms)$
-is dual to the basis of classes $[p',(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$, i.e., the
+is also \emph{dual} to the basis of classes $[p',(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$, i.e., the
 matrix considered in the end of the previous proof is not only upper-triangular but also
@@
 Let $g=1$, $m=2$, $p=1$, $p'=2$ and consider the classes $[a]=\epsilon\in H_1(C_1(\D))$,
-$[b]=\u_i\in \H=H_1(C_1(\S'))$; moreover let $[p',(\alpha'_j),u'_1,v'_1]$ be defined by
-$p'=2$, $\alpha'_1=1$, $u'_1=v'_1=0$. Represent $[a]$ by a point in $\D$, for example
-the point $(1/2,3/4)$; represent $[b]$ by a simple closed curve $b\in\S'$
-that intersects only once, transversely, the vertical segment passing through $a$, namely
-$\set{1/2}\times(0,1)$.
+$[b]=\u_1\in \H=H_1(C_1(\S'))$; moreover let $[p',(\alpha'_j),u'_1,v'_1]$ be defined by
+$p'=2$, $\alpha'_1=1$, $u'_1=v'_1=0$.
 
-Then the cycle $|p',(\alpha'_j),u'_1,v'_1|$ consists only of the cell $e^{\tup'}$ with
-$\tup'=(1,(x'_1=2),(u'_1=v'_1=0))$, and this cell intersects once, transversely,
+Represent $[a]$ by a point in $a\in\D$, for example
+the point $(1/2,3/4)$; represent $[b]$ by a simple closed curve $b\subset\S'$
+that intersects only once, transversely, the vertical segment passing through $a$, i.e.
+$\set{1/2}\times(0,1)$. See figure \ref{fig:counterexample}.
+
+\begin{figure}\centering
+ \includegraphics[scale=0.6]{figures/counterexample.png}
+ \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}
+\label{fig:counterexample}
+\end{figure}
+
+Then the cycle $|p',(\alpha'_j),u'_1,v'_1|$ consists uniquely of one cell $e^{\tup'}$, where
+$\tup'=(1,(x'_1=2),(u'_1=v'_1=0))$; and this cell intersects once, transversely,
 the cycle $\mu_*(a\otimes b)$, which is represented by the curve of configurations of two
@@
 of lemma \ref{lem:oplussplitting}.
+
+The proof of theorem \ref{thm:Hbms*as*ggrep} can be easily generalised to surfaces
+with more than one boundary component. Let $\Sigma_{g,n}$ be a surface of genus
+$g$ with $n\geq 1$ parametrised boundary components and let $\Gamma_{g,n}$ be
+the group of connected components of the topological group
+$\Diff(\Sigma_{g,n};\partial\Sigma_{g,n})$:
+then there is an isomorphism of bigraded $\Z_2-$representations
+\[
+ \bigoplus_{m\geq 0}H_*\pa{C_m(\Sigma_{g,n})}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(H_1(\Sigma_{g,m})),
+\]
+where the action of $\Gamma_{g,n}$ on the right-hand side is induced by the natural
+action on $H_1(\Sigma_{g,n})$; for $n\geq 2$ the $\Z_2-$vector space $H_1(\Sigma_{g,n})$
+doesn't carry a \emph{non-degenerate} symplectic form which is invariant under the action of
+$\Gamma_{g,n}$, hence we can't any longer use the word \emph{symplectic} for this representation.
+
+The proof is word by word the same; the main difference is in the construction of the
+model $\T(\Sigma_{g,n})$ for $\mathring{\Sigma}_{g,n}$: we divide the
+vertical segments $\set{0,1}\times[0,1]\subset[0,1]^2$
+into $2g+n-1$ equal parts, that we call $I_i^l$ and $I_i^r$ according to their order;
+we identify, for each $i>2g$, the interval $I_i^l$ with the interval $I_i^r$;
+the other couples of intervals, yielding the genus, are identified just as before.
+
+One can further generalise to non-orientable surfaces with non-empty boundary: it suffices,
+in the above construction, to glue some of the intervals $I_i^l$ and $I_i^r$ reversing their
+orientation. We leave all details of these generalisations to the interested reader.

### 2018-11-05 GenusOne.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

\section{Computation in genus one}
We can apply the results from the previous sections, in particular theorems \ref{thm:main}
and \ref{thm:Hbms*as*ggrep}, to compute the groups $H_*(\gone^m)$ explicitly.

Recall that $\gone$ is generated by two Dehn twists $\tu$ and $\tv$ about two simple
closed curves $\u,\v\subset\Sigma_{1,1}$ that intersect once transversely; a presentation
for $\gone$ is given by
\[
 \gone=\left<\tu,\tv\, |\, \tu\tv\tu=\tv\tu\tv \right>
\]
and therefore $\gone$ is isomorphic to the braid group on three strands $\beta_3=\beta_3(\D)$.

### 2018-11-05 HBraidSurf.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 \label{sec:HBraidSurf}
-We want to understand the homology groups $H_*(\bms)=H_*(C_m(\sg))$: we
-are interested in these groups as they appear as the homology of the fiber
+We want to study $H_*(\bms)=H_*(\cms)$,
+which appears as the homology of the fiber
 in the Leray-Serre spectral sequence associated to
@@
 
-In \cite{LM} L\"offler and Milgram proved that this must be a $\Z_2-$symplectic
+In \cite{LM} L\"offler and Milgram implicitly proved that $H_*(\bms)=H_*(\cms)$ is a $\Z_2-$symplectic
 representation of the mapping class group. By \emph{$\Z_2-$symplectic} we mean the following:
@@
 
-In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(C_m(\sg))$ \emph{as a graded $\Z_2-$vector space}.
-Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced, as
-their descripition of $H_*(C_m(\sg))$ depends on a handle decomposition of $\sg$, which is not preserved,
-not even up to isotopy, by diffeomorphisms of $\sg$.
+In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(\cms)$ \emph{as a graded $\Z_2-$vector space}.
+Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced:
+their descripition of $H_*(\cms)$ depends on a handle decomposition of $\S$, which is not preserved
+by diffeomorphisms of $\S$, not even up to isotopy,.
 
@@
  \[
-  \bigoplus_{m\geq 0} H_*\pa{C_m(\sg)}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
+  \bigoplus_{m\geq 0} H_*\pa{\cms}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
  \]
@@
   and write $(*,m)$ for the bidegree;
-  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\sg))$ of a generator
+  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
   of the group $H_{2^j-1}(C_{2^j}(\D))\simeq \Z_2$ 
@@
   %\item degrees and weights are extended on right by the usual multiplicativity-additivity rule;
-  \item the action of $\gg$ on right is the diagonal action on the product: it is a trivial action
-  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$ and it is the $\Z_2-$symplectic action on the
-  factor $\Sym_{\bullet}(\H)$ induced by the $\Z_2-$symplectic action on $\H$.
+  \item the action of $\gg$ on right is the tensor product of the trivial action
+  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$, and of the action on
+  $\Sym_{\bullet}(\H)$ which is induced by the $\Z_2-$symplectic action on $\H$.
   \end{itemize}
@@
 \label{defn:Tsg}
-If $g=0$, hence $\S=\sg$ is the disc, we set $\T(\S)=(0,1)^2$, the interior of the unit square. Assume
-now $g\geq 1$.
+If $g=0$, hence $\S=\Sigma_{0,1}$ is the disc, we set $\T(\S)=(0,1)^2$, the interior of the unit square. Assume
+now $g\geq 1$, and see figure \ref{fig:defTS} to visualize the following construction.
 
@@
 
-Let $\D\subset\mrS$ be the open square $(1/4;3/4)\times(1/2,1)$:
+In this model we fix $\D\subset\mrS$ to be the open square $(1/4;3/4)\times(1/2,1)$:
 $\D$ is an open disc in $\mrS$ near $\partial\S$ and 
-it is disjoint from all $\U_i$'s and $\V_i$'s.
+it is disjoint from all $\U_i$'s and $\V_i$'s. The interior of the surface $\S'$
+is then identified with $\T(\S)\setminus [1/4,3/4]\times[1/2,1)$.
+
 \end{defn}
+
+\begin{figure}\centering
+ \includegraphics[scale=0.7]{figures/defTS.png}
+ \caption{The space $\T(\S)$ for $\S=\Sigma_{3,1}$.}
+\label{fig:defTS}
+\end{figure}
 
@@
 % $\bsymb=\set{(U_i,V_i)_{i\leq g},(W_i)_{i\leq n-1}}\subset\symb$.
-
+In symbols we write $\tup=(l,(x_i)_{1\leq i\leq l},(u_i,v_i)_{1\leq i\leq g})$.
 The \emph{degree} of $\tup$ is defined as $m+l$.
@@
 For a tuple $\tup$ let $e^{\tup}$ be the subspace
-of $\cms$ of configurations of $m$ points in $\mrS$ such that
+of $\cms$ of configurations of $m$ points in $\mrS$ such that the following conditions hold
+(see picture \ref{fig:defetup}):
 \begin{itemize}
@@
 \end{defn}
+
+\begin{figure}\centering
+ \includegraphics[scale=0.7]{figures/defetup.png}
+ \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,0,2,3,0,2))$. }
+\label{fig:defetup}
+\end{figure}
+
 The space $e^{\tup}$ is homeomorphic to \emph{the interior} of the following product of simplices:
@@
 
-Therefore the collection of the $e^{\tup}$, together with the $0-$cell given by $\infty$,
-give a cell decomposition of $\cms^c$, with characteristic maps of cells $\phi^{\tup}$.
+Therefore the collection of the $e^{\tup}$'s, together with the $0-$cell $\infty$,
+gives a cell decomposition of $\cms^c$, with characteristic maps of cells $\phi^{\tup}$.
 
-We can easily use these characteristic maps to compute the \emph{reduced} cellular chain complex of $\cms^c$
+We can use these characteristic maps to compute the \emph{reduced} cellular chain complex of $\cms^c$
 with coefficients in $\Z_2$, which
@@
  \]
- This binomial coefficient may be anyway equal to $0\in\Z_2$.
+ This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
  \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
@@
  \]
- This binomial coefficient may be anyway equal to $0\in\Z_2$.
+ This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
 \end{itemize}
@@
  of the lemma for the cases $1\leq i\leq l-1$,
- $i=0$ and $i=l$ respectively, and no other open cell.
- We are working in $\Z_2$ so we don't have to be careful about orientations of simplices
+ $i=0$ and $i=l$ respectively, and no other open cell of dimension $m+l-1$.
+ We are working in $\Z_2$ so we don't have to be careful about orientations of cells
  and therefore about signs.
@@
 We can filter the chain complex $\tilde\Ch_*(\cms^c)$ by giving norm $\sum_{i=1}^lx_i$ to
-the cell $e^{\tup}$, with $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$. By lemma \ref{lem:doperatoropenmodtwo}
-the norm is weakly decreasing along boundaries. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
+the cell $e^{\tup}$, with $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$; for example
+the cell in figure \ref{fig:defetup} has norm 6. By lemma \ref{lem:doperatoropenmodtwo}
+the norm is weakly decreasing along differentials. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
 of norm $\leq p$, and let $F_p/F_{p-1}$ be the $p-$th filtration stratum.
@@
 Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1}))$,
-one copy for each way of splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$; the isomorphism
+one copy for each way of splitting $(m-p)=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$. The isomorphism
 doesn't preserve the degrees but shifts them by $p$.
@@
 differentials vanish (see lemma \ref{lem:doperatoropenmodtwo}
-for the definition), in particular the numbers $u_i,v_i$
-don't change along the boundary map of $F_p/F_{p-1}$, and this latter chain complex splits as a direct sum
-of chain complexes indexed by splittings $m-p=\sum_{i=1}^g (u_i+v_i)$ as above.
+for the definition), and in particular the numbers $u_i,v_i$
+don't change along the differentials of $F_p/F_{p-1}$. Therefore $F_p/F_{p-1}$ splits as a direct sum
+of chain complexes indexed by splittings $(m-p)=\sum_{i=1}^g (u_i+v_i)$ as above.
 It is then immediate to identify the inner differentials
@@
 Given a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_j$,
-the associated symmetric chain in $\tCh_*(C_p(\Sigma_{0,1})^c)$
+the associated \emph{symmetric chain} in $\tCh_*(C_p(\Sigma_{0,1})^c)$
 is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
@@
 In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\Sigma_{0,1})^c)$
-is given by the collection of all classes $[(\alpha_j)_{j\geq 0}]$ satisfying $\sum\alpha_j2^j$.
+is given by the collection of all classes $[(\alpha_j)_{j\geq 0}]$ satisfying $p=\sum\alpha_j2^j$.
 
-Using Poincaré-Lefschetz duality we obtain a basis for the cohomology of $C_p(\Sigma_{0,1})$;
-the dual basis of $H_*(C_p(\Sigma_{0,1}))$
+By Poincaré-Lefschetz duality this is a basis for the cohomology $H^*(C_p(\Sigma_{0,1}))$.
+The \emph{dual} basis of $H_*(C_p(\Sigma_{0,1}))$
 happens to be the basis of monomials
-$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\Sigma_{0,1}))$ of weight $m$, using
+$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\Sigma_{0,1}))$, that is, all monomials
+of weight $p$, using
 the isomorphism \ref{eq:Cohen} in its full meaning (i.e. as an isomorphism of rings,
@@
 
-We will not need this finer result in this article, so the expression
+We will not need this finer result in this article, so in the following the expression
 $\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_{p-l}(C_p(\Sigma_{0,1}))$ will only denote
@@
 \]
-we define a chain in $\tCh_{m+l}(\cms)$, denoted by $|p,(\alpha_j)_j,(u_i,v_i)_{i\leq g}|$: it
+let $l=\sum_{j=1}^{\infty}\alpha_j$. We define a chain in
+$\tCh_{m+l}(\cms)$, denoted by $|p,(\alpha_j)_j,(u_i,v_i)_{i\leq g}|$: it
 is obtained by summing all $e^{\tup}$ where $\tup$
@@
 
-It is straightforward to check that a generalised symmetric chain $|p,(\alpha_j),(u_i,v_i)|$
+We notice now that a generalised symmetric chain $|p,(\alpha_j),(u_i,v_i)|$
 is not only
@@
 in the chain complex $\tCh_*(\cms^c)$ itself. Indeed also all outer
-differentials of a generalised symmetric chain cancel out, since
+differentials of a generalised symmetric chain cancel out: in particular
 every left differential
@@
 
-In particular the spectral sequence collapses on its first page and
+Therefore the spectral sequence collapses on its first page and
 % $H_*(\cms^c)$
@@
 The homology $H_*(\cms^c)$ has a graded basis given by the classes $[p,(\alpha_j),(u_i,v_i)]$
-associated to generalised symmetric chains.
+associated to generalised symmetric chains of weight $m$.
 % this basis is bijection with
@@
 We can see the $\U_i$'s and $\V_i$'s as properly embedaded $1-$manifolds in $\mrS$;
-by Poincaré-Lefschetz duality they represent classes in $H_1\pa{(\mrS)^c}\simeq H^1(\S)$,
+by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^c}\simeq H^1(\S)$,
 and in particular they form a basis
@@
 \begin{equation}\label{eq:isovectorspaces}
-  \bigoplus_{m\geq 0} H^*(\cms)\simeq \Z_2\left[Q^i\epsilon\,|\, i\geq 0\right]\otimes\Sym_{\bullet}(\H),
+  \bigoplus_{m\geq 0} H^*(\cms)\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H),
 \end{equation}
-
-and using the canonical identification of $\Z_2-$cohomology as the dual of homology, we can
-conclude that there exists an isomorphism as in theorem \ref{thm:Hbms*as*ggrep}
-at least \emph{as bigraded $\Z_2-$vector spaces}.
+from which we conclude that
+there exists an isomorphism as in theorem \ref{thm:Hbms*as*ggrep}
+at least \emph{as bigraded $\Z_2-$vector spaces}: the two bigraded
+vector spaces have the same dimension in all bidegrees.

### 2018-11-05 Introduction.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 Let $\gg$ be the mapping class group of $\sg$, i.e. the group of isotopy classes of diffeomorphisms of $\sg$:
-diffeomorphisms are required to fix $\partial\sg$ pointwise; similarly let $\ggm$ be the mapping class group of $\sgm$, i.e.
+diffeomorphisms are required to fix $\partial\sg$ pointwise. Similarly let $\ggm$ be the mapping class group of $\sgm$, i.e.
 the group of isotopy classes of diffeomorphisms of $\sgm$ that fix $\partial\sgm$ pointwise and \emph{permute} the $m$ punctures.
 
-Forgetting punctures yields a surjective map $\ggm\to\gg$ with kernel $\bms$,
-the $m-$th \emph{braid group} of the surface $\sg$: we get the Birman exact sequence (see \cite{Birman:mcgbr}):
+Forgetting the punctures gives a surjective map $\ggm\to\gg$ with kernel $\bms$,
+the $m-$th \emph{braid group} of the surface $\sg$: we obtain the Birman exact sequence (see \cite{Birman:mcgbr}):
 \begin{equation}
@@
 
-Our main theorem is that the spectral sequence has already collapsed in its second page.
+The main theorem of this article states that this spectral sequence has already collapsed in its second page.
 \begin{thm}
@@
 Thus the computation of $H_*(\ggm;\Z_2)$ reduces to the computation of the homology of $\gg$ with
-twisted coefficients in the representation $H_*(\bms;\Z_2)$: as we will see the $\ggm-$representation
-$H_*(\bms;\Z_2)$ splits as a direct sum of symmetric powers of $H_1(\sg)$.
-In particular this representation is symplectic: this had already benn stated in \cite{LM}.
+twisted coefficients in the representation $H_*(\bms;\Z_2)$. We will see that the $\ggm-$representation
+$H_*(\bms;\Z_2)$ splits as a direct sum of symmetric powers of $H_1(\sg)$ with symplectic action
+(see theorem \ref{thm:Hbms*as*ggrep} in section \ref{sec:HBraidSurf}).
+
+%In particular this representation is symplectic: this had already benn stated in \cite{LM}.
 
@@
 computation of the right-hand side of equation \ref{eq:main}
-should be more tractable.
-WE DO THIS COMPUTATION FOR $g=1$.
+may be more tractable.
+%WE DO THIS COMPUTATION FOR $g=1$.

### 2018-11-05 Preliminaries.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 In this section we recollect some classical definitions and results about braid groups and mapping class groups.
-By $\sg$ we will always denote a compact, smooth, orientable surface of genus $g$ with one parametrised
-boundary component $\partial \sg$; by $\sgm$ we denote a surface $\sg$ with a choice of $m$ distinct
-points in the interior: these points are called \emph{punctures} and are not assumed to be ordered
-or labelled.
+Let $\sg$ be a compact, smooth, orientable surface of genus $g$ with one parametrised
+boundary component $\partial \sg$. Let $\sgm$ be $\sg$ with a choice of $m$ distinct
+points in the interior: these points are called \emph{punctures} and are not assumed to be ordered.
 
@@
 \[
- F_m(\sg)=\set{(P_1,\dots,P_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  P_i\neq P_j  \,\forall i\neq j}
+ F_m(\sg)=\set{(P_1,\dots,P_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  P_i\neq P_j  \;\forall i\neq j}.
 \]
- Notice that we require the points of the configuration to lie in the interior of $\sg$.
- Thus $F_m(\sg)$ is naturally endowed with the structure of a smooth, orientable $2m-$dimensional
- manifold.
+
+ Notice that we require the points of the configuration to lie in the interior of $\sg$;
+ $F_m(\sg)$ is a smooth, orientable $2m-$dimensional manifold.
 
@@
  is called the \emph{$m-$th unordered configuration space}
- of $\sg$ and denoted by $C_m(\sg)$; this space is also a $2m-$dimensional orientable manifold.
+ of $\sg$ and is denoted by $C_m(\sg)$; this space is also a $2m-$dimensional orientable manifold
+ (see figure \ref{fig:unordered}).
 
@@
 \end{defn}
+
+\begin{figure}\centering
+ \includegraphics[scale=0.8]{figures/unordered.png}
+ \caption{A configuration in the space $C_5(\Sigma_{1,1})$}
+\label{fig:unordered}
+\end{figure}
 
@@
  endowed with the Whitney $C^{\infty}-$topology. We denote by $\gg=\pi_0(\Diff(\sg;\partial\sg))$ its group of connected
- components, called also the \emph{mapping class group} of $\sg$.
+ components, which is called the \emph{mapping class group} of $\sg$.
 
  Similarly let $\Diff(\sgm;\partial\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
- pointwise and permute the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$ be its
+ pointwise and restrict to a permutation of the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$ be its
  group of connected components, called the mapping class group of $\sgm$.
@@
  \[
-  \sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
+  \cF_{g,1}\colon=\sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
  \]
@@
 group $\ggm$ and the Birman exact sequence \ref{eq:Birman} is obtained by taking fundamental
-groups of the spaces
+groups of the aspherical spaces
 \begin{equation}\label{eq:Birmanbundle}
@@
 $\Diff(\S;\partial\S)$ with its subgroup $\Diff(\S;\partial\S\cup\D)$ of diffeomorphisms
-of $\S$ that fix pointwise both $\partial\S$ and $\D$: here when we say that $\D$ is embedded
-\emph{near $\partial\S$} we mean that there is a compact subsurface $\S'\subset\S$
+of $\S$ that fix pointwise both $\partial\S$ and $\D$. Saying that $\D$ is embedded
+\emph{near $\partial\S$} means that there is a compact subsurface $\S'\subset\S$
 such that $\S$ is the boundary connected sum $\S'\natural\bar\D$.
 From now on we suppose such an embedding to be fixed and we consider $\D$ as a subspace
-of $\mrS$. In section \ref{sec:HBraidSurf} we will specify an
-embedding in a particular model for $\mrS$ that we will define.
+of $\mrS$ (see figure \ref{fig:SS'D}). In section \ref{sec:HBraidSurf}, definition \ref{defn:Tsg},
+we will specify an embedding $\D\hookrightarrow\mrS$ using a particular model for $\mrS$.
 
-The construction of definition \ref{defn:cms} can be specialised to the open
+\begin{figure}\centering
+ \includegraphics[scale=0.8]{figures/SS'D.png}
+ \caption{A splitting of $\S$ as $\S'\natural\bar\D$.}
+\label{fig:SS'D}
+\end{figure}
+
+The construction of definition \ref{defn:cms} can be specialised to the
 surfaces $\D$ and $\S'$, yielding spaces $C_m(\D)$ and $C_m(\S')$ respectively.
@@
 The inclusion $\Diff(\S;\partial\S\cup\D)\subset\Diff(\S;\partial\S)$ is a homotopy
-equivalence, hence the previous constructions are equivalent to the ones that we
-give in the next definition.
+equivalence, hence also the induced map
+\[
+B\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S)
+\]
+is a homotopy equivalence. We will replace the previous construction with the following,
+equivalent ones.
 \begin{defn}
@@
  \[
-  \S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
+  \cF_{\S,\D}\colon=\S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
  \]
-The group $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms;
-the space $\cF_{\S,\D}$ contains subspaces
+Notice that $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms, and acts trivially
+on $\D$; therefore $\cF_{\S,\D}$ contains subspaces
 \[
@@
 \]
-and $\D\times B\Diff(\S;\partial\S\cup\D)$, fibering over $B\Diff(\S;\partial\S\cup\D)$
+and $\D\times B\Diff(\S;\partial\S\cup\D)$; these subspaces fiber over $B\Diff(\S;\partial\S\cup\D)$
 with fibers $\S'$ and $\D$ respectively.
@@
 $C_m(\cF_{\S,\D})$, $C_m(\cF_{\S'})$ and $C_m(\D)\times B\Diff(\S;\partial\S\cup\D)$,
-all fibering over $B\Diff(\S;\partial\S\cup\D)$ fibers, respectively, $\cms$,
+all fibering over $B\Diff(\S;\partial\S\cup\D)$ with fibers, respectively, $\cms$,
 $C_m(\S')$ and $C_m(\D)$.
+\end{defn}
 
+\begin{defn}
 For all $0\leq p\leq m$ there is a natural map
@@
 \]
-which consists on the union of configurations:
+which takes the union of configurations (see figure \ref{fig:defmu}):
  \[
-  \mu(\set{P_1,\dots,P_p};\set{P'_1,\dots,P'_{m-p}})=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
+  \mu\pa{\set{P_1,\dots,P_p};\set{P'_1,\dots,P'_{m-p}}}=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
  \]
+\begin{figure}\centering
+ \includegraphics[scale=0.7]{figures/defmu.png}
+ \caption{The product of two configurations in $C_2(\D)$ and $C_4(\S)$.}
+\label{fig:defmu}
+\end{figure}
+
 We can apply this construction to each couple of fibers of $C_m(\cF_{\S,\D})$ and $C_{m-p}(\cF_{\S'})$
@@
 \]
-If we see the domain of $\mu^{\cF}$ as a fibered product
+If we see the domain of $\mu^{\cF}$ as a fibered product of bundles over $B\Diff(\S;\partial\S\cup\D)$
 \[
@@
 \]
-then the map $\mu^{\cF}$ is a map of bundles over $B\Diff(\S;\partial\S\cup\D)$, and corresponding
+then the map $\mu^{\cF}$ is also a map of bundles over $B\Diff(\S;\partial\S\cup\D)$, and the corresponding
 map on fibers is precisely $\mu$.
@@
 The cohomology of $C_m(\D)$ with coefficients in $\Z_2$ was first computed by Fuchs in
-\cite{Fuchs:CohomBraidModtwo}: we will recall this computation more in detail in section
-\ref{sec:HBraidSurf}, where we will make in detail a similar computation.
+\cite{Fuchs:CohomBraidModtwo}: we will recall this computation in section
+\ref{sec:HBraidSurf}, where we will make a similar computation in detail.
 
-In \cite[Chap.III]{CLM} Cohen considers $\coprod_{m\geq 0}C_m(\D)$ as an
-algebra over the operad of little $2-$cubes, and in particular describes its
-homology with coefficients in $\Z_2$ as follows:
+In \cite[Chap.III]{CLM} Cohen considers the space $\coprod_{m\geq 0}C_m(\D)$ as an
+algebra over the operad of little $2-$cubes, and describes its
+$\Z_2-$homology as follows:
 \begin{equation}
@@
 the first Dyer-Lashof operation. In particular $Q^j\epsilon$ is the generator of
-$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$.
+$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$. See figure \ref{fig:monomial}
 
 The isomorphism in equation \ref{eq:Cohen} is
-an isomorphism of bigraded rings: the left-hand side is a ring with the Pontryagin product,
-and the right-hand side is a polynomial ring in the infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
-the bidegree is given by
-$*$, called \emph{degree}, which is the homological degree,
-and $m$, called \emph{weight}, which is the index of the connected component
-on which the homology class is supported, or informally the number of points
-involved in the construction of a homology class.
+an isomorphism of bigraded rings. The left-hand side is a ring with the Pontryagin product
+coming from the action of the little $2-$cubes,
+and the right-hand side is a polynomial ring in infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$.
+The bidegree is given by the homological degree
+$*$, that we simply call the \emph{degree},
+and by the index $m$ of the connected component
+on which the homology class is supported (informally, the number of points
+involved in the construction of a homology class), that we call the \emph{weight}.
+
+ \begin{figure}\centering
+ \includegraphics[scale=0.7]{figures/monomial.png}
+ \caption{The class in $H_5(C_9(\D))$ corresponding to the monomial $\epsilon\cdot(Q\epsilon)^2\cdot Q^2\epsilon$.}
+\label{fig:monomial}
+\end{figure}
 
@@
 In particular for all choices of natural numbers $(\alpha_j)_{j\geq 0}$ with all but finitely
-many $\alpha_j=0$, we will consider the monomial $\prod_j\pa{Q^j\epsilon}^{\alpha_j}$, which is
+many $\alpha_j=0$, we will consider the monomial $\prod_j\pa{Q^j\epsilon}^{\alpha_j}$, corresponding to
 a homology class in some bidegree $(*,m)$: all that

### 2018-11-05 Rational.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 In this section we prove that a statement as in theorem \ref{thm:Hbms*as*ggrep}
-cannot hold if we consider homology with coefficients in $\Q$.
+cannot hold if we consider homology with coefficients in $\Q$. We still don't know
+if the analogue of theorem \ref{thm:main} holds in homology with coefficients in $\Q$
+or in fields of odd characteristic.
 
-We recall that $H_*(\cms;\Q)$ has been computed \emph{as a bigraded $\Q-$vector space}
-in \cite{BCM} and, more recently, in \cite{Knudson}; a description of it as a
+We recall that $H_*(C_m(\Sigma_{g,1});\Q)$ has been computed \emph{as a bigraded $\Q-$vector space}
+by B\"odigheimer, Cohen and Milgram in \cite{BCM}, and more recently by Knudson in \cite{Knudson}. A description of
+these homology groups as a
 $\gg-$representation seems to be still missing in the literature.
 
-In particular we prove the following fact.
+We will prove the following theorem:
 \begin{thm}
  \label{thm:counterexample}
- Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\S);\Q)$ is not a symplectic
+ Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\Sigma_{g,1});\Q)$ is not a symplectic
  representation of $\gg$.
@@
 \begin{proof}
- We will use the following strategy:
+ Let again $\S=\Sigma_{g,1}$. We will use the following strategy:
  \begin{itemize}
@@
   $[a]$ with a homology class
-  in $\tH_2(\cms^c;\Q)$; here it is worth stressing that for all $m$ the manifold
+  in $\tH_2(\cms^c;\Q)$; here it is worth stressing that the manifold
   $\cms$ is orientable: therefore we can apply Poincaré-Lefschetz duality
@@
 
-Let $\tup=(0,(u_i,v_i))$ with $u_1=2$ and all other $u_i$'s, as well as all $v_i$'s, equal to zero;
-then the map
+Let $\tup=(0,(u_i,v_i))$ with $u_1=2$ and all other $u_i$'s, as well as all $v_i$'s, equal to zero.
+Then the map
 \[
@@
 maps the fundamental class in $H_2(\Delta^{\tup},\partial\Delta^{\tup};\Q)$ to a homology
-class in $H_2(C_2(\S);\Q)$; if we see $e^{\tup}$ as a properly embedded, orientable submanifold
-of $C_2(\S)$, then we are just considering its fundamental class $[e^{\tup}]\in H_2(C_2(\S)^c)$.
+class in $\tH_2(C_2(\S)^c;\Q)$.
+% ; if we see $e^{\tup}$ as a properly embedded, orientable submanifold
+% of $C_2(\S)$, then we are just considering its fundamental class $[e^{\tup}]\in H_2(C_2(\S)^c)$.
 
-The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore $[a]\cap[e^{\tup}]=\pm 1$,
+The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore the
+algebraic intersection $[a]\cap[e^{\tup}]$ is $\pm 1$,
 where the signs depends on how we have chosen orientations on $a$, $e^{\tup}$ and $C_2(\S)$ itself;
@@
 the construction of the torus $a$ with any other copy of parallel, non-separating simple closed curves
-$c$ and $c'$, and the resulting class $[a]$ will always be non-zero in $H_2(C_2(\S);\Q)$.
+$c$ and $c'$, and the resulting class $[a]$ will always be non-trivial in $H_2(C_2(\S);\Q)$.
 
-Let now $d$ and $d'$ be disjoint, non-separating simple closed curves such that
+See figure \ref{fig:rational} to visualize the following discussion.
+Let $d$ and $d'$ be disjoint, non-separating simple closed curves such that
 cutting $\S$ along $d$ and $d'$ we obtain a subsurface $\S'\simeq\Sigma_{1,2}$ of genus 1, with boundary
@@
 \end{itemize}
-See picture PICTURE.
+
+\begin{figure}\centering
+ \includegraphics[scale=1.0]{figures/rational.png}
+ \caption{The curves $d,d',c,c',c'',\gamma,\gamma''$ and the subsurfaces $\S',\S''$.}
+\label{fig:rational}
+\end{figure}
 
@@
 Consider now an element of the Torelli group that fixes $c$ and maps $d$ to $d'$ preserving
-the orientation, for example the bounding pair represented in picture PICTURE:
+the orientation, for example the bounding pair $D_{\gamma}\circ D_{\gamma'}^{-1}$, where
+$\gamma$ and $\gamma'$ are represented in figure \ref{fig:rational}:
 then the class $b$ is mapped to $b'=b+2a\neq b$.
 \end{proof}
+
+The previous proof works word by word if we replace $\Q$ by any field of odd characteristic. Moreover
+all the arguments in the previous proof can be adapted to show that $H_*(F_2(\S);\mathbb{F})$ is
+not a symplectic representation of $\gg$, where $F_2(\S)$ is the ordered configuration space
+and $\mathbb{F}$ is \emph{any} field, including $\Z_2$. The result of theorem \ref{thm:Hbms*as*ggrep}
+is peculiar of unordered configurations and characteristic 2.

### 2018-11-05 Ssequence.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

@@
 We will prove theorem \ref{thm:main} by induction on $m$.
-The case $m=0$ is trivial. Suppose now $m\geq 1$.
+The case $m=0$ is trivial, so suppose $m\geq 1$.
 
-For all $m\geq$ we let $E(m)$ be the Leray-Serre spectral sequence associated with
+For all $m\geq 0$ we let $E(m)$ be the Leray-Serre spectral sequence associated with
 the bundle \ref{eq:BirmanbundleD}: its second page has the form
@@
 \label{eq:Hksplitting}
- \bigoplus_{p=l}^m H_k\pa{\gg;\mu_*\pa{H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}};
+ \bigoplus_{p=l}^m H_k\pa{\gg;\mu_*\pa{H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}}.
 \end{equation} 
-we will now fix $l\leq p\leq m$ and show the vanishing of all differentials $\partial_r$ exiting from the
-summand with label $p$ in equation \ref{eq:Hksplitting}.
+We fix now $l\leq p\leq m$ and show the vanishing of all differentials $\partial_r$ exiting from the
+summand with label $p$ in the previous equation.
 
@@
 from the second page on, to the tensor product
-of $H_*(C_p(\D))$ and the spectral sequence associated with the bundle $(\cF_{\S'})$; the
-latter is isomorphic, in our notation, to the spectral sequence $E(m-p)$.
+of $H_*(C_p(\D))$ and the spectral sequence associated with the bundle $C_{m-p}(\cF_{\S'})$; the
+latter spectral sequence is isomorphic, in our notation, to the spectral sequence $E(m-p)$.
 In particular $\mu^{\cF}$ induces a map of spectral
@@
 \[
-H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}\subset H_k\pa{\gg;H_{m-l}(\cms)}.
+H_k\pa{\gg;\mu_*(H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S')))}\subset H_k\pa{\gg;H_{m-l}(\cms)}.
 \]
@@
 in the first spectral sequence, then also all differentials
-$\partial_r$ exiting from the summand $H_k\pa{\gg;H_{m-l}(\cms)}$ in the
-second spectral sequence must vanish.
+$\partial_r$ exiting from this direct summand in the
+second spectral sequence $E(m)$ must vanish.
 
@@
 that the latter vanish. Theorem \ref{thm:main} is proved.
+
+One can generalise theorem \ref{thm:main} to orientable or non-orientable
+surfaces with at least one boundary component, following the generalisation
+of theorem \ref{thm:Hbms*as*ggrep} discussed at the end of section \ref{sec:Actiongg}.

### 2019-02-18 Actiongg.tex
Implemented some corrections by cfb, created gitignore file

@@
 \section{Action of $\gg$}
-We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of theorem
-\ref{thm:Hbms*as*ggrep} in bidigrees $(*,m)$ with $*=m$, and then extend it to other bidegrees.
+\label{sec:Actiongg}
+We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of Theorem
+\ref{thm:Hbms*as*ggrep} in bigradings $(*,m)$ with $*=m$, and then extend it to other bigradings.
 
-\subsection{Bidigrees of the form $(m,m)$}
-In this subsection we construct a $\gg-$~equivariant isomorphism 
+\subsection{Bigradings of the form $(m,m)$}
+In this subsection we construct a $\gg$-~equivariant isomorphism 
 \[
@@
 
-We already know that these $\Z_2-$vector spaces have the same dimension:
-indeed $\Sym_m(\H)$ is precisely the summand in bidegree $(m,m)$ in equation
-\ref{eq:isovectorspaces}, using that a monomial
+We already know that these $\Z_2$-vector spaces have the same dimension:
+indeed $\Sym_m(\H)$ is precisely the summand in bigrading $(m,m)$ in equation
+\eqref{eq:isovectorspaces}, using that a monomial
  whose
@@
 % \]
-% which carries the fundamental class of the $m-$fold torus to some class in
+% which carries the fundamental class of the $m$-fold torus to some class in
 % $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
@@
 \begin{defn}
+\label{defn:DinTS} 
+Recall Definition \ref{defn:TS}. We denote by $\D\subset\T(\S)\cong\mrS$ the open square $]1/4;3/4[\times]1/2,1[$.
+Note that $\D$ is an open disc in $\mrS$ near $\partial\S$ and 
+it is disjoint from all $\overline{\U}_i$'s and $\overline{\V}_i$'s. The interior of the surface $\S'$
+is then identified with $\T(\S)\setminus [1/4,3/4]\times[1/2,1[$.
+\end{defn}
+
+\begin{defn}
 \label{defn:cCm}
-Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
+Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
 such that
@@
 Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
-images.
+images. Here and in the following $\Sone$ is the unit circle in $\C$.
 
-Two $m-$tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
-transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.
+Two $m$-tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
+transforming one $m$-tuple into the other. In particular $\cC^m$ is more than countable.
 
@@
 
-We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
+We denote by $\ZcC{m}$ the free $\Z_2$-vector space with basis $\cC^m$.
 There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
@@
 The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
-on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg-$equivariant.
+on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg$-equivariant.
 \end{defn}
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defcCm.png}
+ \includegraphics[scale=0.7]{figures/defcCm.pdf}
  \caption{A multicurve in $\cC^5$.}
@@
 
-% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
-% factors through $\pr$ to give the desired $\gg-$equivariant map $\psi$; we will then show that
+% Our aim is to define a $\gg$-equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
+% factors through $\pr$ to give the desired $\gg$-equivariant map $\psi$; we will then show that
 % $\psi$ is an isomorphism.
@@
  \label{defn:cmsD}
- The space $\cmsD$ is the subspace of $\SP^{m}(\mrS)$ of configurations where all points
+ Recall Definition \ref{defn:symprod}.
+ The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
  having multiplicity $\geq 2$ lie inside $D$.
+ See Figure \ref{fig:defcmsD}.
 
- Recall that the $m-$fold symmetric product of an open surface is a $2m-$dimensional manifold;
- the space $\cmsD$ is an open subset
- of $\SP^m(\mrS)$, hence $\cmsD$ is also a manifold.
- See figure \ref{fig:defcmsD}.
-
- There is a natural, open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
-  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
+  Note that $\cmsD$ is open in $\SP^m(\mrS)$. There is an open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
+  \[
+  \j_m\colon \ZcC{m}\to H_m(\cmsD)
+  \]
+  defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
   the composition
@@
 of the fundamental class
-of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
-if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple in the same multicurve.
+of the $m$-fold torus in $H_m(\cmsD)$. The result does not change
+if we substitute $(c_1,\dots,c_m)$ with another isotopic $m$-tuple of curvesin the same multicurve.
 
-We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in $\cmsD$ of the fundamental cycle of the
-$m-$fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
+We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in the singular chain complex of $\cmsD$ of
+the fundamental cycle of the
+$m$-fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
 on the union $c_1\cup\dots\cup c_m$, meaning that it hits configurations in $\cmsD$ of points of $\mrS$
@@
 The group $\Diff(\S,\D\cup\partial\S)$ acts on $\cmsD$, and there is an induced action of $\gg$
-on $H_m(\cmsD)$. The map $\j_m$ is $\gg-$equivariant.
+on $H_m(\cmsD)$. The map $\j_m$ is $\gg$-equivariant.
 \end{defn}
@@
  It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
- or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$
- is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
- $\cmsD^c\setminus\cms$ to $\infty$.
+ or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$
+ is surjective: this last map is induced by the map $\cmsD^{\infty}\to\cms^{\infty}$ collapsing the subspace
+ $\cmsD^{\infty}\setminus\cms$ to $\infty$.
 
- Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $\tH_m(\cms^c)$
- is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
- and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$ for
- such classes.
+ Recall Lemma \ref{lem:gensymchain} and Definition \ref{defn:gensymchain}. A basis for $\tH_m(\cms^{\infty})$
+ is given by classes $[\kappa(0,\uu,\uv)]$, with
+ $\uu=(u_1,\dots, u_g)$, $\uv=(v_1,\dots,v_g)$ and
+ $\sum_{i=1}^g (u_i+v_i)=m$.
 
- The class $[0,(u_i,v_i)]$ is represented by a generalised symmetric chain consisting of only one cell,
- the cell $e^{\tup}$
- with $\tup=(0,(u_i,v_i))$.
+ The class $[\kappa(\uu,\uv)]$ is represented by a generalised symmetric chain consisting of only one tuple $\tup=(0,\uu,\uv)$.
 
  In particular there is a map of pairs
- $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^c,\infty)$, and $[0,(u_i,v_i)]$
+ $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^{\infty},\infty)$, and the class $[\kappa(\uu,\uv)]$
  is the image along this map of the fundamental class of $H_m\pa{\Delta^{\tup},\partial\Delta^{\tup}}$.
@@
  It is straightforward to check that the map $\phi^{\tup}$ factors through
- $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$ follows.
+ $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$ follows.
 \end{proof}
@@
   & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
-  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*"hook] & H_m\pa{\cmsD}
+  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*",hook] & H_m\pa{\cmsD}.
  \end{tikzcd}
@@
 We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
-to a $\gg-$equivariant map $\tpsi_m$ as in the diagram.
-Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
-lands in the image of $(\iota_m)_*$, and this last statement doesn't depend on how $\gg$ acts on these groups.
+to a $\gg$-equivariant map $\tpsi_m$ as in the diagram.
+Since $(\iota_m)_*$ is injective by Lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
+lands in the image of $(\iota_m)_*$, and this last statement does not depend on how $\gg$ acts on these groups.
 
@@
  \label{lem:tpsiwithproperties}
-For each $m-$tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
+For each $m$-tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
 and for each open neighborhood $\N\subset\mrS$ of
@@
 are connected by a homology in $\cmsD$ which is supported on $\N$ (the word
-\emph{homology} denotes here a $(m+1)-$singular chain whose boundary is the difference between the two cycles).
+\emph{homology} denotes here a $(m+1)$-singular chain whose boundary is the difference between the two cycles).
 \end{itemize}
@@
 For $m=0$ both $\ZcC{0}$ and $H_0(C_0(\S))$ are
-isomorphic to $\Z_2$ and there is nothing to show; for $m=1$
+isomorphic to $\Z_2$ and there is nothing to show. For $m=1$
 we have a canonical identification $H_1(C_1(\S))\simeq\H\simeq\Sym_1(\H)$, so we take $\tpsi_1=\pr_1$;
@@
 
-Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^m$.
+Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.
 % and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
@@
 \label{defn:variationsCm}
-We introduce a few variations of the notion of configuration space; see figure \ref{fig:defcmsD}.
+We introduce several variations of the notion of configuration space; see Figure \ref{fig:defcmsD}.
 \begin{itemize} 
@@
 \begin{figure}\centering
- \includegraphics{figures/defcmsD.png}
- \caption{A configuration in each of the space introduced in definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
+ \includegraphics{figures/defcmsD.pdf}
+ \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
  Whenever a multiplicity is not specified, it is equal to 1.}
@@
 
-We have the following natural inclusions:
+We have the following inclusions:
 \[
-C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
+C_{1,m}(\S)\subset\fmstwoD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
 \]
@@
 $[c_1]\cdot \fc$ and $[c_1]\cdot\ldots\cdot[c_{m+1}]$ in $C_{m+1}^{\D}(\S)$: both cycles and the
-homology between them are supported on $\N$. Notice now that the
+homology between them are supported on $\N$. Note now that the
 cycle $[c_1]\cdot\fc$ lives in $\cmstwoD$, so the first lift is done and we can now
@@
 a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
-% the latter being even a $m+1-$fold covering map;
+% the latter being even a $m+1$-fold covering map;
 we have a commutative diagram
@@
 \]
-We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
+We note that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
 and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
-a $2-$fold ramified covering between their respective normal bundles.
+a $2$-fold ramified covering between their respective normal bundles.
 \end{defn}
 
-Diagram \ref{eq:cmstwodiagram} induces a commutative diagram in homology
+Diagram \eqref{eq:cmstwodiagram} induces a commutative diagram in homology
 \begin{equation}
@@
 
-We first notice that there is a lift $[c_1]\otimes\fc$ to a cycle in $\pa{\fmstwoD}$,
-which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
-to be white. We then notice that the right vertical map
+We first note that there is a lift of $[c_1]\cdot\fc$ to a cycle $[c_1]\otimes\fc$ in $\pa{\fmstwoD}$:
+this is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
+to be white. We then note that the right vertical map
 \[
@@
 Therefore the image of the cycle $[c_1]\otimes \fc$ along the diagonal of the square
-in diagram \ref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
-is zero, hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$: more
-precisely there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
+in diagram \eqref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
+is zero; hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$. More
+precisely, there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
 to $[c_1]\cdot\fc$.
 
-To prove lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
+To prove Lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
 a good homology, namely two that are supported on $\N$: a priori both $\fc'$ and the homology
@@
 This can be done by replacing, in the whole argument of the proof, the surface $\mrS$ with the surface $\N$.
-We can define configuration spaces as in definition \ref{defn:variationsCm} also for the open surface
+We can define configuration spaces as in Definition \ref{defn:variationsCm} also for the open surface
 $\N$, and we can repeat the argument considering $\N$ as the \emph{ambient surface}:
@@
 
-It is crucial that the action of $\gg$ is not involved in the statement of lemma
+It is crucial that the action of $\gg$ is not involved in the statement of Lemma
 \ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
@@
 
-We now have to prove the following lemma to conclude the proof of theorem \ref{thm:Hbms*as*ggrep}
-in bidigrees $(m,m)$.
+We now have to prove the following lemma to conclude the proof of Theorem \ref{thm:Hbms*as*ggrep}
+in bigradings $(m,m)$.
 \begin{lem}
@@
 $\pr_m$ and $\tpsi_m$
-are $\gg-$equivariant, also the induced map $\Sym_m(\H)\to H_m(\cms)$ will automatically be
-$\gg-$equivariant.
+are $\gg$-equivariant, also the induced map of vector spaces
+\[
+\Sym_m(\H)=\ZcC{m}/\ker\pr_m\to H_m(\cms)
+\]
+will automatically be
+$\gg$-equivariant.
 
-Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $\tH_m(\cms^c)\simeq H^m(\cms)$, which is
-the dual of $H_m(\cms)$, is given by the classes $[0,(u_i,v_i)]$, represented by generalised
-symmetric chains consisting of only one cell $e^{\tup}$, with $\tup=(0,(u_i,v_i))$.
-The corresponding linear functionals are the algebraic intersection products
-coming from Poincaré-Lefschetz duality:
+Recall from the proof of Lemma \ref{lem:cms->cmsDinj} that a basis for $\tH_m(\cms^{\infty})$ 
+is given by the classes $[\kappa(\uu,\uv)]$, represented by generalised
+symmetric chains consisting of only one tuple $\tup=(0,\uu,\uv)$, for some vectors
+$\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ satisfying $\sum_{i=1}^g(u_i+v_i)=m$.
+
+The homology class
+$[\kappa(\uu,\uv)]$ is the fundamental class
+of the sphere $e^{\tup}\cup\set{\infty}\subset\cms^{\infty}$: the inclusion of this sphere in $\cms^{\infty}$
+restricts to a proper embedding $e^{\tup}\subset\cms$.
+
+By Poincaré Lefschetz duality $\tH_m(\cms^{\infty})\simeq H^m(\cms)$, and the latter is
+the dual of $H_m(\cms)$.
+
+We can therefore associate to $[\kappa(\uu,\uv)]$ a linear functional
+on $H_m(\cms)$, which is the intersection product
 \[
@@
 and it suffices to check that $\ker\pr_m\subseteq\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m}$
-for all $\tup$ of the form $(0,(u_i,v_i))$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.
+for all $\tup$ of the form $(0,\uu,\uv)$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.
 
-Recall from the proof of lemma \ref{lem:cms->cmsDinj} that the cohomology class $(\cdot\cap e^{\tup})$ is a pullback
-of a cohomology class of $\cmsD$, that by abuse of notation we still call $(\cdot\cap e^{\tup})$:
-then we can compute the map $(\cdot\cap e^{\tup})\circ\tpsi_m$ as the map $(\cdot\cap e^{\tup})\circ\j_m$.
+Recall from the proof of Lemma \ref{lem:cms->cmsDinj} that the cohomology class $(\cdot\cap e^{\tup})$ on $\cms$ is a pullback
+of a cohomology class of $\cmsD$, that we call $(\cdot\cap e^{\tup})^{\D}$. Alternatively,
+note that the inclusion $e^{\tup}\cap \set{\infty}\to\cms^{\infty}$ is the composition of the inclusion $e^{\tup}\cap\set{\infty}\to\cmsD^{\infty}$
+and the quotient map $\cmsD^{\infty}\to\cms^{\infty}$, and consider the fundamental class of the sphere $e^{\tup}\cup\set{\infty}$ and its images.
 
-Looking at the geometric intersection between the cycle $[c_1]\cdot\dots[c_m]$ in
-$\cmsD$, and the cycle $e^{\tup}$ in $\cmsD^c$, we see that the map
-\[
-(\cdot\cap e^{\tup})\circ\j_m\colon \ZcC{m}\to\Z_2
-\]
-coincides with the composition
-\[
+We can therefore compute the map $(\cdot\cap e^{\tup})\circ\tpsi_m$ as the map 
+\begin{equation}
+\label{eq:badformula}
+(\cdot\cap e^{\tup})^{\D}\circ\j_m\colon \ZcC{m}\to\Z_2.
+\end{equation}
+The latter map coincides with the composition
+\begin{equation}
+ \label{eq:productformula}
 \pa{ \prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}}\circ\pr_m,
-\]
-where $\prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}\in\Sym_m(\Hom(\H;\Z_2))=\Hom\pa{\Sym_m(\H);\Z_2}$;
-in particular it factors through $\pr_m$.
+\end{equation}
 
-To show surjectivity of $\tpsi_m$, choose a tuple $\tup$ of the form $(0,(u_i,v_i))$
-and an $m-$tuple of curves $(c_1,\dots,c_m)$ containing, for every $1\leq i\leq g$, $u_i$ parallel
+where $\prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}\in\Sym_m(\Hom(\H;\Z_2))=\Hom\pa{\Sym_m(\H);\Z_2}$.
+
+This can be checked on every generator $(c_1,\dots, c_m)\in\ZcC{m}$ by chosing in the isotopy class a representative $(c_1,\dots, c_m)$
+with all curves $c_i$ transverse to all segments $\U_j$ and $\V_j$.
+
+Consider again the map $c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to\cmsD$ that we used to define the cycle $[c_1]\cdot\dots[c_m]$
+representing the class $\j_m(c_1,\dots,c_m)$ (see Definition \ref{defn:cmsD}):
+this map is an embedding near $e^{\tup}$ and is transverse to $e^{\tup}$.
+
+The equality of the maps in equations \eqref{eq:badformula} and \eqref{eq:productformula} on the generator $(c_1,\dots,c_m)\in\ZcC{m}$
+follows from a straightforward
+computation (in $\Z_2$) of the cardinality of the set $[c_1]\cdot\dots[c_m]\cap e^{\tup}$ in terms
+of the cardinalities of all sets of the form $c_i\cap\U_j$ and $c_i\cap \V_j$.
+In particular $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.
+
+To show surjectivity of $\tpsi_m$, choose a tuple $\tup$ of the form $(0,\uu,\uv)$
+and an $m$-tuple of curves $(c_1,\dots,c_m)$ containing, for every $1\leq i\leq g$, $u_i$ parallel
 copies of some curve representing $\u_i$ and $v_i$ parallel copies of some curve representing
-$\v_i$ (see definition \ref{defn:dualHbasis}), such that all intersections between these curves
+$\v_i$ (see Definition \ref{defn:dualHbasis}), such that all intersections between these curves
 lie in $\D$.
@@
 Then $j_m(c_1,\dots,c_m)\cap e^{\tup}=1$ and for all
-other tuples $\tup'=(0,(u'_i,v'_i))$ we have instead $j_m(c_1,\dots,c_m)\cap e^{\tup'}=0$.
+other tuples $\tup'$ of the form $(0,\uu',\uv)$ we have instead $j_m(c_1,\dots,c_m)\cap e^{\tup'}=0$.
 
@@
 
-Theorem \ref{thm:Hbms*as*ggrep} is now proved in all bidegrees of the form $(m,m)$.
+Theorem \ref{thm:Hbms*as*ggrep} is now proved in all bigradings of the form $(m,m)$.
 
-\subsection{General bidegrees $(m-l,m)$.} In this subsection we fix a bidegree $(m-l,m)$
-with $0\leq l\leq m$.
-Our aim is to prove theorem \ref{thm:Hbms*as*ggrep} for this bidegree.
-
+\subsection{General bigradings $(m-l,m)$.} Fix $0\leq l\leq m$ for the whole subsection: our next aim is to prove Theorem
+\ref{thm:Hbms*as*ggrep} for the bigrading $(m-l,m)$.
 % \begin{defn}
@@
 % %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
-% %  Notice that $\S'$ is also a surface of type $\sg$.
+% %  Note that $\S'$ is also a surface of type $\sg$.
 % %  
@@
 %  
-%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
+%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in Definition \ref{defn:cms},
 %  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
@@
 % \end{defn}
-Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
-$\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.
+% Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
+% $\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.
 
@@
 and on $\cms$, and the map $\mu$ is equivariant with respect to this action
-(see definition \ref{defn:universalSbundle});
-hence, using the K\"{u}nneth formula, there is an induced $\gg-$equivariant map in homology
+(see Definition \ref{defn:universalSbundle});
+hence, using the K\"{u}nneth formula, there is an induced $\gg$-equivariant map in homology
 \begin{equation}
@@
 
-Notice that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
+Note that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
 is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
@@
 
-We will prove the following lemma, from which theorem \ref{thm:Hbms*as*ggrep} follows:
+We will prove the following lemma, from which Theorem \ref{thm:Hbms*as*ggrep} follows:
 \begin{lem}
  \label{lem:oplussplitting}
-For all $l\leq p\leq m$ the map $\mu_*$ in equation \ref{eq:mu*} is injective, and the collection
+For all $l\leq p\leq m$ the map $\mu_*$ in equation \eqref{eq:mu*} is injective, and the collection
 of all these maps yields a splitting
@@
 \begin{proof}
-Notice that the statement of the lemma doesn't depend on the the action of $\gg$:
-we have a map from the right-hand side to the left-hand side of equation \ref{eq:musplitting},
-we already know that it is $\gg-$equivariant,
-we only need to show that it is a linear isomorphism. Notice also that lemma \ref{lem:gensymchain}
+Note that the statement of the lemma does not depend on the the action of $\gg$:
+we have a map from the right-hand side to the left-hand side of equation \eqref{eq:musplitting},
+we already know that it is $\gg$-equivariant,
+we only need to show that it is a linear isomorphism. Note also that Lemma \ref{lem:gensymchain}
 implies that the two vector spaces have the same dimension.
@@
 % \]
-% Notice that the natural target for the cohomology map $\mu^*$ is the whole group
+% Note that the natural target for the cohomology map $\mu^*$ is the whole group
 % \[
@@
 % \[
-% \mu^c_* \colon \tH_{m-l}(\cms^c)\to \tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}
+% \mu^{\infty}_* \colon \tH_{m-l}(\cms^{\infty})\to \tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}
 % \]
-% are surjective and exhibit $\tH_{m-l}(\cms^c)$ as a product
+% are surjective and exhibit $\tH_{m-l}(\cms^{\infty})$ as a product
 % \[
-% \tH_{m-l}(\cms^c)=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}.
+% \tH_{m-l}(\cms^{\infty})=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}.
 % \]
 % Here a few explanations are required:
-% we are considering the map $\mu^c\colon \cms^c\to \pa{C_{m-p}(\S')\times C_p(\D)}^c$ which collapses the complement
+% we are considering the map $\mu^{\infty}\colon \cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$ which collapses the complement
 % of the open subspace $C_{m-p}(\S')\times C_p(\D)\subset\cms$ to the point at infinity; it induces a map
-% $\mu^c_*\colon \tH_{m-l}(\cms^c) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}$;
+% $\mu^{\infty}_*\colon \tH_{m-l}(\cms^{\infty}) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}$;
 % applying Poincaré-Lefschetz duality
@@
 % \[
-%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
+%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
 % \]
-% we are considering the composition of the map $\mu^c_*$ with the projection onto the factor $q=p$, and by
-% abuse of notation we call this map simply $\mu^c_*$.
+% we are considering the composition of the map $\mu^{\infty}_*$ with the projection onto the factor $q=p$, and by
+% abuse of notation we call this map simply $\mu^{\infty}_*$.
 %
 
-Fix $l\leq p\leq m$ and let $[a]=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
-$H_{p-l}(C_p(\D))$, hence $l=\sum_j\alpha_j$ and $p=\sum_j\alpha_j2^j$. Let also
-$[b]=\prod_{i=1}^g (\u_i^{u_i}\v_i^{v_i})$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
-isomorphism proved in the previous subsection, with $(m-p)=\sum_i(u_i+v_i)$. Here $a$ and
+Fix $l\leq p\leq m$ and $\ualpha=(\alpha_j)_{j\geq 0}$, and let $[a]=Q^{\ualpha}\epsilon=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
+$H_{p-l}(C_p(\D))$, hence $l=\sum_j\alpha_j$ and $p=\sum_j\alpha_j2^j$.
+
+Fix also $\uu=(u_1,\dots, u_g)$ and $\uv=(v_1,\dots,v_g)$,
+and let $[b]=\u^{\uu}\cdot\v^{\uv}=\prod_{i=1}^g (\u_i^{u_i}\v_i^{v_i})$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
+isomorphism proved in the previous subsection, with $(m-p)=\sum_i(u_i+v_i)$.
+
+Here $a$ and
 $b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
-and $b$ supported on $\S'$. See figure \ref{fig:axb}
+and $b$ supported on $\S'$. See Figure \ref{fig:axb}
 
 \begin{figure}\centering
- \includegraphics[scale=0.65]{figures/axb.png}
- \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_3(\D))$;
+ \includegraphics[scale=0.65]{figures/axb.pdf}
+ \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_5(\D))$;
  the class $[b]=\v_1\cdot\u_2\cdot\v_3^2\in H_4(C_4(\S'))$; and the product class
@@
 Then $[a]\otimes [b]$ is a generator of $H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))$,
-and we are interested in the homology class $\mu_*([a]\otimes [b])$: there is one such
-class for any choice of $[a]$ and $[b]$ as above, that is, for any choice of a string of numbers
-$(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
-such that $l=\sum_j\alpha_j$, $p=\sum_j\alpha_j2^j$ and $(m-p)=\sum_i(u_i+v_i)$; our aim
-is to show that the collection of all the corresponding classes of the form $\mu_*([a]\otimes [b])$
+by the K\"unneth formula, and we are interested in the homology class $\mu_*([a]\otimes [b])$.
+
+There is one such
+class for any choice of $[a]$ and $[b]$ as above, that is, for any choice of
+$p$, $\ualpha$, $\uu$ and $\uv$ satisfying the conditions
+$l=\sum_{j\geq 0}\alpha_j$, $p=\sum_{j\geq 0}\alpha_j2^j$ and $(m-p)=\sum_{i=1}^g(u_i+v_i)$, where
+we use the notation above.
+
+We want to show that the collection of all the corresponding classes of the form $\mu_*([a]\otimes [b])$
 gives a basis for $H_{m-l}(\cms)$.
@@
 the intersection of $\mu_*([a]\otimes [b])$ with cohomology classes of $\cms$ represented by
-generalised symmetric chains in $\cms^c$.
+generalised symmetric chains in $\cms^{\infty}$.
 
-% $[a]\otimes [b]$ has a dual class in $\tH_{p-l}(C_p(\D)^c)\otimes \tH_p(C_{m-p}(\S')^c)$
+% $[a]\otimes [b]$ has a dual class in $\tH_{p-l}(C_p(\D)^{\infty})\otimes \tH_p(C_{m-p}(\S')^{\infty})$
 % represented by the product of generalised symmetric chains
@@
 % the \emph{geometric} intersection between the cycle $|p',(\alpha'_j),(u'_i,v'_i)|$
-% in $\cms^c$ and the cycle $\mu_*(a\otimes b)$ in $\cms$ is empty: indeed the first cycle is supported
-% on $\infty\in\cms^c$ and on configurations where at least $p'$ points lie in the union
+% in $\cms^{\infty}$ and the cycle $\mu_*(a\otimes b)$ in $\cms$ is empty: indeed the first cycle is supported
+% on $\infty\in\cms^{\infty}$ and on configurations where at least $p'$ points lie in the union
 % $\bigcup_i(\U_i\cup\V_i)\subset\S'$,
@@
 To compute the algebraic intersection
-between $\mu_*([a]\otimes [b])$ and $[p',(\alpha'_j),(u'_i,v'_i)]$ we
+between $\mu_*([a]\otimes [b])$ and $[\kappa(p',\ualpha',\uu',\uv')]$ we
 consider the map
 \[
- \mu^c\colon \cms^c\to \pa{C_p(\D)\times C_{m-p}(\S')}^c
+ \mu^{\infty}\colon \cms^{\infty}\to \pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}
 \]
-which collapses to $\infty$ the complement in $\cms^c$ of $C_p(\D)\times C_{m-p}(\S')$.
+which collapses to $\infty$ the complement in $\cms^{\infty}$ of the open submanifold $C_p(\D)\times C_{m-p}(\S')$.
 
-By Poincaré-Lefschetz duality, the map $\mu^c_*$ in reduced homology corresponds to the cohomology map
+By Poincaré-Lefschetz duality, the map $\mu^{\infty}_*$ in reduced homology corresponds to the cohomology map
 \[
@@
 
-We give $\pa{C_p(\D)\times C_{m-p}(\S')}^c$ the cell complex structure of the smash product
-$C_p(\D)^c\wedge C_{m-p}(\S')^c$. Here $C_p(\D)^c$ is given the cell structure of $C_p((0,1)^2)^c$
-coming from the natural identification $\D=(1/4,3/4)\times(1/2,1)\cong(0,1)^2$, which is obtained by rescaling
+We give $\pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}$ the cell complex structure of the smash product
+$C_p(\D)^{\infty}\wedge C_{m-p}(\S')^{\infty}$. Here $C_p(\D)^{\infty}$ is given the cell structure of $C_p((0,1)^2)^{\infty}$
+coming from the natural identification $\D=]1/4,3/4[\times]1/2,1[\cong]0,1[^2$, which is obtained by rescaling
 and translating. Moreover we choose any diffeomorphism $\mrS'\cong\mrS$ that restricts to the identity
-on all $\U_i$'s and $\V_i$'s, and give $C_{m-p}(\S')^c$ the cell structure of $C_{m-p}(\S)^c$.
+on all $\U_i$'s and $\V_i$'s, and give $C_{m-p}(\S')^{\infty}$ the cell structure of $C_{m-p}(\S)^{\infty}$.
 
-Recall that $\cms^c$ can be filtered according to the norm of cells: a cell $e^{\tup}$ with
-$\tup=(l,(x_i),(u_i,v_i))$ has norm $\sum x_i$, and the norm is weakly decreasing along boundaries.
+Recall that $\cms^{\infty}$ can be filtered according to the norm of cells: a cell $e^{\tup}$ associated with
+the tuple $\tup=(l,\ux,\uu,\uv)$ has norm $\sum_{i=1}^l x_i$, and the norm is weakly decreasing along boundaries.
 In the previous section we just considered
-the associated filtration of the reduced chain complex $\tCh_*(\cms^c)$, whereas now we
-consider the closed subcomplex $F_p\cms^c\subset\cms^c$, which is the union
+the associated filtration of the reduced chain complex $\tCh_*(\cms^{\infty})$, whereas now we
+consider the closed subcomplex $F_p\cms^{\infty}\subset\cms^{\infty}$, which is the union
 of all cells of norm $\leq p$.
 
-The crucial observation is that $\mu^c$ restricts to a cellular map 
+The crucial observation is that $\mu^{\infty}$ restricts to a cellular map 
 \[
-F_p\cms^c\to \pa{C_p(\D)\times C_{m-p}(\S')}^c.
+F_p\cms^{\infty}\to \pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}.
 \]
 
-To see this, consider an open cell
-cell $e^{\tup}\subset\cms$ for some tuple $\tup=(l', (x_i),(u'_i,v'_i))$ of norm $p'\leq p$.
-If $p'<p$, then
-$e^{\tup}\cap (C_p(\D)\times C_{m-p}(\S'))$ is empty. If $p'=p$, then
+To see this, fix a tuple $\tilde{\tup}=(\tilde{l}, \tilde{\ux},\tilde{\uu},\tilde{\uv})$ of norm $\tilde{p}\leq p$
+and of dimension $\tilde{l}+m$, and
+consider the open cell
+cell $e^{\tilde{\tup}}\subset F_p\cms^{\infty}$.
+
+If $\tilde{p}<p$, then
+$e^{\tilde{\tup}}\cap (C_p(\D)\times C_{m-p}(\S'))$ is empty. If $\tilde{p}=p$, then
 \[
-e^{\tup}\cap (C_p(\D)\times C_{m-p}(\S'))=e^{\tup'}\times e^{\tup''},
+e^{\tilde{\tup}}\cap (C_p(\D)\times C_{m-p}(\S'))=e^{\tup'}\times e^{\tup''},
 \]
-where $\tup'=(l',(x_i))$ and $\tup''=(0,(u'_i,v'_i))$.
+where $\tup'=(\tilde{l},\tilde{\ux})$ and $\tup''=(0,\tilde{\uu},\tilde{\uv})$.
 
 Therefore
-$\mu^c(e^{\tup})$ is $\set{\infty}$ in the first case, and in the second case it
+$\mu^{\infty}(e^{\tilde{\tup}})$ is $\set{\infty}$ in the first case, and in the second case it
 is contained in the union $\set{\infty}\cup e^{\tup'}\times e^{\tup''}$, which is also
-contained in the $p'-$skeleton of $\pa{C_p(\D)\times C_{m-p}(\S')}^c$.
+contained in the $(\tilde{l}+m)$-skeleton of $\pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}$.
 
-% We replace the map $\mu^c$ by a cellular approximation that agrees with it on the $p-$skeleton,
-% and by abuse of notation we still call $\mu^c$ the new map.
+% We replace the map $\mu^{\infty}$ by a cellular approximation that agrees with it on the $p$-skeleton,
+% and by abuse of notation we still call $\mu^{\infty}$ the new map.
 % 
-Consider now any generalised symmetric chain $|p',(\alpha'_j),(u'_i,v'_i)|$ representing
-a class in $\tH_{m+l}(\cms^c)=H^{m-l}(\cms)$, hence $l=\sum_j\alpha'_j$; and suppose $p'\leq p$.
+Consider now the generalised symmetric chain $\kappa(p',\ualpha',\uu',\uv')$ representing
+a class in $\tH_{m+l}(\cms^{\infty})=H^{m-l}(\cms)$, with $\ualpha'=(\alpha'_j)_{j\geq 0}$,
+$\uu'=(u_1,\dots,u'_g)$ and $\uv'=(v'_1,\dots,v'_g)$; in particular $l=\sum_{j\geq 0}\alpha'_j$.
+Suppose moreover $p'\leq p$.
 
-If $p'<p$, the previous argument shows that $\mu^c_*\pa{|p',(\alpha'_j),(u'_i,v'_i)|}=\infty$,
+If $p'<p$, the previous argument shows that $\mu^{\infty}_*\pa{\kappa(p',\ualpha',\uu',\uv')}=0$
+in the reduced cellular chain complex,
 and in particular the corresponding homology class is mapped to zero.
 
-Suppose now $p'=p$: then the homology class $[p,(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$
-is mapped along $\mu^c_*$ to the class
+Suppose now $p'=p$: then the previous argument shows that
+the homology class $[\kappa(p',\ualpha',\uu',\uv')]\in \tH_{m+l}(\cms^{\infty})$
+is mapped along $\mu^{\infty}_*$ to the class
 \[
-[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]\in\tH\pa{C_p(\D)^c\wedge C_{m-p}(\S')^c}.
+[\kappa(\ualpha')]\otimes [\kappa(\uu',\uv')]\in\tH\pa{C_p(\D)^{\infty}\wedge C_{m-p}(\S')^{\infty}}.
 \]
-Indeed each cell appearing in the cycle $|p,(\alpha'_j),(u'_i,v'_i)|$ is mapped with degree
-1 onto a corresponding cell of the cycle $|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|$, and hits no other cell
-in dimension $m+l$: hence even at the level of chains we have 
+Indeed each tuple $\tup$ in the cycle $\kappa(p',\ualpha',\uu',\uv')|$ is mapped by $\mu^{\infty}_*$
+to a corresponding pair of tuples $\tup'\otimes\tup''$ in the cycle $\kappa(\ualpha')\otimes \kappa(\uu',\uv')$,
+so even at the level of chains we have
 \[
-\mu^c_*\pa{|p,(\alpha'_j),(u'_i,v'_i)|}=|l,(\alpha'_j)|\otimes |0,(u'_i,v'_i)|.
+\mu^{\infty}_*\pa{\kappa(p',\ualpha',\uu',\uv')}=\kappa(\ualpha')\otimes \kappa(\uu',\uv').
 \]
@@
 We can now compute the algebraic intersection of $\mu_*([a]\otimes [b])$ with
-any cohomology class $[p',(\alpha'_j),(u'_i,v'_i)]$ as the intersection between
-$[a]\otimes [b]$ and $\mu^c_*[p',(\alpha'_j),(u'_i,v'_i)]$.
+the cohomology class $[\kappa(p',\ualpha',\uu',\uv')]$ as the algebraic intersection between
+$[a]\otimes [b]$ and $\mu^{\infty}_*\pa{[\kappa(p',\ualpha',\uu',\uv')]}$.
 
@@
 
-For $p'=p$ we see that the intersection between
-$[a]\otimes [b]$ and $\mu^c_*([p,(\alpha'_j),(u'_i,v'_i)])=[l,(\alpha'_j)]\otimes [0,(u'_i,v'_i)]$
-is $1\in\Z_2$ exactly when $u_i=u'_i$ and $v_i=v'_i$ for all $i$, and $\alpha_j=\alpha'_j$
-for all $j$; otherwise it is $0$.
+For $p'=p$ the intersection between
+$[a]\otimes [b]$ and $\mu^{\infty}_*([\kappa(p,\ualpha',\uu',\uv')])=[\kappa(\ualpha')]\otimes [\kappa(\uu',\uv')]$
+is $1\in\Z_2$ exactly when $\ualpha=\ualpha'$, $\uu=\uu'$ and $\uv=\uv'$; otherwise it is $0$.
 
-We consider the collection of all strings of the form $(p,(\alpha_j)_{j\geq 0},(u_i,v_i)_{1\leq i\leq g})$
+To finish the proof we consider the collection of all strings of the form
+\[
+\pa{p,\ualpha=(\alpha_j)_{j\geq 0},\uu=(u_1,\dots,u_g),\uv=(v_1,\dots,v_g)}
+\]
 satisfying $l=\sum_j\alpha_j$,$p=\sum_j \alpha_j2^j$ and $(m-p)=\sum_i(u_i+v_i)$;
-we order these strings lexicographically,
-so that in particular $p$ is weakly increasing; we associate
+we choose a total order on the set of these strings,
+such that the parameter $p$ is weakly increasing along this order; we associate
 to each string its corresponding class in $H_{m-l}(\cms)$ of the form $\mu_*([a]\otimes [b])$ and its
-corresponding class $[p,(\alpha_j),(u_i,v_i)]\in\tH_{m+l}(\cms^c)$.
+corresponding class $[\kappa(p,\ualpha',\uu',\uv')]\in\tH_{m+l}(\cms^{\infty})$.
 
@@
 with $1$'s on the diagonal, and in particular
-it is invertible. This shows that the set of classes $\mu_*([a]\otimes [b])$
+it is invertible. This shows that the set of classes of the form $\mu_*([a]\otimes [b])$
 is a basis for $H_{m-l}(\cms)$.
@@
 % $\tup=(l,(x_i)_{i\leq l}, (u_i,v_i)_{i\leq g})$, and consider the restriction
-% of $\mu^c\colon\cms^c\to \pa{C_{m-p}(\S')\times C_p(\D)}^c$
-% to $e^{\tup}\subset\cms^c$
+% of $\mu^{\infty}\colon\cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$
+% to $e^{\tup}\subset\cms^{\infty}$
 % 
@@
 % to $\infty$: indeed if $p<p'$, then a configuration in $e^{\tup}$ contains
-% at least $p'$ points inside $\S'$, so it can't lie in $C_p(\D)\times C_{m-p}(\S')$;
+% at least $p'$ points inside $\S'$, so it cannot lie in $C_p(\D)\times C_{m-p}(\S')$;
 % if $p'<p$ then a configuration in $e^{\tup}$ contains at least $m-p$ points 
@@
 
-One could expect that the basis given by all elements $[a]\otimes [b]\in H_{m-l}(\cms)$
-is also \emph{dual} to the basis of classes $[p',(\alpha'_j),(u'_i,v'_i)]\in \tH_{m+l}(\cms^c)$, i.e., the
+One could expect that the basis given by classes of the form $[a]\otimes [b]\in H_{m-l}(\cms)$
+is also \emph{dual} to the basis of classes $[\kappa(p,\ualpha,\uu,\uv)]\in \tH_{m+l}(\cms^{\infty})$, i.e., the
 matrix considered in the end of the previous proof is not only upper-triangular but also
@@
 
-Let $g=1$, $m=2$, $p=1$, $p'=2$ and consider the classes $[a]=\epsilon\in H_1(C_1(\D))$,
-$[b]=\u_1\in \H=H_1(C_1(\S'))$; moreover let $[p',(\alpha'_j),u'_1,v'_1]$ be defined by
-$p'=2$, $\alpha'_1=1$, $u'_1=v'_1=0$.
+Let $g=1$, $m=2$, $p=1$, $p'=2$ and consider the classes $[a]=\epsilon\in H_0(C_1(\D))$,
+$[b]=\u_1\in \H=H_1(C_1(\S'))$. Moreover let the generalised symmetric chain
+$\kappa(p',\ualpha',\uu',\uv')$ be defined by
+$\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\alpha'_1=1$ and all other $\alpha_j=0$, $\uu'=(u'_1=0)$
+and $\uv'=(v'_1=0)$.
 
@@
 that intersects only once, transversely, the vertical segment passing through $a$, i.e.
-$\set{1/2}\times(0,1)$. See figure \ref{fig:counterexample}.
+$\set{1/2}\times]0,1[$. See Figure \ref{fig:counterexample}.
 
 \begin{figure}\centering
- \includegraphics[scale=0.6]{figures/counterexample.png}
+ \includegraphics[scale=0.6]{figures/counterexample.pdf}
  \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}
@@
 
-Then the cycle $|p',(\alpha'_j),u'_1,v'_1|$ consists uniquely of one cell $e^{\tup'}$, where
-$\tup'=(1,(x'_1=2),(u'_1=v'_1=0))$; and this cell intersects once, transversely,
+Then the cycle $\kappa(p',\ualpha',\uu',\uv')$ consists uniquely of one tuple
+\[
+\tup'=(1,\ux'=(x'_1=2),\uu'=(u'_1=0),\uv'=(v'_1=0)).
+\]
+The corresponding cell $e^{\tup'}$ intersects once, transversely,
 the cycle $\mu_*(a\otimes b)$, which is represented by the curve of configurations of two
-points in $\mrS$, one of which is fixed at $a$ whereas the other runs along $b$: indeed
-there is exactly one position for the second point to lie under $a$.
+points in $\mrS$, one of which is fixed at $a$ whereas the other runs along $b$:
+there is exactly one position on $b$ lying under $a$.
 
@@
 this is an entry strictly above the diagonal in the matrix considered in the proof
-of lemma \ref{lem:oplussplitting}.
+of Lemma \ref{lem:oplussplitting}.
 
-The proof of theorem \ref{thm:Hbms*as*ggrep} can be easily generalised to surfaces
+The proof of Theorem \ref{thm:Hbms*as*ggrep} can be easily generalised to surfaces
 with more than one boundary component. Let $\Sigma_{g,n}$ be a surface of genus
@@
 $\Diff(\Sigma_{g,n};\partial\Sigma_{g,n})$:
-then there is an isomorphism of bigraded $\Z_2-$representations
-\[
- \bigoplus_{m\geq 0}H_*\pa{C_m(\Sigma_{g,n})}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(H_1(\Sigma_{g,m})),
-\]
+then there is an isomorphism of bigraded $\Z_2$-representations
+\begin{equation}
+\label{eq:sgn}
+\bigoplus_{m\geq 0}H_*\pa{C_m(\Sigma_{g,n})}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(H_1(\Sigma_{g,n})),
+\end{equation}
+
 where the action of $\Gamma_{g,n}$ on the right-hand side is induced by the natural
-action on $H_1(\Sigma_{g,n})$; for $n\geq 2$ the $\Z_2-$vector space $H_1(\Sigma_{g,n})$
-doesn't carry a \emph{non-degenerate} symplectic form which is invariant under the action of
-$\Gamma_{g,n}$, hence we can't any longer use the word \emph{symplectic} for this representation.
+action on $H_1(\Sigma_{g,n})$.
 
-The proof is word by word the same; the main difference is in the construction of the
+For $n\geq 2$ the intersection form on the vector space $H_1(\Sigma_{g,n})$
+is \emph{degenerate}, but it is still invariant under the action of
+$\Gamma_{g,n}$, so there is still a map from $\Gamma_{g,n}$ to the subgroup
+of $GL_{2g+n-1}(\Z_2)$ fixing this bilinear form, and in this sense we can say
+that the representation in \eqref{eq:sgn} is \emph{symplectic}.
+
+The proof of the isomorphism \eqref{eq:sgn} is almost verbatim the same; the main difference is in the construction of the
 model $\T(\Sigma_{g,n})$ for $\mathring{\Sigma}_{g,n}$: we divide the

### 2019-02-18 HBraidSurf.tex
Implemented some corrections by cfb, created gitignore file

@@
 in the Leray-Serre spectral sequence associated to
-\ref{eq:Birman}, \ref{eq:Birmanbundle} or \ref{eq:BirmanbundleD}.
+\eqref{eq:Birman}, \eqref{eq:Birmanbundle} or \eqref{eq:BirmanbundleD}.
 In particular we are interested in $H_*(C_m(\S))$
-as a $\Z_2-$representation of the group $\gg$.
+as a $\Z_2$-representation of the group $\gg$.
 
-In \cite{LM} L\"offler and Milgram implicitly proved that $H_*(\bms)=H_*(\cms)$ is a $\Z_2-$symplectic
-representation of the mapping class group. By \emph{$\Z_2-$symplectic} we mean the following:
+In \cite{LM} L\"offler and Milgram implicitly proved that $H_*(\bms)=H_*(\cms)$ is a $\Z_2$-symplectic
+representation of the mapping class group. By \emph{$\Z_2$-symplectic} we mean the following:
 \begin{defn}
@@
  The natural action of $\gg$ on $\H$ induces a surjective map
- $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{$\Z_2-$symplectic}
+ $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{$\Z_2$-symplectic}
  if it is a pull-back of a representation of $Sp_{2g}(\Z_2)$ along this map.
@@
 
-In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(\cms)$ \emph{as a graded $\Z_2-$vector space}.
+In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(\cms)$ \emph{as a graded $\Z_2$-vector space}.
 Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced:
 their descripition of $H_*(\cms)$ depends on a handle decomposition of $\S$, which is not preserved
-by diffeomorphisms of $\S$, not even up to isotopy,.
+by diffeomorphisms of $\S$, not even up to isotopy.
 
 In this section and in the next one we will prove the following theorem; to the best of the author's knowledge
-it doesn't appear in the literature.
+it does not appear in the literature.
 \begin{thm}
  \label{thm:Hbms*as*ggrep}
- There is an isomorphism of bigraded $\Z_2-$representations of $\gg$
+ There is an isomorphism of bigraded $\Z_2$-representations of $\gg$
  \[
@@
  \begin{itemize}
-  \item the bidegree is given on left by homological degree $*$ and by the direct summand,
+  \item[(i)] on the left-hand side the bigrading is given by homological degree $*$ and by the direct summand,
   indexed by $m$,
@@
   involved in constructing the homology class; we call $*$ the \emph{degree} and $m$ the \emph{weight},
-  and write $(*,m)$ for the bidegree;
-  \item for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
+  and write $(*,m)$ for the bigrading;
+  \item[(ii)] for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
   of the group $H_{2^j-1}(C_{2^j}(\D))\simeq \Z_2$ 
@@
   infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
-  \item $\H=H_1(\sg)$ is identified with $H_1(C_1(\S))$ in a natural way, and $\Sym_{\bullet}(\H)$ is the
-  symmetric algebra on $\H$;
-  %\item degrees and weights are extended on right by the usual multiplicativity-additivity rule;
-  \item the action of $\gg$ on right is the tensor product of the trivial action
+  \item[(iii)] $\H=H_1(\sg)$ is identified with $H_1(C_1(\S))$ in a natural way, and lives in degree $1$ and weight $1$;
+  $\Sym_{\bullet}(\H)$ denotes the symmetric algebra on $\H$;
+  \item[(iv)] degrees and weights are extended on the right-hand side by the usual multiplicativity rule;
+  \item[(v)] the action of $\gg$ on the right is the tensor product of the trivial action
   on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$, and of the action on
-  $\Sym_{\bullet}(\H)$ which is induced by the $\Z_2-$symplectic action on $\H$.
+  $\Sym_{\bullet}(\H)$ which is induced by the $\Z_2$-symplectic action on $\H$.
   \end{itemize}
 \end{thm}
-Notice that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
-the degree: indeed factors of the form $Q^j\epsilon$ have weight strictly higher than the degree,
-whereas factors in $\H$ and its symmetric powers have equal weight and degree.
+Note that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
+the degree: indeed factors of the form $Q^j\epsilon$ have weight strictly higher than their degree,
+whereas factors belonging to $\H$ or to its symmetric powers have equal weight and degree.
 
-Notice that in the case $g=0$ the group $\Gamma_{0,1}$ is trivial and the previous theorem
-reduces to equation \ref{eq:Cohen}.
+Note that in the case $g=0$ the group $\Gamma_{0,1}$ is trivial and the previous theorem
+reduces to equation \eqref{eq:Cohen}.
+
+We point out that in \cite{BoT} B\"{o}digheimer and Tillmann have essentially proved that for a field
+$\mathbb{F}$ of any characteristic the $\mathbb{F}$-vector space
+\[
+  \bigoplus_{m\geq 0} H_*\pa{\cms;\mathbb{F}}
+\]
+is isomorphic, as a bigraded $\gg$-representation over $\mathbb{F}$, to the tensor product
+of the ring $\mathbb{F}[\epsilon]$, with trivial action, and some other bigraded representation:
+here $\epsilon$ denotes, in analogy with the notation of Theorem \ref{thm:Hbms*as*ggrep}, the standard generator of $H_0(C_1(\D))$.
+
 % isomorphism of bigraded vector spaces
@@
 % \label{eq:genuszero}
-% \bigoplus_{m\geq 0} H_*\pa{C_m(\Sigma_{0,1})}\simeq \Z_2[Q^j\epsilon\,|\, j\geq 0].
+% \bigoplus_{m\geq 0} H_*\pa{C_m(\D)}\simeq \Z_2[Q^j\epsilon\,|\, j\geq 0].
 % \end{equation}
 % This isomorphism was essentially proved by Fuchs in \ref{Fuchs:CohomBraidModtwo}. Moreover the left hand
-% side is the homology of the space $\coprod_{m\geq 0} C_m(\Sigma_{0,1})$, which has a natural structure
-% of algebra over the operad $E^2$ of little $2-$cubes. The isomorphism in equation \ref{eq:genuszero}
+% side is the homology of the space $\coprod_{m\geq 0} C_m(\D)$, which has a natural structure
+% of algebra over the operad $E^2$ of little $2$-cubes. The isomorphism in equation \eqref{eq:genuszero}
 % is an isomorphism of rings (using the Pontryagin product on the left hand side), and $Q$ represents here
 % the Araki-Kudo-Dyer-Lashof operation $Q\colon H_*(X;\Z_2)\to H_{2*+1;\Z_2}(X)$,
-% which is defined for any $E^2-$algebra $X$; $\epsilon\in H_0(C_1(\Sigma_{0,1}))$ is the generator.
+% which is defined for any $E^2$-algebra $X$; $\epsilon\in H_0(C_1(\D))$ is the generator.
 % This explains our notation; we refer to Cohen \cite[Chap.3]{CLM} for more details.
 
-In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2-$vector spaces}
-as in theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
+In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2$-vector spaces}
+as in Theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
 $\gg$.
@@
 
-Since we work with coefficients in the field $\Z_2$, it is equivalent to compute homology or cohomology,
-and in this section we will prefer to compute $H^*(\cms)$ for all bidegrees $(*,m)$.
+Since we work with coefficients in a field, it is equivalent to compute homology or cohomology,
+and in this section we will prefer to compute $H^*(\cms)$ for all bigradings $(*,m)$.
 % the dual graded vector space $H_*(\cms)$ is abstractly isomorphic,
@@
 
-We will mimic the method used by Fuchs (\cite{Fuchs:CohomBraidModtwo}) to compute the $\Z_2-$cohomology
+We will mimic the method used by Fuchs \cite{Fuchs:CohomBraidModtwo} to compute the $\Z_2$-cohomology
 of $C_m(\D)$.
-As already said this computation recovers a known result, but it has the advantage of
+As already mentioned in section \ref{sec:Preliminaries}, our computation recovers a known result, but it has the advantage of
 being quite elementary and of providing a part of
@@
 In the whole section we assume $m\geq 0$ to be fixed.
+Since the space $\cms$ is homeomorphic to the interior of a compact
+$2m$-manifold with boundary, by Poincaré-Lefschetz
+duality we have
+\[
+ H^*(\cms)\simeq \tilde H_{2m-*}(\cms^{\infty}),
+\]
+where in the right hand side we consider reduced homology of the one-point compactification (see Definition \ref{defn:opc}).
 
-We first introduce a space $\T(\S)$ which is homeomorphic to $\mrS$, the interior of $\S$. The construction
-corresponds to a handle decomposition of $\S$ with one $0-$handle and $2g$ $1-$handles.
+We introduce a space $\T(\S)$ which is homeomorphic to $\mrS$, the interior of $\S$. The construction
+corresponds to a handle decomposition of $\S$ with one $0$-handle and $2g$ $1$-handles.
 
 \begin{defn}
-\label{defn:Tsg}
-If $g=0$, hence $\S=\Sigma_{0,1}$ is the disc, we set $\T(\S)=(0,1)^2$, the interior of the unit square. Assume
-now $g\geq 1$, and see figure \ref{fig:defTS} to visualize the following construction.
+\label{defn:TS}
+If $g=0$, hence $\S=\Sigma_{0,1}$ is the disc, we set $\T(\S)=]0,1[^2$, the interior of the unit square. Assume
+now $g\geq 1$, and see Figure \ref{fig:defTS} to visualize the following construction.
 
@@
 
-Let $Q\subset[0,1]^2$ be the union of $(0,1)^2$ and the vertical open intervals
-$I_i^l=\set{0}\times (\cP_i,\cP_{i+1})$ and $I_i^r=\set{1}\times (\cP_i,\cP_{i+1})$ for $1\leq i\leq 2g$;
-equivalently, we remove from $[0,1]^2$ the two horizontal sides
-$[0,1]\times\set{0,1}$ and the $4g-2$ points $\set{0,1}\times \set{\cP_1,\dots,\cP_{2g-1}}$.
+Consider on the vertical sides of $[0,1]^2$ the intervals
+$I_i^l=\set{0}\times [\cP_i,\cP_{i+1}]$ and $I_i^r=\set{1}\times [\cP_i,\cP_{i+1}]$ for $1\leq i\leq 2g$:
+all these intervals are
+canonically diffeomorphic
+to $[0,1]$ by projecting on the second coordinate, rescaling linearly by a factor $2g$
+and translating.
 
-Notice that all intervals $I_i^l$ and $I_j^r$ are
-canonically diffeomorphic
-to $(0,1)$ by projecting on the second coordinate, rescaling linearly by a factor $2g$
-and translating; therefore we will specify a bijection
-between the two sets of left and right intervals, and then $\T(\S)$ will be obtained from $Q$
-by identifying in the canonical way the two intervals in each couple.
+We define a bijection
+between the two sets of left and right intervals:
+for $1\leq i\leq g$, the interval $I^l_{2i-1}$ corresponds to $I^r_{2i}$,
+and the interval $I^r_{2i-1}$ corresponds $I^l_{2i}$.
 
-For $1\leq i\leq g$, we identify $I^l_{2i-1}$ with $I^r_{2i}$, obtaining an open interval $\U_i\subset\T(\S)$,
-and we identify $I^r_{2i-1}$ with $I^l_{2i}$, obtaining an open interval $\V_i\subset\T(\S)$.
-Each interval $\U_i,\V_i$ has a natural parametrisation by $(0,1)$.
+The space $Q(\S)$ is the quotient of the square $[0,1]^2$ obtained by identifying each couple
+of corresponding intervals in the canonical way. For $1\leq i\leq g$ we call $\overline{\U}_i$ the image
+of $I^l_{2i-1}$ in the quotient $Q(\S)$, and we call $\overline{\V}_i$ the image of $I^l_{2g}$ in $Q(\S)$.
 
-The space $\T(\S)$ is homeomorphic to $\mrS$. From now
-on we will identify the two open surfaces and in particular we will identify $\cms$ with the
+Note that the image in $Q(\S)$ of the set $\set{0,1}\times\set{\cP_0,\dots,\cP_{2g}}$ consists of two
+points $\cP_{odd}$ and $\cP_{even}$: for $\epsilon\in\set{0,1}$ and $0\leq i\leq 2g$ the point
+$(\epsilon,\cP_i)$ is mapped to $\cP_{even}$ if $\epsilon+i$ is even, and is mapped to $\cP_{odd}$
+otherwise.
+
+The spaces $\overline{\U}_i\subset Q(\S)$ and $\overline{\V}_i\subset Q(\S)$ are intervals with endpoints $\cP_{even}$ and $\cP_{odd}$;
+the interiors of these intervals are disjoint. Each $\overline{\U}_i$ and $\overline{\V}_i$ is the homeomorphic image of
+some left interval $I^l_j$, and inherits from the latter a parametrisation by $[0,1]$.
+
+We call $\U_i$ and $\V_i$ the interiors of the intervals $\overline{\U}_i$ and $\overline{\V}_i$ respectively.
+
+The space $Q(\S)$ is homeomorphic to the compact surface $\S$; we call $\T(\S)$ the interior of
+$Q(\S)$.
+
+From now on we will identify $\S$ with $Q(\S)$ and $\mrS$ with $\T(\S)$;
+consequently we will identify $\cms$ with the
 space of configurations of $m$ points in $\T(\S)$.
-
-In this model we fix $\D\subset\mrS$ to be the open square $(1/4;3/4)\times(1/2,1)$:
-$\D$ is an open disc in $\mrS$ near $\partial\S$ and 
-it is disjoint from all $\U_i$'s and $\V_i$'s. The interior of the surface $\S'$
-is then identified with $\T(\S)\setminus [1/4,3/4]\times[1/2,1)$.
-
 \end{defn}
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defTS.png}
+ \includegraphics[scale=0.7]{figures/defTS.pdf}
  \caption{The space $\T(\S)$ for $\S=\Sigma_{3,1}$.}
@@
 
-In the following we will consider the one-point compactification of some spaces that
-happen to be open manifolds. For a space $\mathcal{M}$ we denote by $\mathcal{M}^c$ its one-point
-compactification; the basepoint is the point at infinity, that we always denote by $\infty$, even
-if it is a different point for different spaces $\mathcal{M}^c$.
-
-We consider in particular the one-point-compactification $\cms^c$ of $\cms$.
-Since the space $\cms$ is homeomorphic to the interior of a compact
-$2m-$manifold with boundary, by Poincaré-Lefschetz
-duality we have
-\[
- H^*(\cms)\simeq \tilde H_{2m-*}(\cms^c).
-\]
-Our next aim is to define a cell structure on the space $\cms^c$.
+Our next aim is to define a structure of $CW$-complex on the space $\cms^{\infty}$:
+the only $0$-cell will be the point $\infty$, whereas the other cells will be introduced
+in the following definition.
 \begin{defn}
@@
   \item a natural number $0\leq l\leq m$;
-  \item integers $x_1,\dots,x_l\geq 1$;
-  \item integers $u_1,\dots,u_g$ and $v_1,\dots,v_g$, all $\geq 0$.
+  \item a vector $\ux=(x_1,\dots, x_l)$ of integers $\geq 1$;
+  \item vectors $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ of integers $\geq 0$,
  \end{itemize}
-with the condition that
+satisfying the following equality
 \[
@@
 % $\bsymb=\set{(U_i,V_i)_{i\leq g},(W_i)_{i\leq n-1}}\subset\symb$.
-In symbols we write $\tup=(l,(x_i)_{1\leq i\leq l},(u_i,v_i)_{1\leq i\leq g})$.
-The \emph{degree} of $\tup$ is defined as $m+l$.
+In symbols we write $\tup=(l,\ux,\uu,\uv)$. We omit from the notation $\ux$ if it vanishes: this happens precisely when $l=0$.
+The \emph{dimension} of $\tup$ is defined as $m+l$.
 
@@
  and exactly $v_i$ points lie on $\V_i$;
- \item there are exactly $l$ vertical lines in $(0,1)^2\subset\mrS$ of the
- form $\set{s_i}\times(0,1)$ for some $0<s_1<\dots<s_l<1$, containing at least one
+ \item there are exactly $l$ vertical lines in the open square $]0,1[^2\subset\mrS$ of the
+ form $\set{s_i}\times]0,1[$ for some $0<s_1<\dots<s_l<1$, containing at least one
  point of the configuration. From left to right, these lines contain exactly $x_1,\dots,x_l$ points
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defetup.png}
- \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,0,2,3,0,2))$. }
+ \includegraphics[scale=0.7]{figures/defetup.pdf}
+ \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,2,0),(0,3,2))$. Points
+ lying on the $\U_i$'s and the $\V_i$'s are depicted twice.}
 \label{fig:defetup}
@@
 
-The space $e^{\tup}$ is homeomorphic to \emph{the interior} of the following product of simplices:
+The space $e^{\tup}$ is homeomorphic to \emph{the interior} of the following multisimplex:
 
 % We want to show that the collection of the $e_h$ for varying $h$, together with
-% the 0-cell $*$, give a cell decomposition of $C_k(\S)^c$. Denote by $\Delta^h$ the following product of simplices
+% the 0-cell $*$, give a cell decomposition of $C_k(\S)^{\infty}$. Denote by $\Delta^h$ the following product of simplices
 \[
@@
 \begin{itemize}
- \item the local coordinates of the $\Delta^l-$factor correspond to the positions $s_1,\dots,s_l$ of the vertical
+ \item the local coordinates of the $\Delta^l$-factor correspond to the positions $s_1,\dots,s_l$ of the vertical
  lines in $(0,1)^2$ containing points of the configuration;
- \item the local coordinates of the $\Delta^{x_i}-$factor correspond to the positions of the $x_i$ points
+ \item the local coordinates of the $\Delta^{x_i}$-factor correspond to the positions of the $x_i$ points
  lying on the vertical line $\set{s_i}\times(0,1)$;
- \item the local coordinates of the $\Delta^{u_i}-$factor correspond to the positions of the $u_i$ points
- lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}-$factor,
- with $v_i$ and $\V_i$ instead of $u_i$ and $\U_i$.
+ \item the local coordinates of the $\Delta^{u_i}$-factor correspond to the positions of the $u_i$ points
+ lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}$-factor,
+ with $v_i$ and $\V_i$ replacing $u_i$ and $\U_i$.
 \end{itemize}
-Notice that the dimension of $\Delta^{\tup}$ is equal to the degree of $\tup$.
-The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^c$ extends to a continuous map
-$\phi^{\tup}\colon\Delta^{\tup}\to\cms^c$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
-all subspaces $e^{\tup'}$ for tuples $\tup'$ with \emph{lower} degree than $\tup$, together with the point $\infty$.
+Note that the dimension of $\Delta^{\tup}$ agrees with the dimension of $\tup$ from Definition \ref{defn:ehopen}.
+The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^{\infty}$ extends to a continuous map
+$\Phi^{\tup}\colon\Delta^{\tup}\to\cms^{\infty}$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
+all subspaces $e^{\tup'}$ for tuples $\tup'$ of \emph{lower} dimension than $\tup$, together with the $0$-cell $\infty$.
 
-The construction of the map $\phi^{\tup}$ is as follows:
+The construction of the map $\Phi^{\tup}$ is as follows:
 \begin{enumerate}
 \item we identify the one-point compactification
-$(\mrS)^c$ of $\mrS=\T(\S)$ as the quotient of $[0,1]^2$ that identifies left and right intervals as in
-in definition \ref{defn:Tsg} of $\T(\S)$, and moreover collapses $[0,1]^2\setminus Q$ to the point $\infty$;
-\item we consider the $m-$fold symmetric product $\SP^m\pa{(\mrS)^c}$: it contains 
-as an \emph{open} subspace $\cms$, so we can identify $\cms^c$ as the quotient of $\SP^m\pa{(\mrS)^c}$ collapsing
-the subspace $\SP^m\pa{(\mrS)^c}\setminus\cms$ to $\infty$;
+$(\mrS)^{\infty}$ of $\mrS=\T(\S)$ as the quotient of $Q(\S)$ collapsing the boundary to a point $\infty$;
+\item we consider the $m$-fold symmetric product $\SP^m\pa{(\mrS)^{\infty}}$ (see Definition \ref{defn:SP}): it contains 
+as an \emph{open} subspace $\cms$, so we can identify $\cms^{\infty}$ as the quotient of $\SP^m\pa{(\mrS)^{\infty}}$ collapsing
+the subspace $\SP^m\pa{(\mrS)^{\infty}}\setminus\cms$ to $\infty$;
 \item the homeomorphism $\mathring{\Delta}^{\tup}\to e^{\tup}\subset\cms$ extends
-now to a map $\Delta^{\tup}\to\SP^m\pa{[0,1]^2}$, that we can then further project to
-$ \SP^m\pa{(\mrS)^c}$ and then to $\cms^c$: the composition
-is the map $\phi^{\tup}$.
+now to a map $\Delta^{\tup}\to\SP^m\pa{[0,1]^2}$, that we can then further project to $\SP^m(Q(\S))$, then to
+$ \SP^m\pa{(\mrS)^{\infty}}$ and then to $\cms^{\infty}$: the composition
+is the map $\Phi^{\tup}$.
 \end{enumerate}
+In the last step we used the open inclusion $\cms\subset\SP^m\pa{\mrS}\subset\SP^m\pa{(\mrS)^{\infty}}$, which
+induces a map $ \SP^m\pa{(\mrS)^{\infty}}\to\cms^{\infty}$ by quotienting to $\infty$ the closed subspace
+$ \SP^m\pa{(\mrS)^{\infty}}\setminus \cms^{\infty}$.
 
-Therefore the collection of the $e^{\tup}$'s, together with the $0-$cell $\infty$,
-gives a cell decomposition of $\cms^c$, with characteristic maps of cells $\phi^{\tup}$.
+We conclude that the collection of the $e^{\tup}$'s, together with the $0$-cell $\infty$,
+gives a cell decomposition of $\cms^{\infty}$, with characteristic maps of cells $\Phi^{\tup}$.
 
-We can use these characteristic maps to compute the \emph{reduced} cellular chain complex of $\cms^c$
-with coefficients in $\Z_2$, which
-reads as follows.
+We compute the \emph{reduced} cellular chain complex of $\cms^{\infty}$
+with coefficients in $\Z_2$. It is a chain complex in $\Z_2$-vector spaces with basis given
+by tuples $\tup$, which correspond to the cells $e^{\tup}$.
 \begin{lem}
 \label{lem:doperatoropenmodtwo}
-Let $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$ and
-$\tup'=(l-1,(x'_i)_{i\leq l-1},(u'_i,v'_i)_{i\leq g})$
-be tuples in consecutive degrees $m+l$ and $m+l-1$, and let $[\partial \tup\colon\tup']\in\Z_2$ denote the coefficient
-of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^c)$.
-Then $[\partial \tup:\tup']=0$ unless one (and exactly one) of the following situations occur:
+Let $\tup=(l,\ux,\uu, \uv)$ and
+$\tup'=(l-1,\ux',\uu',\uv')$
+be tuples in consecutive dimensions $m+l$ and $m+l-1$. Denote by $[\tup\colon\tup']\in\Z_2$ the coefficient
+of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$.
+Then $[\tup:\tup']=0$ unless one, and exactly one, of the following situations occurs:
 \begin{itemize}
@@
  for one value $1\leq i\leq l-1$, and shifting the values
- $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. We call this way to obtain $\tup'$ from $\tup$ an \emph{inner differential}.
- In this case
+ $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case we say that $\tup'$ is an \emph{inner boundary} of $\tup$, and
+ we have
  \[
-  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}.
+  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2.
  \]
- This binomial coefficient may be anyway equal to $0\in\Z_2$.
+%  This binomial coefficient may be anyway equal to $0\in\Z_2$.
  \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
- in integers $\delta u_i,\delta v_i\geq 0$
+ in integers $\tildeu_i,\tildev_i\geq 0$
  \[
-  x_1=\sum_{i=1}^g \pa{\delta u_i+\delta v_i},
+  x_1=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
  \]
- setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and shifting the values
- $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. We call this a \emph{left, outer differential}. In this case
+ setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and shifting the values
+ $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case we say that $\tup'$ is a \emph{left, outer boundary} of $\tup$, and we have
  \[
-  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}.
+  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
  \]
- This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
+%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
  \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
- in integers $\delta u_i,\delta v_i\geq 0$
+ in integers $\tildeu_i,\tildev_i\geq 0$
  \[
-  x_l=\sum_{i=1}^g \pa{\delta u_i+\delta v_i},
+  x_l=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
  \]
- setting $u'_i=u_i+\delta u_i$ and $v'_i=v_i+\delta v_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
- We call this a \emph{right, outer differential}. Also in this case
+ setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
+ In this case we say that $\tup'$ is a \emph{right, outer boundary} of $\tup$, and we have
  \[
-  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\delta u_i}{u_i}\binom{v_i+\delta v_i}{v_i}.
+  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
  \]
- This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
+%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
 \end{itemize}
+
 \end{lem}
+It may indeed happen that $\tup'$ is both a left and a right outer boundary of $\tup$,
+namely when all numbers $(x_i)_{1\leq i\leq l}$ are equal; then the two contributions cancel each other,
+so that $[\tup\colon\tup']=0$ as stated.
 \begin{proof}
- If we restrict the map $\phi^{\tup}\colon\Delta^{\tup}\to\cms^c$ to any face of the multisimplex
- $\Delta^{\tup}$ coming from a factor different from $\Delta^l$,
- we obtain a constant map to $\infty$, hence these faces don't contribute
- to the boundary of $e^{\tup}$ in the reduced cellular chain complex. Instead for $0\leq i\leq l$ the restriction
+ For $1\leq i\leq l$ denote by $\partial^{\hor}_i\Delta^{\tup}$ the face
  \[
-  \phi^{\tup}\colon\partial_i\Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}}\to\cms^c
+ \partial_i\Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}}\subset\Delta^{\tup}.
  \]
+ This is also referred as the $i$-th \emph{horizontal face}. All other faces of codimension 1 of
+ the multisimplex $\Delta^{\tup}$ are called \emph{vertical}.
+
+ We note that $\Phi^{\tup}$ restricts to a cellular map $\partial_i^{\hor}\Delta^{\tup}\to\cms^{\infty}$ on every
+ horizontal face, where $\partial_i^{\hor}\Delta^{\tup}$
+ is given the cell structure coming from the multisimplicial structure: every subface of $\partial_i^{\hor}\Delta^{\tup}$ of dimension
+ $k\leq m+l-1$ is mapped to the $k$-skeleton of $\cms^{\infty}$. Therefore the map $\Phi^{\tup}\colon\partial_i^{\hor}\Delta^{\tup}\to\cms^{\infty}$
+ has a well-defined local index over the cell $e^{\tup'}$, that
+ we call $[\partial^{\hor}_i\tup\colon\tup']\in\Z_2$.
+
+ The same holds for vertical faces, where the restriction of $\Phi^{\tup}$ is the constant map to $\infty$: in this case the local index
+ over $e^{\tup'}$ is zero. The index $[\tup\colon\tup']$ of the map $\Phi^{\tup}\colon\partial\Delta^{\tup}\to\cms^{\infty}$ on
+ $e^{\tup'}$ splits as a sum of local indices:
+ \[
+  [\tup\colon\tup']=\sum_{i=0}^l[\partial_i^{\hor}\tup\colon\tup']\in\Z_2.
+ \]
+ For $0\leq i\leq l$ the restriction $\Phi^{\tup}\colon\partial^{\hor}_i\Delta^{\tup}\to\cms^{\infty}$
  hits homeomorphically the open cell $e^{\tup'}$
@@
  of the lemma for the cases $1\leq i\leq l-1$,
- $i=0$ and $i=l$ respectively, and no other open cell of dimension $m+l-1$.
- We are working in $\Z_2$ so we don't have to be careful about orientations of cells
- and therefore about signs.
+ $i=0$ and $i=l$ respectively.
+
+ The only possibility in which the same cell $e^{\tup'}$ is hit by different faces $\partial_i^{\hor}\Delta^{\tup}$
+ and $\partial_j^{\hor}\Delta^{\tup}$ is the one described in the remark preceding this proof: in this case
+ there are two equal contributions $[\partial_i^{\hor}\tup\colon\tup']$ and $[\partial_j^{\hor}\tup\colon\tup']$
+ that cancel each other.
+%  We are working in $\Z_2$ so we do not have to be careful about orientations of cells
+%  and therefore about signs.
 \end{proof}
 
-We can filter the chain complex $\tilde\Ch_*(\cms^c)$ by giving norm $\sum_{i=1}^lx_i$ to
-the cell $e^{\tup}$, with $\tup=(l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g})$; for example
-the cell in figure \ref{fig:defetup} has norm 6. By lemma \ref{lem:doperatoropenmodtwo}
-the norm is weakly decreasing along differentials. Let $F_p\subset\tCh_*(\cms^c)$ be the subcomplex of cells
-of norm $\leq p$, and let $F_p/F_{p-1}$ be the $p-$th filtration stratum.
+We can filter the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$ by giving filtration norm $\sum_{i=1}^lx_i$ to
+the tuple $\tup=(l,\ux,\uu,\uv)$, with $\ux=(x_1,\dots,x_l)$. For example
+the tuple in Figure \ref{fig:defetup} has norm 6.
 
-Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1}))$,
-one copy for each way of splitting $(m-p)=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$. The isomorphism
-doesn't preserve the degrees but shifts them by $p$.
+By Lemma \ref{lem:doperatoropenmodtwo}
+the norm is weakly decreasing along differentials. Denote by $F_p\subset\tCh_*(\cms^{\infty})$ the subcomplex generated by
+tuples of norm $\leq p$, and let $F_p/F_{p-1}$ be the $p$-th filtration stratum.
+Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1})^{\infty})$:
+there is one copy for each partition
+\[
+(m-p)=\sum_{i=1}^g (u_i+v_i)
+\]
+with $u_i,v_i\geq 0$. The isomorphism
+does not preserve the degrees but shifts them by $p$.
 
 Indeed in $F_p/F_{p-1}$ all outer
-differentials vanish (see lemma \ref{lem:doperatoropenmodtwo}
-for the definition), and in particular the numbers $u_i,v_i$
-don't change along the differentials of $F_p/F_{p-1}$. Therefore $F_p/F_{p-1}$ splits as a direct sum
-of chain complexes indexed by splittings $(m-p)=\sum_{i=1}^g (u_i+v_i)$ as above.
-It is then immediate to identify the inner differentials
-with the ones one would have in the case $g=0$, i.e. for the surface $\Sigma_{0,1}$.
+differentials vanish (see Lemma \ref{lem:doperatoropenmodtwo}): in particular the numbers $u_i,v_i$
+do not change along the differentials of $F_p/F_{p-1}$. Therefore $F_p/F_{p-1}$ splits as a direct sum
+of chain complexes indexed by all partitions $(m-p)=\sum_{i=1}^g (u_i+v_i)$ as above.
+It is then immediate to identify the inner faces
+with the ones one would have in the case $g=0$, i.e. for the surface $\Sigma_{0,1}$, after shifting
+degrees by $p$.
 
-We notice that $\tCh_*(C_p(\Sigma_{0,1})^c)$ is exactly the chain complex described by Fuchs in
-\cite{Fuchs:CohomBraidModtwo}, so we pause for a moment and
-recall some classical facts about the cohomology of configuration
-spaces of the disc.
+We note that $\tCh_*(C_p(\Sigma_{0,1})^{\infty})$ is exactly the chain complex described by Fuchs in
+\cite{Fuchs:CohomBraidModtwo}: we recall Fuchs' computation of the cohomology of configuration
+spaces of the disc, and abbreviate $\D$ for $\mathring{Sigma}_{0,1}$ as in section \ref{sec:Preliminaries}.
 \begin{defn}
 \label{defn:symchain}
-Given a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ of $p$ into powers of 2, with multiplicities $\alpha_j$,
-the associated \emph{symmetric chain} in $\tCh_*(C_p(\Sigma_{0,1})^c)$
-is the sum of all $e^{\tup}$ where $\tup$ ranges among all tuples
-$\pa{l,(x_i)_{i\leq l}}$ such that
+Consider a partition of $p$ into powers of 2
+\[
+ p=\sum_{j\geq 0}\alpha_j2^j,
+\]
+and let $\ualpha=(\alpha_j)_{j\geq 0}$ be the sequence of multiplicities. We understand that only finitely
+many $\alpha_j$'s are strictly positive.
+
+The associated \emph{symmetric chain} in $\tCh_*(C_p(\D)^{\infty})$, denoted by $\kappa(\ualpha)$,
+is the sum of all tuples $\tup=\pa{l,(x_i)_{i\leq l}}$ such that
 \begin{itemize}
- \item $l=\sum_{j=1}^{\infty}\alpha_j$;
+ \item $l=\sum_{j\geq 1}\alpha_j$;
  \item every $x_i$ is a power of 2;
@@
 \end{itemize}
-Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
+% Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
+A symmetric chain $\kappa(\ualpha)$ is a cycle in the chain complex $\tCh_*(C_p(\D)^{\infty})$ (see \cite{Fuchs:CohomBraidModtwo}).
+We denote by $[\kappa(\ualpha)]\in\tH_{p+l}(C_p(\D)^{\infty})$ the associated homology class.
+\end{defn}
+In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\D)^{\infty})$
+is given by the collection of all classes $[\kappa(\ualpha)]$ associated to sequences $\ualpha=(\alpha_j)_{j\geq 0}$
+which satisfy the equality $p=\sum_{j\geq 0}\alpha_j2^j$.
 
-We denote by $|(\alpha_j)_{j\geq 0}|\in\tCh_{p+l}(C_p(\Sigma_{0,1})^c)$ the symmetric chain
-and by $[(\alpha_j)_{j\geq 0}]\in\tH_{p+l}(C_p(\Sigma_{0,1})^c)$ the associated homology class.
-\end{defn}
-In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\Sigma_{0,1})^c)$
-is given by the collection of all classes $[(\alpha_j)_{j\geq 0}]$ satisfying $p=\sum\alpha_j2^j$.
-
-By Poincaré-Lefschetz duality this is a basis for the cohomology $H^*(C_p(\Sigma_{0,1}))$.
-The \emph{dual} basis of $H_*(C_p(\Sigma_{0,1}))$
+By Poincaré-Lefschetz duality this corresponds to a basis for $H^*(C_p(\D))$.
+The \emph{dual} basis of $H_*(C_p(\D))$
 happens to be the basis of monomials
-$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\Sigma_{0,1}))$, that is, all monomials
+\[
+Q^{\ualpha}\epsilon\colon =\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\D)).
+\]
+This basis consists of all monomials
 of weight $p$, using
-the isomorphism \ref{eq:Cohen} in its full meaning (i.e. as an isomorphism of rings,
+the isomorphism \eqref{eq:Cohen} in its full meaning (i.e. as an isomorphism of rings,
 where $Q$ denotes the first Dyer-Lashof operation).
 
-% This notation is compatible, after interpreting equation \ref{eq:genuszero} as an isomorphism
+% This notation is compatible, after interpreting equation \eqref{eq:genuszero} as an isomorphism
 % of rings, with the results of \cite[Chap.3]{CLM}: the monomials constructed there, using
@@
 We will not need this finer result in this article, so in the following the expression
-$\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_{p-l}(C_p(\Sigma_{0,1}))$ will only denote
-the (unique) homology class whose algebraic intersection with $[(\alpha'_j)]\in\tH_{p+l}(C_p(\Sigma_{0,1})^c)$
-is $1\in\Z_2$ if and only if the equality $\alpha_j=\alpha'_j$ holds for all $j\geq 0$.
+$Q^{\ualpha}$ will only denote
+the (unique) homology class in $H_{p-l}(C_p(\D)$ such that the following holds:
+for all $\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\sum_{j\geq 0}\alpha'_j=p$, the
+algebraic intersection between $Q^{\ualpha}\epsilon\in H_{p-l}(C_p(\D))$
+and $[\kappa(\ualpha')]\in\tH_{p+l}(C_p(\D)^{\infty})$
+is $1\in\Z_2$ if and only $\ualpha=\ualpha'$.
 % i.e. we \emph{define} monomials as the dual basis of the basis of symmetric chains, and keep considering
-% \ref{eq:Cohen} only as an isomorphism of bigraded vector spaces.
+% \eqref{eq:Cohen} only as an isomorphism of bigraded vector spaces.
 
-We now go back to the filtered chain complex $\tCh_*(\cms^c)$:
-if we run the associated Leray spectral sequence,
-the $E^1-$page contains on the $p-$th column the homology of $F_p/F_{p-1}$; as we have seen this consists of
-many copies of the homology of $\tCh_*(C_p(\Sigma_{0,1}))$, each generated by symmetric chains.
+We now go back to the filtered chain complex $\tCh_*(\cms^{\infty})$.
+The $E^1$-page of the associated Leray spectral sequence contains on the $p$-th column
+the homology of $F_p/F_{p-1}$; as we have seen the homology of this filtration stratum is
+the direct sum of several copies of the homology of $\tCh_*(C_p(\D))$, one copy for each
+partition $(m-p)=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$.
 
@@
 \label{defn:gensymchain}
-For a splitting
+Consider a partition
 \[
@@
 \]
-let $l=\sum_{j=1}^{\infty}\alpha_j$. We define a chain in
-$\tCh_{m+l}(\cms)$, denoted by $|p,(\alpha_j)_j,(u_i,v_i)_{i\leq g}|$: it
-is obtained by summing all $e^{\tup}$ where $\tup$
-ranges among all tuples of the form $\pa{l,(x_i)_{i\leq l},(u_i,v_i)_{i\leq g}}$
-satisfying the three properties listed in definition \ref{defn:symchain}.
+and let $\uu=(u_1,\dots,u_g)$, $\uv=(v_1,\dots, v_g)$, $\ualpha=(\alpha_j)_{j\geq 0}$; denote also $l=\sum_{j=1}^{\infty}\alpha_j$.
+
+We define a chain $\kappa(p,\ualpha,\uu,\uv)$ in
+$\tCh_{m+l}(\cms)$: it
+is the sum of all tuples $\tup$ of the form $\pa{l,\ux,\uu,\uv}$, for varying $\ux$,
+which satisfy the three properties listed in Definition \ref{defn:symchain}.
 
 We call such a chain a
-\emph{generalised symmetric chain}, and its homology class in $\tH_{m+l}(\cms^c)$ is denoted
-by $[p,(\alpha_j),(u_i,v_i)]$
+\emph{generalised symmetric chain}; we will see in the following that it is a cycle, and we will
+denote by $[\kappa(p,\ualpha,\uu,\uv)]$ the associated homology class in $\tH_{m+l}(\cms^{\infty})$.
+
+If any of $\ualpha$, $\uu$ and $\uv$ vanishes, we omit it from the notation.
 \end{defn}
 
-We notice now that a generalised symmetric chain $|p,(\alpha_j),(u_i,v_i)|$
+A generalised symmetric chain $\kappa(p,\ualpha,\uu,\uv)$
 is not only
-a cycle when projected to its filtration quotient $F_p/F_{p-1}$, as the $E^1-$page
+a cycle when projected to its filtration quotient $F_p/F_{p-1}$, as the $E^1$-page
 of the spectral sequence tells us, but also
-in the chain complex $\tCh_*(\cms^c)$ itself. Indeed also all outer
-differentials of a generalised symmetric chain cancel out: in particular
-every left differential
-of one cell $e^{\tup}$ in the generalised symmetric chain cancels out with a right differential
-of possibly another cell $e^{\tup'}$.
+in the chain complex $\tCh_*(\cms^{\infty})$ itself.
 
-Therefore the spectral sequence collapses on its first page and
-% $H_*(\cms^c)$
+To prove this fact, first note that an inner boundary of a tuple
+$\tup$ preserves the norm; hence the fact that $\kappa(p,\ualpha,\uu,\uv)$ is a cycle
+in $F_p/F_{p-1}$ guarantees the fact that all inner boundaries of $\kappa(p,\ualpha,\uu,\uv)$
+cancel each other in $\tCh_*(\cms^{\infty})$, that is,
+\[
+\partial\kappa(p,\ualpha,\uu,\uv)\in \tCh_*(\cms^{\infty})
+\]
+is equal to the sum of all outer boundaries of the tuples involved.
+
+Note now that outer boundaries of a generalised symmetric chain also cancel out:
+the left outer boundary
+of a tuple $\tup=(l,\ux,\uu,\uv)$ in the generalised symmetric chain cancels against the right outer boundary
+of the tuple $\tup'=(l,\ux',\uu,\uv)$, with $x'_l=x_1$ and $x'_i=x_{i+1}$ for $1\leq i\leq l-1$. If all
+$x_i$ happen to be equal, then $\tup=\tup'$ and
+we are in the situation described before the proof of Lemma \ref{lem:doperatoropenmodtwo}.
+
+The spectral sequence considered above collapses on its first page and
+% $H_*(\cms^{\infty})$
 % has a $\Z_2$-basis given by generalised symmetric chains.
@@
 \label{lem:gensymchain}
-The homology $H_*(\cms^c)$ has a graded basis given by the classes $[p,(\alpha_j),(u_i,v_i)]$
+The homology $H_*(\cms^{\infty})$ has a graded basis given by the classes $[\kappa(p,\ualpha,\uu,\uv)]$
 associated to generalised symmetric chains of weight $m$.
@@
 % \end{itemize}
-% The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^c)$ has homological degree
+% The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^{\infty})$ has homological degree
 % $m+\sum_{j=1}^{\infty}\alpha_j$.
@@
 % To interpret this class
-% as a monomial in the tensor product of theorem \ref{thm:Hbms*as*ggrep} we need the following definition.
+% as a monomial in the tensor product of Theorem \ref{thm:Hbms*as*ggrep} we need the following definition.
 
-% We are almost ready to prove an isomorphism of bigraded $\Z_2-$vector spaces as in theorem \ref{thm:Hbms*as*ggrep}.
+% We are almost ready to prove an isomorphism of bigraded $\Z_2$-vector spaces as in Theorem \ref{thm:Hbms*as*ggrep}.
 
@@
  \label{defn:dualHbasis}
-We can see the $\U_i$'s and $\V_i$'s as properly embedaded $1-$manifolds in $\mrS$;
-by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^c}\simeq H^1(\S)$,
+We can see the $\U_i$'s and $\V_i$'s as properly embedded $1$-manifolds in $\mrS$;
+by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^{\infty}}\simeq H^1(\S)$,
 and in particular they form a basis
@@
 \end{defn}
-We establish a bijection between monomials in the tensor product of theorem  \ref{thm:Hbms*as*ggrep}
-and the basis of $H^*(\cms)\simeq H_*(\cms^c)$ in lemma \ref{lem:gensymchain}:
+We establish a bijection between monomials in the tensor product of Theorem \ref{thm:Hbms*as*ggrep}
+and the basis of $H^*(\cms)\simeq H_*(\cms^{\infty})$ in Lemma \ref{lem:gensymchain}:
 the class 
 \[
-[p,(u_i,v_i),(\alpha_j)]\in H_{m+\sum\alpha_j}(\cms^c)\simeq H^{m-\sum\alpha_j}(\cms)
+[\kappa(p,\ualpha,\uu,\uv)]\in H_{m+\sum\alpha_j}(\cms^{\infty})\simeq H^{m-\sum\alpha_j}(\cms)
 \]
-is associated with the monomial $\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g([\u_i]^{u_i}[\v_i]^{v_i})$.
+is associated with the monomial
+\[
+Q^{\ualpha}\epsilon\cdot\u^{\uu}\cdot\v^{\uv}\colon =\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g(\u_i^{u_i}\v_i^{v_i}),
+\]
+where $\ualpha=(\alpha_j)_{j\geq 0}$, $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$.
 
-This shows an isomorphism of bigraded $\Z_2-$vector spaces
+This shows an isomorphism of bigraded $\Z_2$-vector spaces
 \begin{equation}\label{eq:isovectorspaces}
@@
 from which we conclude that
-there exists an isomorphism as in theorem \ref{thm:Hbms*as*ggrep}
-at least \emph{as bigraded $\Z_2-$vector spaces}: the two bigraded
-vector spaces have the same dimension in all bidegrees.
+there exists an isomorphism as in Theorem \ref{thm:Hbms*as*ggrep}
+at least \emph{as bigraded $\Z_2$-vector spaces}: the two bigraded
+vector spaces have the same dimension in all bigradings.

### 2019-02-18 Introduction.tex
Implemented some corrections by cfb, created gitignore file

@@
 Forgetting the punctures gives a surjective map $\ggm\to\gg$ with kernel $\bms$,
-the $m-$th \emph{braid group} of the surface $\sg$: we obtain the Birman exact sequence (see \cite{Birman:mcgbr}):
+the $m$-th \emph{braid group} of the surface $\sg$. We obtain the Birman exact sequence (see \cite{Birman:mcgbr})
 \begin{equation}
@@
 
-The associated Leray-Serre spectral sequence $E(m)$ in $\Z_2-$homology has a second page $E(m)^2_{k,q}=H_k(\gg;H_q(\bms;\Z_2))$,
+The associated Leray-Serre spectral sequence $E(m)$ in $\Z_2$-homology has a second page $E(m)^2_{k,q}=H_k(\gg;H_q(\bms;\Z_2))$,
 and converges to $H_{k+q}(\ggm;\Z_2)$.
 
-The main theorem of this article states that this spectral sequence has already collapsed in its second page.
+The main result of this article is that this spectral sequence collapses in its second page.
 \begin{thm}
@@
 \label{eq:main}
-H_l(\ggm;\Z_2)\cong \bigoplus_{k+q=l} H_k(\gg;H_q(\bms;\Z_2)).
+H_l\pa{\ggm;\Z_2}\cong \bigoplus_{k+q=l} H_k\pa{\gg;H_q\pa{\bms;\Z_2}}.
 \end{equation}
 \end{thm}
-Thus the computation of $H_*(\ggm;\Z_2)$ reduces to the computation of the homology of $\gg$ with
-twisted coefficients in the representation $H_*(\bms;\Z_2)$. We will see that the $\ggm-$representation
-$H_*(\bms;\Z_2)$ splits as a direct sum of symmetric powers of $H_1(\sg)$ with symplectic action
-(see theorem \ref{thm:Hbms*as*ggrep} in section \ref{sec:HBraidSurf}).
+Thus the computation of $H_*\pa{\ggm;\Z_2}$ reduces to the computation of the homology of $\gg$ with
+twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\ggm$-representation
+splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action
+(see Theorem \ref{thm:Hbms*as*ggrep}).
+% in section \ref{sec:HBraidSurf}).
 
@@
 
-The group $\gg$ is significantly smaller than $\ggm$, so the
-computation of the right-hand side of equation \ref{eq:main}
-may be more tractable.
+% The group $\gg$ is significantly smaller than $\ggm$, so the
+% computation of the right-hand side of equation \eqref{eq:main}
+% may be more tractable.
 %WE DO THIS COMPUTATION FOR $g=1$.
+
+The strategy of the proof does not generalize to fields of characteristic different from 2 or to the \emph{pure}
+mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In section \ref{sec:rational}
+we describe in detail a counterexample with coefficients in $\Q$, which can be generalized both to coefficients
+in a field $\F_p$ of odd characteristic and to the pure mapping class group.

### 2019-02-18 Preliminaries.tex
Implemented some corrections by cfb, created gitignore file

@@
 In this section we recollect some classical definitions and results about braid groups and mapping class groups.
-Let $\sg$ be a compact, smooth, orientable surface of genus $g$ with one parametrised
-boundary component $\partial \sg$. Let $\sgm$ be $\sg$ with a choice of $m$ distinct
-points in the interior: these points are called \emph{punctures} and are not assumed to be ordered.
+% Let $\sg$ be a compact, smooth, orientable surface of genus $g$ with one parametrised
+% boundary component $\partial \sg$. Let $\sgm$ be $\sg$ with a choice of $m$ distinct
+% points in the interior: these points are called \emph{punctures} and are not assumed to be ordered.
 
@@
 
- Notice that we require the points of the configuration to lie in the interior of $\sg$;
- $F_m(\sg)$ is a smooth, orientable $2m-$dimensional manifold.
+ Note that we require the points of the configuration to lie in the interior of $\sg$; the space
+ $F_m(\sg)$ is a smooth, orientable $2m$-dimensional manifold.
 
- The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the points of a configuration;
+ The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the labels $1,\dots,n$ of a configuration;
  the orbit space
@@
  \]
- is called the \emph{$m-$th unordered configuration space}
- of $\sg$ and is denoted by $C_m(\sg)$; this space is also a $2m-$dimensional orientable manifold
+ is called the \emph{$m$-th unordered configuration space}
+ of $\sg$ and is denoted by $C_m(\sg)$; this space is also a $2m$-dimensional orientable manifold
  (see figure \ref{fig:unordered}).
@@
 %  
-%  We denote by $C_k(\sg)^c$ the one-point compactification of the space $C_k(\sg)$; the point
+%  We denote by $C_k(\sg)^{\infty}$ the one-point compactification of the space $C_k(\sg)$; the point
 %  at infinity is denoted by $*$ and serves as basepoint.
@@
 \begin{figure}\centering
- \includegraphics[scale=0.8]{figures/unordered.png}
+\includegraphics[scale=0.8]{figures/unordered.pdf}
 \caption{A configuration in the space $C_5(\Sigma_{1,1})$}
@@
  Let $\Diff(\sg;\partial\sg)$ be the group of diffeomorphisms of $\sg$ that fix $\partial\sg$ pointwise,
- endowed with the Whitney $C^{\infty}-$topology. We denote by $\gg=\pi_0(\Diff(\sg;\partial\sg))$ its group of connected
+ endowed with the Whitney $C^{\infty}$-topology. We denote by $\gg=\pi_0(\Diff(\sg;\partial\sg))$ its group of connected
  components, which is called the \emph{mapping class group} of $\sg$.
@@
  Similarly let $\Diff(\sgm;\partial\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
- pointwise and restrict to a permutation of the $m$ punctures, and let $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$ be its
- group of connected components, called the mapping class group of $\sgm$.
+ pointwise and restrict to a permutation of the $m$ punctures: the \emph{mapping class group} of $\sgm$ is
+ $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$.
  \end{defn}
- A classical result by Earle and Schatz (\cite{EarleSchatz}) ensures that the connected components
+ A classical result by Earle and Schatz \cite{EarleSchatz} ensures that the connected components
  of $\Diff(\sg)$ are contractible. In particular $B\Diff(\sg;\partial\sg)\simeq B\gg$ is
  a classifying space for
- the group $\gg$, i.e. a space of type $K(\gg,1)$; we call $\cF_{g,1}\to B\Diff(\sg;\partial\sg)$ the universal
- $\sg-$bundle
+ the group $\gg$, i.e. an Eilenberg-MacLane space of type $K(\gg,1)$.
+
+ We call $\cF_{g,1}\to B\Diff(\sg;\partial\sg)$ the universal
+ $\sg$-bundle
  \[
-  \cF_{g,1}\colon=\sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
+  \sg\to\cF_{g,1}=\sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
  \]
-Applying the construction of the $m-$th unordered configuration space fiberwise we obtain a bundle
+Applying the construction of the $m$-th unordered configuration space fiberwise we obtain a bundle
 $C_m(\cF_{g,1})\to B\Diff(\sg;\partial\sg)$ with fiber $C_m(\sg)$.
@@
 The space $C_m(\cF_{g,1})$ is a classifying space for the
-group $\ggm$ and the Birman exact sequence \ref{eq:Birman} is obtained by taking fundamental
+group $\ggm$ and the Birman exact sequence \eqref{eq:Birman} is obtained by taking fundamental
 groups of the aspherical spaces
@@
 
-In the whole article the genus $g$ of the surfaces that we consider is supposed to be fixed,
-unless explicitly stated otherwise, and we will abbreviate $\S=\sg$.
+In the whole article the genus $g\geq 0$ of the surfaces that we consider is supposed to be fixed,
+% unless explicitly stated otherwise,
+and we will abbreviate $\S=\sg$.
 We denote by $\D$ the open disc $\mathring{\Sigma}_{0,1}$.
@@
 It will be useful, for many constructions, to choose an embedding
-$\D\hookrightarrow\mrS$ near $\partial\S$ and to replace
+$\D\hookrightarrow\mrS$ \emph{near} $\partial\S$ and to replace
 $\Diff(\S;\partial\S)$ with its subgroup $\Diff(\S;\partial\S\cup\D)$ of diffeomorphisms
@@
 \emph{near $\partial\S$} means that there is a compact subsurface $\S'\subset\S$
-such that $\S$ is the boundary connected sum $\S'\natural\bar\D$.
+such that $\S=\S'\natural\bar\D$ is the union along a segment of $\S'$ and the closure of $\D$ in $\S$.
+
 From now on we suppose such an embedding to be fixed and we consider $\D$ as a subspace
-of $\mrS$ (see figure \ref{fig:SS'D}). In section \ref{sec:HBraidSurf}, definition \ref{defn:Tsg},
-we will specify an embedding $\D\hookrightarrow\mrS$ using a particular model for $\mrS$.
+of $\mrS$ (see Figure \ref{fig:SS'D}). In Section \ref{sec:HBraidSurf}, Definition \ref{defn:Tsg},
+we will introduce a convenient model $\T(\S)$ for the space $\mrS$, and in Section \ref{sec:Actiongg},
+Definition \ref{defn:DinTS}
+we will specify an embedding $\D\hookrightarrow\mrS$ using the model $\T(\S)$.
 
 \begin{figure}\centering
- \includegraphics[scale=0.8]{figures/SS'D.png}
+ \includegraphics[scale=0.8]{figures/SS'D.pdf}
  \caption{A splitting of $\S$ as $\S'\natural\bar\D$.}
@@
 
-The construction of definition \ref{defn:cms} can be specialised to the
+The construction in Definition \ref{defn:cms} can be specialised to the
 surfaces $\D$ and $\S'$, yielding spaces $C_m(\D)$ and $C_m(\S')$ respectively.
+We take configurations of points in the interior of the surfaces $\D$ and $\S'$ respectively
+(since $\D$ is already an open surface this remark applies in particular to $\S'$).
 
 The inclusion $\Diff(\S;\partial\S\cup\D)\subset\Diff(\S;\partial\S)$ is a homotopy
-equivalence, hence also the induced map
+equivalence of topological groups, hence also the induced map
 \[
@@
 is a homotopy equivalence. We will replace the previous construction with the following,
-equivalent ones.
+homotopy equivalent ones.
 \begin{defn}
@@
 We call $\cF_{\S,\D}\to B\Diff(\S;\partial\S\cup\D)$
-the universal $\S-$bundle
+the universal $\S$-bundle
  \[
-  \cF_{\S,\D}\colon=\S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
+  \S\to\cF_{\S,\D}\colon=\S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
  \]
-Notice that $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms, and acts trivially
+Note that $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms, and acts trivially
 on $\D$; therefore $\cF_{\S,\D}$ contains subspaces
@@
 
-Applying the construction of the $m-$th unordered configuration space fiberwise, we obtain spaces
+Applying the construction of the $m$-th unordered configuration space fiberwise, we obtain spaces
 $C_m(\cF_{\S,\D})$, $C_m(\cF_{\S'})$ and $C_m(\D)\times B\Diff(\S;\partial\S\cup\D)$,
-all fibering over $B\Diff(\S;\partial\S\cup\D)$ with fibers, respectively, $\cms$,
+all fibering over $B\Diff(\S;\partial\S\cup\D)$, respectively with fiber $\cms$,
 $C_m(\S')$ and $C_m(\D)$.
 \end{defn}
+The fiber bundle \eqref{eq:Birmanbundle} corresponding to the Birman exact sequence \eqref{eq:Birman}
+can now be replaced with the following one
+\begin{equation}\label{eq:BirmanbundleD}
+\cms\to C_m(\cF_{\S,\D})\to B\Diff(\S;\partial\S\cup\D).
+\end{equation}
 
 \begin{defn}
-For all $0\leq p\leq m$ there is a natural map
+For all $0\leq p\leq m$ there is a map
 \[
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defmu.png}
- \caption{The product of two configurations in $C_2(\D)$ and $C_4(\S)$.}
+ \includegraphics[scale=0.7]{figures/defmu.pdf}
+ \caption{The product of two configurations in $C_2(\D)$ and $C_4(\S')$.}
 \label{fig:defmu}
@@
 
-We can apply this construction to each couple of fibers of $C_m(\cF_{\S,\D})$ and $C_{m-p}(\cF_{\S'})$
-over the same point of $ B\Diff(\S;\partial\S\cup\D)$, obtaining a map $\mu^{\cF}$:
+We can apply this construction at the same time to each couple of fibers of the bundles $C_m(\cF_{\S,\D})$ and $C_{m-p}(\cF_{\S'})$
+lying over the same point of $ B\Diff(\S;\partial\S\cup\D)$.
+
+We obtain a map
 \[
@@
 \end{defn}
-The fiber bundle \ref{eq:Birmanbundle} corresponding to the Birman exact sequence \ref{eq:Birman}
-can now be replaced by the following one
-\begin{equation}\label{eq:BirmanbundleD}
-\cms\to C_m(\cF_{\S,\D})\to B\Diff(\S;\partial\S\cup\D).
-\end{equation}
-We conclude this section by recalling the structure of $H_*(C_m(\D))$.
 
+We recall now the structure of $H_*(C_m(\D))$.
 The cohomology of $C_m(\D)$ with coefficients in $\Z_2$ was first computed by Fuchs in
-\cite{Fuchs:CohomBraidModtwo}: we will recall this computation in section
-\ref{sec:HBraidSurf}, where we will make a similar computation in detail.
+\cite{Fuchs:CohomBraidModtwo}: in section
+\ref{sec:HBraidSurf} we will generalise  Fuchs' argument to surfaces with boundary of positive genus.
 
 In \cite[Chap.III]{CLM} Cohen considers the space $\coprod_{m\geq 0}C_m(\D)$ as an
-algebra over the operad of little $2-$cubes, and describes its
-$\Z_2-$homology as follows:
+algebra over the operad of little $2$-cubes, and describes its
+$\Z_2$-homology as follows:
 \begin{equation}
@@
 
-The isomorphism in equation \ref{eq:Cohen} is
-an isomorphism of bigraded rings. The left-hand side is a ring with the Pontryagin product
-coming from the action of the little $2-$cubes,
+The isomorphism in equation \eqref{eq:Cohen} is
+an isomorphism of bigraded rings. The left-hand side is a ring with the Pontryagin product,
+% coming from the action of the little $2$-cubes,
 and the right-hand side is a polynomial ring in infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$.
-The bidegree is given by the homological degree
-$*$, that we simply call the \emph{degree},
+The bigrading is given by the homological degree
+$*$, that we call the \emph{degree},
 and by the index $m$ of the connected component
 on which the homology class is supported (informally, the number of points
-involved in the construction of a homology class), that we call the \emph{weight}.
+involved in the construction of the homology class), that we call the \emph{weight}.
 
  \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/monomial.png}
+ \includegraphics[scale=0.7]{figures/monomial.pdf}
  \caption{The class in $H_5(C_9(\D))$ corresponding to the monomial $\epsilon\cdot(Q\epsilon)^2\cdot Q^2\epsilon$.}
@@
 
-In this article we will only need the isomorphism in equation \ref{eq:Cohen} to hold as
-an isomorphism of bigraded $\Z_2-$vector spaces.
+In this article we will only need the isomorphism in equation \eqref{eq:Cohen} to hold as
+an isomorphism of bigraded $\Z_2$-vector spaces.
 In particular for all choices of natural numbers $(\alpha_j)_{j\geq 0}$ with all but finitely
 many $\alpha_j=0$, we will consider the monomial $\prod_j\pa{Q^j\epsilon}^{\alpha_j}$, corresponding to
-a homology class in some bidegree $(*,m)$: all that
-we are going to need is that the set of these monomials is a bigraded basis for the
-left-hand side of equation \ref{eq:Cohen}.
+a homology class in some bigrading $(*,m)$: we will only use the fact
+that the set of these monomials is a bigraded basis for the
+left-hand side of equation \eqref{eq:Cohen}.
+
+We end this section recalling the classical notions of symmetric product and one-point compactification.
+
+\begin{defn}
+ \label{defn:SP}
+ The $m$-fold symmetric product of a topological space $\mathcal{M}$, denoted by $\SP^m(\mathcal{M})$, is
+ the quotient of $\mathcal{M}^m$ by the (non-free) action of the symmetric group $\mathfrak{S}_m$
+ which permutes the coordinates.
+\end{defn}
+We regard a point of $\SP^m(\mathcal{M})$ as a configuration of $m$ points in $\mathcal{M}$ with multiplicities.
+
+In the case of an open surface $\mrS$, it is known that $\SP^m(\mrS)$ is a non-compact manifold of dimension $2m$;
+it contains $\cms$ as an open submanifold.
+
+\begin{defn}
+ \label{defn:opc}
+For a space $\mathcal{M}$ we denote by $\mathcal{M}^{\infty}$ its one-point
+compactification; the basepoint is the point at infinity, that we denote by $\infty$.
+\end{defn}
+
+We consider in particular the one-point-compactification $\cms^{\infty}$ of $\cms$.

### 2019-02-18 Rational.tex
Implemented some corrections by cfb, created gitignore file

--- 2018-10-29 Rational.tex
@@
 \section{A rational counterexample}
-In this section we prove that a statement as in theorem \ref{thm:Hbms*as*ggrep}
-cannot hold if we consider homology with coefficients in $\Q$.
+\label{sec:rational}
+In this section we prove that a statement as in Theorem \ref{thm:Hbms*as*ggrep}
+cannot hold if we consider homology with coefficients in $\Q$. We do not know
+if the analogue of Theorem \ref{thm:main} holds in homology with coefficients in $\Q$
+or in fields of odd characteristic.
 
-We recall that $H_*(\cms;\Q)$ has been computed \emph{as a bigraded $\Q-$vector space}
-in \cite{BCM} and, more recently, in \cite{Knudson}; a description of it as a
-$\gg-$representation seems to be still missing in the literature.
+We point out that $H_*(C_m(\Sigma_{g,1});\Q)$ has been computed \emph{as a bigraded $\Q$-vector space}
+by B\"odigheimer, Cohen and Milgram in \cite{BCM}, and more recently by Knudson in \cite{Knudson}. A description of
+these homology groups as a
+$\gg$-representation seems to be still missing in the literature.
 
-In particular we prove the following fact.
+We will prove the following theorem:
 \begin{thm}
  \label{thm:counterexample}
- Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\S);\Q)$ is not a symplectic
+ Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\Sigma_{g,1});\Q)$ is not a symplectic
  representation of $\gg$.
@@
   $[a]$ with a homology class
-  in $\tH_2(\cms^c;\Q)$; here it is worth stressing that for all $m$ the manifold
-  $\cms$ is orientable: therefore we can apply Poincaré-Lefschetz duality
+  in $\tH_2(\cms^{\infty};\Q)$: note that the manifold
+  $\cms$ is orientable, hence Poincaré-Lefschetz duality holds
   also with rational coefficients;
@@
 Recall that the Torelli group $\mathcal{I}_{g,1}$ is the kernel of the surjective homomorphism
-$\gg\to Sp_{2g}(\Z)$; if an element of the Torelli group acts non-trivially on
-some class in $H_2(C_2(\S);\Q)$, then this is not a symplectic representation of $\gg$.
+$\gg\to Sp_{2g}(\Z)$ induced by the action of $\gg$ on $H_1\pa{\sg;\Z}$; if an element of the Torelli group acts non-trivially on
+some class in $H_2(C_2(\S);\Q)$, the latter cannot be a symplectic representation of $\gg$.
 
@@
 We consider the torus $a=c\times c'$ of configurations in $C_2(\S)$ having one point lying on $c$ and
-one lying on $c'$: by abuse of notation, denote also one of its fundamental cycles by $a$
-(there are two possible choices, corresponding to the two possible orientations of $a$).
+one lying on $c'$; we let $[a]\in H_1(C_2(\S);\Q)$ be the fundamental class of $a$ associated with one
+orientation of $a$.
 
-Let $\tup=(0,(u_i,v_i))$ with $u_1=2$ and all other $u_i$'s, as well as all $v_i$'s, equal to zero;
-then the map
+Let $\tup=(0,\uu,\uv)$ with $\uu=(2,0)$ and $\uv=(0,0)$.
+Then the map
 \[
-\phi_{\tup}\colon (\Delta^{\tup},\partial\Delta^{\tup})\to (C_2(\S)^c,\infty)
+\Phi_{\tup}\colon (\Delta^{\tup},\partial\Delta^{\tup})\to (C_2(\S)^{\infty},\infty)
 \]
-maps the fundamental class in $H_2(\Delta^{\tup},\partial\Delta^{\tup};\Q)$ to a homology
-class in $H_2(C_2(\S);\Q)$; if we see $e^{\tup}$ as a properly embedded, orientable submanifold
-of $C_2(\S)$, then we are just considering its fundamental class $[e^{\tup}]\in H_2(C_2(\S)^c)$.
+maps the fundamental class in $H_2(\Delta^{\tup},\partial\Delta^{\tup};\Q)$ to a
+class in $\tH_2(C_2(\S)^{\infty};\Q)$, that we call $[\kappa(0,\uu,\uv)]_{\Q}$.
+% ; if we see $e^{\tup}$ as a properly embedded, orientable submanifold
+% of $C_2(\S)$, then we are just considering its fundamental class $[e^{\tup}]\in H_2(C_2(\S)^{\infty})$.
 
-The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore $[a]\cap[e^{\tup}]=\pm 1$,
-where the signs depends on how we have chosen orientations on $a$, $e^{\tup}$ and $C_2(\S)$ itself;
-in particular $[a]\neq 0$.
+The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore the
+algebraic intersection $[a]\cap[\kappa(0,\uu,\uv)]_{\Q}$ is $\pm 1$,
+where the signs depends on how we have chosen orientations on $a$, $e^{\tup}$ and $C_2(\S)$ itself.
+In particular $[a]\neq 0$.
 
-The action of $\gg$ on isotopy classes of simple closed curves is transitive, so we can repeat
+The action of $\gg$ on isotopy classes of non-separating simple closed curves is transitive, so we can repeat
 the construction of the torus $a$ with any other copy of parallel, non-separating simple closed curves
-$c$ and $c'$, and the resulting class $[a]$ will always be non-zero in $H_2(C_2(\S);\Q)$.
+$c$ and $c'$, and the resulting class $[a]$ will always be non-trivial in $H_2(C_2(\S);\Q)$.
 
-Let now $d$ and $d'$ be disjoint, non-separating simple closed curves such that
-cutting $\S$ along $d$ and $d'$ we obtain a subsurface $\S'\simeq\Sigma_{1,2}$ of genus 1, with boundary
-components $d$ and $d'$ (here we need that the genus of $\S$ is at least 2);
-suppose moreover that $c$ is a non-separating simple closed curve in $\S'$,
+See Figure \ref{fig:rational} to visualize the following discussion.
+Let $d$ and $d'$ be disjoint, non-separating simple closed curves such that
+the following holds: if we cut $\S$ along $d$ and $d'$ we separate $\S$ into two pieces, one of
+which is a subsurface $\S'\simeq\Sigma_{1,2}$ of genus 1, with boundary
+components $d$ and $d'$. Here we use that the genus of $\S$ is at least 2.
+
+Suppose now that $c$ is a non-separating simple closed curve in $\S'$,
 and let $c',c''$ be two parallel copies of $c$ in $\Sigma_{1,2}$, one on each side of a small tubular
@@
 \end{itemize}
-See picture PICTURE.
+
+\begin{figure}\centering
+ \includegraphics[scale=1.0]{figures/rational.pdf}
+ \caption{The curves $d,d',c,c',c'',\gamma,\gamma''$ and the subsurfaces $\S',\S''$.}
+\label{fig:rational}
+\end{figure}
 
@@
 
-Let again $a=[c\times c']$ and let $b=[c\times d]$: then the class $b'=[c\times d']$
-is equal to $b+2a\in H_2(C_2(\S);\Q)$.
+Let again $[a]=[c\times c']$ and let $[b]=[c\times d]$: then the class $[b']=[c\times d']$
+is equal to $[b]+2[a]\in H_2(C_2(\S);\Q)$.
 
 Consider now an element of the Torelli group that fixes $c$ and maps $d$ to $d'$ preserving
-the orientation, for example the bounding pair represented in picture PICTURE:
+the orientation, for example the bounding pair $D_{\gamma}\circ D_{\gamma'}^{-1}$, where
+$\gamma$ and $\gamma'$ are represented in Figure \ref{fig:rational}:
 then the class $b$ is mapped to $b'=b+2a\neq b$.
 \end{proof}
+
+The previous proof works verbatim if we replace $\Q$ by any field of odd characteristic. Moreover
+all the arguments in the previous proof can be adapted to show that $H_*(F_2(\S);\mathbb{F})$ is
+not a symplectic representation of $\gg$ (see Definition \ref{defn:cms}), where
+$\mathbb{F}$ is \emph{any} field, including $\Z_2$.
+
+Therefore, at least for orientable surfaces $\sg$ of genus $g\geq 2$,
+being a symplectic representation of the mapping class group seems to be a peculiarity
+of the homology of \emph{unordered} configuration space and of the characteristic 2.

### 2019-02-18 Ssequence.tex
Implemented some corrections by cfb, created gitignore file

@@
-\section{Proof of theorem \ref{thm:main}}
-We will prove theorem \ref{thm:main} by induction on $m$.
-The case $m=0$ is trivial, so suppose $m\geq 1$.
+\section{Proof of Theorem \ref{thm:main}}
+We will prove Theorem \ref{thm:main} by induction on $m$.
+The case $m=0$ is trivial.
 
 For all $m\geq 0$ we let $E(m)$ be the Leray-Serre spectral sequence associated with
-the bundle \ref{eq:BirmanbundleD}: its second page has the form
+the bundle \eqref{eq:BirmanbundleD}: its second page has the form
 \[
- E(m)^2_{k,q}=H_k(B\Diff(\S;\partial\S\cup\D);H_q(\cms)).
+ E(m)^2_{k,q}=H_k(B\Diff(\S;\partial\S\cup\D);H_q(\cms))=H_k(\gg;H_q(\cms)).
 \]
-From theorem \ref{thm:Hbms*as*ggrep} we know that this spectral sequence is concentrated
+From Theorem \ref{thm:Hbms*as*ggrep} we know that this spectral sequence is concentrated
 on the rows $q=0,\dots, m$.
@@
 We want to prove the vanishing of all differentials appearing in the pages
-$E(m)^r$ with $r\geq 2$; the $r-$th differential takes the form
+$E(m)^r$ with $r\geq 2$; the $r$-th differential takes the form
 \[
@@
 
-Fix now $q=m-l<m$, in particular $l\geq 1$; by theorem \ref{thm:Hbms*as*ggrep}, and
-in particular by lemma \ref{lem:oplussplitting}, we have a splitting of $H_k(\gg;H_{m-l}(\cms))$ as
+Fix now $q=m-l<m$, in particular $l\geq 1$; by Theorem \ref{thm:Hbms*as*ggrep}, and
+in particular by Lemma \ref{lem:oplussplitting}, we have a splitting of $H_k(\gg;H_{m-l}(\cms))$ as
 \begin{equation}
@@
 
-Consider the map $\mu^{\cF}$ from definition \ref{defn:universalSbundle} as a map of
+Consider the map $\mu^{\cF}$ from Definition \ref{defn:universalSbundle} as a map of
 bundles over the space $B\Diff(\S;\partial\S\cup\D)$:
@@
 \]
-Notice that the first bundle $C_p(\D)\times C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$
+Note that the first bundle $C_p(\D)\times C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$
 is the product of the space $C_p(\D)$ with the bundle $C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$;
@@
 \]
-that in the second page, on the $(m-l)-$th row, restricts to
-the inclusion of one of the direct summands in equation \ref{eq:Hksplitting}:
+that in the second page, on the $(m-l)$-th row and $k$-th column, restricts to
+the inclusion of one of the direct summands in equation \eqref{eq:Hksplitting}:
 \[
@@
 
-One can generalise theorem \ref{thm:main} to orientable or non-orientable
-surfaces with at least one boundary component, following the generalisation
-of theorem \ref{thm:Hbms*as*ggrep} discussed at the end of section \ref{sec:Actiongg}.
+One can generalise Theorem \ref{thm:main} to orientable or non-orientable
+surfaces with non-empty boundary, following the generalisation
+of Theorem \ref{thm:Hbms*as*ggrep} discussed at the end of section \ref{sec:Actiongg}.

### 2019-02-28 Actiongg.tex
Drawn and redrawn figures

@@
 \label{sec:Actiongg}
-We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of Theorem
-\ref{thm:Hbms*as*ggrep} in bigradings $(*,m)$ with $*=m$, and then extend it to other bigradings.
+We now turn back to \emph{homology} of $\cms$. In the first subsection we describe geometrically some
+homology classes, in order to give some intuition for the following subsections.
+In the second subsection we prove Theorem
+\ref{thm:Hbms*as*ggrep} in bigradings $(*,m)$ with $*=m$.
+In the third subsection we extend the proof to all other bigradings.
+
+\input{simpleexample.tex}
 
@@
 
-We call \emph{multicurve} an element of $\cC^m$; by abuse of notation
+An element of $\cC^m$ is called \emph{multicurve}; by abuse of notation the $m$-tuple
 $(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
@@
  \label{defn:cmsD}
- Recall Definition \ref{defn:symprod}.
+ Recall Definition \ref{defn:SP}.
  The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
@@
 of the $m$-fold torus in $H_m(\cmsD)$. The result does not change
-if we substitute $(c_1,\dots,c_m)$ with another isotopic $m$-tuple of curvesin the same multicurve.
+if we substitute $(c_1,\dots,c_m)$ with another isotopic $m$-tuple of curves in the same multicurve.
 
@@
 We can therefore associate to $[\kappa(\uu,\uv)]$ a linear functional
-on $H_m(\cms)$, which is the intersection product
+on $H_m(\cms)$. This is the algebraic intersection product with the cell $e^{\tup}$, seen as a proper submanifold of $\cms$:
+we denote it by
 \[
- (\cdot\cap e^{\tup})\colon H_m(\cms)\to\Z_2.
+ \cdot\cap e^{\tup}\colon H_m(\cms)\to\Z_2.
 \]

### 2019-02-28 simpleexample.tex
Drawn and redrawn figures

\subsection{Geometric examples of homology classes}
In this subsection we consider the surface $\Sigma_{2,1}$, which we abbreviate with $\S$, and
construct some homology classes in $H_2(C_2(\S))$: our aim is to get a first understanding
of why this homology group is isomorphic to $\Sym_2(\H)=\H^{\otimes 2}/\mathfrak{S}_2$.

### 2019-07-12 Actiongg.tex
Started the comparison with BoT, reformulated two main theorems

@@
 The proof of Theorem \ref{thm:Hbms*as*ggrep} can be easily generalised to surfaces
-with more than one boundary component. Let $\Sigma_{g,n}$ be a surface of genus
-$g$ with $n\geq 1$ parametrised boundary components and let $\Gamma_{g,n}$ be
+with more than one boundary curve. Let $\Sigma_{g,n}$ be a surface of genus
+$g$ with $n\geq 1$ parametrised boundary curves and let $\Gamma_{g,n}$ be
 the group of connected components of the topological group

### 2019-07-12 Comparison.tex
Started the comparison with BoT, reformulated two main theorems

\section{Comparison with the work of B\"{o}digheimer and Tillmann}
\label{sec:comparison}
In this section we will compare Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep}
with the results in \cite{BoT}, in particular the following two theorems.
\begin{thm}[Corollary 1.2 in \cite{BoT}]
Let $\F$ be a field, and let $m\geq 0$, $g\geq 0$. Then the following graded $\F$-vector spaces are isomorphic in degrees $*\leq \frac 23 g-\frac 23$:
\[
 H_*\pa{\ggm;\F}\cong H_*\pa{\gg;\F}\otimes_{\F} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]};
\]
here each $x_1,\dots,x_m$ have degree 2 and the symmetric group $\mathfrak{S}_m$ acts on the polynomial
ring $\F[x_1,\dots,x_m]$ by permuting the variables.
\end{thm}
We point out that the range of degrees $*\leq \frac 23 g-\frac 23$ is the stable range for the homology $H_*(\gg;\F)$
of the mapping class group, see \cite{Harer} for the original stability theorem and \cite{Boldsen, ORW:resolutions_homstab} for the improved
stability range.
\begin{thm}[Theorem 1.3 in \cite{BoT}]
For all $m\geq 0$ the map $\mu^{\cF}\colon C_1(\D)\times C_m(\cF_{\S'})\to C_{m+1}(\cF_{\S,\D})$ from Definition \ref{defn:muF}
induces a split-injective map in homology
\[
 \mu^{\cF}_*\colon H_*\pa{C_m(\cF_{\S'})}\cong H_0\pa{C_1(\D)}\otimes_{\F} H_*\pa{C_m(\cF_{\S'})}\hookrightarrow H_*\pa{C_{m+1}(\cF_{\S,\D})};
\]
in other words, the inclusion of groups $\ggm\hookrightarrow\gg^{m+1}$ given by adding a puncture near the boundary induces a split
injective map $H_*(\ggm;\F)\hookrightarrow H_*(\gg^{m+1};\F)$.
\end{thm}
\subsection{Reformulation of our results}
\label{subsec:reformulation}
In this subsection we reformulate Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep} in a convenient way.
\begin{defn}
 \label{defn:Cbullet}
 We introduce some abbreviations for the following disjoint unions:
 \[
 C_{\bullet}(\D)=\coprod_{m\geq 0}C_m(\D);
 \]
 \[
 C_{\bullet}(\S)=\coprod_{m\geq 0}C_m(\S); 
 \]
 \[
 C_{\bullet}(\cF_{\S,\D})=\coprod_{m\geq 0}C_m(\cF_{\S,\D}).
 \]
\end{defn}
Then $C_{\bullet}(\D)$ is a (homotopy associative) topological monoid, with associated
Pontryiagin ring $H_*\pa{C_{\bullet}(\D)}\cong\Z_2[Q^j\epsilon|j\geq 0]$ (see Equation \ref{eq:Cohen}).
This is a bigraded ring, where the two gradings are the homological degree and the weight.

The maps $\mu$ from Definition \ref{defn:muF} make $C_{\bullet}(\S)$ into a (homotopy associative)
module over the monoid $C_{\bullet}(\D)$; correspondingly $H_*\pa{C_{\bullet}(\S)}$ is a bigraded
module over $H_*\pa{C_{\bullet}(\D)}$. The structure of this module is described by the following
reformulation of Theorem \ref{thm:Hbms*as*ggrep}, see in particular the proof of Lemma \ref{lem:oplussplitting}.

\begin{thm}
 \label{thm:firstreformulation}
 There is an isomorphism of $\gg$-representations in (bigraded) modules over the ring $H_*\pa{C_{\bullet}(\D)}$
 \[
  H_*\pa{C_{\bullet}(\S)}\cong   H_*\pa{C_{\bullet}(\D)}\otimes_{\Z_2}\Sym_{\bullet}(\H).
 \]
\end{thm}

Similarly, Theorem \ref{thm:main} can be reformulated as follows, considering at the same time all values of $m\geq 0$.
\begin{thm}
 \label{thm:mainreformulation}
 There is an isomorphism of (bigraded) modules over $H_*\pa{C_{\bullet}(\D)}$
 \[
 H_*\pa{C_{\bullet}(\cF_{\S,\D})}\cong   H_*\pa{C_{\bullet}(\D)}\otimes_{\Z_2}H_*\pa{\gg;\Sym_{\bullet}(\H)}.
 \] 
\end{thm}
The proof follows from the arguments used in Section \ref{sec:sseq}: we have actually shown that the direct
sum of the spectral sequences $\bigoplus_{m\geq 0} E(m)$ is itself a free module over $H_*\pa{C_{\bullet}(\D)}$
on the second page, with the same description as above.

\subsection{Comparison between results}
\label{subsec:comparison}

### 2019-07-12 HBraidSurf.tex
Drawn and redrawn figures
Started the comparison with BoT, reformulated two main theorems

@@
  \[
-  \bigoplus_{m\geq 0} H_*\pa{\cms}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
+  \bigoplus_{m\geq 0} H_*\pa{\cms}\cong \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
  \]
@@
 of $C_m(\D)$.
-As already mentioned in section \ref{sec:Preliminaries}, our computation recovers a known result, but it has the advantage of
+As already mentioned in Section \ref{sec:Preliminaries}, our computation recovers a known result, but it has the advantage of
 being quite elementary and of providing a part of
@@
  \includegraphics[scale=0.7]{figures/defTS.pdf}
- \caption{The space $\T(\S)$ for $\S=\Sigma_{3,1}$.}
+ \caption{The space $Q(\S)$ for $\S=\Sigma_{3,1}$. The black pentagons represent the point $\cP_{even}$, whereas the black
+ squares represent $\cP_{odd}$.}
 \label{fig:defTS}
@@
 \cite{Fuchs:CohomBraidModtwo}: we recall Fuchs' computation of the cohomology of configuration
-spaces of the disc, and abbreviate $\D$ for $\mathring{Sigma}_{0,1}$ as in section \ref{sec:Preliminaries}.
+spaces of the disc, and abbreviate $\D$ for $\mathring{Sigma}_{0,1}$ as in Section \ref{sec:Preliminaries}.
 \begin{defn}

### 2019-07-12 Introduction.tex
Started the comparison with BoT, reformulated two main theorems

@@
 \section{Introduction}
-Let $\sg$ be a smooth orientable surface of genus $g$ with one boundary component $\partial\sg$, and let $\sgm$ be $\sg$ with
+Let $\sg$ be a smooth orientable surface of genus $g$ with one boundary curve $\partial\sg$, and let $\sgm$ be $\sg$ with
 a choice of $m$ distinct points in the interior, called \emph{punctures}.
@@
 \label{thm:main}
-For all $l\geq 0$ there is an isomorphism
+For all $l\geq 0$ there is an isomorphism of vector spaces
 \begin{equation}
@@
 Thus the computation of $H_*\pa{\ggm;\Z_2}$ reduces to the computation of the homology of $\gg$ with
-twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\ggm$-representation
-splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action
-(see Theorem \ref{thm:Hbms*as*ggrep}).
+twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\gg$-representation
+splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action: this is done in
+Theorem \ref{thm:Hbms*as*ggrep}, which together with Theorem \ref{thm:main} is the main result of the article.
 % in section \ref{sec:HBraidSurf}).
@@
 in a field $\F_p$ of odd characteristic and to the pure mapping class group.
+
+I would like to thank my PhD advisor Carl-Friedrich B\"odigheimer for his precious suggestions and his continuous encouragement
+during the preparation of this work.

### 2019-07-12 Preliminaries.tex
Started the comparison with BoT, reformulated two main theorems

@@
 From now on we suppose such an embedding to be fixed and we consider $\D$ as a subspace
-of $\mrS$ (see Figure \ref{fig:SS'D}). In Section \ref{sec:HBraidSurf}, Definition \ref{defn:Tsg},
+of $\mrS$ (see Figure \ref{fig:SS'D}). In Section \ref{sec:HBraidSurf}, Definition \ref{defn:TS},
 we will introduce a convenient model $\T(\S)$ for the space $\mrS$, and in Section \ref{sec:Actiongg},
@@
 \begin{defn}
+\label{defn:muF}
 For all $0\leq p\leq m$ there is a map
@@
  \label{eq:Cohen}
-H_*\pa{\coprod_{m\geq 0}C_m(\D)}\simeq \Z_2\left[Q^j\epsilon \, |\, j\geq 0\right].
+H_*\pa{\coprod_{m\geq 0}C_m(\D)}\cong \Z_2\left[Q^j\epsilon \, |\, j\geq 0\right].
 \end{equation}

### 2019-07-12 Rational.tex
Started the comparison with BoT, reformulated two main theorems

@@
 the following holds: if we cut $\S$ along $d$ and $d'$ we separate $\S$ into two pieces, one of
-which is a subsurface $\S'\simeq\Sigma_{1,2}$ of genus 1, with boundary
-components $d$ and $d'$. Here we use that the genus of $\S$ is at least 2.
+which is a subsurface $\tilde{\S}\simeq\Sigma_{1,2}$ of genus 1, with boundary
+curves $d$ and $d'$. Here we use that the genus of $\S$ is at least 2.
 
-Suppose now that $c$ is a non-separating simple closed curve in $\S'$,
+Suppose now that $c$ is a non-separating simple closed curve in $\tilde{\S}$,
 and let $c',c''$ be two parallel copies of $c$ in $\Sigma_{1,2}$, one on each side of a small tubular
 neighborhood of $c$.
-Then $d,d',c',c''$ are the boundary of a subsurface $\S''\simeq\Sigma_{0,4}\subset\S'$.
+Then $d,d',c',c''$ are the boundary of a subsurface $\check{\S}\simeq\Sigma_{0,4}\subset\tilde{\S}$.
 Orient all curves $d,d',c,c',c''$ in such a way that the following equalities hold:
 \begin{itemize}
- \item $[d]=[d']\in H_1(\S;\Q)$, as witnessed by the subsurface $\S'$;
+ \item $[d]=[d']\in H_1(\S;\Q)$, as witnessed by the subsurface $\tilde{\S}$;
  \item $[c]=[c']=[c'']\in H_1(\S;\Q)$;
- \item $[d]-[d']+[c']-[c'']=0\in H_1(\S\setminus c;\Q)$, as witnessed by the subsurface $\S''$.
+ \item $[d]-[d']+[c']-[c'']=0\in H_1(\S\setminus c;\Q)$, as witnessed by the subsurface $\check{\S}$.
 \end{itemize}
@@
  \includegraphics[scale=1.0]{figures/rational.pdf}
- \caption{The curves $d,d',c,c',c'',\gamma,\gamma''$ and the subsurfaces $\S',\S''$.}
+ \caption{The curves $d,d',c,c',c'',\gamma,\gamma''$ and the subsurfaces $\tilde{\S},\check{\S}$.}
 \label{fig:rational}
@@
 \]
-holds, as witnessed by the homology $c\times \S''\subset C_2(\S)$.
+holds, as witnessed by the homology $c\times \check{\S}\subset C_2(\S)$.
 

### 2019-07-12 simpleexample.tex
Started the comparison with BoT, reformulated two main theorems

@@
 \subsection{Geometric examples of homology classes}
-In this subsection we consider the surface $\Sigma_{2,1}$, which we abbreviate with $\S$, and
-construct some homology classes in $H_2(C_2(\S))$: our aim is to get a first understanding
+% \begin{flushright}
+%  \textit{An meinen Doktorvater Herr B\"{o}digheimer}
+% \end{flushright}
+
+In this subsection we consider the case $g=2$, hence $\S$ denotes the surface $\Sigma_{2,1}$.
+We construct some homology classes in $H_2(C_2(\S))$: our aim is to get a first understanding
 of why this homology group is isomorphic to $\Sym_2(\H)=\H^{\otimes 2}/\mathfrak{S}_2$.
+
+In the following $c$ and $d$ will always denote two simple closed curves on $\S$, with corresponding homology classes $[c],[d]\in\H$.
+\begin{ex}
+ \label{ex:one}
+Suppose that $c$ and $d$ are as in Figure \ref{fig:exampleone}: $c$ and $d$ are disjoint and non-separating.
+We can consider inside $C_2(\S)$
+the torus $c\times d$ of configurations in which one of the two points runs along $c$, while the other runs
+along $d$. We associate to the fundamental class $\tau_1\in H_2(C_2(\S))$ of this torus the tensor product $[c]\otimes[d]\in\H^{\otimes 2}$.
+
+Since our configurations are unordered and since we work with coefficients in $\Z_2$, the same
+class $\tau_1$ can be obtained as a tensor product $[d]\otimes[c]$, i.e. exchanging the order of the two
+curves: we can neglect the sign $-1$ that this operation would generate. We represent $\tau_1$ as $[c]\cdot[d]\in\Sym_2(\H)$.
+\end{ex}
+\begin{figure}[h]\centering
+ \includegraphics[scale=1.0]{figures/exampleone.pdf}
+ \caption{}
+\label{fig:exampleone}
+\end{figure}
+
+\begin{ex}
+ \label{ex:two}
+Suppose that $c$ and $d$ are as in Figure \ref{fig:exampletwo}:
+$c$ and $d$ are disjoint, and
+ $c$ bounds a subsurface $\check{\S}$ of $\S$ (the shaded region), such that $d$ is not contained in $\check{\S}$.
+
+ Then the homology class $\tau_2=[c]\cdot[d]$ vanishes, because the torus $c\times d$ is the boundary in $C_2(\S)$
+ of the 3-manifold $\check{\S}\times d$ containing all configurations in which one point lies on $\check{\S}$ and
+ the other on $d$. This is consistent with the representation $\tau_2=[c]\cdot[d]$, because $[c]=0\in \H$.
+\end{ex}
+\begin{figure}[h]\centering
+ \includegraphics[scale=1.0]{figures/exampletwo.pdf}
+ \caption{}
+\label{fig:exampletwo}
+\end{figure}
+
+\begin{ex}
+ \label{ex:three}
+ Suppose that $c$ and $d$ are as in Figure \ref{fig:examplethree}: $c$ and $d$ are non-separating and 
+intersect transversely in one point $P\in\S$. In this case the torus $c\times d$ is a subspace
+of $\SP^2(\S)$, but $c\times d$ is not contained in $C_2(\S)$ as in Example \ref{ex:one}.
+
+To solve this problem, we consider a small open neighborhood $P\in \U\subset\S$, and we remove from the torus $c\times d$
+all configurations in which both points lie in $\U$: these removed configurations form an open disc inside the torus $c\times d$.
+
+We obtain an embedding $\Sigma_{1,1}\hookrightarrow C_2(\S)$;
+the boundary $\partial\Sigma_{1,1}$,
+seen as a curve in $C_2(\S)$, is homotopic to a curve $\gamma$ in which one point spins $360^{\circ}$ around the other.
+We note that the curve $\gamma\subset C_2(\S)$ is homotopic to a double covering of a curve $\gamma'\subset C_2(\S)$,
+in which the two points exchange their positions after spinning $180^{\circ}$ around each other. All curves
+$\partial\Sigma_{1,1}$, $\gamma$, $\gamma'$ and the homotopies relating them
+are supported on the closure of $C_2(\U)$ in $C_2(\S)$.
+
+We can therefore find a map from a M\"{o}bius band $\M$ to the closure of $C_2(\U)$ in $C_2(\S)$,
+such that the images of the curves $\partial\M$ and $\partial\Sigma_{1,1}$ in $C_2(\S)$ coincide.
+The union $\Sigma_{1,1}\cup_{\partial}\M$ along the boundary is then a closed
+non-orientable surface of genus $3$, i.e. the connected sum of a torus and a projective plane.
+
+The surface $\Sigma_{1,1}\cup_{\partial}\M$ is equipped with a map to $C_2(\S)$, hence its fundamental class with coefficients
+in $\Z_2$ yields a homology class $\tau_3\in H_2(C_2(\S))$; thus we have managed to adapt
+the construction from Example \ref{ex:one} to the case of two intersecting curves. We represent $\tau_3$
+as $[c]\cdot[d]\in\Sym_2(\H)$.
+\end{ex}
+\begin{figure}[h]\centering
+ \includegraphics[scale=1.0]{figures/examplethree.pdf}
+ \caption{}
+\label{fig:examplethree}
+\end{figure}
+
+\begin{ex}
+ \label{ex:four}
+ Suppose that $c$ and $d$ are as in Figure \ref{fig:examplefour}: $c$ and $d$ are disjoint, and
+ $c$ bounds a subsurface $\check{\S}$ of $\S$ (the shaded region), such that $d$ is contained in $\check{\S}$.
+
+ The torus $c\times d$ is contained in $C_2(\S)$, but it is not obvious, at first glance, why its corresponding homology
+ class $\tau_4\in H_2(C_2(\S))$ should vanish: the argument used in Example \ref{ex:two} does not work immediately, because the
+ 3-manifold $\check{\S}\times d$ is only contained in $\SP^2(\S)$ and not in $C_2(\S)$.
+
+ We try to modify this 3-manifold in the same spirit of Example \ref{ex:three}. Let $\U\tilde{\times} d\subset \check{\S}\times d$
+ be an open tubular neighborhood of $d\times d$: note that this tubular neighborhood is a trivial bundle over
+ $d$ with fiber an open disc $\U$. The notation $\tilde{\times}$ means that abstractly we are dealing with a product $\U\times d$,
+ but not geometrically: fibers over different points of $d$ are naturally identified with different discs in $\check{\S}$.
+
+ We consider the 3-manifold $\check{\S}\times d\setminus \U\tilde{\times} d$: its boundary is the disjoint union
+ of the torus $c\times d$ and the torus $\partial \pa{\U\tilde{\times} d}$. The latter torus contains configurations of
+ two points in $\S$, one of which spins around $d$, whereas the other is a \emph{satellite} of the first and spins
+ around it $360^{\circ}$.
+
+ We regard $\partial \U\tilde{\times} d$ as a trivial bundle $\gamma\tilde{\times} d$ over $d$, with fiber the curve $\gamma$
+ considered in Example \ref{ex:three}. After a homotopy the latter torus
+ becomes a double covering of a torus $\gamma'\tilde{\times} d$: this is a trivial bundle over $d$ with fiber the curve $\gamma'$;
+ we can see $\gamma'\tilde{\times} d\subset C_2(\S)$ as a torus of configurations in which the two points
+ exchange their positions spinning $180^{\circ}$ around each other, while their barycenter spins around $d$.
+
+ We can fill the second boundary of the $3$-manifold $\check{\S}\times d\setminus \U\tilde{\times} d$
+ with the product $\M\times d$: the output is a non-orientable $3$-manifold with boundary $c\times d$, witnessing
+ that $\tau_4=0$.
+
+ This is consistent with the representation $\tau_4=[c]\cdot[d]$, because $[c]=0\in \H$.
+\end{ex}
+
+\begin{figure}[h]\centering
+ \includegraphics[scale=0.8]{figures/examplefour.pdf}
+ \caption{}
+\label{fig:examplefour}
+\end{figure}
+
+\begin{ex}
+ \label{ex:five}
+ Suppose that $c$ and $d$ are as in Figure \ref{fig:examplefive}:
+ $c$ bounds a subsurface $\check{\S}$, and $d$ is cut by $c$ into two arcs, one of which, denoted by $e$, lies
+ in $\check{\S}$.
+
+ In this example the complications from Examples \ref{ex:three} and \ref{ex:four} arise at the same time.
+
+ To define the class $\tau_5\in H_2(C_2(\S))$, we start with the torus $c\times d$ and we perform two surgeries
+ with two different M\"{o}bius bands to solve the two intersections of $c$ and $d$: the homology
+ class that we obtain is represented by a non-orientable surface of genus 4, i.e. the connected
+ sum of a torus and 2 projective planes.
+
+ To show that $\tau_5$ vanishes, we start with the 3-manifold with boundary $\check{\S}\times d\subset\SP^2(\S)$ and we perform
+ a surgery. We identify with $e$ the arc
+ \[
+  \check{\S}\times d\cap\pa{\SP^2(\S)\setminus C_2(\S)};
+ \]
+ this is a properly embedded arc in the 3-manifold with boundary $\check{\S}\times d$, and a tubular neighborhood
+ of it, after a suitable homotopy, can be identified with a trivial $\U$-bundle $\U\tilde{\times} e$. We remove
+ this solid cylinder from the 3-manifold and glue a trivial $\M$-bundle $\M\tilde{\times} e$, by applying an argument
+ similar as the one in Examples \ref{ex:three} and \ref{ex:four}.
+
+ We obtain a 3-manifold with boundary endowed with a map to $C_2(\S)$; the boundary of this 3-manifold
+ is precisely the surface used to represent $\tau_5$: therefore $\tau_5=0\in H_2(C_2(\S))$,
+ and this is consistent with the representation $\tau_5=[c]\cdot[d]$, because $[c]=0\in \H$.
+\end{ex}
+
+\begin{figure}[h]\centering
+ \includegraphics[scale=1]{figures/examplefive.pdf}
+ \caption{}
+\label{fig:examplefive}
+\end{figure}
+
+If we consider more than 2 points, the examples become more and more complicated, and proving the theorem with this geometric
+approach seems rather difficult.

### 2019-07-12 splitting.tex
Implemented some corrections by cfb, created gitignore file
Started the comparison with BoT, reformulated two main theorems

--- 2018-10-19 MAIN.tex
@@
 \usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
+\usepackage{enumitem}
 \usepackage[hyphens]{url}
@@
 \usepackage{hyperref}
+\usepackage{tikz-cd}
+
+\setlength\parindent{0pt}
+
+\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}
+% \setlength{\parindent}{6pt}
 
@@
     \newtheorem{conj}[thm]{Conjecture}
+    \newtheorem{ex}{Example}
 
@@
 
-%\keywords{Topological complexity, aspherical spaces, Lusternik-Schnirelmann category, cohomological dimension, topological robotics, braid groups}
-%\subjclass[2010]{55M99, 55P20 (Primary); 55M30, 20J06, 68T40 (Secondary).}
+\keywords{Mapping class group, braid group, symplectic coefficients.}
+\subjclass[2010]{20F36, %Braid groups; Artin groups
+55N25, %Homology with local coefficients, equivariant cohomology
+55R20, %Spectral sequences and homology of fiber spaces 
+55R35, %Classifying spaces of groups and $H$-spaces
+55R40, %Homology of classifying spaces, characteristic classes
+55R80, %Discriminantal varieties, configuration spaces
+55T10, %Serre spectral sequences
+}
 
 \begin{abstract}
-Let $\ggm$ be the mapping class group of an orientable surface $\sgm$ of genus $g$ with one parametrised
-boundary component and $m$ permutable punctures, and let $\bms$ be the braid group on $m$ strands
-of the surface $\sg$. We prove that $H_*(\ggm;\Z_2)\cong H_*(\gg;H_*(\bms;\Z_2))$.
-We compute (?) the right hand side explicitly for $g=1$.
+Let $\Gamma_{g,1}^m$ be the mapping class group of the orientable surface $\Sigma_{g,1}^m$ of genus $g$ with one parametrised
+boundary curve and $m$ permutable punctures; when $m=0$ we omit it from the notation.
+Let $\beta_{m}(\Sigma_{g,1})$ be the braid group on $m$ strands
+of the surface $\Sigma_{g,1}$.
+
+\noindent We prove that $H_*(\Gamma_{g,1}^m;\mathbb{Z}_2)\cong H_*(\Gamma_{g,1};H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2))$. The main ingredient
+is the computation of $H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2)$ as a symplectic representation of $\Gamma_{g,1}$.
+%We compute (?) the right hand side explicitly for $g=1$.
+
+% 
+% Let $\Gamma^m_{g,1}$ be the mapping class group of an orientable surface $\Sigma^m_{g,1}$ of genus $g$ with one parametrised
+% boundary curve and $m$ permutable punctures, and let $\beta_m(\Sigma_{g,1})$ be the braid group on $m$ strands
+% of the surface $\Sigma_{g,1}$ of genus $g$ with one boundary curve.
+% 
+% We prove that $H_*(\Gamma^m_{g,1};\mathbb{Z}_2)\cong H_*(\Gamma_{g,1};H_*(\beta_m(\Sigma_{g,1});\mathbb{Z}_2))$. The main ingredient
+% is the computation of $H_*(\beta_m(\Sigma_{g,1});\mathbb{Z}_2)$ as a symplectic representation of $\Gamma_{g,1}$.
 \end{abstract}
@@
 \maketitle
+THIS VERSION IS NOT ANYMORE UP TO DATE! SEE ARXIVE VERSION
 
@@
 
-\input{ThmModTwo.tex}
+\input{Actiongg.tex}
 
-\input{Sink.tex}
+\input{Ssequence.tex}
 
-\input{ThmQ.tex}
+\input{Comparison.tex}
 
-\input{GenusOne.tex}
-
-\input{Conclusions.tex}
+% \input{Rational.tex}
 

### 2020-05-01 referee report/Bianchi_answer_referee.tex
Final version ready to appear in JTop

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[a4article]{geometry}

\newcommand{\N}{\mathbb{N}}    
\newcommand{\C}{\mathbb{C}}
% \newcommand{\cC}{\mathcal{C}}
% \newcommand{\D}{\mathbb{D}}
% \newcommand{\F}{\mathbb{F}}
% \newcommand{\cF}{\mathcal{F}}
% \newcommand{\cK}{\mathcal{K}}
% \newcommand{\ff}{\mathfrak{f}}
\newcommand{\fM}{\mathfrak{M}}
% \newcommand{\Q}{\mathbb{Q}}
% \newcommand{\R}{\mathbb{R}}
% \renewcommand{\H}{\mathcal{H}}
% \newcommand{\fH}{\mathfrak{H}}
% \newcommand{\V}{\mathcal{V}}
% \newcommand{\Z}{\mathbb{Z}}
% \newcommand{\cS}{\mathcal{S}}
% \newcommand{\bS}{\mathbb{S}}
\newcommand{\fS}{\mathfrak{S}}
% \newcommand{\T}{\mathcal{T}}

% \newcommand{\pa}[1]{\left(#1\right)}
% \newcommand{\abs}[1]{\left|#1\right|}
% \newcommand{\set}[1]{\left\{#1\right\}}

\renewcommand{\phi}{\varphi}

\hyphenation{sta-bi-li-ty}

\begin{document}
\title{
Answer to the referee report for ``Splitting of the homology of the
punctured mapping class group''
}
\author{Andrea Bianchi}
\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}
\email{bianchi@math.uni-bonn.de}
\date{\today}
\maketitle

First, I would like to thank the referee for the meticulous work he/she made
in reviewing my article.

I have accepted almost all corrections and suggested improvements, but in the following cases
I have followed a different way than the one proposed by the referee; I will describe briefly
the reasons for my choice, and I remain open for further discussion.
\begin{itemize}
 \item The referee writes \emph{p.3, line 2: The condition that $g\geq 2$ is needed for $\mathrm{Diff}(\Sigma_{g,1})$ to be contractible.}\\
 Since we are working with surfaces with non-empty boundary, this result should also be true in genera $0$ and $1$; differently
 the components of $\mathrm{Diff}^+(\Sigma_g)$ are not contractible for $g=0,1$. I have briefly reminded the reader of this fact in the article.
 \item The referee writes \emph{p.19, line −11: ``$\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$''
 Since $\mathcal{N}$ is a neighborhood of $\mathcal{D} \cup c_1 \cup\dots\cup c_m$, it contains the neighborhood of
 $c_1$. It seems that the author wants to take the complement of the neighborhood
 of $c_1$ in the definition of $\mathcal{N}'$.}\\
 This is not necessary: $\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$ is already an open set, contains
 $\mathcal{D} \cup c_2 \cup\dots\cup c_m$ and is disjoint from $c_1\setminus\mathcal{D}$. One can reduce further
 $\mathcal{N}'$ by deleting an entire neighborhood of $c_1$, and then taking again the union with $\mathcal{D}$, but
 this is not necessary for the argument to work. I have stressed in the article that $\mathcal{N}' = (\mathcal{N} \setminus c_1 ) \cup \mathcal{D}$
 is already open.
 \item The referee writes \emph{Section 6: How about mentioning the articles ``Stable cohomology of the mapping
 class group with symplectic coefficients and the universal Abel-Jacobi map'' written by Eduard Looijenga
 and ``Characteristic classes of surface bundles'' written
 by Shigeyuki Morita. They compute the integral stable cohomology ring of the
 mapping class group of punctured surface with boundary.}\\
 It seems to me that Morita's article (BULLETIN   (New  Series)  OF  THE  AMERICAN  MATHEMATICAL   SOCIETY
 Volume  11,  Number  2, pages 386-388,  October   1984) only deals with characteristic classes of
 $\Sigma_g$-bundles (i.e., no puncture, no boundary) over $\mathbb{Q}$. Looijenga's article
 (Journal of Algebraic Geometry. 5. January 1994) also deals mainly with coefficients in $\mathbb{Q}$.
 It seems to me that the only with coefficients in $\mathbb{Z}$ in Looijenga's article is Proposition 2.1:
 however, only the case of \emph{ordered} punctures is treated, whereas the case of unordered
 punctures follows easily only over $\mathbb{Q}$, but, in my opinion, not over $\mathbb{Z}_2$.
 The stable ($g=\infty$), unordered ($m<\infty$ unordered punctures) case with coefficients
 in $\mathbb{Z}_2$ (actually any field) is rather treated in the article of B\"odigheimer and Tillmann,
 which I already cite. Therefore I would leave these two suggested references out. 
\end{itemize}
\end{document}

### 2020-05-01 Rational homology/rational_homology.tex
Final version ready to appear in JTop

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tikz-cd}

\setlength\parindent{0pt}

\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}
% \setlength{\parindent}{6pt}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}
    \newtheorem{ex}{Example}

\input{macros.tex}

\begin{document}

\title{Rational homology of configuration spaces of surfaces}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

\maketitle
Let $C_n(\Sg)$ denote the unordered configuration spaces of $n$ points
on the surface $\Sg$. In this note I list all the properties I know about $H_*\pa{C_n(\Sg);\Q}$.
Many of these properties can be found somewhere in the literature,
From now on rational coefficients in homology are always understood.

Let $C_{\bullet}(\Sg):=\coprod_{n\geq 0}C_n(\Sg)$, which is a module over the 
(homotopy) associative H-space $C_{\bullet}(\C)=\coprod_{n\geq 0}C_n(\C)$.

Then $\M:=H_*(C_{\bullet}(\Sg)$ is a bigraded $\Gamma_{g,1}$-representation over the bigraded Pontryiagin
ring $H_*(C_{\bullet}(\C))\cong\Q[\epsilon]\otimes \Lambda (\alpha)$, where $\epsilon$ has
bidegree (0,1) and $\alpha$ has bidegree (1,2). The bidegree $(*,\bullet)$ is given by
the homological degree $*$ and the weight (number of points used) $\bullet$.

Let $\bH=H_1(\Sg)$, concentrated in bidegree (2,2), and let $\cH=H_1(\Sg)$, concentrated in bidegree (1,1).
The following hold:

\begin{enumerate}
 \item $\M$ is in general neither a free $\Q[\epsilon]\otimes \Lambda (\alpha)$-module (only for $g=0$),
 nor a symplectic $\Gamma_{g,1}$-representation (only for $g\leq 1$, see the \emph{rational counterexample}
 in my article).
 \item Nevertheless $\M$ is a free $\Q[\epsilon]$-module (see work of B\"odigheimer and Tillmann), and the Johnson's kernel
 $\mathcal{J}_{g,1}\subset\Gamma_{g,1}$ acts trivially on it (one can check that
 separating Dehn twists act trivially on a generatring set of $\M$).
 \item Let $\Sym(\bH)$ denote the symmetric algebra on $\bH$, which is a bigraded ring with a
 symplectic action of $\Gamma_{g,1}$. Then $\M$ has a natural structure of (equivariant)
 $\Sym(\bH)$-module, i.e. there is a $\Gamma_{g,1}$-equivariant map
 \[
  \mu\colon\Sym(\bH)\otimes \M\to \M
 \]
 which satisfies the usual associativity, it preserves the bidigree and it is compatible with
 the action of the ring $\Q[\epsilon]\otimes \Lambda (\alpha)$ (this seems true to me, although
 I have never written down a formal proof of it).
 \item One can filter $\M$ by the submodules
 \[
  \M\supset\bH\cdot\M\supset\bH^2\cdot\M\supset\bH^3\cdot \M.
 \]
 Every $\bH^k\cdot\M$ is bigraded, $\Gamma_{g,1}$-invariant, and an equivariant module
 over $\Q[\epsilon]\otimes \Lambda (\alpha)\otimes \Sym(\bH)$. The associated graded,
 denoted $grad\M$, is also a bigraded $\Gamma_{g,1}$-equivariant module over $\Q[\epsilon]\otimes \Lambda (\alpha)\otimes \Sym(\bH)$.
 \item As such, $grad\M$ is isomorphic to the following
 \[
  grad\M\cong \Q[\epsilon]\otimes \Sym(\bH)\otimes [\alpha\otimes\coker(\cdot\wedge\omega)\oplus\Ker(\cdot\wedge\omega)].
 \]
 Here $(\cdot\wedge\omega)\colon\Lambda(\cH)\to\Lambda(\cH)$ is a map of bidegree (-2,-2), given by contracting
 an element of the exterior algebra $\Lambda(\cH)$ with the intersection form of the surface $\omega\in\Lambda_2(\cH^*)$.
 The action of $\Gamma_{g,1}$ is symplectic on $grad\M$. Multiplication by $\alpha$ is trivial on the summand
 $\alpha\otimes\coker(\cdot\wedge\omega)$ and is induced by the natural map
 $\Ker(\cdot\wedge\omega)\subset \Lambda(\cH)\to\coker(\cdot\wedge\omega)$ on the other summand (this should
 also be entailed in the work of Pagaria).
\end{enumerate}
\end{document}

### 2020-05-01 arxiv version/Actiongg.tex
Final version ready to appear in JTop

--- 2019-07-12 Actiongg.tex
@@
 \]
-% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
-% both by Poincaré-Lefschetz duality.
 
@@
 $Q^i\epsilon$.
-% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
-% is $m$ it must lie in $\Sym_m(\H)$.
-% 
-% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
-% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
-% closed curves; we consider the product map 
-% \[
-% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
-% \]
-% which carries the fundamental class of the $m$-fold torus to some class in
-% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
-% in $H_m(\cms)$.
 
@@
 Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
-such that
-%\begin{itemize}
-%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
-% \item
- any two curves $c_i,c_j$ intersect each other only inside $\D$.
+such that any two curves $c_i,c_j$ intersect each other only inside $\D$.
 
-% \item each curve $c_i$ is transverse to $\partial\D$.
-%\end{itemize}
 Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
@@
 See picture \ref{fig:defcCm}
-% For each isotopy class we
-% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
-% also transverse to all $\U_i$'s and $\V_i$'s. We write
-% $[c_1,\dots,c_m]$ for the isotopy class whose preferred representative is
-% $(c_1,\dots,c_m)$.
 
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defcCm.pdf}
+ \includegraphics[scale=0.7]{defcCm.pdf}
  \caption{A multicurve in $\cC^5$.}
@@
 
-% Our aim is to define a $\gg$-equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
-% factors through $\pr$ to give the desired $\gg$-equivariant map $\psi$; we will then show that
-% $\psi$ is an isomorphism.
-
-% The map $\tpsi_m$ that we construct will have two additional properties. The first
-% is the following: for any
-% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
-% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
-% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
-% this singular cycle only hits configurations in $\cms$ whose points actually
-% lie in $\D\cup c_1\cup \dots\cup c_m$.
-% 
-% For the second property we need a definition
 \begin{defn}
@@
 Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.
-% and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
-% with the aforementioned property.
 
@@
  from each other.
-%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
-%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
-%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
-%  Again $q$ is thewhite point
  \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
@@
  Again $\bar P$ is called the white point.
- %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
  \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
@@
 \begin{figure}\centering
- \includegraphics{figures/defcmsD.pdf}
+ \includegraphics{defcmsD.pdf}
  \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
@@
 a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
-% the latter being even a $m+1$-fold covering map;
 we have a commutative diagram
@@
   \begin{CD}
-   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
-\\   @V\p VV @V\p VV %@V\p VV
-\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
+   C_{1,m}(\S) @>\subset >> \fmstwoD
+\\   @V\p VV @V\p VV
+\\   C_{m+1}(\S) @>\subset >> \cmstwoD
    \end{CD}
@@
 \ref{thm:Hbms*as*ggrep} for the bigrading $(m-l,m)$.
-% \begin{defn}
-% %  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
-% %  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
-% %  $\D\subset\D'$.
-% %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
-% %  Note that $\S'$ is also a surface of type $\sg$.
-% %  
-%  Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
-%  $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
-%  
-%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in Definition \ref{defn:cms},
-%  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
-%  
-%  For every splitting $m=p+(m-p)$ there is an open embedding
-%  \[
-%  \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
-%  \]
-%  given by taking the union of configurations:
-% 
-% %  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
-% %  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
-% \end{defn}
-% Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
-% $\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.
 
@@
 implies that the two vector spaces have the same dimension.
-% 
-% We can equivalently consider the maps in cohomology
-%  \[
-%    \mu^*\colon  H^{m-l}\pa{\cms}\to H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)},
-%  \]
-% and show that they are surjective for $0\leq p\leq m-l$ and together exhibit a product splitting
-% \[
-%   H^{m-l}(\cms)=\prod_{p=0}^{m-l} H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)}.
-% \]
-% Note that the natural target for the cohomology map $\mu^*$ is the whole group
-% \[
-%  H^{m-l}\pa{C_{m-p}(\S')\times C_p(\D)}\simeq \prod_{q=0}^p H^q\pa{C_{m-p}(\S')}\otimes H^{m-l-q}\pa{C_p(\D)},
-% \]
-% where we have used the K\"unneth formula; we consider here, by abuse of notation,
-% the composition of this map with the projection
-% on the factor corresponding to $q=p$.
-% 
-% Using Poincaré-Lefschetz duality and again the K\"unneth formula we reduce to show that the maps
-% \[
-% \mu^{\infty}_* \colon \tH_{m-l}(\cms^{\infty})\to \tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}
-% \]
-% are surjective and exhibit $\tH_{m-l}(\cms^{\infty})$ as a product
-% \[
-% \tH_{m-l}(\cms^{\infty})=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^{\infty}}\otimes \tH_{p-l}\pa{C_p(\D)^{\infty}}.
-% \]
-% Here a few explanations are required:
-% we are considering the map $\mu^{\infty}\colon \cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$ which collapses the complement
-% of the open subspace $C_{m-p}(\S')\times C_p(\D)\subset\cms$ to the point at infinity; it induces a map
-% $\mu^{\infty}_*\colon \tH_{m-l}(\cms^{\infty}) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}$;
-% applying Poincaré-Lefschetz duality
-% and the K\"unneth formula to the second group we get an isomorphism
-% \[
-%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
-% \]
-% we are considering the composition of the map $\mu^{\infty}_*$ with the projection onto the factor $q=p$, and by
-% abuse of notation we call this map simply $\mu^{\infty}_*$.
-%
 
@@
 \begin{figure}\centering
- \includegraphics[scale=0.65]{figures/axb.pdf}
+ \includegraphics[scale=0.65]{axb.pdf}
  \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_5(\D))$;
@@
 generalised symmetric chains in $\cms^{\infty}$.
-
-% $[a]\otimes [b]$ has a dual class in $\tH_{p-l}(C_p(\D)^{\infty})\otimes \tH_p(C_{m-p}(\S')^{\infty})$
-% represented by the product of generalised symmetric chains
-% \[
-%  (\alpha_j)_j\otimes (p,(u_i,v_i)_{i\leq g}).
-% \]
-% If $\tup=(l,(x_i)_{i\leq l})$ is a tuple such that $e^{\tup}$ appears in the symmetric
-% chain $(\alpha_j)_j$, and if $\tup'=(0,(u_i,v_i)_{i\leq g})$ is the only tuple such that
-% $e^{\tup'}$ appears in the generalised symmetric chain $(p,(u_i,v_i)_{i\leq g})$, then
-% the product $e^{\tup}\times e^{\tup'}\subset C_p(\D)\times C_{m-p}(\S')\subset\cms$ is
-% contained in the closure of the generalised symmetric chain $e^{\tup}$, where...
-
-% If $p<p'$ then
-% the \emph{geometric} intersection between the cycle $|p',(\alpha'_j),(u'_i,v'_i)|$
-% in $\cms^{\infty}$ and the cycle $\mu_*(a\otimes b)$ in $\cms$ is empty: indeed the first cycle is supported
-% on $\infty\in\cms^{\infty}$ and on configurations where at least $p'$ points lie in the union
-% $\bigcup_i(\U_i\cup\V_i)\subset\S'$,
-% whereas the second cycle is supported on configurations where at least $m-p$ points lie
-% inside $\D$. The algebraic intersection between the homology classes $\mu_*([a]\otimes [b])$
-% and $[p',(\alpha'_j),(u'_i,v'_i)]$ is therefore also zero.
-% 
-% Consider now the case $p=p'$.
 
@@
 
-% We replace the map $\mu^{\infty}$ by a cellular approximation that agrees with it on the $p$-skeleton,
-% and by abuse of notation we still call $\mu^{\infty}$ the new map.
-% 
 Consider now the generalised symmetric chain $\kappa(p',\ualpha',\uu',\uv')$ representing
@@
 is a basis for $H_{m-l}(\cms)$.
-
-% 
-% 
-% 
-% , and let $e^{\tup}$
-% be one of the cells appearing in this generalised symmetric chain, with
-% $\tup=(l,(x_i)_{i\leq l}, (u_i,v_i)_{i\leq g})$, and consider the restriction
-% of $\mu^{\infty}\colon\cms^{\infty}\to \pa{C_{m-p}(\S')\times C_p(\D)}^{\infty}$
-% to $e^{\tup}\subset\cms^{\infty}$
-% 
-% If $p\neq p'$, the whole cell $e^{\tup}$ is mapped
-% to $\infty$: indeed if $p<p'$, then a configuration in $e^{\tup}$ contains
-% at least $p'$ points inside $\S'$, so it cannot lie in $C_p(\D)\times C_{m-p}(\S')$;
-% if $p'<p$ then a configuration in $e^{\tup}$ contains at least $m-p$ points 
-% 
-% 
-% then the algebraic
-% intersection between $(p',(u'_i,v'_i),(\alpha'_j))$ and $\mu_*(a\otimes b)$ is $0\in\Z_2$,
-% unless $p=p'$, $\alpha_j=\alpha'_j$ for all $j$, and $u_i=u'_i$ and $v_i=v'_i$ for all $1\leq i\leq g$.
-% 
-% Indeed the generalised symmetric chain $(p',(u'_i,v'_i),(\alpha'_j))$ hits only configurations
-% in $\cms$ where for all $1\leq i\leq g$ at least $u'_i$ points lie on $\U_i$ and at least $v'_i$
-% points lie on $\V_i$
 \end{proof}
@@
 \begin{figure}\centering
- \includegraphics[scale=0.6]{figures/counterexample.pdf}
+ \includegraphics[scale=0.6]{counterexample.pdf}
  \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}

### 2020-05-01 arxiv version/Comparison.tex
Final version ready to appear in JTop

\section{Comparison with the work of B\"{o}digheimer and Tillmann}
\label{sec:comparison}
In this section we will compare Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep}
with the results in \cite{BoT}. In particular we consider the following two theorems, that
we formulate in an equivalent, but different way as in \cite{BoT}, fitting in our
framework. In the entire section homology is taken with coefficients in $\Z_2$ unless explicitly
stated otherwise.
\begin{thm}[Corollary 1.2 in \cite{BoT}]
\label{thm:BoTone}
Let $\F$ be a field, and let $m\geq 0$, $g\geq 0$. Then the following graded vector spaces are isomorphic in degrees $*\leq \frac 23 g-\frac 23$
\[
 H_*\pa{\ggm;\F}\cong H_*\pa{\gg;\F}\otimes_{\F} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}.
\]
Here each variable $x_1,\dots,x_m$ has degree 2 and the symmetric group $\mathfrak{S}_m$ acts on the polynomial
ring $\F[x_1,\dots,x_m]$ by permuting the variables.
\end{thm}
We point out that the range of degrees $*\leq \frac 23 g-\frac 23$ is the stable range for the homology $H_*(\gg;\F)$
of the mapping class group, see \cite{Harer} for the original stability theorem and \cite{Boldsen, ORW:resolutions_homstab} for the improved
stability range.
\begin{thm}[Theorem 1.3 in \cite{BoT}]
\label{thm:BoTtwo}
For all $m\geq 0$ the map $\mu^{\cF}\colon C_1(\D)\times C_m(\cF_{\S'})\to C_{m+1}(\cF_{\S,\D})$ from Definition \ref{defn:muF}
induces a split-injective map in homology
\[
 \mu^{\cF}_*\colon H_*\pa{C_m(\cF_{\S'})}\cong H_0\pa{C_1(\D)}\otimes_{\F} H_*\pa{C_m(\cF_{\S'})}\hookrightarrow H_*\pa{C_{m+1}(\cF_{\S,\D})}.
\]
In other words, the inclusion of groups $\ggm\hookrightarrow\gg^{m+1}$ given by adding a puncture near the boundary induces a split
injective map $H_*(\ggm;\F)\hookrightarrow H_*(\gg^{m+1};\F)$.
\end{thm}

We first reformulate Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep} in a convenient way.
\begin{defn}
 \label{defn:Cbullet}
 We introduce some abbreviations for the following disjoint unions:
 \[
 C_{\bullet}(\D)=\coprod_{m\geq 0}C_m(\D);
 \]
 \[
 C_{\bullet}(\S)=\coprod_{m\geq 0}C_m(\S); 
 \]
 \[
 C_{\bullet}(\cF_{\S,\D})=\coprod_{m\geq 0}C_m(\cF_{\S,\D}).
 \]
\end{defn}
Then $C_{\bullet}(\D)$ is a (homotopy associative) topological monoid, with associated
Pontryagin ring $H_*\pa{C_{\bullet}(\D)}\cong\Z_2[Q^j\epsilon|j\geq 0]$ (see Equation \ref{eq:Cohen}).
This is a bigraded ring, where the two gradings are the homological degree and the weight.

The maps $\mu$ from Definition \ref{defn:muF} make $C_{\bullet}(\S)$ into a (homotopy associative)
module over the monoid $C_{\bullet}(\D)$; correspondingly $H_*\pa{C_{\bullet}(\S)}$ is a bigraded
module over $H_*\pa{C_{\bullet}(\D)}$. The structure of this module is described by the following
reformulation of Theorem \ref{thm:Hbms*as*ggrep}, see in particular the proof of Lemma \ref{lem:oplussplitting}.

\begin{thm}
 \label{thm:firstreformulation}
 There is an isomorphism of $\gg$-representations in (bigraded) modules over the ring $H_*\pa{C_{\bullet}(\D)}$
 \[
  H_*\pa{C_{\bullet}(\S)}\cong   H_*\pa{C_{\bullet}(\D)}\otimes\Sym_{\bullet}(\H).
 \]
\end{thm}

Similarly, Theorem \ref{thm:main} can be reformulated as follows, considering at the same time all values of $m\geq 0$.
\begin{thm}
 \label{thm:mainreformulation}
 There is an isomorphism of (bigraded) modules over $H_*\pa{C_{\bullet}(\D)}$
 \[
 H_*\pa{C_{\bullet}(\cF_{\S,\D})}\cong   H_*\pa{C_{\bullet}(\D)}\otimes H_*\pa{\gg;\Sym_{\bullet}(\H)}.
 \] 
\end{thm}
The proof follows from the arguments used in Section \ref{sec:sseq}: we have actually shown that the direct
sum of the spectral sequences $\bigoplus_{m\geq 0} E(m)$ is itself a free module over $H_*\pa{C_{\bullet}(\D)}$
on the second page, with the same description as above.

By virtue of homological stability in $g$ for the sequences of groups $\pa{\gg}_{g\geq 0}$ and 
$\pa{\ggm}_{g\geq 0}$, Theorem \ref{thm:BoTone} can be equivalently rephrased as an isomorphism of graded vector spaces
\[
  H_*\pa{\Gamma_{\infty,1}^m;\F}\cong H_*\pa{\Gamma_{\infty,1};\F}\otimes_{\F} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}.
\]
Here $\Gamma_{\infty,1}=\colim_{g\to\infty}\gg$ and $\Gamma_{\infty,1}^m=\colim_{g\to\infty}\ggm$.

Let $\Gamma_{\infty,1}^{\bullet}$ denote the disjoint union $\coprod_{m\geq 0}\Gamma_{\infty,1}^m$, which we
consider as a groupoid. Then Theorem \ref{thm:BoTone} is equivalent to the isomorphism of bigraded vector spaces
\[
  H_*\pa{\Gamma_{\infty,1}^{\bullet};\F}\cong H_*\pa{\Gamma_{\infty,1};\F}\otimes_{\F} \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}}.
\]
We consider $\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}$ as a free module over the Pontryagin ring
$H_*(\fS_{\bullet};\F)$, where $\fS_{\bullet}=\coprod_{m\geq 0}\fS_m$ and the Pontryagin product is induced by
the natural maps of groups $\fS_m\times\fS_{m'}\to\fS_{m+m'}$. This module structure is induced by the natural maps
\[
H_*\pa{\fS_m;\F[x_1,\dots,x_m]}\otimes_{\F} H_*\pa{\fS_{m'};\F}\to  H_*\pa{\fS_m\times\fS_{m'};\F[x_1,\dots,x_m]\otimes_{\F}\F} \to
\]
\[
\to H_*\pa{\fS_{m+m'};\F[x_1,\dots,x_{m+m'}]}. 
\]
From now on we assume $\F=\Z_2$.
Let $\beta_m=\beta_m(\D)$ denote the braid group on $m$ strands of the disc, and let $\beta_{\bullet}=\coprod_{m\geq 0}\beta_m$.
Then the natural projections $\beta_m\to\fS_m$ induce an \emph{inclusion} of Pontryagin rings (see \cite{CLM})
\[
 H_*(C_{\bullet}(\D))\cong H_*(\beta_{\bullet})\hookrightarrow H_*(\fS_{\bullet}).
\]
Hence $\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}$ becomes a free module over $H_*(C_{\bullet}(\D))$.
Theorem \ref{thm:mainreformulation} gives then the following isomorphism, which turns out to be an isomorphism of free
$H_*(C_{\bullet}(\D))$-modules.
\begin{equation}
 \label{eq:doubleiso}
\begin{split}
  H_*\pa{\Gamma_{\infty,1}^{\bullet}} & \cong H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H_{\infty})}\\
  & \cong H_*\pa{\Gamma_{\infty,1}}\otimes \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}}.
  \end{split}
\end{equation}
Here $\H_{\infty}=\colim_{g\to\infty} H_1(\Sigma_{g,1})$ is the first homology of the surface of infinite genus with one
boundary component.

Recall that $H_*\pa{\Gamma_{\infty,1}}$ is also a Pontryagin ring: there are natural maps of groups
$\Gamma_{g,1}\times\Gamma_{g',1}\to\Gamma_{g+g',1}$ inducing a Pontryagin product on the homology 
$H_*\pa{\Gamma_{\infty,1}}=\colim_{g\to\infty} H_*\pa{\gg}$.

Equation \ref{eq:doubleiso} gives
an isomorphism of modules over the ring
$H_*\pa{\Gamma_{\infty,1}}\otimes H_*(C_{\bullet}(\D))$. The action of the ring $H_*\pa{\Gamma_{\infty,1}}$
on $H_*\pa{\Gamma_{\infty,1}^{\bullet}}$ is induced on the colimit by 
the maps of groups $\gg\times\Gamma_{g',1}^m\to\Gamma_{g+g',1}^m$.
The corresponding action of $H_*\pa{\Gamma_{\infty,1}}$ on $H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$
comes from the action on $ H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$ which is induced on the colimit by the natural maps
\[
 H_*(\gg)\otimes H_*(\Gamma_{g',1};\Sym_m(H_1(\Sigma_{g',1})))\to H_*(\Gamma_{g+g',1};\Sym_m(H_1(\Sigma_{g+g',1}))).
\]

% Note also that $H_*\pa{\Gamma_{\infty,1}}\subset H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$ in a natural way.
The right-hand side $H_*\pa{\Gamma_{\infty,1}}\otimes \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}}$
in Equation \ref{eq:doubleiso} is a free module over the ring $H_*\pa{\Gamma_{\infty,1}}\otimes H_*(C_{\bullet}(\D))$.
As a consequence we have that $H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H_{\infty})}$ is a free module
over $H_*\pa{\Gamma_{\infty,1}}$; this last statement is of independent interest, and regards only the homology
of the group $\Gamma_{\infty,1}$ with twisted coefficients in $\Sym_{\bullet}(\H_{\infty})$.

% Both modules are free over the ring $H_*\pa{\Gamma_{\infty,1}}$, so we obtain an isomorphism of $H_*(C_{\bullet}(\D))$-modules
% \[
%  H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{>0}(\H)}\cong\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}.
% \]
% Note also that the right-hand side of Equation \ref{eq:doubleiso} does not depend on the stable mapping class group $\Gamma_{\infty,1}$.
% Here $\Sym_{>0}(\H)=\bigoplus_{m\geq 1}\Sym_m(\H)$.

We turn now to Theorem \ref{thm:BoTtwo}, and for the moment we assume again that $\F$ is a generic field and that the genus $g$
is fixed and finite. Considering
all values of $m$ at the same time, we can equivalently write
\begin{equation}
\label{eq:epsilonsplitting}
 H_*\pa{C_{\bullet}(\cF_{\S,\D});\F}\simeq \F[\epsilon]\otimes_{\F}\pa{\bigoplus_{m\geq 0} H_*\pa{C_m(\cF_{\S,\D}),C_{m-1}(\cF_{\S,\D})};\F}.
 \end{equation}

Here we use the convention $C_{-1}(\cF_{\S,\D})=\emptyset$, and we regard $C_{m-1}(\cF_{\S,\D})$ as a subspace of $C_m(\cF_{\S,\D})$ by using the map
$\mu^{\cF}$ in the statement of Theorem \ref{thm:BoTtwo}, with first input any fixed point $*\in C_1(\D)$. The class
$\epsilon$ is the canonical generator of $H_0(C_1(\D);\F)$.

In the case $\F=\Z_2$, Theorem \ref{thm:mainreformulation} improves Equation \ref{eq:epsilonsplitting} by exhibiting
$ H_*\pa{C_{\bullet}(\cF_{\S,\D})}$ as a free module over the ring $H_*(C_{\bullet}(\D))\cong\Z_2[Q^j\epsilon|j\geq 0]$,
rather than its subring $\Z_2[\epsilon]$.

### 2020-05-01 arxiv version/HBraidSurf.tex
Final version ready to appear in JTop

--- 2019-07-12 HBraidSurf.tex
@@
 
-% isomorphism of bigraded vector spaces
-% \begin{equation}
-% \label{eq:genuszero}
-% \bigoplus_{m\geq 0} H_*\pa{C_m(\D)}\simeq \Z_2[Q^j\epsilon\,|\, j\geq 0].
-% \end{equation}
-% This isomorphism was essentially proved by Fuchs in \ref{Fuchs:CohomBraidModtwo}. Moreover the left hand
-% side is the homology of the space $\coprod_{m\geq 0} C_m(\D)$, which has a natural structure
-% of algebra over the operad $E^2$ of little $2$-cubes. The isomorphism in equation \eqref{eq:genuszero}
-% is an isomorphism of rings (using the Pontryagin product on the left hand side), and $Q$ represents here
-% the Araki-Kudo-Dyer-Lashof operation $Q\colon H_*(X;\Z_2)\to H_{2*+1;\Z_2}(X)$,
-% which is defined for any $E^2$-algebra $X$; $\epsilon\in H_0(C_1(\D))$ is the generator.
-% This explains our notation; we refer to Cohen \cite[Chap.3]{CLM} for more details.
-
 In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2$-vector spaces}
@@
 $\gg$.
-%In these two sections we abbreviate $\S=\sg$.
 
@@
 and in this section we will prefer to compute $H^*(\cms)$ for all bigradings $(*,m)$.
-% the dual graded vector space $H_*(\cms)$ is abstractly isomorphic,
-% i.e. it has the same dimension.
 
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defTS.pdf}
+ \includegraphics[scale=0.7]{defTS.pdf}
  \caption{The space $Q(\S)$ for $\S=\Sigma_{3,1}$. The black pentagons represent the point $\cP_{even}$, whereas the black
@@
 \]
-% We will generically use the letter $m$ to denote any of the numbers $l$, $x_i$,$y_i$,$z$,$u_i$,$v_i$ or $w_i$
-% in the above sum, so $m\in\Z$; instead .
-% Similarly we will use the letter $\M$ to denote any of the intervals $\U_i,\V_i,\W_i$;
-% they will correspond to the letters $U_i,V_i,W_i$, so we also define a subset
-% $\bsymb=\set{(U_i,V_i)_{i\leq g},(W_i)_{i\leq n-1}}\subset\symb$.
 In symbols we write $\tup=(l,\ux,\uu,\uv)$. We omit from the notation $\ux$ if it vanishes: this happens precisely when $l=0$.
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defetup.pdf}
+ \includegraphics[scale=0.7]{defetup.pdf}
  \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,2,0),(0,3,2))$. Points
@@
 
-% We want to show that the collection of the $e_h$ for varying $h$, together with
-% the 0-cell $*$, give a cell decomposition of $C_k(\S)^{\infty}$. Denote by $\Delta^h$ the following product of simplices
 \[
@@
  \Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}},
-%  =\prod_{m\in\symb_l}\Delta^m.
 \]
@@
  \[
-  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2.
+  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2,
  \]
-%  This binomial coefficient may be anyway equal to $0\in\Z_2$.
+ where $\binom{x_i+x_{i+1}}{x_i}$ denotes the binomial coefficient.
  \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
@@
  \]
-%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
  \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
@@
  \]
-%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
 \end{itemize}
@@
  that cancel each other.
-%  We are working in $\Z_2$ so we do not have to be careful about orientations of cells
-%  and therefore about signs.
 \end{proof}
@@
 \end{itemize}
-% Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
 A symmetric chain $\kappa(\ualpha)$ is a cycle in the chain complex $\tCh_*(C_p(\D)^{\infty})$ (see \cite{Fuchs:CohomBraidModtwo}).
@@
 
-% This notation is compatible, after interpreting equation \eqref{eq:genuszero} as an isomorphism
-% of rings, with the results of \cite[Chap.3]{CLM}: the monomials constructed there, using
-% the Pontryagin product and the operation $Q$, are really the dual basis of the basis of
-% symmetric chains for cohomology.
-
 We will not need this finer result in this article, so in the following the expression
@@
 is $1\in\Z_2$ if and only $\ualpha=\ualpha'$.
-% i.e. we \emph{define} monomials as the dual basis of the basis of symmetric chains, and keep considering
-% \eqref{eq:Cohen} only as an isomorphism of bigraded vector spaces.
 
@@
 The spectral sequence considered above collapses on its first page and
-% $H_*(\cms^{\infty})$
-% has a $\Z_2$-basis given by generalised symmetric chains.
 we have the following lemma:
@@
 associated to generalised symmetric chains of weight $m$.
-% this basis is bijection with
-% the choices of the following set of data:
-% \begin{itemize}
-%  \item a number $0\leq p\leq m$;
-%  \item a splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$;
-%  \item a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ with $\alpha_j\geq 0$.
-% \end{itemize}
-% The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^{\infty})$ has homological degree
-% $m+\sum_{j=1}^{\infty}\alpha_j$.
 \end{lem}
-% To interpret this class
-% as a monomial in the tensor product of Theorem \ref{thm:Hbms*as*ggrep} we need the following definition.
-
-% We are almost ready to prove an isomorphism of bigraded $\Z_2$-vector spaces as in Theorem \ref{thm:Hbms*as*ggrep}.
 
@@
 of the latter cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
-
-% We fix simple closed curves $\u_i,\v_i$ on $\S$, for $1\leq i\leq g$, whose fundamental classes
-% $[\u_i],[\v_i]$ form the dual basis of $H_1(\S)$. We assume that, apart from the following necessary exceptions,
-% all curves $\U_i,\V_i,\u_i,\v_i$ for $1\leq i\leq g$ are disjoint:
-% \begin{itemize}
-%  \item $\u_i$ and $\v_i$ intersect once, transversely;
-%  \item $\u_i$ and $\U_i$ intersect once, transversely;
-%  \item $\v_i$ and $\V_i$ intersect once, transversely;
-%  \end{itemize}
-% SEE PICTURE.
 \end{defn}

### 2020-05-01 arxiv version/Introduction.tex
Final version ready to appear in JTop

--- 2019-07-12 Introduction.tex
@@
 \end{equation}
-% This short exact sequence corresponds to 
-% \begin{equation}
-%  \cms\to\mgm\to\mg.
-% \end{equation}
-% Here $\cms$ is the \emph{unordered} configuration space of $m$ points in the interior of $\Sigma$, i.e.
-% \begin{equation}
-%  \cms=\fms/\S_m=\set{(p_1,\dots,p_m)\in\pa{\mathring\sg}^m|p_i\neq p_j \;\forall i\neq j}/\S_m,
-% \end{equation}
-% the quotient of the \emph{ordered} configuration space of $m$ points in the interior of $\Sigma$ by the natural
-% action of the symmetric group $\S_m$ that permutes the (ordered list of) points in the configuration; $\mg$ and $\mgm$
-% are the moduli spaces of Riemann structures on $\sg$ and $\sgm$ respectively.
 
@@
 Theorem \ref{thm:Hbms*as*ggrep}, which together with Theorem \ref{thm:main} is the main result of the article.
-% in section \ref{sec:HBraidSurf}).
-
-%In particular this representation is symplectic: this had already benn stated in \cite{LM}.
-
-% The group $\gg$ is significantly smaller than $\ggm$, so the
-% computation of the right-hand side of equation \eqref{eq:main}
-% may be more tractable.
-%WE DO THIS COMPUTATION FOR $g=1$.
 
 The strategy of the proof does not generalize to fields of characteristic different from 2 or to the \emph{pure}
-mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In section \ref{sec:rational}
+mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In Section \ref{sec:rational}
 we describe in detail a counterexample with coefficients in $\Q$, which can be generalized both to coefficients

### 2020-05-01 arxiv version/Preliminaries.tex
Final version ready to appear in JTop

--- 2019-07-12 Preliminaries.tex
@@
 In this section we recollect some classical definitions and results about braid groups and mapping class groups.
-% Let $\sg$ be a compact, smooth, orientable surface of genus $g$ with one parametrised
-% boundary component $\partial \sg$. Let $\sgm$ be $\sg$ with a choice of $m$ distinct
-% points in the interior: these points are called \emph{punctures} and are not assumed to be ordered.
 
@@
 
- The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the labels $1,\dots,n$ of a configuration;
+ The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the labels $1,\dots,m$ of a configuration;
  the orbit space
@@
 
-%  
-%  We denote by $C_k(\sg)^{\infty}$ the one-point compactification of the space $C_k(\sg)$; the point
-%  at infinity is denoted by $*$ and serves as basepoint.
 \end{defn}
@@
 \begin{figure}\centering
-\includegraphics[scale=0.8]{figures/unordered.pdf}
+\includegraphics[scale=0.8]{unordered.pdf}
 \caption{A configuration in the space $C_5(\Sigma_{1,1})$}
@@
  A classical result by Earle and Schatz \cite{EarleSchatz} ensures that the connected components
- of $\Diff(\sg)$ are contractible. In particular $B\Diff(\sg;\partial\sg)\simeq B\gg$ is
+ of $\Diff(\sg)$ are contractible: note that for $g=1$ we are using the fact
+ that our surfaces have non-empty boundary. In particular $B\Diff(\sg;\partial\sg)\simeq B\gg$ is
  a classifying space for
@@
 In the whole article the genus $g\geq 0$ of the surfaces that we consider is supposed to be fixed,
-% unless explicitly stated otherwise,
 and we will abbreviate $\S=\sg$.
@@
 \begin{figure}\centering
- \includegraphics[scale=0.8]{figures/SS'D.pdf}
+ \includegraphics[scale=0.8]{SS'D.pdf}
  \caption{A splitting of $\S$ as $\S'\natural\bar\D$.}
@@
 \begin{defn}
-\label{defn:muF}
 For all $0\leq p\leq m$ there is a map
@@
 \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/defmu.pdf}
+ \includegraphics[scale=0.7]{defmu.pdf}
  \caption{The product of two configurations in $C_2(\D)$ and $C_4(\S')$.}
@@
 the first Dyer-Lashof operation. In particular $Q^j\epsilon$ is the generator of
-$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$. See figure \ref{fig:monomial}
+$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$. See figure \ref{fig:monomial}.
 
@@
 an isomorphism of bigraded rings. The left-hand side is a ring with the Pontryagin product,
-% coming from the action of the little $2$-cubes,
 and the right-hand side is a polynomial ring in infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$.
@@
  \begin{figure}\centering
- \includegraphics[scale=0.7]{figures/monomial.pdf}
+ \includegraphics[scale=0.7]{monomial.pdf}
  \caption{The class in $H_5(C_9(\D))$ corresponding to the monomial $\epsilon\cdot(Q\epsilon)^2\cdot Q^2\epsilon$.}
@@
 
-We consider in particular the one-point-compactification $\cms^{\infty}$ of $\cms$.
+We will consider in particular the one-point-compactification $\cms^{\infty}$ of $\cms$.

### 2020-05-01 arxiv version/Ssequence.tex
Final version ready to appear in JTop

--- 2019-02-18 Ssequence.tex
@@
 In particular if we prove the vanishing of all differentials $\partial_r$
-% exiting from the summand
-% $H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}$
 in the first spectral sequence, then also all differentials
@@
 
-% We observe that there is an isomorphism
-% \[
-%  H_k\pa{\gg;H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}\simeq H_{p-l}(C_p(\D))\otimes H_k\pa{\gg; H_{m-p}(C_{m-p}(\S'))};
-% \]
 The differentials in the spectral sequence $H_*(C_p(\D))\otimes E(m-p)$
@@
 surfaces with non-empty boundary, following the generalisation
-of Theorem \ref{thm:Hbms*as*ggrep} discussed at the end of section \ref{sec:Actiongg}.
+of Theorem \ref{thm:Hbms*as*ggrep} discussed at the end of Section \ref{sec:Actiongg}.

### 2020-05-01 arxiv version/splitting.tex
Final version ready to appear in JTop

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tikz-cd}

\setlength\parindent{0pt}

\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}
    \newtheorem{ex}{Example}

\newcommand{\C}{\mathbb{C}}
\newcommand{\cC}{\mathfrak{C}}
\newcommand{\fc}{\mathfrak{c}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\cF}{\mathcal{F}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\p}{p}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\fS}{\mathfrak{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\U}{\mathcal{U}}
\renewcommand{\u}{\mathfrak{u}}
\newcommand{\V}{\mathcal{V}}
\renewcommand{\v}{\mathfrak{v}}
\newcommand{\Z}{\mathbb{Z}}

\newcommand{\sgm}{\Sigma_{g,1}^m}
\newcommand{\sg}{\Sigma_{g,1}}

\newcommand{\Sg}{\mathcal{S}_{g,1}}
\newcommand{\Sgm}{\Sg^m}

\newcommand{\mg}{\mathfrak{M}_{g,1}}
\newcommand{\mgm}{\mg^m}
\newcommand{\Ig}{\mathcal{I}_{g,1}}

\renewcommand{\gg}{\Gamma_{g,1}}
\newcommand{\ggm}{\gg^m}

\newcommand{\bms}{\beta_{m}(\sg)}
\newcommand{\fms}{F_m(\sg)}
\newcommand{\cms}{C_m(\S)}
\newcommand{\fmsD}{C_{1,m}^{\D}(\S)}
\newcommand{\cmsD}{C_{m}^{\D}(\S)}
\newcommand{\fmstwoD}{C_{1,m}^{(2),\D}(\S)}
\newcommand{\cmstwoD}{C_{m+1}^{(2),\D}(\S)}
\newcommand{\Dmone}{\triangle_{m+1}}
\newcommand{\Donem}{\triangle_{1,m}}
\renewcommand{\i}{\iota}
\renewcommand{\j}{j}
\newcommand{\pr}{\mathfrak{p}}

\newcommand{\tup}{\mathfrak{h}}

\newcommand{\bH}{\mathbb{H}}
\newcommand{\cH}{\mathcal{H}}

\newcommand{\Ch}{\mathcal{C}}
\newcommand{\tCh}{\tilde\Ch}
\newcommand{\tH}{\tilde{H}}

\newcommand{\ZcC}[1]{\Z_2\left<\cC^{#1}\right>}
\newcommand{\tpsi}{\tilde\psi}

\newcommand{\pa}[1]{\left(#1\right)}
\newcommand{\qua}[1]{\left[#1\right]}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\set}[1]{\left\{#1\right\}}

\newcommand{\Sone}{\mathbb{S}^1}
\newcommand{\mrS}{\mathring{\S}}
\newcommand{\gone}{\Gamma_{1,1}}
\newcommand{\tu}{\tau_{\u}}
\newcommand{\tv}{\tau_{\v}}

\newcommand{\ux}{\underline{x}}
\newcommand{\uu}{\underline{u}}
\newcommand{\uv}{\underline{v}}
\newcommand{\ualpha}{\underline{\alpha}}

\newcommand{\tildeu}{\tilde u}
\newcommand{\tildev}{\tilde v}

\newcommand{\SP}{S\!P}

\renewcommand{\phi}{\varphi}
\renewcommand{\epsilon}{\varepsilon}

\newcommand{\hor}{hor}
\newcommand{\ver}{ver}

\renewcommand{\L}{\Lambda}
\DeclareMathOperator{\Diff}{Diff}
\DeclareMathOperator{\Imm}{Imm}
\DeclareMathOperator{\rk}{rk}
\DeclareMathOperator{\Ker}{ker}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\Hom}{Hom}
\def\colim{\mathop{\mathrm{colim}}\nolimits}

\begin{document}

\title{Splitting of the homology of the punctured mapping class group}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

\keywords{Mapping class group, braid group, symplectic coefficients.}
\subjclass[2010]{20F36, %Braid groups; Artin groups
55N25, %Homology with local coefficients, equivariant cohomology
55R20, %Spectral sequences and homology of fiber spaces 
55R35, %Classifying spaces of groups and $H$-spaces
55R40, %Homology of classifying spaces, characteristic classes
55R80, %Discriminantal varieties, configuration spaces
55T10. %Serre spectral sequences
}

\begin{abstract}
Let $\Gamma_{g,1}^m$ be the mapping class group of the orientable surface $\Sigma_{g,1}^m$ of genus $g$ with one parametrised
boundary curve and $m$ permutable punctures; when $m=0$ we omit it from the notation.
Let $\beta_{m}(\Sigma_{g,1})$ be the braid group on $m$ strands
of the surface $\Sigma_{g,1}$.

\noindent We prove that $H_*(\Gamma_{g,1}^m;\mathbb{Z}_2)\cong H_*(\Gamma_{g,1};H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2))$. The main ingredient
is the computation of $H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2)$ as a symplectic representation of $\Gamma_{g,1}$.
\end{abstract}

\maketitle

\section{Introduction}
Let $\sg$ be a smooth orientable surface of genus $g$ with one boundary curve $\partial\sg$, and let $\sgm$ be $\sg$ with
a choice of $m$ distinct points in the interior, called \emph{punctures}.

Let $\gg$ be the mapping class group of $\sg$, i.e. the group of isotopy classes of diffeomorphisms of $\sg$:
diffeomorphisms are required to fix $\partial\sg$ pointwise. Similarly let $\ggm$ be the mapping class group of $\sgm$, i.e.
the group of isotopy classes of diffeomorphisms of $\sgm$ that fix $\partial\sgm$ pointwise and \emph{permute} the $m$ punctures.

Forgetting the punctures gives a surjective map $\ggm\to\gg$ with kernel $\bms$,
the $m$-th \emph{braid group} of the surface $\sg$. We obtain the Birman exact sequence (see \cite{Birman:mcgbr})
\begin{equation}
\label{eq:Birman}
1\to\bms\to\ggm\to\gg\to 1.
\end{equation}

The associated Leray-Serre spectral sequence $E(m)$ in $\Z_2$-homology has a second page $E(m)^2_{k,q}=H_k(\gg;H_q(\bms;\Z_2))$,
and converges to $H_{k+q}(\ggm;\Z_2)$.

The main result of this article is that this spectral sequence collapses in its second page.
\begin{thm}
\label{thm:main}
For all $l\geq 0$ there is an isomorphism of vector spaces
\begin{equation}
\label{eq:main}
H_l\pa{\ggm;\Z_2}\cong \bigoplus_{k+q=l} H_k\pa{\gg;H_q\pa{\bms;\Z_2}}.
\end{equation}
\end{thm}
Thus the computation of $H_*\pa{\ggm;\Z_2}$ reduces to the computation of the homology of $\gg$ with
twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\gg$-representation
splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action: this is done in
Theorem \ref{thm:Hbms*as*ggrep}, which together with Theorem \ref{thm:main} is the main result of the article.

The strategy of the proof does not generalize to fields of characteristic different from 2 or to the \emph{pure}
mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In Section \ref{sec:rational}
we describe in detail a counterexample with coefficients in $\Q$, which can be generalized both to coefficients
in a field $\F_p$ of odd characteristic and to the pure mapping class group.

I would like to thank my PhD advisor Carl-Friedrich B\"odigheimer for his precious suggestions and his continuous encouragement
during the preparation of this work.

\section{Preliminaries}
\label{sec:Preliminaries}
In the whole article $\Z_2$-coefficients for homology and cohomology will be understood,
unless explicitly stated otherwise.

In this section we recollect some classical definitions and results about braid groups and mapping class groups.

\begin{defn}
\label{defn:cms}
 The $m$-th \emph{ordered configuration space} of a surface $\sg$ is the space
\[
 F_m(\sg)=\set{(P_1,\dots,P_m) \in \pa{\mathring{\Sigma}_{g,1}}^{\times m}  \,|\,  P_i\neq P_j  \;\forall i\neq j}.
\]

 Note that we require the points of the configuration to lie in the interior of $\sg$; the space
 $F_m(\sg)$ is a smooth, orientable $2m$-dimensional manifold.

 The symmetric group $\mathfrak{S}_m$ acts freely on $F_m(\sg)$ by permuting the labels $1,\dots,m$ of a configuration;
 the orbit space
 \[
 F_m(\sg)/\mathfrak{S}_m
 \]
 is called the \emph{$m$-th unordered configuration space}
 of $\sg$ and is denoted by $C_m(\sg)$; this space is also a $2m$-dimensional orientable manifold
 (see Figure \ref{fig:unordered}).

\end{defn}

\begin{figure}[ht]\centering
\includegraphics[scale=0.8]{unordered.pdf}
\caption{A configuration in the space $C_5(\Sigma_{1,1})$}
\label{fig:unordered}
\end{figure}

A classical result by Fadell and Neuwirth (\cite{FadellNeuwirth}) ensures
that $C_m(\sg)$ is aspherical; the fundamental group $\pi_1(C_m(\sg))$ is
called the \emph{braid group on $m$ strands of $\sg$} and is denoted by $\bms$.

\begin{defn}
 \label{def:mcg}
 Let $\Diff(\sg;\partial\sg)$ be the group of diffeomorphisms of $\sg$ that fix $\partial\sg$ pointwise,
 endowed with the Whitney $C^{\infty}$-topology. We denote by $\gg=\pi_0(\Diff(\sg;\partial\sg))$ its group of connected
 components, which is called the \emph{mapping class group} of $\sg$.

 Similarly let $\Diff(\sgm;\partial\sgm)$ be the group of diffeomorphisms of $\sgm$ that fix $\partial(\sg)$
 pointwise and restrict to a permutation of the $m$ punctures: the \emph{mapping class group} of $\sgm$ is
 $\ggm=\pi_0(\Diff(\sgm;\partial\sgm))$.
 \end{defn}
 A classical result by Earle and Schatz \cite{EarleSchatz} ensures that the connected components
 of $\Diff(\sg;\partial\sg)$ are contractible: note that for $g=0$ and $g=1$ this result holds
 because we consider surfaces with non-empty boundary, and the bounday must be fixed pointwise
 by our diffeomorphisms. In particular $B\Diff(\sg;\partial\sg)\simeq B\gg$ is
 a classifying space for
 the group $\gg$, i.e. an Eilenberg-MacLane space of type $K(\gg,1)$.

 We call $\cF_{g,1}\to B\Diff(\sg;\partial\sg)$ the universal
 $\sg$-bundle
 \[
  \sg\to\cF_{g,1}=\sg\times_{\Diff(\sg;\partial\sg)}E\Diff(\sg;\partial\sg)\to B\Diff(\sg;\partial\sg).
 \]
Applying the construction of the $m$-th unordered configuration space fiberwise we obtain a bundle
$C_m(\cF_{g,1})\to B\Diff(\sg;\partial\sg)$ with fiber $C_m(\sg)$.

The space $C_m(\cF_{g,1})$ is a classifying space for the
group $\ggm$ and the Birman exact sequence \eqref{eq:Birman} is obtained by taking fundamental
groups of the aspherical spaces
\begin{equation}\label{eq:Birmanbundle}
C_m(\sg)\to C_m(\cF_{g,1})\to B\Diff(\sg;\partial\sg).
\end{equation}

In the whole article the genus $g\geq 0$ of the surfaces that we consider is supposed to be fixed,
and we will abbreviate $\S=\sg$.
We denote by $\D$ the open disc $\mathring{\Sigma}_{0,1}$.

It will be useful, for many constructions, to choose an embedding
$\D\hookrightarrow\mrS$ \emph{near} $\partial\S$ and to replace
$\Diff(\S;\partial\S)$ with its subgroup $\Diff(\S;\partial\S\cup\D)$ of diffeomorphisms
of $\S$ that fix pointwise both $\partial\S$ and $\D$. Saying that $\D$ is embedded
\emph{near $\partial\S$} means that there is a compact subsurface $\S'\subset\S$
such that $\S=\S'\natural\bar\D$ is the union along a segment of $\S'$ and the closure of $\D$ in $\S$.

From now on we suppose such an embedding to be fixed and we consider $\D$ as a subspace
of $\mrS$ (see Figure \ref{fig:SS'D}). In Section \ref{sec:HBraidSurf}, Definition \ref{defn:TS},
we will introduce a convenient model $\T(\S)$ for the space $\mrS$, and in Section \ref{sec:Actiongg},
Definition \ref{defn:DinTS}
we will specify an embedding $\D\hookrightarrow\mrS$ using the model $\T(\S)$.

\begin{figure}[ht]\centering
 \includegraphics[scale=0.8]{SSprimeD.pdf}
 \caption{A splitting of $\S$ as $\S'\natural\bar\D$.}
\label{fig:SS'D}
\end{figure}

The construction in Definition \ref{defn:cms} can be specialised to the
surfaces $\D$ and $\S'$, yielding spaces $C_m(\D)$ and $C_m(\S')$ respectively.
We take configurations of points in the interior of the surfaces $\D$ and $\S'$ respectively
(since $\D$ is already an open surface this remark applies in particular to $\S'$).

The inclusion $\Diff(\S;\partial\S\cup\D)\subset\Diff(\S;\partial\S)$ is a homotopy
equivalence of topological groups, hence also the induced map
\[
B\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S)
\]
is a homotopy equivalence. We will replace the previous construction with the following,
homotopy equivalent ones.
\begin{defn}
\label{defn:universalSbundle}
We call $\cF_{\S,\D}\to B\Diff(\S;\partial\S\cup\D)$
the universal $\S$-bundle
 \[
  \S\to\cF_{\S,\D}\colon=\S\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)\to B\Diff(\S;\partial\S\cup\D).
 \]
Note that $\Diff(\S;\partial\S\cup\D)$ acts on $\S'$ by restriction of diffeomorphisms, and acts trivially
on $\D$; therefore $\cF_{\S,\D}$ contains subspaces
\[
 \cF_{\S'}= \S'\times_{\Diff(\S;\partial\S\cup\D)}E\Diff(\S;\partial\S\cup\D)
\]
and $\D\times B\Diff(\S;\partial\S\cup\D)$; these subspaces fiber over $B\Diff(\S;\partial\S\cup\D)$
with fibers $\S'$ and $\D$ respectively.

Applying the construction of the $m$-th unordered configuration space fiberwise, we obtain spaces
$C_m(\cF_{\S,\D})$, $C_m(\cF_{\S'})$ and $C_m(\D)\times B\Diff(\S;\partial\S\cup\D)$,
all fibering over $B\Diff(\S;\partial\S\cup\D)$, respectively with fiber $\cms$,
$C_m(\S')$ and $C_m(\D)$.
\end{defn}
The fiber bundle \eqref{eq:Birmanbundle} corresponding to the Birman exact sequence \eqref{eq:Birman}
can now be replaced with the following one
\begin{equation}\label{eq:BirmanbundleD}
\cms\to C_m(\cF_{\S,\D})\to B\Diff(\S;\partial\S\cup\D).
\end{equation}

\begin{defn}
\label{defn:muF}
For all $0\leq p\leq m$ there is a map
\[
 \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
\]
which takes the union of configurations (see Figure \ref{fig:defmu}):
 \[
  \mu\pa{\set{P_1,\dots,P_p};\set{P'_1,\dots,P'_{m-p}}}=\set{P_1,\dots,P_p,P'_1,\dots,P'_{m-p}}\in\cms.
 \]
\begin{figure}[ht]\centering
 \includegraphics[scale=0.7]{defmu.pdf}
 \caption{The product of two configurations in $C_2(\D)$ and $C_4(\S')$.}
\label{fig:defmu}
\end{figure}

We can apply this construction at the same time to each couple of fibers of the bundles $C_m(\cF_{\S,\D})$ and $C_{m-p}(\cF_{\S'})$
lying over the same point of $ B\Diff(\S;\partial\S\cup\D)$.

We obtain a map
\[
 \mu^{\cF}\colon C_p(\D)\times C_{m-p}(\cF_{\S'})\to C_m(\cF_{\S,\D}).
\]
If we see the domain of $\mu^{\cF}$ as a fibered product of bundles over $B\Diff(\S;\partial\S\cup\D)$
\[
 \pa{C_m(\D)\times B\Diff(\S;\partial\S\cup\D)}\times_{B\Diff(\S;\partial\S\cup\D)}C_{m-p}(\cF_{\S'}),
\]
then the map $\mu^{\cF}$ is also a map of bundles over $B\Diff(\S;\partial\S\cup\D)$, and the corresponding
map on fibers is precisely $\mu$.
\end{defn}

We recall now the structure of $H_*(C_m(\D))$.
The cohomology of $C_m(\D)$ with coefficients in $\Z_2$ was first computed by Fuchs in
\cite{Fuchs:CohomBraidModtwo}: in section
\ref{sec:HBraidSurf} we will generalise  Fuchs' argument to surfaces with boundary of positive genus.

In the appendix of \cite[Chap.III]{CLM}, Cohen considers the space $\coprod_{m\geq 0}C_m(\D)$ as an
algebra over the operad of little $2$-cubes, and describes its
$\Z_2$-homology as follows:
\begin{equation}
 \label{eq:Cohen}
H_*\pa{\coprod_{m\geq 0}C_m(\D)}\cong \Z_2\left[Q^j\epsilon \, |\, j\geq 0\right].
\end{equation}

Here $\epsilon\in H_0(C_1(\D))$ is the fundamental class, and for all
$k,m\geq 0$ we denote by $Q\colon H_k(C_m(\D))\to H_{2k+1}(C_{2m}(\D))$
the first Dyer-Lashof operation. In particular $Q^j\epsilon$ is the generator of
$H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$. See Figure \ref{fig:monomial}.

The isomorphism in equation \eqref{eq:Cohen} is
an isomorphism of bigraded rings. The left-hand side is a ring with the Pontryagin product,
and the right-hand side is a polynomial ring in infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$.
The bigrading is given by the homological degree
$*$, that we call the \emph{degree},
and by the index $m$ of the connected component
on which the homology class is supported (informally, the number of points
involved in the construction of the homology class), that we call the \emph{weight}.

 \begin{figure}[ht]\centering
 \includegraphics[scale=0.7]{monomial.pdf}
 \caption{The class in $H_5(C_9(\D))$ corresponding to the monomial $\epsilon\cdot(Q\epsilon)^2\cdot Q^2\epsilon$.}
\label{fig:monomial}
\end{figure}

In this article we will only need the isomorphism in equation \eqref{eq:Cohen} to hold as
an isomorphism of bigraded $\Z_2$-vector spaces.
In particular for all choices of natural numbers $(\alpha_j)_{j\geq 0}$ with all but finitely
many $\alpha_j=0$, we will consider the monomial $\prod_j\pa{Q^j\epsilon}^{\alpha_j}$, corresponding to
a homology class in some bigrading $(*,m)$: we will only use the fact
that the set of these monomials is a bigraded basis for the
left-hand side of equation \eqref{eq:Cohen}.

We end this section recalling the classical notions of symmetric product and one-point compactification.

\begin{defn}
 \label{defn:SP}
 The $m$-fold symmetric product of a topological space $\mathcal{M}$, denoted by $\SP^m(\mathcal{M})$, is
 the quotient of $\mathcal{M}^m$ by the (non-free) action of the symmetric group $\mathfrak{S}_m$
 which permutes the coordinates.
\end{defn}
We regard a point of $\SP^m(\mathcal{M})$ as a configuration of $m$ points in $\mathcal{M}$ with multiplicities.

In the case of an open surface $\mrS$, it is known that $\SP^m(\mrS)$ is a non-compact manifold of dimension $2m$;
it contains $\cms$ as an open submanifold.

\begin{defn}
 \label{defn:opc}
For a space $\mathcal{M}$ we denote by $\mathcal{M}^{\infty}$ its one-point
compactification; the basepoint is the point at infinity, that we denote by $\infty$.
\end{defn}

We will consider in particular the one-point-compactification $\cms^{\infty}$ of $\cms$.

\section{Homology of configuration spaces of surfaces}
\label{sec:HBraidSurf}
We want to study $H_*(\bms)=H_*(\cms)$,
which appears as the homology of the fiber
in the Leray-Serre spectral sequence associated to
\eqref{eq:Birman}, \eqref{eq:Birmanbundle} or \eqref{eq:BirmanbundleD}.
In particular we are interested in $H_*(C_m(\S))$
as a $\Z_2$-representation of the group $\gg$.

In \cite{LM} L\"offler and Milgram implicitly proved that $H_*(\bms)=H_*(\cms)$ is a $\Z_2$-symplectic
representation of the mapping class group. By \emph{$\Z_2$-symplectic} we mean the following:
\begin{defn}
 \label{defn:symplrep}
 Let $\H=H_1(\S)\simeq\Z_2^{2g}$.
 The natural action of $\gg$ on $\H$ induces a surjective map
 $\gg\to Sp_{2g}(\Z_2)$. A representation of $\gg$ over $\Z_2$ is called \emph{$\Z_2$-symplectic}
 if it is a pull-back of a representation of $Sp_{2g}(\Z_2)$ along this map.
\end{defn}

In \cite{BCT} B\"odigheimer, Cohen and Taylor computed $H_*(\cms)$ \emph{as a graded $\Z_2$-vector space}.
Their method provides all Betti numbers, but the action of $\gg$ cannot be easily deduced:
their descripition of $H_*(\cms)$ depends on a handle decomposition of $\S$, which is not preserved
by diffeomorphisms of $\S$, not even up to isotopy.

In this section and in the next one we will prove the following theorem; to the best of the author's knowledge
it does not appear in the literature.
\begin{thm}
 \label{thm:Hbms*as*ggrep}
 There is an isomorphism of bigraded $\Z_2$-representations of $\gg$
 \[
  \bigoplus_{m\geq 0} H_*\pa{\cms}\cong \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H).
 \]
 Here we mean the following:
 \begin{itemize}
  \item[(i)] on the left-hand side the bigrading is given by homological degree $*$ and by the direct summand,
  indexed by $m$,
  on which the homology class is supported, i.e. by the number $m$ of points
  involved in constructing the homology class; we call $*$ the \emph{degree} and $m$ the \emph{weight},
  and write $(*,m)$ for the bigrading;
  \item[(ii)] for $j\geq 0$, $Q^j\epsilon$ is the image in $H_{2^j-1}(C_{2^j}(\S))$ of a generator
  of the group $H_{2^j-1}(C_{2^j}(\D))\simeq \Z_2$ 
  under the natural map induced by the embedding $\D\hookrightarrow\mrS$,
  and $\Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]$ is the polynomial ring on
  infinitely many variables $\epsilon,Q\epsilon,Q^2\epsilon,\dots$;
  \item[(iii)] $\H=H_1(\sg)$ is identified with $H_1(C_1(\S))$ in a natural way, and lives in degree $1$ and weight $1$;
  $\Sym_{\bullet}(\H)$ denotes the symmetric algebra on $\H$;
  \item[(iv)] degrees and weights are extended on the right-hand side by the usual multiplicativity rule;
  \item[(v)] the action of $\gg$ on the right is the tensor product of the trivial action
  on the factor $\Z_2[Q^j\epsilon\,|\, j\geq 0]$, and of the action on
  $\Sym_{\bullet}(\H)$ which is induced by the $\Z_2$-symplectic action on $\H$.
  \end{itemize}
\end{thm}
Note that for any bi-homogeneus element in the right-hand side, the weight is greater or equal than
the degree: indeed factors of the form $Q^j\epsilon$ have weight strictly higher than their degree,
whereas factors belonging to $\H$ or to its symmetric powers have equal weight and degree.

Note that in the case $g=0$ the group $\Gamma_{0,1}$ is trivial and the previous theorem
reduces to equation \eqref{eq:Cohen}.

We point out that in \cite{BoT} B\"{o}digheimer and Tillmann have essentially proved that for a field
$\mathbb{F}$ of any characteristic the $\mathbb{F}$-vector space
\[
  \bigoplus_{m\geq 0} H_*\pa{\cms;\mathbb{F}}
\]
is isomorphic, as a bigraded $\gg$-representation over $\mathbb{F}$, to the tensor product
of the ring $\mathbb{F}[\epsilon]$, with trivial action, and some other bigraded representation:
here $\epsilon$ denotes, in analogy with the notation of Theorem \ref{thm:Hbms*as*ggrep}, the standard generator of $H_0(C_1(\D))$.
In Section \ref{sec:comparison} we will compare in detail Theorems \ref{thm:Hbms*as*ggrep}
and \ref{thm:main} with the results in \cite{BoT}.

In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2$-vector spaces}
as in Theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
$\gg$.

Since we work with coefficients in a field, it is equivalent to compute homology or cohomology,
and in this section we will prefer to compute $H^*(\cms)$ for all bigradings $(*,m)$.

We will mimic the method used by Fuchs \cite{Fuchs:CohomBraidModtwo} to compute the $\Z_2$-cohomology
of $C_m(\D)$.
As already mentioned in Section \ref{sec:Preliminaries}, our computation recovers a known result, but it has the advantage of
being quite elementary and of providing a part of
the geometric insight that we will need in the next section.

In the whole section we assume $m\geq 0$ to be fixed.
Since the space $\cms$ is homeomorphic to the interior of a compact
$2m$-manifold with boundary, by Poincaré-Lefschetz
duality we have
\[
 H^*(\cms)\simeq \tilde H_{2m-*}(\cms^{\infty}),
\]
where in the right hand side we consider reduced homology of the one-point compactification (see Definition \ref{defn:opc}).

We introduce a space $\T(\S)$ which is homeomorphic to $\mrS$, the interior of $\S$. The construction
corresponds to a handle decomposition of $\S$ with one $0$-handle and $2g$ $1$-handles.

\begin{defn}
\label{defn:TS}
If $g=0$, hence $\S=\Sigma_{0,1}$ is the disc, we set $\T(\S)=]0,1[^2$, the interior of the unit square. Assume
now $g\geq 1$, and see Figure \ref{fig:defTS} to visualize the following construction.

Dissect the interval $[0,1]$ into $2g$ equal subintervals through the points $\cP_i=\frac{i}{2g}$ for $0\leq i\leq 2g$
(for $i=0,2g$ we get the two endpoints of $[0,1]$).

Consider on the vertical sides of $[0,1]^2$ the intervals
$I_i^l=\set{0}\times [\cP_i,\cP_{i+1}]$ and $I_i^r=\set{1}\times [\cP_i,\cP_{i+1}]$ for $1\leq i\leq 2g$:
all these intervals are
canonically diffeomorphic
to $[0,1]$ by projecting on the second coordinate, rescaling linearly by a factor $2g$
and translating.

We define a bijection
between the two sets of left and right intervals:
for $1\leq i\leq g$, the interval $I^l_{2i-1}$ corresponds to $I^r_{2i}$,
and the interval $I^r_{2i-1}$ corresponds $I^l_{2i}$.

The space $Q(\S)$ is the quotient of the square $[0,1]^2$ obtained by identifying each couple
of corresponding intervals in the canonical way. For $1\leq i\leq g$ we call $\overline{\U}_i$ the image
of $I^l_{2i-1}$ in the quotient $Q(\S)$, and we call $\overline{\V}_i$ the image of $I^l_{2g}$ in $Q(\S)$.

Note that the image in $Q(\S)$ of the set $\set{0,1}\times\set{\cP_0,\dots,\cP_{2g}}$ consists of two
points $\cP_{odd}$ and $\cP_{even}$: for $\epsilon\in\set{0,1}$ and $0\leq i\leq 2g$ the point
$(\epsilon,\cP_i)$ is mapped to $\cP_{even}$ if $\epsilon+i$ is even, and is mapped to $\cP_{odd}$
otherwise.

The spaces $\overline{\U}_i\subset Q(\S)$ and $\overline{\V}_i\subset Q(\S)$ are intervals with endpoints $\cP_{even}$ and $\cP_{odd}$;
the interiors of these intervals are disjoint. Each $\overline{\U}_i$ and $\overline{\V}_i$ is the homeomorphic image of
some left interval $I^l_j$, and inherits from the latter a parametrisation by $[0,1]$.

We call $\U_i$ and $\V_i$ the interiors of the intervals $\overline{\U}_i$ and $\overline{\V}_i$ respectively.

The space $Q(\S)$ is homeomorphic to the compact surface $\S$; we call $\T(\S)$ the interior of
$Q(\S)$.

From now on we will identify $\S$ with $Q(\S)$ and $\mrS$ with $\T(\S)$;
consequently we will identify $\cms$ with the
space of configurations of $m$ points in $\T(\S)$.
\end{defn}

\begin{figure}[ht]\centering
 \includegraphics[scale=0.7]{defTS.pdf}
 \caption{The space $Q(\S)$ for $\S=\Sigma_{3,1}$. The black pentagons represent the point $\cP_{even}$, whereas the black
 squares represent $\cP_{odd}$.}
\label{fig:defTS}
\end{figure}

Our next aim is to define a structure of $CW$-complex on the space $\cms^{\infty}$:
the only $0$-cell will be the point $\infty$, whereas the other cells will be introduced
in the following definition.
\begin{defn}
\label{defn:ehopen}
A \emph{tuple} $\tup$ is a choice of the following set of data:
 \begin{itemize}
  \item a natural number $0\leq l\leq m$;
  \item a vector $\ux=(x_1,\dots, x_l)$ of integers $\geq 1$;
  \item vectors $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ of integers $\geq 0$,
 \end{itemize}
satisfying the following equality
\[
 m=\sum_{i=1}^lx_i+\sum_{i=1}^g(u_i+v_i).
\]
In symbols we write $\tup=(l,\ux,\uu,\uv)$. We omit from the notation $\ux$ if it vanishes: this happens precisely when $l=0$.
The \emph{dimension} of $\tup$ is defined as $m+l$.

For a tuple $\tup$ let $e^{\tup}$ be the subspace
of $\cms$ of configurations of $m$ points in $\mrS$ such that the following conditions hold
(see picture \ref{fig:defetup}):
\begin{itemize}
 \item for all $1\leq i\leq g$, exactly $u_i$ points lie on $\U_i$
 and exactly $v_i$ points lie on $\V_i$;
 \item there are exactly $l$ vertical lines in the open square $]0,1[^2\subset\mrS$ of the
 form $\set{s_i}\times]0,1[$ for some $0<s_1<\dots<s_l<1$, containing at least one
 point of the configuration. From left to right, these lines contain exactly $x_1,\dots,x_l$ points
 respectively.
\end{itemize}
\end{defn}

\begin{figure}[ht]\centering
 \includegraphics[scale=0.7]{defetup.pdf}
 \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,2,0),(0,3,2))$. Points
 lying on the $\U_i$'s and the $\V_i$'s are depicted twice.}
\label{fig:defetup}
\end{figure}

The space $e^{\tup}$ is homeomorphic to \emph{the interior} of the following multisimplex:

\[
 \Delta^{\tup}\colon =
 \Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}},
\]
where the simplex $\Delta^r$ is the subspace of $[0,1]^r$ of sequences $0\leq \tau_1\leq\dots\leq\tau_r\leq 1$
(the numbers $\tau_1,\dots,\tau_r$ are the \emph{local coordinates} of the simplex). The homeomorphism is given
as follows:
\begin{itemize}
 \item the local coordinates of the $\Delta^l$-factor correspond to the positions $s_1,\dots,s_l$ of the vertical
 lines in $(0,1)^2$ containing points of the configuration;
 \item the local coordinates of the $\Delta^{x_i}$-factor correspond to the positions of the $x_i$ points
 lying on the vertical line $\set{s_i}\times(0,1)$;
 \item the local coordinates of the $\Delta^{u_i}$-factor correspond to the positions of the $u_i$ points
 lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}$-factor,
 with $v_i$ and $\V_i$ replacing $u_i$ and $\U_i$.
\end{itemize}
Note that the dimension of $\Delta^{\tup}$ agrees with the dimension of $\tup$ from Definition \ref{defn:ehopen}.
The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^{\infty}$ extends to a continuous map
$\Phi^{\tup}\colon\Delta^{\tup}\to\cms^{\infty}$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
all subspaces $e^{\tup'}$ for tuples $\tup'$ of \emph{lower} dimension than $\tup$, together with the $0$-cell $\infty$.

The construction of the map $\Phi^{\tup}$ is as follows:
\begin{enumerate}
\item we identify the one-point compactification
$(\mrS)^{\infty}$ of $\mrS=\T(\S)$ as the quotient of $Q(\S)$ collapsing the boundary to a point $\infty$;
\item we consider the $m$-fold symmetric product $\SP^m\pa{(\mrS)^{\infty}}$ (see Definition \ref{defn:SP}): it contains 
as an \emph{open} subspace $\cms$, so we can identify $\cms^{\infty}$ as the quotient of $\SP^m\pa{(\mrS)^{\infty}}$ collapsing
the subspace $\SP^m\pa{(\mrS)^{\infty}}\setminus\cms$ to $\infty$;
\item the homeomorphism $\mathring{\Delta}^{\tup}\to e^{\tup}\subset\cms$ extends
now to a map $\Delta^{\tup}\to\SP^m\pa{[0,1]^2}$, that we can then further project to $\SP^m(Q(\S))$, then to
$ \SP^m\pa{(\mrS)^{\infty}}$ and then to $\cms^{\infty}$: the composition
is the map $\Phi^{\tup}$.
\end{enumerate}
In the last step we used the open inclusion $\cms\subset\SP^m\pa{\mrS}\subset\SP^m\pa{(\mrS)^{\infty}}$, which
induces a map $ \SP^m\pa{(\mrS)^{\infty}}\to\cms^{\infty}$ by quotienting to $\infty$ the closed subspace
$ \SP^m\pa{(\mrS)^{\infty}}\setminus \cms^{\infty}$.

We conclude that the collection of the $e^{\tup}$'s, together with the $0$-cell $\infty$,
gives a cell decomposition of $\cms^{\infty}$, with characteristic maps of cells $\Phi^{\tup}$.

We compute the \emph{reduced} cellular chain complex of $\cms^{\infty}$
with coefficients in $\Z_2$. It is a chain complex in $\Z_2$-vector spaces with basis given
by tuples $\tup$, which correspond to the cells $e^{\tup}$.
\begin{lem}
\label{lem:doperatoropenmodtwo}
Let $\tup=(l,\ux,\uu, \uv)$ and
$\tup'=(l-1,\ux',\uu',\uv')$
be tuples in consecutive dimensions $m+l$ and $m+l-1$. Denote by $[\tup\colon\tup']\in\Z_2$ the coefficient
of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$.
Then $[\tup:\tup']=0$ unless one, and exactly one, of the following situations occurs:
\begin{itemize}
 \item $l\geq 2$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, setting $x'_i=x_i+x_{i+1}$
 for one value $1\leq i\leq l-1$, and shifting the values
 $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case we say that $\tup'$ is an \emph{inner boundary} of $\tup$, and
 we have
 \[
  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2,
 \]
 where $\binom{x_i+x_{i+1}}{x_i}$ denotes the binomial coefficient.
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_1=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and shifting the values
 $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case we say that $\tup'$ is a \emph{left, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_l=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
 In this case we say that $\tup'$ is a \emph{right, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
\end{itemize}

\end{lem}
It may indeed happen that $\tup'$ is both a left and a right outer boundary of $\tup$,
namely when all numbers $(x_i)_{1\leq i\leq l}$ are equal; then the two contributions cancel each other,
so that $[\tup\colon\tup']=0$ as stated.
\begin{proof}
 For $1\leq i\leq l$ denote by $\partial^{\hor}_i\Delta^{\tup}$ the face
 \[
 \partial_i\Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}}\subset\Delta^{\tup}.
 \]
 This is also referred as the $i$-th \emph{horizontal face}. All other faces of codimension 1 of
 the multisimplex $\Delta^{\tup}$ are called \emph{vertical}.

 We note that $\Phi^{\tup}$ restricts to a cellular map $\partial_i^{\hor}\Delta^{\tup}\to\cms^{\infty}$ on every
 horizontal face, where $\partial_i^{\hor}\Delta^{\tup}$
 is given the cell structure coming from the multisimplicial structure: every subface of $\partial_i^{\hor}\Delta^{\tup}$ of dimension
 $k\leq m+l-1$ is mapped to the $k$-skeleton of $\cms^{\infty}$. Therefore the map $\Phi^{\tup}\colon\partial_i^{\hor}\Delta^{\tup}\to\cms^{\infty}$
 has a well-defined local index over the cell $e^{\tup'}$, that
 we call $[\partial^{\hor}_i\tup\colon\tup']\in\Z_2$.

 The same holds for vertical faces, where the restriction of $\Phi^{\tup}$ is the constant map to $\infty$: in this case the local index
 over $e^{\tup'}$ is zero. The index $[\tup\colon\tup']$ of the map $\Phi^{\tup}\colon\partial\Delta^{\tup}\to\cms^{\infty}$ on
 $e^{\tup'}$ splits as a sum of local indices:
 \[
  [\tup\colon\tup']=\sum_{i=0}^l[\partial_i^{\hor}\tup\colon\tup']\in\Z_2.
 \]
 For $0\leq i\leq l$ the restriction $\Phi^{\tup}\colon\partial^{\hor}_i\Delta^{\tup}\to\cms^{\infty}$
 hits homeomorphically the open cell $e^{\tup'}$
 exactly as many times as specified in the statement
 of the lemma for the cases $1\leq i\leq l-1$,
 $i=0$ and $i=l$ respectively.

 The only possibility in which the same cell $e^{\tup'}$ is hit by different faces $\partial_i^{\hor}\Delta^{\tup}$
 and $\partial_j^{\hor}\Delta^{\tup}$ is the one described in the remark preceding this proof: in this case
 there are two equal contributions $[\partial_i^{\hor}\tup\colon\tup']$ and $[\partial_j^{\hor}\tup\colon\tup']$
 that cancel each other.
\end{proof}

We can filter the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$ by giving filtration norm $\sum_{i=1}^lx_i$ to
the tuple $\tup=(l,\ux,\uu,\uv)$, with $\ux=(x_1,\dots,x_l)$. For example
the tuple in Figure \ref{fig:defetup} has norm 6.

By Lemma \ref{lem:doperatoropenmodtwo}
the norm is weakly decreasing along differentials. Denote by $F_p\subset\tCh_*(\cms^{\infty})$ the subcomplex generated by
tuples of norm $\leq p$, and let $F_p/F_{p-1}$ be the $p$-th filtration stratum.
Then $F_p/F_{p-1}$ is isomorphic, as a chain complex, to a direct sum of copies of $\tCh_*(C_p(\Sigma_{0,1})^{\infty})$:
there is one copy for each partition
\[
(m-p)=\sum_{i=1}^g (u_i+v_i)
\]
with $u_i,v_i\geq 0$. The isomorphism
does not preserve the degrees but shifts them by $p$.

Indeed in $F_p/F_{p-1}$ all outer
differentials vanish (see Lemma \ref{lem:doperatoropenmodtwo}): in particular the numbers $u_i,v_i$
do not change along the differentials of $F_p/F_{p-1}$. Therefore $F_p/F_{p-1}$ splits as a direct sum
of chain complexes indexed by all partitions $(m-p)=\sum_{i=1}^g (u_i+v_i)$ as above.
It is then immediate to identify the inner faces
with the ones one would have in the case $g=0$, i.e. for the surface $\Sigma_{0,1}$, after shifting
degrees by $p$.

We note that $\tCh_*(C_p(\Sigma_{0,1})^{\infty})$ is exactly the chain complex described by Fuchs in
\cite{Fuchs:CohomBraidModtwo}: we recall Fuchs' computation of the cohomology of configuration
spaces of the disc, and abbreviate $\D$ for $\mathring{\Sigma}_{0,1}$ as in Section \ref{sec:Preliminaries}.
\begin{defn}
\label{defn:symchain}
Consider a partition of $p$ into powers of 2
\[
 p=\sum_{j\geq 0}\alpha_j2^j,
\]
and let $\ualpha=(\alpha_j)_{j\geq 0}$ be the sequence of multiplicities. We understand that only finitely
many $\alpha_j$'s are strictly positive.

The associated \emph{symmetric chain} in $\tCh_*(C_p(\D)^{\infty})$, denoted by $\kappa(\ualpha)$,
is the sum of all tuples $\tup=\pa{l,(x_i)_{i\leq l}}$ such that
\begin{itemize}
 \item $l=\sum_{j\geq 0}\alpha_j$;
 \item every $x_i$ is a power of 2;
 \item for all $j\geq 0$ there are exactly $\alpha_j$ indices $i$ such that $x_i=2^j$.
\end{itemize}
A symmetric chain $\kappa(\ualpha)$ is a cycle in the chain complex $\tCh_*(C_p(\D)^{\infty})$ (see \cite{Fuchs:CohomBraidModtwo}).
We denote by $[\kappa(\ualpha)]\in\tH_{p+l}(C_p(\D)^{\infty})$ the associated homology class.
\end{defn}
In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\D)^{\infty})$
is given by the collection of all classes $[\kappa(\ualpha)]$ associated to sequences $\ualpha=(\alpha_j)_{j\geq 0}$
which satisfy the equality $p=\sum_{j\geq 0}\alpha_j2^j$.

By Poincaré-Lefschetz duality this corresponds to a basis for $H^*(C_p(\D))$.
The \emph{dual} basis of $H_*(C_p(\D))$
happens to be the basis of monomials
\[
Q^{\ualpha}\epsilon\colon =\prod_{j\geq 0}(Q^i\epsilon)^{\alpha_j}\in H_*(C_p(\D)).
\]
This basis consists of all monomials
of weight $p$, using
the isomorphism \eqref{eq:Cohen} in its full meaning (i.e. as an isomorphism of rings,
where $Q$ denotes the first Dyer-Lashof operation).

We will not need this finer result in this article, so in the following the expression
$Q^{\ualpha}$ will only denote
the (unique) homology class in $H_{p-l}(C_p(\D))$ such that the following holds:
for all $\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\sum_{j\geq 0}\alpha'_j=p$, the
algebraic intersection between $Q^{\ualpha}\epsilon\in H_{p-l}(C_p(\D))$
and $[\kappa(\ualpha')]\in\tH_{p+l}(C_p(\D)^{\infty})$
is $1\in\Z_2$ if and only $\ualpha=\ualpha'$.

We now go back to the filtered chain complex $\tCh_*(\cms^{\infty})$.
The $E^1$-page of the associated Leray spectral sequence contains on the $p$-th column
the homology of $F_p/F_{p-1}$; as we have seen the homology of this filtration stratum is
the direct sum of several copies of the homology of $\tCh_*(C_p(\D))$, one copy for each
partition $(m-p)=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$.

\begin{defn}
\label{defn:gensymchain}
Consider a partition
\[
 m=p+(m-p)=\pa{\sum_{j=0}^{\infty}\alpha_j2^j}+\pa{\sum_{i=1}^g(u_i+v_i)}
\]
and let $\uu=(u_1,\dots,u_g)$, $\uv=(v_1,\dots, v_g)$, $\ualpha=(\alpha_j)_{j\geq 0}$; denote also $l=\sum_{j=0}^{\infty}\alpha_j$.

We define a chain $\kappa(p,\ualpha,\uu,\uv)$ in
$\tCh_{m+l}(\cms)$: it
is the sum of all tuples $\tup$ of the form $\pa{l,\ux,\uu,\uv}$, for varying $\ux$,
which satisfy the three properties listed in Definition \ref{defn:symchain}.

We call such a chain a
\emph{generalised symmetric chain}; we will see in the following that it is a cycle, and we will
denote by $[\kappa(p,\ualpha,\uu,\uv)]$ the associated homology class in $\tH_{m+l}(\cms^{\infty})$.

If any of $\ualpha$, $\uu$ and $\uv$ vanishes, we omit it from the notation.
\end{defn}

A generalised symmetric chain $\kappa(p,\ualpha,\uu,\uv)$
is not only
a cycle when projected to its filtration quotient $F_p/F_{p-1}$, as the $E^1$-page
of the spectral sequence tells us, but also
in the chain complex $\tCh_*(\cms^{\infty})$ itself.

To prove this fact, first note that an inner boundary of a tuple
$\tup$ preserves the norm; hence the fact that $\kappa(p,\ualpha,\uu,\uv)$ is a cycle
in $F_p/F_{p-1}$ guarantees the fact that all inner boundaries of $\kappa(p,\ualpha,\uu,\uv)$
cancel each other in $\tCh_*(\cms^{\infty})$, that is,
\[
\partial\kappa(p,\ualpha,\uu,\uv)\in \tCh_*(\cms^{\infty})
\]
is equal to the sum of all outer boundaries of the tuples involved.

Note now that outer boundaries of a generalised symmetric chain also cancel out:
the left outer boundary
of a tuple $\tup=(l,\ux,\uu,\uv)$ in the generalised symmetric chain cancels against the right outer boundary
of the tuple $\tup'=(l,\ux',\uu,\uv)$, with $x'_l=x_1$ and $x'_i=x_{i+1}$ for $1\leq i\leq l-1$. If all
$x_i$ happen to be equal, then $\tup=\tup'$ and
we are in the situation described before the proof of Lemma \ref{lem:doperatoropenmodtwo}.

The spectral sequence considered above collapses on its first page and
we have the following lemma:
\begin{lem}
\label{lem:gensymchain}
The homology $H_*(\cms^{\infty})$ has a graded basis given by the classes $[\kappa(p,\ualpha,\uu,\uv)]$
associated to generalised symmetric chains of weight $m$.
\end{lem}

\begin{defn}
 \label{defn:dualHbasis}
We can see the $\U_i$'s and $\V_i$'s as properly embedded $1$-manifolds in $\mrS$;
by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^{\infty}}\simeq H^1(\S)$,
and in particular they form a basis
of the latter cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
\end{defn}
We establish a bijection between monomials in the tensor product of Theorem \ref{thm:Hbms*as*ggrep}
and the basis of $H^*(\cms)\simeq H_*(\cms^{\infty})$ in Lemma \ref{lem:gensymchain}:
the class 
\[
[\kappa(p,\ualpha,\uu,\uv)]\in H_{m+\sum\alpha_j}(\cms^{\infty})\simeq H^{m-\sum\alpha_j}(\cms)
\]
is associated with the monomial
\[
Q^{\ualpha}\epsilon\cdot\u^{\uu}\cdot\v^{\uv}\colon =\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g(\u_i^{u_i}\v_i^{v_i}),
\]
where $\ualpha=(\alpha_j)_{j\geq 0}$, $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$.

This shows an isomorphism of bigraded $\Z_2$-vector spaces
\begin{equation}\label{eq:isovectorspaces}
  \bigoplus_{m\geq 0} H^*(\cms)\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(\H),
\end{equation}
from which we conclude that
there exists an isomorphism as in Theorem \ref{thm:Hbms*as*ggrep}
at least \emph{as bigraded $\Z_2$-vector spaces}: the two bigraded
vector spaces have the same dimension in all bigradings.

\section{Action of \texorpdfstring{$\gg$}{Gamma(g,1)}}
\label{sec:Actiongg}
We now turn back to \emph{homology} of $\cms$. In the first subsection we describe geometrically some
homology classes, in order to give some intuition for the following subsections.
In the second subsection we prove Theorem
\ref{thm:Hbms*as*ggrep} in bigradings $(*,m)$ with $*=m$.
In the third subsection we extend the proof to all other bigradings.

\subsection{Geometric examples of homology classes}

In this subsection we consider the case $g=2$, hence $\S$ denotes the surface $\Sigma_{2,1}$.
We construct some homology classes in $H_2(C_2(\S))$: our aim is to get a first understanding
of why this homology group is isomorphic to $\Sym_2(\H)=\H^{\otimes 2}/\mathfrak{S}_2$;
the notation $\H=H_1(\S)$ was introduced in Definition \ref{defn:symplrep}.

In the following $c$ and $d$ will always denote two simple closed curves on $\S$, with corresponding homology classes $[c],[d]\in\H$.
\begin{ex}
 \label{ex:one}
Suppose that $c$ and $d$ are as in Figure \ref{fig:exampleone}: $c$ and $d$ are disjoint and non-separating.
We can consider inside $C_2(\S)$
the torus $c\times d$ of configurations in which one of the two points runs along $c$, while the other runs
along $d$. We associate to the fundamental class $\tau_1\in H_2(C_2(\S))$ of this torus the tensor product $[c]\otimes[d]\in\H^{\otimes 2}$.

Since our configurations are unordered and since we work with coefficients in $\Z_2$, the same
class $\tau_1$ can be obtained as a tensor product $[d]\otimes[c]$, i.e. exchanging the order of the two
curves: we can neglect the sign $-1$ that this operation would generate. We represent $\tau_1$ as $[c]\cdot[d]\in\Sym_2(\H)$.
\end{ex}
\begin{figure}[ht]\centering
 \includegraphics[scale=1.0]{exampleone.pdf}
 \caption{}
\label{fig:exampleone}
\end{figure}

\begin{ex}
 \label{ex:two}
Suppose that $c$ and $d$ are as in Figure \ref{fig:exampletwo}:
$c$ and $d$ are disjoint, and
 $c$ bounds a subsurface $\check{\S}$ of $\S$ (the shaded region), such that $d$ is not contained in $\check{\S}$.

 Then the homology class $\tau_2=[c]\cdot[d]$ vanishes, because the torus $c\times d$ is the boundary in $C_2(\S)$
 of the 3-manifold $\check{\S}\times d$ containing all configurations in which one point lies on $\check{\S}$ and
 the other on $d$. This is consistent with the representation $\tau_2=[c]\cdot[d]$, because $[c]=0\in \H$.
\end{ex}
\begin{figure}[ht]\centering
 \includegraphics[scale=1.0]{exampletwo.pdf}
 \caption{}
\label{fig:exampletwo}
\end{figure}

\begin{ex}
 \label{ex:three}
 Suppose that $c$ and $d$ are as in Figure \ref{fig:examplethree}: $c$ and $d$ are non-separating and 
intersect transversely in one point $P\in\S$. In this case the torus $c\times d$ is a subspace
of $\SP^2(\S)$, but $c\times d$ is not contained in $C_2(\S)$ as in Example \ref{ex:one}.

To solve this problem, we consider a small open neighborhood $P\in \U\subset\S$, and we remove from the torus $c\times d$
all configurations in which both points lie in $\U$: these removed configurations form an open disc inside the torus $c\times d$.

We obtain an embedding $\Sigma_{1,1}\hookrightarrow C_2(\S)$;
the boundary $\partial\Sigma_{1,1}$,
seen as a curve in $C_2(\S)$, is homotopic to a curve $\gamma$ in which one point spins $360^{\circ}$ around the other.
We note that the curve $\gamma\subset C_2(\S)$ is homotopic to a double covering of a curve $\gamma'\subset C_2(\S)$,
in which the two points exchange their positions after spinning $180^{\circ}$ around each other. All curves
$\partial\Sigma_{1,1}$, $\gamma$, $\gamma'$ and the homotopies relating them
are supported on the closure of $C_2(\U)$ in $C_2(\S)$.

We can therefore find a map from a M\"{o}bius band $\M$ to the closure of $C_2(\U)$ in $C_2(\S)$,
such that the images of the curves $\partial\M$ and $\partial\Sigma_{1,1}$ in $C_2(\S)$ coincide.
The union $\Sigma_{1,1}\cup_{\partial}\M$ along the boundary is then a closed
non-orientable surface of genus $3$, i.e. the connected sum of a torus and a projective plane.

The surface $\Sigma_{1,1}\cup_{\partial}\M$ is equipped with a map to $C_2(\S)$, hence its fundamental class with coefficients
in $\Z_2$ yields a homology class $\tau_3\in H_2(C_2(\S))$; thus we have managed to adapt
the construction from Example \ref{ex:one} to the case of two intersecting curves. We represent $\tau_3$
as $[c]\cdot[d]\in\Sym_2(\H)$.
\end{ex}
\begin{figure}[ht]\centering
 \includegraphics[scale=1.0]{examplethree.pdf}
 \caption{}
\label{fig:examplethree}
\end{figure}

\begin{ex}
 \label{ex:four}
 Suppose that $c$ and $d$ are as in Figure \ref{fig:examplefour}: $c$ and $d$ are disjoint, and
 $c$ bounds a subsurface $\check{\S}$ of $\S$ (the shaded region), such that $d$ is contained in $\check{\S}$.

 The torus $c\times d$ is contained in $C_2(\S)$, but it is not obvious, at first glance, why its corresponding homology
 class $\tau_4\in H_2(C_2(\S))$ should vanish: the argument used in Example \ref{ex:two} does not work immediately, because the
 3-manifold $\check{\S}\times d$ is only contained in $\SP^2(\S)$ and not in $C_2(\S)$.

 We try to modify this 3-manifold in the same spirit of Example \ref{ex:three}. Let $\U\tilde{\times} d\subset \check{\S}\times d$
 be an open tubular neighborhood of $d\times d$: note that this tubular neighborhood is a trivial bundle over
 $d$ with fiber an open disc $\U$. The notation $\tilde{\times}$ means that abstractly we are dealing with a product $\U\times d$,
 but not geometrically: fibers over different points of $d$ are naturally identified with different discs in $\check{\S}$.

 We consider the 3-manifold $\check{\S}\times d\setminus \U\tilde{\times} d$: its boundary is the disjoint union
 of the torus $c\times d$ and the torus $\partial \pa{\U\tilde{\times} d}$. The latter torus contains configurations of
 two points in $\S$, one of which spins around $d$, whereas the other is a \emph{satellite} of the first and spins
 around it $360^{\circ}$.

 We regard $\partial \U\tilde{\times} d$ as a trivial bundle $\gamma\tilde{\times} d$ over $d$, with fiber the curve $\gamma$
 considered in Example \ref{ex:three}. After a homotopy the latter torus
 becomes a double covering of a torus $\gamma'\tilde{\times} d$: this is a trivial bundle over $d$ with fiber the curve $\gamma'$;
 we can see $\gamma'\tilde{\times} d\subset C_2(\S)$ as a torus of configurations in which the two points
 exchange their positions spinning $180^{\circ}$ around each other, while their barycenter spins around $d$.

 We can fill the second boundary of the $3$-manifold $\check{\S}\times d\setminus \U\tilde{\times} d$
 with the product $\M\times d$: the output is a non-orientable $3$-manifold with boundary $c\times d$, witnessing
 that $\tau_4=0$.

 This is consistent with the representation $\tau_4=[c]\cdot[d]$, because $[c]=0\in \H$.
\end{ex}

\begin{figure}[ht]\centering
 \includegraphics[scale=0.8]{examplefour.pdf}
 \caption{}
\label{fig:examplefour}
\end{figure}

\begin{ex}
 \label{ex:five}
 Suppose that $c$ and $d$ are as in Figure \ref{fig:examplefive}:
 $c$ bounds a subsurface $\check{\S}$, and $d$ is cut by $c$ into two arcs, one of which, denoted by $e$, lies
 in $\check{\S}$.

 In this example the complications from Examples \ref{ex:three} and \ref{ex:four} arise at the same time.

 To define the class $\tau_5\in H_2(C_2(\S))$, we start with the torus $c\times d$ and we perform two surgeries
 with two different M\"{o}bius bands to solve the two intersections of $c$ and $d$: the homology
 class that we obtain is represented by a non-orientable surface of genus 4, i.e. the connected
 sum of a torus and 2 projective planes.

 To show that $\tau_5$ vanishes, we start with the 3-manifold with boundary $\check{\S}\times d\subset\SP^2(\S)$ and we perform
 a surgery. We identify with $e$ the arc
 \[
  \check{\S}\times d\cap\pa{\SP^2(\S)\setminus C_2(\S)};
 \]
 this is a properly embedded arc in the 3-manifold with boundary $\check{\S}\times d$, and a tubular neighborhood
 of it, after a suitable homotopy, can be identified with a trivial $\U$-bundle $\U\tilde{\times} e$. We remove
 this solid cylinder from the 3-manifold and glue a trivial $\M$-bundle $\M\tilde{\times} e$, by applying an argument
 similar as the one in Examples \ref{ex:three} and \ref{ex:four}.

 We obtain a 3-manifold with boundary endowed with a map to $C_2(\S)$; the boundary of this 3-manifold
 is precisely the surface used to represent $\tau_5$: therefore $\tau_5=0\in H_2(C_2(\S))$,
 and this is consistent with the representation $\tau_5=[c]\cdot[d]$, because $[c]=0\in \H$.
\end{ex}

\begin{figure}[ht]\centering
 \includegraphics[scale=1]{examplefive.pdf}
 \caption{}
\label{fig:examplefive}
\end{figure}

If we consider more than 2 points, the examples become more and more complicated, and proving the theorem with this geometric
approach seems rather difficult.

\subsection{Bigradings of the form \texorpdfstring{$(m,m)$}{(m,m)}}
In this subsection we construct a $\gg$-~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]

We already know that these $\Z_2$-vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bigrading $(m,m)$ in equation
\eqref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.

The construction of $\psi_m$ is rather long and technical and involves a few definitions.

\begin{defn}
\label{defn:DinTS} 
Recall Definition \ref{defn:TS}. We denote by $\D\subset\T(\S)\cong\mrS$ the open square $]1/4;3/4[\times]1/2,1[$.
Note that $\D$ is an open disc in $\mrS$ near $\partial\S$ and 
it is disjoint from all $\overline{\U}_i$'s and $\overline{\V}_i$'s. The interior of the surface $\S'$
is then identified with $\T(\S)\setminus [1/4,3/4]\times[1/2,1[$.
\end{defn}

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that any two curves $c_i,c_j$ intersect each other only inside $\D$.

Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images. Here and in the following $\Sone$ is the unit circle in $\C$.

Two $m$-tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m$-tuple into the other. In particular $\cC^m$ is more than countable.

An element of $\cC^m$ is called \emph{multicurve}; by abuse of notation the $m$-tuple
$(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
See picture \ref{fig:defcCm}

We denote by $\ZcC{m}$ the free $\Z_2$-vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg$-equivariant.
\end{defn}

\begin{figure}[ht]\centering
 \includegraphics[scale=0.7]{defcCm.pdf}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

\begin{defn}
 \label{defn:cmsD}
 Recall Definition \ref{defn:SP}.
 The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.
 See Figure \ref{fig:defcmsD}.

  Note that $\cmsD$ is open in $\SP^m(\mrS)$. There is an open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  \[
  \j_m\colon \ZcC{m}\to H_m(\cmsD)
  \]
  defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
  the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> (\mrS)^{\times m} @>>> \SP^m(\mrS)
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m$-fold torus in $H_m(\cmsD)$. The result does not change
if we substitute $(c_1,\dots,c_m)$ with another isotopic $m$-tuple of curves in the same multicurve.

We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in the singular chain complex of $\cmsD$ of
the fundamental cycle of the
$m$-fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
on the union $c_1\cup\dots\cup c_m$, meaning that it hits configurations in $\cmsD$ of points of $\mrS$
lying in this union.

The group $\Diff(\S,\D\cup\partial\S)$ acts on $\cmsD$, and there is an induced action of $\gg$
on $H_m(\cmsD)$. The map $\j_m$ is $\gg$-equivariant.
\end{defn}

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$
 is surjective: this last map is induced by the map $\cmsD^{\infty}\to\cms^{\infty}$ collapsing the subspace
 $\cmsD^{\infty}\setminus\cms$ to $\infty$.

 Recall Lemma \ref{lem:gensymchain} and Definition \ref{defn:gensymchain}. A basis for $\tH_m(\cms^{\infty})$
 is given by classes $[\kappa(0,\uu,\uv)]$, with
 $\uu=(u_1,\dots, u_g)$, $\uv=(v_1,\dots,v_g)$ and
 $\sum_{i=1}^g (u_i+v_i)=m$.

 The class $[\kappa(\uu,\uv)]$ is represented by a generalised symmetric chain consisting of only one tuple $\tup=(0,\uu,\uv)$.

 In particular there is a map of pairs
 $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^{\infty},\infty)$, and the class $[\kappa(\uu,\uv)]$
 is the image along this map of the fundamental class of $H_m\pa{\Delta^{\tup},\partial\Delta^{\tup}}$.

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$ follows.
\end{proof}

We have the following diagram of $\gg$-equivariant maps, where the full arrows are those
that we have already constructed, and we still have to prove the existence of the dashed arrows
\begin{equation}
\label{eq:fulldasheddiagram}
\begin{tikzcd}[column sep=8em,row sep=5em]
  \ZcC{m} \ar[r,"\pr_m",two heads]\ar[d,dashed, swap, "\tpsi_m"]\ar[dr,"\j_m",near start]
  & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*",hook] & H_m\pa{\cmsD}.
 \end{tikzcd}
\end{equation}

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg$-equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by Lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and this last statement does not depend on how $\gg$ acts on these groups.

We will prove by induction on $m$ the following technical lemma:

\begin{lem}
 \label{lem:tpsiwithproperties}
For each $m$-tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
and for each open neighborhood $\N\subset\mrS$ of
$\D\cup c_1\cup\dots\cup c_m$, there is a singular cycle $\fc=\tpsi_m(c_1,\dots,c_m)$
in $\cms$ with the following properties:
\begin{itemize}
 \item the cycle $\fc$ is \emph{supported} on $\N$, i.e.,
this singular cycle only hits configurations of $m$ distinct points of $\mrS$ that actually lie in $\N$;
\item $\pa{\iota_m}_*(\fc)$ represents the homology class $\j_m(c_1,\dots,c_m)\in H_m\pa{\cmsD}$;
\item the two cycles $\pa{\iota_m}_*(\fc)$ and $[c_1]\cdot\ldots\cdot[c_m]$
are connected by a homology in $\cmsD$ which is supported on $\N$ (the word
\emph{homology} denotes here a $(m+1)$-singular chain whose boundary is the difference between the two cycles).
\end{itemize}
\end{lem}

For $m=0$ both $\ZcC{0}$ and $H_0(C_0(\S))$ are
isomorphic to $\Z_2$ and there is nothing to show. For $m=1$
we have a canonical identification $H_1(C_1(\S))\simeq\H\simeq\Sym_1(\H)$, so we take $\tpsi_1=\pr_1$;
obviously for all $c_1$ representing a class in $\cC^1$, the homology class $\pr_1(c_1)\in\H$
is represented by a cycle supported on $c_1$,
and in this case the cycles $\iota_*\pa{\tpsi(c_1)}$ and $\pr_1(c_1)=\j_m(c_1)$ coincide.

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.

\begin{defn}
\label{defn:variationsCm}
We introduce several variations of the notion of configuration space; see Figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}[ht]\centering
 \includegraphics{defcmsD.pdf}
 \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

We have the following inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}(\mrS).
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

In particular there is a sequence of maps
\[
 H_{m+1}\pa{C_{m+1}(\S)}\to H_{m+1}\pa{\cmstwoD}\to H_{m+1}\pa{C_{m+1}^{\D}(\S)}
\]
and we will first lift the homology class $j_{m+1}(c_1,\dots,c_{m+1})$ to $H_m\pa{\cmstwoD}$ and
then to $H_{m+1}(C_{m+1}(\S))$, each time controlling the support of our representing
cycles and of the homologies between them.

Fix a neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.

For the first lift, let $\N'=\pa{\N\setminus c_1}\cup\D$; note that $\N'$ is open
in $\mrS$ and contains $\D\cup c_2\cup\dots\cup c_{m+1}$, so $\N'$ is an open
neighborhood of $\D\cup c_2\cup\dots\cup c_{m+1}$ in $\mrS$. By inductive hypothesis
there is a cycle $\fc=\tpsi_m(c_2,\dots,c_{m+1})$ in $\cms$ which is supported on $\N'$,
and such that $(\iota_m)_*(\fc)$ is homologous to $[c_2]\cdot\ldots\cdot[c_{m+1}]$
along a homology in $\cmsD$ supported on $\N'$ as well.

We can multiply both
cycles and the homology between them by the cycle $[c_1]$: the result are the two homologous cycles
$[c_1]\cdot \fc$ and $[c_1]\cdot\ldots\cdot[c_{m+1}]$ in $C_{m+1}^{\D}(\S)$: both cycles and the
homology between them are supported on $\N$. Note now that the
cycle $[c_1]\cdot\fc$ lives in $\cmstwoD$, so the first lift is done and we can now
deal with the second lift.

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD
\\   @V\p VV @V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We note that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
a $2$-fold ramified covering between their respective normal bundles.
\end{defn}

Diagram \eqref{eq:cmstwodiagram} induces a commutative diagram in homology
\begin{equation}
 \label{eq:fivediagram}
\minCDarrowwidth15pt
 \begin{CD}
  @. H_{m+1}\pa{\fmstwoD} @>>> H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)}\\
  @. @V\p_*VV @V\p_*VV\\
  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwoD} @>>> H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
 \end{CD}
\end{equation}

Recall that we want to lift the homology class represented by the cycle $[c_1]\cdot\fc$
from the bottom central group to the bottom left group.

We first note that there is a lift of $[c_1]\cdot\fc$ to a cycle $[c_1]\otimes\fc$ in $\pa{\fmstwoD}$:
this is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be white. We then note that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
\]
can be rewritten, after using excision to tubular neighborhoods of $\Donem$ and $\Dmone$ respectively,
and the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone).
\]
The latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

Therefore the image of the cycle $[c_1]\otimes \fc$ along the diagonal of the square
in diagram \eqref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
is zero; hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$. More
precisely, there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
to $[c_1]\cdot\fc$.

To prove Lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
a good homology, namely two that are supported on $\N$: a priori both $\fc'$ and the homology
between $\pa{\iota_{m+1}}_*(\fc')$ and $[c_1]\cdot\fc$ are only supported on $\mrS$.

This can be done by replacing, in the whole argument of the proof, the surface $\mrS$ with the surface $\N$.
We can define configuration spaces as in Definition \ref{defn:variationsCm} also for the open surface
$\N$, and we can repeat the argument considering $\N$ as the \emph{ambient surface}:
indeed we only needed a surface containing $\D$ and all curves $c_1,\dots,c_{m+1}$.

It is crucial that the action of $\gg$ is not involved in the statement of Lemma
\ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
by diffeomorphisms of $\S$.
Lemma \ref{lem:tpsiwithproperties} is proved.

We now have to prove the following lemma to conclude the proof of Theorem \ref{thm:Hbms*as*ggrep}
in bigradings $(m,m)$.
\begin{lem}
 \label{lem:tpsi->psi}
The map $\tpsi_m\colon\ZcC{m}\to H_m(\cms)$ is surjective and factors through the map $\pr_m$.
\end{lem}
\begin{proof}
The factorisation is equivalent to the inclusion $\ker\pr_m\subseteq\ker\tpsi_m$: since both
$\pr_m$ and $\tpsi_m$
are $\gg$-equivariant, also the induced map of vector spaces
\[
\Sym_m(\H)=\ZcC{m}/\ker\pr_m\to H_m(\cms)
\]
will automatically be
$\gg$-equivariant.

Recall from the proof of Lemma \ref{lem:cms->cmsDinj} that a basis for $\tH_m(\cms^{\infty})$ 
is given by the classes $[\kappa(\uu,\uv)]$, represented by generalised
symmetric chains consisting of only one tuple $\tup=(0,\uu,\uv)$, for some vectors
$\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ satisfying $\sum_{i=1}^g(u_i+v_i)=m$.

The homology class
$[\kappa(\uu,\uv)]$ is the fundamental class
of the sphere $e^{\tup}\cup\set{\infty}\subset\cms^{\infty}$: the inclusion of this sphere in $\cms^{\infty}$
restricts to a proper embedding $e^{\tup}\subset\cms$.

By Poincaré Lefschetz duality $\tH_m(\cms^{\infty})\simeq H^m(\cms)$, and the latter is
the dual of $H_m(\cms)$.

We can therefore associate to $[\kappa(\uu,\uv)]$ a linear functional
on $H_m(\cms)$. This is the algebraic intersection product with the cell $e^{\tup}$, seen as a proper submanifold of $\cms$:
we denote it by
\[
 \cdot\cap e^{\tup}\colon H_m(\cms)\to\Z_2.
\]
Therefore
\[
 \ker\tpsi_m=\bigcap_{\tup}\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m},
\]
and it suffices to check that $\ker\pr_m\subseteq\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m}$
for all $\tup$ of the form $(0,\uu,\uv)$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.

Recall from the proof of Lemma \ref{lem:cms->cmsDinj} that the cohomology class $(\cdot\cap e^{\tup})$ on $\cms$ is a pullback
of a cohomology class of $\cmsD$, that we call $(\cdot\cap e^{\tup})^{\D}$. Alternatively,
note that the inclusion $e^{\tup}\cap \set{\infty}\to\cms^{\infty}$ is the composition of the inclusion $e^{\tup}\cap\set{\infty}\to\cmsD^{\infty}$
and the quotient map $\cmsD^{\infty}\to\cms^{\infty}$, and consider the fundamental class of the sphere $e^{\tup}\cup\set{\infty}$ and its images.

We can therefore compute the map $(\cdot\cap e^{\tup})\circ\tpsi_m$ as the map 
\begin{equation}
\label{eq:badformula}
(\cdot\cap e^{\tup})^{\D}\circ\j_m\colon \ZcC{m}\to\Z_2.
\end{equation}
The latter map coincides with the composition
\begin{equation}
 \label{eq:productformula}
\pa{ \prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}}\circ\pr_m,
\end{equation}

where $\prod_{i=1}^g(\cdot\cap\U_i)^{u_i}(\cdot\cap\V_i)^{v_i}\in\Sym_m(\Hom(\H;\Z_2))=\Hom\pa{\Sym_m(\H);\Z_2}$.

This can be checked on every generator $(c_1,\dots, c_m)\in\ZcC{m}$ by chosing in the isotopy class a representative $(c_1,\dots, c_m)$
with all curves $c_i$ transverse to all segments $\U_j$ and $\V_j$.

Consider again the map $c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to\cmsD$ that we used to define the cycle $[c_1]\cdot\dots[c_m]$
representing the class $\j_m(c_1,\dots,c_m)$ (see Definition \ref{defn:cmsD}):
this map is an embedding near $e^{\tup}$ and is transverse to $e^{\tup}$.

The equality of the maps in equations \eqref{eq:badformula} and \eqref{eq:productformula} on the generator $(c_1,\dots,c_m)\in\ZcC{m}$
follows from a straightforward
computation (in $\Z_2$) of the cardinality of the set $[c_1]\cdot\dots[c_m]\cap e^{\tup}$ in terms
of the cardinalities of all sets of the form $c_i\cap\U_j$ and $c_i\cap \V_j$.
In particular $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.

To show surjectivity of $\tpsi_m$, choose a tuple $\tup$ of the form $(0,\uu,\uv)$
and an $m$-tuple of curves $(c_1,\dots,c_m)$ containing, for every $1\leq i\leq g$, $u_i$ parallel
copies of some curve representing $\u_i$ and $v_i$ parallel copies of some curve representing
$\v_i$ (see Definition \ref{defn:dualHbasis}), such that all intersections between these curves
lie in $\D$.

Then $j_m(c_1,\dots,c_m)\cap e^{\tup}=1$ and for all
other tuples $\tup'$ of the form $(0,\uu',\uv)$ we have instead $j_m(c_1,\dots,c_m)\cap e^{\tup'}=0$.

This shows that $\psi_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
one of the generating monomials.
\end{proof}

Theorem \ref{thm:Hbms*as*ggrep} is now proved in all bigradings of the form $(m,m)$.

\subsection{General bigradings \texorpdfstring{$(m-l,m)$}{(m-l,m)}.} Fix $0\leq l\leq m$ for the whole subsection: our next aim is to prove Theorem
\ref{thm:Hbms*as*ggrep} for the bigrading $(m-l,m)$.

For all $0\leq p\leq m$, the group $\Diff(\S;\partial\S\cup\D)$ acts both on $C_p(\D)\times C_{m-p}(\S')$
and on $\cms$, and the map $\mu$ is equivariant with respect to this action
(see Definition \ref{defn:universalSbundle});
hence, using the K\"{u}nneth formula, there is an induced $\gg$-equivariant map in homology
\begin{equation}
 \label{eq:mu*}
 \mu_*\colon H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}\to H_{m-l}\pa{\cms}.
\end{equation}

Note that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
of the representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
section is isomorphic to the symplectic representation $\Sym_{m-p}(\H)$.

We will prove the following lemma, from which Theorem \ref{thm:Hbms*as*ggrep} follows:
\begin{lem}
 \label{lem:oplussplitting}
For all $l\leq p\leq m$ the map $\mu_*$ in equation \eqref{eq:mu*} is injective, and the collection
of all these maps yields a splitting
\begin{equation}
  \label{eq:musplitting}
 H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}.
\end{equation}
\end{lem}
\begin{proof}
Note that the statement of the lemma does not depend on the the action of $\gg$:
we have a map from the right-hand side to the left-hand side of equation \eqref{eq:musplitting},
we already know that it is $\gg$-equivariant,
we only need to show that it is a linear isomorphism. Note also that Lemma \ref{lem:gensymchain}
implies that the two vector spaces have the same dimension.

Fix $l\leq p\leq m$ and $\ualpha=(\alpha_j)_{j\geq 0}$, and let $[a]=Q^{\ualpha}\epsilon=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
$H_{p-l}(C_p(\D))$, hence $l=\sum_j\alpha_j$ and $p=\sum_j\alpha_j2^j$.

Fix also $\uu=(u_1,\dots, u_g)$ and $\uv=(v_1,\dots,v_g)$,
and let $[b]=\u^{\uu}\cdot\v^{\uv}=\prod_{i=1}^g (\u_i^{u_i}\v_i^{v_i})$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
isomorphism proved in the previous subsection, with $(m-p)=\sum_i(u_i+v_i)$.

Here $a$ and
$b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
and $b$ supported on $\S'$. See Figure \ref{fig:axb}

\begin{figure}[ht]\centering
 \includegraphics[scale=0.65]{axb.pdf}
 \caption{From left to right, the class $[a]=\epsilon\cdot(Q\epsilon)^2\in H_2(C_5(\D))$;
 the class $[b]=\v_1\cdot\u_2\cdot\v_3^2\in H_4(C_4(\S'))$; and the product class
 $\mu_*([a]\otimes[b])\in H_6(C_9(\S))$.}
\label{fig:axb}
\end{figure}

Then $[a]\otimes [b]$ is a generator of $H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))$,
by the K\"unneth formula, and we are interested in the homology class $\mu_*([a]\otimes [b])$.

There is one such
class for any choice of $[a]$ and $[b]$ as above, that is, for any choice of
$p$, $\ualpha$, $\uu$ and $\uv$ satisfying the conditions
$l=\sum_{j\geq 0}\alpha_j$, $p=\sum_{j\geq 0}\alpha_j2^j$ and $(m-p)=\sum_{i=1}^g(u_i+v_i)$, where
we use the notation above.

We want to show that the collection of all the corresponding classes of the form $\mu_*([a]\otimes [b])$
gives a basis for $H_{m-l}(\cms)$.

We will study
the intersection of $\mu_*([a]\otimes [b])$ with cohomology classes of $\cms$ represented by
generalised symmetric chains in $\cms^{\infty}$.

To compute the algebraic intersection
between $\mu_*([a]\otimes [b])$ and $[\kappa(p',\ualpha',\uu',\uv')]$ we
consider the map
\[
 \mu^{\infty}\colon \cms^{\infty}\to \pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}
\]
which collapses to $\infty$ the complement in $\cms^{\infty}$ of the open submanifold $C_p(\D)\times C_{m-p}(\S')$.

By Poincaré-Lefschetz duality, the map $\mu^{\infty}_*$ in reduced homology corresponds to the cohomology map
\[
 \mu^*\colon H^*(\cms)\to H^*(C_p(\D)\times C_{m-p}(\S'))=H^*(C_p(\D))\otimes H^*(C_{m-p}(\S')).
\]

We give $\pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}$ the cell complex structure of the smash product
$C_p(\D)^{\infty}\wedge C_{m-p}(\S')^{\infty}$. Here $C_p(\D)^{\infty}$ is given the cell structure of $C_p((0,1)^2)^{\infty}$
coming from the natural identification $\D=]1/4,3/4[\times]1/2,1[\cong]0,1[^2$, which is obtained by rescaling
and translating. Moreover we choose any diffeomorphism $\mrS'\cong\mrS$ that restricts to the identity
on all $\U_i$'s and $\V_i$'s, and give $C_{m-p}(\S')^{\infty}$ the cell structure of $C_{m-p}(\S)^{\infty}$.

Recall that $\cms^{\infty}$ can be filtered according to the norm of cells: a cell $e^{\tup}$ associated with
the tuple $\tup=(l,\ux,\uu,\uv)$ has norm $\sum_{i=1}^l x_i$, and the norm is weakly decreasing along boundaries.
In the previous section we just considered
the associated filtration of the reduced chain complex $\tCh_*(\cms^{\infty})$, whereas now we
consider the closed subcomplex $F_p\cms^{\infty}\subset\cms^{\infty}$, which is the union
of all cells of norm $\leq p$.

The crucial observation is that $\mu^{\infty}$ restricts to a cellular map 
\[
F_p\cms^{\infty}\to \pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}.
\]

To see this, fix a tuple $\tilde{\tup}=(\tilde{l}, \tilde{\ux},\tilde{\uu},\tilde{\uv})$ of norm $\tilde{p}\leq p$
and of dimension $\tilde{l}+m$, and
consider the open cell
cell $e^{\tilde{\tup}}\subset F_p\cms^{\infty}$.

If $\tilde{p}<p$, then
$e^{\tilde{\tup}}\cap (C_p(\D)\times C_{m-p}(\S'))$ is empty. If $\tilde{p}=p$, then
\[
e^{\tilde{\tup}}\cap (C_p(\D)\times C_{m-p}(\S'))=e^{\tup'}\times e^{\tup''},
\]
where $\tup'=(\tilde{l},\tilde{\ux})$ and $\tup''=(0,\tilde{\uu},\tilde{\uv})$.

Therefore
$\mu^{\infty}(e^{\tilde{\tup}})$ is $\set{\infty}$ in the first case, and in the second case it
is contained in the union $\set{\infty}\cup e^{\tup'}\times e^{\tup''}$, which is also
contained in the $(\tilde{l}+m)$-skeleton of $\pa{C_p(\D)\times C_{m-p}(\S')}^{\infty}$.

Consider now the generalised symmetric chain $\kappa(p',\ualpha',\uu',\uv')$ representing
a class in $\tH_{m+l}(\cms^{\infty})=H^{m-l}(\cms)$, with $\ualpha'=(\alpha'_j)_{j\geq 0}$,
$\uu'=(u_1,\dots,u'_g)$ and $\uv'=(v'_1,\dots,v'_g)$; in particular $l=\sum_{j\geq 0}\alpha'_j$.
Suppose moreover $p'\leq p$.

If $p'<p$, the previous argument shows that $\mu^{\infty}_*\pa{\kappa(p',\ualpha',\uu',\uv')}=0$
in the reduced cellular chain complex,
and in particular the corresponding homology class is mapped to zero.

Suppose now $p'=p$: then the previous argument shows that
the homology class $[\kappa(p',\ualpha',\uu',\uv')]\in \tH_{m+l}(\cms^{\infty})$
is mapped along $\mu^{\infty}_*$ to the class
\[
[\kappa(\ualpha')]\otimes [\kappa(\uu',\uv')]\in\tH\pa{C_p(\D)^{\infty}\wedge C_{m-p}(\S')^{\infty}}.
\]
Indeed each tuple $\tup$ in the cycle $\kappa(p',\ualpha',\uu',\uv')|$ is mapped by $\mu^{\infty}_*$
to a corresponding pair of tuples $\tup'\otimes\tup''$ in the cycle $\kappa(\ualpha')\otimes \kappa(\uu',\uv')$,
so even at the level of chains we have
\[
\mu^{\infty}_*\pa{\kappa(p',\ualpha',\uu',\uv')}=\kappa(\ualpha')\otimes \kappa(\uu',\uv').
\]

We can now compute the algebraic intersection of $\mu_*([a]\otimes [b])$ with
the cohomology class $[\kappa(p',\ualpha',\uu',\uv')]$ as the algebraic intersection between
$[a]\otimes [b]$ and $\mu^{\infty}_*\pa{[\kappa(p',\ualpha',\uu',\uv')]}$.

For $p'<p$ the previous argument show that this intersection is zero.

For $p'=p$ the intersection between
$[a]\otimes [b]$ and $\mu^{\infty}_*([\kappa(p,\ualpha',\uu',\uv')])=[\kappa(\ualpha')]\otimes [\kappa(\uu',\uv')]$
is $1\in\Z_2$ exactly when $\ualpha=\ualpha'$, $\uu=\uu'$ and $\uv=\uv'$; otherwise it is $0$.

To finish the proof we consider the collection of all strings of the form
\[
\pa{p,\ualpha=(\alpha_j)_{j\geq 0},\uu=(u_1,\dots,u_g),\uv=(v_1,\dots,v_g)}
\]
satisfying $l=\sum_j\alpha_j$,$p=\sum_j \alpha_j2^j$ and $(m-p)=\sum_i(u_i+v_i)$;
we choose a total order on the set of these strings,
such that the parameter $p$ is weakly increasing along this order; we associate
to each string its corresponding class in $H_{m-l}(\cms)$ of the form $\mu_*([a]\otimes [b])$ and its
corresponding class $[\kappa(p,\ualpha',\uu',\uv')]\in\tH_{m+l}(\cms^{\infty})$.

Then the matrix of algebraic intersections between these two sets of
classes is an upper-triangular matrix
with $1$'s on the diagonal, and in particular
it is invertible. This shows that the set of classes of the form $\mu_*([a]\otimes [b])$
is a basis for $H_{m-l}(\cms)$.
\end{proof}

One could expect that the basis given by classes of the form $[a]\otimes [b]\in H_{m-l}(\cms)$
is also \emph{dual} to the basis of classes $[\kappa(p,\ualpha,\uu,\uv)]\in \tH_{m+l}(\cms^{\infty})$, i.e., the
matrix considered in the end of the previous proof is not only upper-triangular but also
diagonal. This is however not true, as the following example shows.

Let $g=1$, $m=2$, $p=1$, $p'=2$ and consider the classes $[a]=\epsilon\in H_0(C_1(\D))$,
$[b]=\u_1\in \H=H_1(C_1(\S'))$. Moreover let the generalised symmetric chain
$\kappa(p',\ualpha',\uu',\uv')$ be defined by
$\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\alpha'_1=1$ and all other $\alpha_j=0$, $\uu'=(u'_1=0)$
and $\uv'=(v'_1=0)$.

Represent $[a]$ by a point in $a\in\D$, for example
the point $(1/2,3/4)$; represent $[b]$ by a simple closed curve $b\subset\S'$
that intersects only once, transversely, the vertical segment passing through $a$, i.e.
$\set{1/2}\times]0,1[$. See Figure \ref{fig:counterexample}.

\begin{figure}[ht]\centering
 \includegraphics[scale=0.6]{counterexample.pdf}
 \caption{$\mu_*(a\otimes b)$ and $e^{\tup'}$ intersect once, transversely inside $C_2(\S)$.}
\label{fig:counterexample}
\end{figure}

Then the cycle $\kappa(p',\ualpha',\uu',\uv')$ consists uniquely of one tuple
\[
\tup'=(1,\ux'=(x'_1=2),\uu'=(u'_1=0),\uv'=(v'_1=0)).
\]
The corresponding cell $e^{\tup'}$ intersects once, transversely,
the cycle $\mu_*(a\otimes b)$, which is represented by the curve of configurations of two
points in $\mrS$, one of which is fixed at $a$ whereas the other runs along $b$:
there is exactly one position on $b$ lying under $a$.

Hence the algebraic intersection between these two classes is $1$, and since $p'>p$
this is an entry strictly above the diagonal in the matrix considered in the proof
of Lemma \ref{lem:oplussplitting}.

The proof of Theorem \ref{thm:Hbms*as*ggrep} can be easily generalised to surfaces
with more than one boundary curve. Let $\Sigma_{g,n}$ be a surface of genus
$g$ with $n\geq 1$ parametrised boundary curves and let $\Gamma_{g,n}$ be
the group of connected components of the topological group
$\Diff(\Sigma_{g,n};\partial\Sigma_{g,n})$:
then there is an isomorphism of bigraded $\Z_2$-representations
\begin{equation}
\label{eq:sgn}
\bigoplus_{m\geq 0}H_*\pa{C_m(\Sigma_{g,n})}\simeq \Z_2\left[Q^j\epsilon\,|\, j\geq 0\right]\otimes\Sym_{\bullet}(H_1(\Sigma_{g,n})),
\end{equation}

where the action of $\Gamma_{g,n}$ on the right-hand side is induced by the natural
action on $H_1(\Sigma_{g,n})$.

For $n\geq 2$ the intersection form on the vector space $H_1(\Sigma_{g,n})$
is \emph{degenerate}, but it is still invariant under the action of
$\Gamma_{g,n}$, so there is still a map from $\Gamma_{g,n}$ to the subgroup
of $GL_{2g+n-1}(\Z_2)$ fixing this bilinear form, and in this sense we can say
that the representation in \eqref{eq:sgn} is \emph{symplectic}.

The proof of the isomorphism \eqref{eq:sgn} is almost verbatim the same; the main difference is in the construction of the
model $\T(\Sigma_{g,n})$ for $\mathring{\Sigma}_{g,n}$: we divide the
vertical segments $\set{0,1}\times[0,1]\subset[0,1]^2$
into $2g+n-1$ equal parts, that we call $I_i^l$ and $I_i^r$ according to their order;
we identify, for each $i>2g$, the interval $I_i^l$ with the interval $I_i^r$;
the other couples of intervals, yielding the genus, are identified just as before.

One can further generalise to non-orientable surfaces with non-empty boundary: it suffices,
in the above construction, to glue some of the intervals $I_i^l$ and $I_i^r$ reversing their
orientation. We leave all details of these generalisations to the interested reader.

\section{Proof of Theorem \ref{thm:main}}
\label{sec:sseq}
We will prove Theorem \ref{thm:main} by induction on $m$.
The case $m=0$ is trivial.

For all $m\geq 0$ we let $E(m)$ be the Leray-Serre spectral sequence associated with
the bundle \eqref{eq:BirmanbundleD}: its second page has the form
\[
 E(m)^2_{k,q}=H_k(B\Diff(\S;\partial\S\cup\D);H_q(\cms))=H_k(\gg;H_q(\cms)).
\]
From Theorem \ref{thm:Hbms*as*ggrep} we know that this spectral sequence is concentrated
on the rows $q=0,\dots, m$.

We want to prove the vanishing of all differentials appearing in the pages
$E(m)^r$ with $r\geq 2$; the $r$-th differential takes the form
\[
 \partial_r\colon E(m)^r_{k,q}\to E(m)^r_{k-r,q+r-1}.
\]
In particular any differential $\partial_r$ exiting from the row $q=m$ is trivial,
because it lands in a higher, hence trivial row.

Fix now $q=m-l<m$, in particular $l\geq 1$; by Theorem \ref{thm:Hbms*as*ggrep}, and
in particular by Lemma \ref{lem:oplussplitting}, we have a splitting of $H_k(\gg;H_{m-l}(\cms))$ as
\begin{equation}
\label{eq:Hksplitting}
 \bigoplus_{p=l}^m H_k\pa{\gg;\mu_*\pa{H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S'))}}.
\end{equation} 
We fix now $l\leq p\leq m$ and show the vanishing of all differentials $\partial_r$ exiting from the
summand with label $p$ in the previous equation.

Consider the map $\mu^{\cF}$ from Definition \ref{defn:muF} as a map of
bundles over the space $B\Diff(\S;\partial\S\cup\D)$:
\[
 \mu^{\cF}\colon C_p(\D)\times C_{m-p}(\cF_{\S'})\to C_m(\cF_{\S,\D}).
\]
Note that the first bundle $C_p(\D)\times C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$
is the product of the space $C_p(\D)$ with the bundle $C_{m-p}(\cF_{\S'})\to B\Diff(\S;\partial\S\cup\D)$;
therefore the spectral sequence associated with $C_p(\D)\times C_{m-p}(\cF_{\S'})$ is isomorphic,
from the second page on, to the tensor product
of $H_*(C_p(\D))$ and the spectral sequence associated with the bundle $C_{m-p}(\cF_{\S'})$; the
latter spectral sequence is isomorphic, in our notation, to the spectral sequence $E(m-p)$.
In particular $\mu^{\cF}$ induces a map of spectral
sequences
\[
\mu^{\cF}_*\colon H_*(C_p(\D))\otimes E(m-p)\to E(m);
\]
that in the second page, on the $(m-l)$-th row and $k$-th column, restricts to
the inclusion of one of the direct summands in equation \eqref{eq:Hksplitting}:
\[
H_k\pa{\gg;\mu_*(H_{p-l}(C_p(\D))\otimes H_{m-p}(C_{m-p}(\S')))}\subset H_k\pa{\gg;H_{m-l}(\cms)}.
\]

In particular if we prove the vanishing of all differentials $\partial_r$
in the first spectral sequence, then also all differentials
$\partial_r$ exiting from this direct summand in the
second spectral sequence $E(m)$ must vanish.

The differentials in the spectral sequence $H_*(C_p(\D))\otimes E(m-p)$
are obtained by tensoring the identity of $H_*(C_p(\D))$ with the differentials
of the spectral sequence $E(m-p)$; as $p\geq l\geq 1$ we know by inductive hypothesis
that the latter vanish. Theorem \ref{thm:main} is proved.

One can generalise Theorem \ref{thm:main} to orientable or non-orientable
surfaces with non-empty boundary, following the generalisation
of Theorem \ref{thm:Hbms*as*ggrep} discussed at the end of Section \ref{sec:Actiongg}.

\section{Comparison with the work of B\"{o}digheimer and Tillmann}
\label{sec:comparison}
In this section we will compare Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep}
with the results in \cite{BoT}. In particular we consider the following two theorems, that
we formulate in an equivalent, but different way as in \cite{BoT}, fitting in our
framework. In the entire section homology is taken with coefficients in $\Z_2$ unless explicitly
stated otherwise.
\begin{thm}[Corollary 1.2 in \cite{BoT}]
\label{thm:BoTone}
Let $\F$ be a field, and let $m\geq 0$, $g\geq 0$. Then the following graded vector spaces are isomorphic in degrees $*\leq \frac 23 g-\frac 23$
\[
 H_*\pa{\ggm;\F}\cong H_*\pa{\gg;\F}\otimes_{\F} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}.
\]
Here each variable $x_1,\dots,x_m$ has degree 2 and the symmetric group $\mathfrak{S}_m$ acts on the polynomial
ring $\F[x_1,\dots,x_m]$ by permuting the variables.
\end{thm}
We point out that the range of degrees $*\leq \frac 23 g-\frac 23$ is the stable range for the homology $H_*(\gg;\F)$
of the mapping class group, see \cite{Harer} for the original stability theorem and \cite{Boldsen, ORW:resolutions_homstab} for the improved
stability range.
\begin{thm}[Theorem 1.3 in \cite{BoT}]
\label{thm:BoTtwo}
For all $m\geq 0$ the map $\mu^{\cF}\colon C_1(\D)\times C_m(\cF_{\S'})\to C_{m+1}(\cF_{\S,\D})$ from Definition \ref{defn:muF}
induces a split-injective map in homology
\[
 \mu^{\cF}_*\colon H_*\pa{C_m(\cF_{\S'})}\cong H_0\pa{C_1(\D)}\otimes_{\F} H_*\pa{C_m(\cF_{\S'})}\hookrightarrow H_*\pa{C_{m+1}(\cF_{\S,\D})}.
\]
In other words, the inclusion of groups $\ggm\hookrightarrow\gg^{m+1}$ given by adding a puncture near the boundary induces a split
injective map $H_*(\ggm;\F)\hookrightarrow H_*(\gg^{m+1};\F)$.
\end{thm}

We first reformulate Theorems \ref{thm:main} and \ref{thm:Hbms*as*ggrep} in a convenient way.
\begin{defn}
 \label{defn:Cbullet}
 We introduce some abbreviations for the following disjoint unions:
 \[
 C_{\bullet}(\D)=\coprod_{m\geq 0}C_m(\D);
 \]
 \[
 C_{\bullet}(\S)=\coprod_{m\geq 0}C_m(\S); 
 \]
 \[
 C_{\bullet}(\cF_{\S,\D})=\coprod_{m\geq 0}C_m(\cF_{\S,\D}).
 \]
\end{defn}
Then $C_{\bullet}(\D)$ is a (homotopy associative) topological monoid, with associated
Pontryagin ring $H_*\pa{C_{\bullet}(\D)}\cong\Z_2[Q^j\epsilon|j\geq 0]$ (see Equation \ref{eq:Cohen}).
This is a bigraded ring, where the two gradings are the homological degree and the weight.

The maps $\mu$ from Definition \ref{defn:muF} make $C_{\bullet}(\S)$ into a (homotopy associative)
module over the monoid $C_{\bullet}(\D)$; correspondingly $H_*\pa{C_{\bullet}(\S)}$ is a bigraded
module over $H_*\pa{C_{\bullet}(\D)}$. The structure of this module is described by the following
reformulation of Theorem \ref{thm:Hbms*as*ggrep}, see in particular the proof of Lemma \ref{lem:oplussplitting}.

\begin{thm}
 \label{thm:firstreformulation}
 There is an isomorphism of $\gg$-representations in (bigraded) modules over the ring $H_*\pa{C_{\bullet}(\D)}$
 \[
  H_*\pa{C_{\bullet}(\S)}\cong   H_*\pa{C_{\bullet}(\D)}\otimes\Sym_{\bullet}(\H).
 \]
\end{thm}

Similarly, Theorem \ref{thm:main} can be reformulated as follows, considering at the same time all values of $m\geq 0$.
\begin{thm}
 \label{thm:mainreformulation}
 There is an isomorphism of (bigraded) modules over $H_*\pa{C_{\bullet}(\D)}$
 \[
 H_*\pa{C_{\bullet}(\cF_{\S,\D})}\cong   H_*\pa{C_{\bullet}(\D)}\otimes H_*\pa{\gg;\Sym_{\bullet}(\H)}.
 \] 
\end{thm}
The proof follows from the arguments used in Section \ref{sec:sseq}: we have actually shown that the direct
sum of the spectral sequences $\bigoplus_{m\geq 0} E(m)$ is itself a free module over $H_*\pa{C_{\bullet}(\D)}$
on the second page, with the same description as above.

By virtue of homological stability in $g$ for the sequences of groups $\pa{\gg}_{g\geq 0}$ and 
$\pa{\ggm}_{g\geq 0}$, Theorem \ref{thm:BoTone} can be equivalently rephrased as an isomorphism of graded vector spaces
\[
  H_*\pa{\Gamma_{\infty,1}^m;\F}\cong H_*\pa{\Gamma_{\infty,1};\F}\otimes_{\F} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}.
\]
Here $\Gamma_{\infty,1}=\colim_{g\to\infty}\gg$ and $\Gamma_{\infty,1}^m=\colim_{g\to\infty}\ggm$.

Let $\Gamma_{\infty,1}^{\bullet}$ denote the disjoint union $\coprod_{m\geq 0}\Gamma_{\infty,1}^m$, which we
consider as a groupoid. Then Theorem \ref{thm:BoTone} is equivalent to the isomorphism of bigraded vector spaces
\[
  H_*\pa{\Gamma_{\infty,1}^{\bullet};\F}\cong H_*\pa{\Gamma_{\infty,1};\F}\otimes_{\F} \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}}.
\]
We consider $\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\F[x_1,\dots,x_m]}$ as a free module over the Pontryagin ring
$H_*(\fS_{\bullet};\F)$, where $\fS_{\bullet}=\coprod_{m\geq 0}\fS_m$ and the Pontryagin product is induced by
the natural maps of groups $\fS_m\times\fS_{m'}\to\fS_{m+m'}$. This module structure is induced by the natural maps
\[
H_*\pa{\fS_m;\F[x_1,\dots,x_m]}\otimes_{\F} H_*\pa{\fS_{m'};\F}\to  H_*\pa{\fS_m\times\fS_{m'};\F[x_1,\dots,x_m]\otimes_{\F}\F} \to
\]
\[
\to H_*\pa{\fS_{m+m'};\F[x_1,\dots,x_{m+m'}]},
\]
where the first arrow is given by the homology cross product and the second by the aforementioned inclusion
$\fS_m\times\fS_{m'}\to\fS_{m+m'}$ together with the (equivariant) inclusion
of coefficients
\[
\F[x_1,\dots,x_m]\otimes_{\F}\F=\F[x_1,\dots,x_m]\subset \F[x_1,\dots,x_{m+m'}].
\]
From now on we assume $\F=\Z_2$.
Let $\beta_m=\beta_m(\D)$ denote the braid group on $m$ strands of the disc, and let $\beta_{\bullet}=\coprod_{m\geq 0}\beta_m$.
Then the natural projections $\beta_m\to\fS_m$ induce an \emph{inclusion} of Pontryagin rings
\[
 H_*(C_{\bullet}(\D))\cong H_*(\beta_{\bullet})\hookrightarrow H_*(\fS_{\bullet}).
\]
See again the appendix to \cite[Chap.III]{CLM} for a computation of $H_*(\beta_{\bullet})$,
and \cite{Nakaoka:infinite} or \cite{GiustiSalvatoreSinha} for a computation of $H_*(\fS_{\bullet})$.

Hence $\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}$ becomes a free module over $H_*(C_{\bullet}(\D))$.
Theorem \ref{thm:mainreformulation} gives then the following isomorphism, which turns out to be an isomorphism of free
$H_*(C_{\bullet}(\D))$-modules.
\begin{equation}
 \label{eq:doubleiso}
\begin{split}
  H_*\pa{\Gamma_{\infty,1}^{\bullet}} & \cong H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H_{\infty})}\\
  & \cong H_*\pa{\Gamma_{\infty,1}}\otimes \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}}.
  \end{split}
\end{equation}
Here $\H_{\infty}=\colim_{g\to\infty} H_1(\Sigma_{g,1})$ is the first homology of the surface of infinite genus with one
boundary component.

Recall that $H_*\pa{\Gamma_{\infty,1}}$ is also a Pontryagin ring: there are natural maps of groups
$\Gamma_{g,1}\times\Gamma_{g',1}\to\Gamma_{g+g',1}$ inducing a Pontryagin product on the homology 
$H_*\pa{\Gamma_{\infty,1}}=\colim_{g\to\infty} H_*\pa{\gg}$.

Equation \ref{eq:doubleiso} gives
an isomorphism of modules over the ring
$H_*\pa{\Gamma_{\infty,1}}\otimes H_*(C_{\bullet}(\D))$. The action of the ring $H_*\pa{\Gamma_{\infty,1}}$
on $H_*\pa{\Gamma_{\infty,1}^{\bullet}}$ is induced on the colimit by 
the maps of groups $\gg\times\Gamma_{g',1}^m\to\Gamma_{g+g',1}^m$.
The corresponding action of $H_*\pa{\Gamma_{\infty,1}}$ on $H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$
comes from the action on $ H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$ which is induced on the colimit by the natural maps
\[
 H_*(\gg)\otimes H_*(\Gamma_{g',1};\Sym_m(H_1(\Sigma_{g',1})))\to H_*(\Gamma_{g+g',1};\Sym_m(H_1(\Sigma_{g+g',1}))).
\]

% Note also that $H_*\pa{\Gamma_{\infty,1}}\subset H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H)}$ in a natural way.
The right-hand side $H_*\pa{\Gamma_{\infty,1}}\otimes \pa{\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}}$
in Equation \ref{eq:doubleiso} is a free module over the ring $H_*\pa{\Gamma_{\infty,1}}\otimes H_*(C_{\bullet}(\D))$.
As a consequence we have that $H_*\pa{\Gamma_{\infty,1};\Sym_{\bullet}(\H_{\infty})}$ is a free module
over $H_*\pa{\Gamma_{\infty,1}}$; this last statement is of independent interest, and regards only the homology
of the group $\Gamma_{\infty,1}$ with twisted coefficients in $\Sym_{\bullet}(\H_{\infty})$.

% Both modules are free over the ring $H_*\pa{\Gamma_{\infty,1}}$, so we obtain an isomorphism of $H_*(C_{\bullet}(\D))$-modules
% \[
%  H_*(C_{\bullet}(\D))\otimes H_*\pa{\Gamma_{\infty,1};\Sym_{>0}(\H)}\cong\bigoplus_{m\geq 0} H_*\pa{\mathfrak{S}_m;\Z_2[x_1,\dots,x_m]}.
% \]
% Note also that the right-hand side of Equation \ref{eq:doubleiso} does not depend on the stable mapping class group $\Gamma_{\infty,1}$.
% Here $\Sym_{>0}(\H)=\bigoplus_{m\geq 1}\Sym_m(\H)$.

We turn now to Theorem \ref{thm:BoTtwo}, and for the moment we assume again that $\F$ is a generic field and that the genus $g$
is fixed and finite. Considering
all values of $m$ at the same time, we can equivalently write
\begin{equation}
\label{eq:epsilonsplitting}
 H_*\pa{C_{\bullet}(\cF_{\S,\D});\F}\simeq \F[\epsilon]\otimes_{\F}\pa{\bigoplus_{m\geq 0} H_*\pa{C_m(\cF_{\S,\D}),C_{m-1}(\cF_{\S,\D})};\F}.
 \end{equation}

Here we use the convention $C_{-1}(\cF_{\S,\D})=\emptyset$, and we regard $C_{m-1}(\cF_{\S,\D})$ as a subspace of $C_m(\cF_{\S,\D})$ by using the map
$\mu^{\cF}$ in the statement of Theorem \ref{thm:BoTtwo}, with first input any fixed point $*\in C_1(\D)$. The class
$\epsilon$ is the canonical generator of $H_0(C_1(\D);\F)$.

In the case $\F=\Z_2$, Theorem \ref{thm:mainreformulation} improves Equation \ref{eq:epsilonsplitting} by exhibiting
$ H_*\pa{C_{\bullet}(\cF_{\S,\D})}$ as a free module over the ring $H_*(C_{\bullet}(\D))\cong\Z_2[Q^j\epsilon|j\geq 0]$,
rather than over its subring $\Z_2[\epsilon]$.

\section{A rational counterexample}
\label{sec:rational}
In this section we prove that a statement as in Theorem \ref{thm:Hbms*as*ggrep}
cannot hold if we consider homology with coefficients in $\Q$. We do not know
if the analogue of Theorem \ref{thm:main} holds in homology with coefficients in $\Q$
or in fields of odd characteristic.

We point out that $H_*(C_m(\Sigma_{g,1});\Q)$ has been computed \emph{as a bigraded $\Q$-vector space}
by B\"odigheimer, Cohen and Milgram in \cite{BCM}, and more recently by Knudson in \cite{Knudson}. A description of
these homology groups as a
$\gg$-representation seems to be still missing in the literature.

We will prove the following theorem:
\begin{thm}
 \label{thm:counterexample}
 Let $g\geq 2$ and $m\geq 2$; then $H_2(C_2(\Sigma_{g,1});\Q)$ is not a symplectic
 representation of $\gg$.
\end{thm}
\begin{proof}
 Let again $\S=\Sigma_{g,1}$. We will use the following strategy:
 \begin{itemize}
  \item we define a homology class $[a]\in H_2(C_2(\S);\Q)$ represented by a cycle $a$;
  \item we prove that $[a]\neq 0$ by computing the algebraic intersection of
  $[a]$ with a homology class
  in $\tH_2(\cms^{\infty};\Q)$: note that the manifold
  $\cms$ is orientable, hence Poincaré-Lefschetz duality holds
  also with rational coefficients;
  \item we define another homology class $[b]\in H_2(\cms;\Q)$ and show that
  $[b]$ is mapped to $[b]+2[a]$ by some element in the Torelli group $\mathcal{I}_{g,1}$.
 \end{itemize}
Recall that the Torelli group $\mathcal{I}_{g,1}$ is the kernel of the surjective homomorphism
$\gg\to Sp_{2g}(\Z)$ induced by the action of $\gg$ on $H_1\pa{\sg;\Z}$; if an element of the Torelli group acts non-trivially on
some class in $H_2(C_2(\S);\Q)$, the latter cannot be a symplectic representation of $\gg$.

We consider again $\T(\S)$ as model for $\mrS$. Let $c$ be an simple closed curve
representing the homology class $\u_1$, and assume that $c$ intersects $\U_1$ once
transversely and is disjoint from all other $\U_i$'s and from all $\V_i$'s. Let $c'$
be a parallel copy of $c$.

We consider the torus $a=c\times c'$ of configurations in $C_2(\S)$ having one point lying on $c$ and
one lying on $c'$; we let $[a]\in H_2(C_2(\S);\Q)$ be the fundamental class of $a$ associated with one
orientation of $a$.

Let $\tup=(0,\uu,\uv)$ with $\uu=(2,0)$ and $\uv=(0,0)$.
Then the map
\[
\Phi_{\tup}\colon (\Delta^{\tup},\partial\Delta^{\tup})\to (C_2(\S)^{\infty},\infty)
\]
maps the fundamental class in $H_2(\Delta^{\tup},\partial\Delta^{\tup};\Q)$ to a
class in $\tH_2(C_2(\S)^{\infty};\Q)$, that we call $[\kappa(0,\uu,\uv)]_{\Q}$.

The cell $e^{\tup}$ intersects once, transversely the torus $a$, therefore the
algebraic intersection $[a]\cap[\kappa(0,\uu,\uv)]_{\Q}$ is $\pm 1$,
where the signs depends on how we have chosen orientations on $a$, $e^{\tup}$ and $C_2(\S)$ itself.
In particular $[a]\neq 0$.

The action of $\gg$ on isotopy classes of non-separating simple closed curves is transitive, so we can repeat
the construction of the torus $a$ with any other copy of parallel, non-separating simple closed curves
$c$ and $c'$, and the resulting class $[a]$ will always be non-trivial in $H_2(C_2(\S);\Q)$.

See Figure \ref{fig:rational} to visualize the following discussion.
Let $d$ and $d'$ be disjoint, non-separating simple closed curves such that
the following holds: if we cut $\S$ along $d$ and $d'$ we separate $\S$ into two pieces, one of
which is a subsurface $\tilde{\S}\simeq\Sigma_{1,2}$ of genus 1, with boundary
curves $d$ and $d'$. Here we use that the genus of $\S$ is at least 2.

Suppose now that $c$ is a non-separating simple closed curve in $\tilde{\S}$,
and let $c',c''$ be two parallel copies of $c$ in $\Sigma_{1,2}$, one on each side of a small tubular
neighborhood of $c$.
Then $d,d',c',c''$ are the boundary of a subsurface $\check{\S}\simeq\Sigma_{0,4}\subset\tilde{\S}$.
Orient all curves $d,d',c,c',c''$ in such a way that the following equalities hold:
\begin{itemize}
 \item $[d]=[d']\in H_1(\S;\Q)$, as witnessed by the subsurface $\tilde{\S}$;
 \item $[c]=[c']=[c'']\in H_1(\S;\Q)$;
 \item $[d]-[d']+[c']-[c'']=0\in H_1(\S\setminus c;\Q)$, as witnessed by the subsurface $\check{\S}$.
\end{itemize}

\begin{figure}[ht]\centering
 \includegraphics[scale=1.0]{rational.pdf}
 \caption{The curves $d,d',c,c',c'',\gamma,\gamma''$ and the subsurfaces $\tilde{\S},\check{\S}$.}
\label{fig:rational}
\end{figure}

The four tori $c\times c'$, $c\times c''$, $c\times d$ and $c\times d'$ are contained in $C_2(\S)$
and the equality
\[
 [c\times d]-[c\times d']+[c\times c']-[c\times c'']=0\in H_2(C_2(\S))
\]
holds, as witnessed by the homology $c\times \check{\S}\subset C_2(\S)$.

Moreover the classes $[c\times c']$ and $-[c\times c'']$ are \emph{equal}:
indeed there is an isotopy of $\S$ mapping $c$ to $c'$ and $c''$ to $c$,
as oriented curves; the class $[c\times c'']$ is mapped to the class $[c'\times c]=-[c\times c']$.

Let again $[a]=[c\times c']$ and let $[b]=[c\times d]$: then the class $[b']=[c\times d']$
is equal to $[b]+2[a]\in H_2(C_2(\S);\Q)$.

Consider now an element of the Torelli group that fixes $c$ and maps $d$ to $d'$ preserving
the orientation, for example the bounding pair $D_{\gamma}\circ D_{\gamma'}^{-1}$, where
$\gamma$ and $\gamma'$ are represented in Figure \ref{fig:rational}:
then the class $b$ is mapped to $b'=b+2a\neq b$.
\end{proof}

The previous proof works verbatim if we replace $\Q$ by any field of odd characteristic. Moreover
all the arguments in the previous proof can be adapted to show that $H_*(F_2(\S);\mathbb{F})$ is
not a symplectic representation of $\gg$ (see Definition \ref{defn:cms}), where
$\mathbb{F}$ is \emph{any} field, including $\Z_2$.

Therefore, at least for orientable surfaces $\sg$ of genus $g\geq 2$,
being a symplectic representation of the mapping class group seems to be a peculiarity
of the homology of \emph{unordered} configuration space and of the characteristic 2.

\bibliography{Bibliography.bib}{}
\bibliographystyle{plain}

\end{document}

### 2020-05-01 slides/bianchi_slides.tex
Final version ready to appear in JTop

\documentclass[10pt]{beamer}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,amscd,amsfonts}
\usepackage{xcolor}
\usepackage{graphicx}
%\usepackage{eulervm}
\usepackage{subfigure}
\usepackage{overpic}

\usepackage{epsfig}
\usepackage{color}

% Aspetto di beamer
\usetheme{CambridgeUS}
\usecolortheme{giove}

\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\title[Splitting of homology of mapping class group]{Splitting of the homology of the punctured mapping class group}
\subtitle{Young Topologists Meeting, EPFL, Lausanne}
\author{Andrea Bianchi}
\date{26th July 2019}
\institute[Universit\"{a}t Bonn]{Mathematisches Institut, Universit\"{a}t Bonn}

\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\phi}{\varphi}

\newcommand{\B}{\mathcal{B}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\ab}{\mathfrak{ab}}
\newcommand{\Sone}{\mathbb{S}^1}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\cF}{\mathcal{F}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\fM}{\mathfrak{M}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\fS}{\mathfrak{S}}
\newcommand{\C}{\mathbb{C}}

\newcommand{\D}{\mathcal{D}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\sg}{\Sigma_{g,1}}
\renewcommand{\gg}{\Gamma_{g,1}}
\newcommand{\ggm}{\Gamma_{g,1}^m}
\newcommand{\Cb}{C_{\bullet}}

\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\pa}[1]{\left(#1\right)}

\DeclareMathOperator{\Diff}{Diff}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Sp}{Sp}

\begin{document}

\maketitle

\begin{frame}{Configuration spaces of surfaces}
$\S:=\sg$ is an orientable surface of genus $g$ with one boundary
component.

\pause
The \emph{ordered} configuration space of $m$ points on $\S$ is
\[
 F_m(\S)=\set{(p_1,\dots,p_m)\in \mathrm{Int}(\S)^m \, | \, p_i\neq p_j \,\forall i\neq j}.
\]
\pause
The \emph{unordered} configuration space of $m$ points on $\S$ is
\[
 C_m(\S)=F_m(\S)/\fS_m.
\]
\pause
\vspace{-0.5cm}
\begin{figure}[h]
\subfigure{
\includegraphics[width=3cm]{ordered.pdf}
}
\hspace{1cm}
\subfigure{
\includegraphics[width=3cm]{unordered.pdf}
}
\caption{A configuration in $F_5(\Sigma_{1,1})$ and the corresponding configuration in $C_5(\Sigma_{1,1})$.}
\end{figure}
\pause
\vspace{-0.5cm}
Want to compute $H_*(C_m(\S);\Z_2)$. From now on $H_*(-):=H_*(-;\Z_2)$. 
\end{frame}

\begin{frame}{Pontryagin rings and modules}
$\D:=\Sigma_{0,1}$ is the disc. Have product
$\mu\colon C_p(\D)\times C_q(\D)\to C_{p+q}(\D)$.
\begin{figure}[h]
\includegraphics[width=8cm]{product.pdf}
\end{figure}
\vspace{-0.3cm}
$\Cb(\D):=\coprod_{m\geq 0}C_m(\D)$ is (homotopy associative) $H$-space.
\pause

\vspace{0.1cm}
Fix $\D\hookrightarrow\S$ near $\partial\S$, and identify $\S\cong\S\setminus\D$. Have product
$\mu\colon C_p(\D)\times C_q(\textcolor{red}{\S})\cong C_p(\D)\times C_q(\textcolor{red}{\S\setminus\D})\to C_{p+q}(\S)$.
\begin{figure}[h]
\includegraphics[width=9cm]{defmu.pdf}
\end{figure}
\vspace{-0.3cm}
$\Cb(\S):=\coprod_{m\geq 0} C_m(\S)$ is (homotopy assoc.) module over $\Cb(\D)$.

\pause
\vspace{0.1cm}
Then $H_*(\Cb(\D))=\bigoplus_{m\geq 0}H_*(C_m(\D))$ is a ring and $H_*(\Cb(\S))=\bigoplus_{m\geq 0}H_*(C_m(\S))$
is a $H_*(\Cb(\D))$-module.
\end{frame}

\begin{frame}{Mapping class group}
Recall $\D\hookrightarrow\S$ near $\partial \S$. $\Diff(\S)$ is the group of
diffeomorphisms $f\colon\S\to\S$ fixing $\textcolor{red}{\partial\S\cup\D}$ pointwise.
The \emph{mapping class group} is $\gg:=\Diff(\S)/$isotopy.

\pause
\vspace{0.1cm}
For all $m\geq 0$, $\Diff(\S)$ acts on $C_m(\S)$, hence $\gg$ acts on $H_*(C_m(\S))$.

\pause
Since $\D$ is fixed by all $f\in\Diff(\S)$, the action of $\gg$ on $H_*(\Cb(\S))$ is compatible
with the action of $H_*(\Cb(\D))$ on $H_*(\Cb(\S))$.
\pause
\begin{figure}[h]
\includegraphics[scale=0.5]{mappingclass.pdf}
\vspace{1mm}
\includegraphics[scale=0.5]{DiffactionCm.pdf}
\caption{A diffeomorphism $f$ acts on the product of two configurations in $C_3(\D)$ and $C_6(\S)$.}
\end{figure}

\pause
\vspace{-0.3cm}
Want to compute $H_*(\Cb(\S))$ as a $\gg$-representation in $H_*(\Cb(\D))$-modules.
\end{frame}

\begin{frame}{First result}
Let $\H:=H_1(\S)\cong H_1(C_1(\S))\simeq\Z_2^{2g}$.
\begin{block}{Theorem A (B.'19)}
There is an isomorphism of $\gg$-representation over $H_*(\Cb(\D))$
\[
 H_*(\Cb(\S))\cong H_*(\Cb(\D))\otimes_{\Z_2}\Sym_{\bullet}(\H).
\]
$\gg$ acts trivially on $H_*(\Cb(\D))$ and \emph{symplectically} on $\Sym_{\bullet}(\H)$.
\end{block}

\pause
Recall: the action of $\gg$ on $\H$ gives a surjective homomorphism $\gg\to\Sp_{2g}(\Z_2)$.
The action of $\gg$ on $H_*(\Cb(\S))$ is \emph{symplectic}, i.e. it
factors through the group $\Sp_{2g}(\Z_2)$.

\pause
\vspace{0.2cm}
\textbf{L\"offler-Milgram '88}. $H_*(\Cb(\S))$ is a \emph{symplectic} representation of $\gg$.

\textbf{B\"odigheimer-Cohen-Taylor '89}. Have isomorphism of \emph{bigraded} $\Z_2$-vector spaces.

\textbf{Cohen '76}. $H_*(\Cb(\D))\cong\Z_2[Q^j\epsilon|j\geq 0]$ as bigraded rings, with $Q^j\epsilon$
generator of $H_{2^j-1}(C_{2^j}(\D))\simeq\Z_2$.
\end{frame}

\begin{frame}{Examples of homology classes}
Here are some homology classes in $H_*(\Cb(\D))$: $\epsilon\in H_0(C_1(\D))$, $Q\epsilon\in H_1(C_2(\D))$, $Q^2\epsilon\in H_3(C_4(\D))$,
$\epsilon^3\cdot (Q\epsilon)^2\cdot Q^2\epsilon\in H_5(C_{11}(\D))$.

\begin{figure}[h]
\includegraphics[scale=0.47]{Qje.pdf}
\end{figure}

\pause
By Theorem A, $H_m(C_m(\S))\cong \Sym_m(\H)$ for all $m\geq 0$.
\pause
Let $c_1,\dots,c_m$ be \emph{disjoint} simple closed curves
on $\S$. Then $c_1\times\dots\times c_m\subset C_m(\S)$ is an embedded torus,
and $[c_1\times\dots\times c_m]\in H_m(C_m(\S))$ corresponds to $[c_1]\cdot\dots\cdot[c_m]\in\Sym_m(\H)$.

\begin{figure}[h]
\includegraphics[scale=0.67]{torus.pdf}
\caption{Three curves $c_1,c_2,c_3$ on $\S$ and the corresponding torus $c_1\times c_2\times c_3\subset C_3(\S)$.}
\end{figure}
\end{frame}

\begin{frame}{More examples of homology classes}
If the curves $c_1,\dots,c_m$ are not disjoint, we can still associate to $[c_1]\cdot\dots\cdot[c_m]\in\Sym_m(\H)$
a class in $H_m(C_m(\S))$. Here is an example.
\begin{figure}[h]
\includegraphics[scale=0.7]{intersection.pdf}
\end{figure}
\pause
% $c_1\times c_2\not\subset C_2(\S)$, but $(c_1\times c_2)\setminus R\subset C_2(\S)$
$R\subset c_1\times c_2$
is the \emph{rectangle} where both points are in the shaded region. $(c_1\times c_2)\setminus R\subset C_2(\S)$. $\partial R=\gamma=2\gamma'=0$.
% Can close $(c_1\times c_2)\setminus R$ in $C_2(\S)$ with a M\"obius strip.
Get class in $H_2(C_2(\S))$.

It is represented by $\mathbb{T}^2\sharp\R P^2\subset C_2(\S)$, it corresponds to $[c_1]\cdot[c_2]\in\Sym_2(\H)$.

\pause
\vspace{0.2cm}
We can complicate the example. Here is $\epsilon^2\cdot Q\epsilon\otimes [c_1]\cdot[c_2]\cdot[c_3]\cdot[c_4]$.
\begin{figure}[h]
\includegraphics[scale=0.7]{complicate.pdf}
\end{figure}
\end{frame}

\begin{frame}{Second result}
Fix $P_1,\dots, P_m\in \mathrm{Int}(\S\setminus\D)$. $\Diff^m(\S)$ is the group of
diffeomorphisms of $\S$ fixing $\partial\S\cup\D$ pointwise 
and \emph{permuting} the set $\set{P_1,\dots, P_m}$.

The \emph{punctured mapping class group} is $\ggm:=\Diff^m(\S)/$isotopy.

\pause
The \emph{$m$-th braid group} of $\S$ is $\beta_m(\S)=\pi_1(C_m(\S))$.

\pause
\vspace{0.1cm}
Birman's short exact sequence relates these groups:
\[
 1\to\beta_m(\S)\to\ggm\to\gg\to 1.
\]

\pause
Want to compute $H_*(\ggm)$ using Leray-Serre spectral sequence
\[
 E(m)^2_{p,q}=H_p\pa{\gg;H_q\pa{\beta_m(\S)}} \Rightarrow H_{p+q}\pa{\ggm}.
\]
\pause
\begin{block}{Theorem B (B.'19)}
The previous spectral sequence collapses on the second page. For all $l\geq 0$
\[
 H_l\pa{\ggm}\cong \bigoplus_{p+q=l} H_p\pa{\gg;H_q\pa{\beta_m(\S)}}.
\]
\end{block}
\end{frame}

\begin{frame}{From groups to classifying spaces}
$B\gg$ is the classifying space of $\gg$.
% The space $B\Diff(\S)=E\Diff(\S)/\Diff(\S)$ is also a $K(\gg,1)$ space. Let
% $\cF(\S):=\S\times_{\Diff(\S)} E\Diff(\S)$.
Have universal $\S$-bundle
\[
\S\hookrightarrow  \cF \overset{\pi}{\to} B\gg.
\]

\pause
For all $m\geq 0$ can apply $C_m(-)$ fiberwise. Get bundle with fiber $C_m(\S)$
\[
 C_m(\S)\hookrightarrow C_m(\pi)\to B\gg
\]
\pause
The previous fiber bundle of aspherical spaces classifies Birman $1\to\beta_m(\S)\to\ggm\to\gg\to 1$.
In particular $C_m(\pi)\simeq B\ggm$.

\pause
The associated spectral sequence is again 
\[
E(m)^2_{p,q}=H_p\pa{B\gg;H_q\pa{C_m(\S)}} \Rightarrow H_{p+q}\pa{C_m(\pi)}.
\]
\pause
Can operate product $\mu\colon C_p(\D)\times C_q(\S)\to C_{p+q}(\S)$ \emph{fiberwise}: get product
\[
 \mu^{\pi}\colon C_p(\D)\times C_q(\pi)\to C_{p+q}(\pi)
\]
\pause
It is a map of bundles over $B\gg$. Get map of spectral sequences $\mu^{\pi}\colon H_*(C_p(\D))\otimes_{\Z_2} E(q)\to E(p+q)$.
Theorem B by induction.
\end{frame}

\begin{frame}{A convenient reformulation}
Let $\Cb(\pi):=\coprod_{m\geq 0} C_m(\pi)$, where
\[
 C_m(\S)\hookrightarrow C_m(\pi)\to B\gg.
\]
\pause
The extended products $ \mu^{\pi}\colon C_p(\D)\times C_q(\pi)\to C_{p+q}(\pi)$
 make $\Cb(\pi)$
into a $\Cb(\D)$-module.

\pause
\vspace{0.3cm}
Passing to homology, $H_*(\Cb(\pi))$ is a $H_*(\Cb(\D))$-module.

\pause
\begin{block}{Reformulation of Theorem B}
There is an isomorphism of $H_*(\Cb(\D))$-modules
\[
 \pa{\bigoplus_{m\geq 0}H_*(\ggm)\cong} H_*(\Cb(\pi))\cong H_*(\Cb(\D))\otimes_{\Z_2} H_*(\gg;\Sym_{\bullet}(\H)).
\]
\end{block}
\vspace{0.5cm}
\pause
\centering{\large{Thank you for your attention!}}

\end{frame}

\end{document}
